---
title: "Establishment level analysis of CIT-IRP5 files"
date: "r format(Sys.time(), "%Y%m%d %R")"
output: html_document
#### path <- "Z:/Workbenches/epguest/seiro_ito/outfiles/"; rmarkdown::render(paste0(path, "program/ReadCITIRP5Data.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
#### path <- "Z:/Workbenches/epguest/seiro_ito/"
#### pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
#### pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```






# HHI

## Condensed chunks sent to Noreen

### Read IRP5_20yr_cleaned.dta (ver 5 beta) and save icount by year

```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```


```{r read irp5 file, eval = F}
#### irp5.qs is created in IRP5Condense.rmd
library(qs)
library(data.table)
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
```

```{r plot number of employees 2009-2015}
library(qs)
library(data.table)
library(ggplot2)
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
geovars <- grepout("^bus.*geo$", colnames(irp5))
irp5[, Num := as.integer(.N), by = c(geovars, "taxrefno", "taxyear")]
irp5ForPlot <- unique(irp5[taxyear >= 2012 & taxyear <= 2015, 
  c(geovars, "taxyear", "taxrefno", "Num"), with = F])
irp5ForPlot[, taxyear := as.character(taxyear)]
g1 <- ggplot(data = irp5ForPlot[busmainplc_geo != "", ], aes(x = Num, group = taxyear, 
  colour = taxyear, fill = taxyear)) +
  geom_density(alpha = .3) + 
  scale_color_viridis_d("Tax year") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits = c(1, 100), trans = "log10") +
  theme_bw()
 g1
g1 + facet_grid(taxyear ~ .)
```

# Location information granularity checks

We need to know how granular the location information is. Based on granularity, we decide at what level one can use the sampled firm.
```{r location granularity}
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
#### Location granularity indicator
irp5[, LocGranular := "none"]
for (yr in 9:22) {
	irp5[Txrf %in% Txrf
		[
		  busprov_geo     != "" & 
		  busdistmuni_geo != "" & 
		  buslocmuni_geo  != "" & 
		  busmainplc_geo  != "" & 
      taxyear == 2000 + yr
		] &
    taxyear == 2000 + yr, 
	  LocGranular := GeoLevel[4]]
	irp5[Txrf %in% Txrf
		[
		  busprov_geo     != "" & 
		  busdistmuni_geo != "" & 
		  buslocmuni_geo  != "" & 
		  busmainplc_geo  == "" &
      taxyear == 2000 + yr
		] &
    taxyear == 2000 + yr, 
	  LocGranular := GeoLevel[3]]
	irp5[Txrf %in% Txrf
		[
		  busprov_geo     != "" & 
		  busdistmuni_geo != "" & 
		  buslocmuni_geo  == "" & 
		  busmainplc_geo  == "" &
      taxyear == 2000 + yr
		] &
    taxyear == 2000 + yr, 
	  LocGranular := GeoLevel[2]]
	irp5[Txrf %in% Txrf
		[
		  busprov_geo     != "" & 
		  busdistmuni_geo == "" & 
		  buslocmuni_geo  == "" & 
		  busmainplc_geo  == "" &
      taxyear == 2000 + yr
		] &
    taxyear == 2000 + yr, 
	  LocGranular := GeoLevel[1]]
}
irp5[, LocGranular := factor(LocGranular, levels = c(GeoLevel, "none"))]
setnames(irp5, "LocGranular", "LocGranularFirm")
irp5[, LocGranularJob := "none"]
irp5[busprov_geo != "", LocGranularJob := GeoLevel[1]]
irp5[busdistmuni_geo != "", LocGranularJob := GeoLevel[2]]
irp5[buslocmuni_geo != "", LocGranularJob := GeoLevel[3]]
irp5[busmainplc_geo != "", LocGranularJob := GeoLevel[4]]

#### Location granularity due to establishment clustering
#### When location info is missing in some rows, one can allocate workers
#### only up to location level where all establishments are clustered
for (mm in 1:length(GeoLevel)) {
  irp5[, (GeoLevel[mm]) := get(geovars[mm])]
  irp5[get(geovars[mm]) == "", (GeoLevel[mm]) := NA]
}
#### Number of unique entries at each level 
#### Subtract 1 because NA is counted as 1 (all NA entries are MainPlNum == 1L)
irp5[, MainPlNum := as.integer(length(unique(Mai)))-1L, by = .(Txrf, taxyear)]
irp5[, LocMunNum := as.integer(length(unique(Loc)))-1L, by = .(Txrf, taxyear)]
irp5[, DisMunNum := as.integer(length(unique(Dis)))-1L, by = .(Txrf, taxyear)]
irp5[, ProvNum   := as.integer(length(unique(Prv)))-1L, by = .(Txrf, taxyear)]
irp5[, MainPlNA := as.integer(all(is.na(Mai))), by = .(Txrf, taxyear)]
irp5[, LocMunNA := as.integer(all(is.na(Loc))), by = .(Txrf, taxyear)]
irp5[, DisMunNA := as.integer(all(is.na(Dis))), by = .(Txrf, taxyear)]
irp5[, ProvNA := as.integer(all(is.na(Prv))), by = .(Txrf, taxyear)]

irp5[, CommonLocality := "none"]
irp5[
      ProvNum == 1L & 
    DisMunNum == 1L & 
    LocMunNum == 1L & 
    MainPlNum == 1L &
    MainPlNA  == 0L, 
  CommonLocality := GeoLevel[4]]
irp5[
    ProvNum   == 1L & 
    DisMunNum == 1L & 
    LocMunNum == 1L & 
    MainPlNum != 1L & 
    LocMunNA  == 0L, 
  CommonLocality := GeoLevel[3]]
irp5[
    ProvNum   == 1L & 
    DisMunNum == 1L & 
    LocMunNum != 1L & 
    MainPlNum != 1L &
    DisMunNA  == 0L, 
  CommonLocality := GeoLevel[2]]
irp5[
    ProvNum   == 1L & 
    DisMunNum != 1L & 
    LocMunNum != 1L & 
    MainPlNum != 1L & 
    ProvNA  == 0L, 
  CommonLocality := GeoLevel[1]]
irp5[, CommonLocality := factor(CommonLocality, levels = c(GeoLevel, "none"))]
irp5[, NARowsMai := as.integer(sum(is.na(Mai))), by = .(Txrf, taxyear)]
irp5[, NARowsLoc := as.integer(sum(is.na(Loc))), by = .(Txrf, taxyear)]
irp5[, NARowsDis := as.integer(sum(is.na(Dis))), by = .(Txrf, taxyear)]
irp5[, NARowsPrv := as.integer(sum(is.na(Prv))), by = .(Txrf, taxyear)]
irp5[, TotalRowsInTxrf := as.integer(.N), by = .(Txrf, taxyear)]
qsave(irp5, paste0(pathdata, "irp5_LocationGranularity.qs"), nthreads = 13)

unique(irp5[Txrf %in% Txrf[grepl("^M", LocGranularFirm) | grepl("^M", CommonLocality)][1]
  & taxyear == 2009, 
  c("Txrf", "UID", geovars, grepout("Num|Rows|Gran|Commo", colnames(irp5))), with = F])

#### LocGranularFirm: the most granular location info at firm level
#### LocGranularJob : the most granular location info at each job level
#### MainPlNum      : the number of unique entries at main place level for firm-taxyear
#### NARowsMai      : the number of NA entries at main place level for firm-taxyear
#### CommonLocality : Lowest area level common among all the jobs for firm-taxyear
####   E.g., if all jobs are in Stellenbosch (main place), it is main place
####   If there are jobs attached to a branch in Capetown, then it is province
####   If location info is missing in some rows, then "none"

#### We know that rows with no location info can be allocated to the area level by:
####   MainPlNum == 1L                    ==> main place
####   MainPlNum >  1L & LocMunNum == 1L  ==> local muni
####   LocMunNum >  1L & DisMunNum == 1L  ==> district muni
####   DisMunNum >  1L & ProvNum == 1L    ==> province 

#### Panel of various levels
#### Main place panel: 
####  a. location entries are complete & location is granular down to main place level, or,
####        NARowsMai == 0L & LocGranularFirm == Mai
####  b. location == NA for some rows & uniquely common locality level is main place
####        NARowsMai > 0L  & MainPlNum == 1L
#### Local municipality panel: (addition to main place panel)
####  a. location entries are complete down to local municipality & location is granular down to local municipality, or,
####        NARowsLoc == 0L & LocGranularFirm == Loc
####  b. location == NA for some rows & uniquely common locality level is local municipality
####        NARowsLoc > 0L  & LocMunNum == 1L
#### District municipality panel: (addition to local municipality panel)
####  a. location entries are complete down to district municipality & location is granular down to district municipality, or,
####        NARowsDis == 0L & LocGranularFirm == Dis
####  b. location == NA for some rows & uniquely common locality level is district municipality
####        NARowsDis > 0L  & DisMunNum == 1L
#### Province panel: (addition to district municipality panel)
####  a. location entries are complete down to province & location is granular down to province, or,
####        NARowsPrv == 0L & LocGranularFirm == Prv
####  b. location == NA for some rows & uniquely common locality level is province
####        NARowsPrv > 0L  & ProvNum == 1L
iiM <- irp5[,
  NARowsMai == 0L & grepl("^M", LocGranularFirm) | 
  NARowsMai >  0L & MainPlNum == 1L]
iiL <- !iiM & irp5[, 
  NARowsLoc == 0L & grepl("^L", LocGranularFirm) | 
  NARowsLoc >  0L & LocMunNum == 1L]
iiD <- !iiM & !iiL & irp5[, 
  NARowsDis == 0L & grepl("^D", LocGranularFirm) | 
  NARowsDis >  0L & DisMunNum == 1L]
iiP <- !iiM & !iiL & !iiD & irp5[, 
  NARowsPrv == 0L & grepl("^P", LocGranularFirm) | 
  NARowsPrv >  0L & ProvNum   == 1L] 
nrow(irp5M <- irp5[iiM, ])
nrow(irp5L <- irp5[iiL, ])
nrow(irp5D <- irp5[iiD, ])
nrow(irp5P <- irp5[iiP, ])
qsave(irp5M, paste0(pathdata, "irp5M.qs"), nthreads = 13)
qsave(irp5L, paste0(pathdata, "irp5L.qs"), nthreads = 13)
qsave(irp5D, paste0(pathdata, "irp5D.qs"), nthreads = 13)
qsave(irp5P, paste0(pathdata, "irp5P.qs"), nthreads = 13)

irp5L[Txrf %in% unique(Txrf)[4]] #### 5 rows missing location, 2 branches in same loc muni
irp5D[Txrf %in% unique(Txrf)[5]] #### 1 row missing location, 2 branches in same dist muni
irp5P[Txrf %in% unique(Txrf)[5]] #### 1 row missing location, 3 branches in same prov


#### Missingness in location information
#### 1. Maximum sample size data sets
####   a. Main Place data = firms with information down to Main Place (irp5M) 
####   b. Loc Muni data   = firms with information down to Loc Muni   (irp5L) + a
####   c. Dis Muni data   = firms with information up to Dist Muni    (irp5D) + a + b
####   d. Prov data       = firms with information up to Prov         (irp5P) + a + b + c
#### 2. Same sample data sets (irp5M) 
####   a. Main Place data = firms with information up to Main Place
####   b. Loc Muni data   = a
####   c. Dis Muni data   = a
####   d. Prov data       = a
```


# Compute HHI


We use main place level sample ~irp5M~ by following the "same sample data sets"


```{r hhi}
#### code "hhi"
library(qs)
library(data.table)
irp5 <- qread(paste0(pathdata, "irp5M.qs"), nthreads = 13)
geovars <- grepout("^bus.*geo$", colnames(irp5))
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
for (yr in 12:22) {
  ipyr <- irp5[taxyear == 2000+yr, ]
  LShare <- NULL
  for (g in 1:length(GeoLevel)) {
    print(GeoLevel[g])
    byvar <- 	c("busprov_geo", "busdistmuni_geo", 
  			           "buslocmuni_geo", "busmainplc_geo")[1:g]
    ipGeo = copy(ipyr)
  	ipGeo[, WorkersAtEstab := as.integer(.N), by = c(byvar, "Txrf")]
	  #### Some firms have missing location information in some (but not all) rows
	  #### Cannot allocate 8 workers with missing location info between branches
	  if (yr == 12) print(
	    ipGeo[Txrf %in% unique(Txrf[busmainplc_geo != ""])[2], 
	      c(geovars, "Txrf", "payereferenceno", "UID",  "dateofbirth", 
		    "periodemployedfrom", "WorkersAtEstab", "kerr_income"), with = F])
	  ipGeo[, WorkersInMarket := as.integer(.N), by = byvar]
    ipGeo[, Share := round(WorkersAtEstab/WorkersInMarket, 8)]
	  ipGeo[busmainplc_geo != "", ]
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    ipGeo[, WorkersAtEstabNoGov := 0L]
    ipGeo[!grepl("gov", Entity), WorkersAtEstabNoGov := WorkersAtEstab]
	  #### Below gives private firm totals for rows of private firms
    ipGeo[, WorkersInMarketNoGov := as.integer(.N), by = c(byvar, "Entity")]
    ipGeo[, ShareNoGov := round(WorkersAtEstabNoGov/WorkersInMarketNoGov, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
	  #### employment shares
    lshare <- unique(ipGeo[, c(
	    byvar, "Entity", "Txrf",  "taxrefno",
        "AreaLevel", grepout("^Wo|^Sha", colnames(ipGeo))
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(Share^(2), na.rm = T), by = byvar]
    lshare[, nHHI := (HHI-1/WorkersInMarket)/(1-1/WorkersInMarket)]
    lshare[, HHIG := sum(ShareNoGov^(2), na.rm = T), by = byvar]
    lshare[, nHHIG := (HHIG-1/WorkersInMarketNoGov)/(1-1/WorkersInMarketNoGov)]
    LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  }
  LShare[, AreaLevel := factor(AreaLevel, levels = GeoLevel)]
  LShare[, (geovars) := lapply(.SD, factor), .SDcols = geovars]
  setkey(LShare, AreaLevel, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"), nthreads = 13)
}
#### LShare:  Txrf, Share, ShareNoGov, HHI, HHINoGov, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

library(ggplot2)
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[is.na(busdistmuni_geo) & grepl('D', AreaLevel), ]
LShare[grepl('M', AreaLevel) & WorkersInMarket < 500, ]
g12 <- ggplot(data = LShare, aes(x = HHI, colour = AreaLevel, fill = AreaLevel)) +
  geom_density(alpha = .2) + 
  scale_fill_manual(values = rep("blue", 4)) +
  scale_color_manual(values = rep("blue", 4)) +
  scale_x_continuous(trans = "log10") +
  theme(legend.position = "none") +
  facet_grid(AreaLevel ~ .)
LShare <- qread(paste0(pathdata, "ShareHHI13.qs"))
g13 <- ggplot(data = LShare, aes(x = HHI, colour = AreaLevel, fill = AreaLevel)) +
  geom_density(alpha = .2) + 
  scale_fill_manual(values = rep("blue", 4)) +
  scale_color_manual(values = rep("blue", 4)) +
  scale_x_continuous(trans = "log10") +
  theme(legend.position = "none") +
  facet_grid(AreaLevel ~ .)
ggsave(
  paste0(pathdata, "HHI2012byAreaLevel.jpg")
  , g12, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
ggsave(
  paste0(pathdata, "HHI2013byAreaLevel.jpg")
  , g13, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
#### HHI in 2012 show almost no observations at large HHI levels, in contrast to
#### 2013 values. We should not use 2012 HHIs.
```

```{r create LShare panel}
LS <- NULL
for (yr in 12:22) {
  ls <- qread(paste0(pathdata, "ShareHHI", yr, ".qs"))
  ls[, taxyear := 2000+yr]
  LS <- rbindlist(list(LS, ls))
}
qsave(LS, paste0(pathdata, "LShareHHI.qs"), nthreads = 13)
```

```{r }

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    WorkersAtEstab, WorkersAtEstabNoGov, 
  	WorkersInMarket, WorkersInMarketNoGov, Share, 
	  HHI)]



ipyr <- qread(paste0(pathdata, "irp512.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)]
print(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
summary(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)])
print(LShare[TotalNeEmp <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
print(LShare[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)

```

# Fraction affected

1. Use IRP5 2012 (`ipyrc`) to count total number of workers, number of affected, FA = affected/total
1. Merge branch weight data () with FAdata
    a) Distribute workers and number of affected workers proportionately to branch weights
    a) Compute FA per branch


```{r define fraction affected}
#### code "compute FA"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### ipyrsClean.qs=irp12.qs+dropped anomalous taxrefno
irp5M <- qread(paste0(pathdata, "irp5M.qs"), nthreads = 13)
ipyrc0 <- irp5M[taxyear == 2012, ]
ipyrc0[taxrefno == "AACBTBBAJZ", ]
#### Choose only unique uid-income rows
ipyrc <- ipyrc0[!duplicated(
    ipyrc0[, .(busmainplc_geo, taxrefno, uid, kerr_income #, ptrs_income
    )]
  ), ]
#### Compute job duration and monthly income
####   (DateEnd-DateStart2012)/30.5
####   a3601_income/DurationMonth
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
ipyrc[, DateStart := as.IDate(periodemployedfrom)]
ipyrc[, DateEnd := as.IDate(periodemployedto)]
ipyrc[, DateStart2012 := DateStart]
ipyrc[DateStart < as.IDate("2011/03/01"), 
  DateStart2012 := as.IDate("2011/03/01")]
ipyrc[, DurationMonth := round((DateEnd-DateStart2012)/30.5, 3)]
#### ipyrc[, IncomeMonth := round(a3601_income/DurationMonth, 0)]
ipyrc[, IncomeMonth := round(kerr_income/DurationMonth, 0)]
setkey(ipyrc, busmainplc_geo, taxrefno, uid, DateStart)
ipyrc[, DateStart2 := shift(DateStart, n = 1L, type = "lead"), 
  by = .(Txrf, taxyear, uid)]
ipyrc[, DateEnd2 := shift(DateEnd, n = 1L, type = "lead"), 
  by = .(Txrf, taxyear, uid)]
#### Double work duration: DateEnd - DateStart2
####  <----job1---->
####           <----job2---->
ipyrc[, DJobDurationMonth := round((DateEnd - DateStart2)/30.5, 3)]
ipyrc[, TDurationMonth := 
  round((DateEnd2 - DateStart2012)/30.5, 3)]
#### If job2 ends before job1 ends: DateEnd > DateEnd2
####  <----------job1---------->
####           <----job2---->
####   (DateEnd2 - DateStart2)/30.5 (double work duration)
####   (DateEnd-DateStart2012)/30.5 (job1+job2 duration)
ipyrc[DateEnd > DateEnd2, 
  DJobDurationMonth := round((DateEnd2 - DateStart2)/30.5, 3)]
ipyrc[DateEnd > DateEnd2, TDurationMonth := 
  round((DateEnd-DateStart2012)/30.5, 3)]
ipyrc[, TDurationMonth := TDurationMonth[!is.na(TDurationMonth)][1], 
  by = .(Txrf, taxyear, uid)]
ipyrc[, DoubleJobRatio := round((DJobDurationMonth/TDurationMonth)*100, 2)]
ipyrc[, DoubleJobRatio := DoubleJobRatio[!is.na(DoubleJobRatio)][1],
  by = .(Txrf, taxyear, uid)]
ipyrc[, NumJobsPerWorker := .N, by = .(Txrf, taxyear, uid)]
ipyrc[, TIncomeMonth := sum(IncomeMonth, na.rm = T), 
  by = .(Txrf, taxyear, uid)]
#### check data
ipyrc[busmainplc_geo != "" & Txrf != "" & NumJobsPerWorker > 1, 
  .(busmainplc_geo, Txrf, uid, NumJobsPerWorker, 
    DateStart, DateEnd,
    DateStart2, DateEnd2, DurationMonth, 
    DJobDurationMonth, TDurationMonth, DoubleJobRatio,
    #a3601_income, 
    kerr_income, IncomeMonth, TIncomeMonth)]
print(ipyrc[NumJobsPerWorker == 5, 
  .(busmainplc_geo, Txrf, uid, NumJobsPerWorker, 
    DateStart, DateEnd,
    DateStart2, DateEnd2, DurationMonth, 
    DJobDurationMonth, TDurationMonth, DoubleJobRatio,
    a3601_income, IncomeMonth, TIncomeMonth)], topn = 30)
ipyrc[busmainplc_geo != "" & Txrf != "", 
  .(busmainplc_geo, taxrefno, DateStart, DateEnd, 
    DurationMonth, a3601_income, IncomeMonth)]
#### Find sub MW workers and define fraction affected
#### We can use IncomeMonth or TIncomeMonth
####  PropToMW: Proportion to MW line
####  SubMW: PropToMW < 1 & IncomeMonth > 0
####  NumSubMW: number of entries with SubMW == 1L
####  Employees: number of entries with IncomeMonth > 0
####  FA: NumSubMW/Employees
ipyrc[, PropToMW := round(IncomeMonth/2274, 6)]
ipyrc[, SubMW := as.integer(PropToMW < 1 & IncomeMonth > 0)]
ipyrc[, NumSubMW := uniqueN(UID[SubMW==1L]), 
  by = .(Txrf, taxyear)]
ipyrc[, Employees := uniqueN(UID[IncomeMonth > 0]), 
  by = .(Txrf, taxyear)]
ipyrc[, FA := round(NumSubMW/Employees, 6)]
ipyrc[grepl("^BCGACCKJZC", Txrf), 
  .(busmainplc_geo, 
    Txrf, uid, UID, DateStart, DurationMonth, 
    IncomeMonth, SubMW, Employees, FA)]
#### Vary sub MW workers by DoubleJobRatio X >= 50
####  SubMWX: PropToMW < 1 & TIncomeMonth > 0 & DoubleJobRatio >= X
####  NumSubMWX: number of entries with SubMW == 1L & DoubleJobRatio >= X
####  EmployeesX: number of entries with IncomeMonth > 0 & DoubleJobRatio >= X
####  FA25: NumSubMWX/EmployeesX with X=25
####  FA50: NumSubMWX/EmployeesX with X=50
####  FA75: NumSubMWX/EmployeesX with X=75
ipyrc[, PropToMWX := round(TIncomeMonth/2274, 6)]
  by = .(Txrf, taxyear)]
ipyrc[, EmployeesX := uniqueN(uid[TIncomeMonth > 0]), 
  by = .(Txrf, taxyear)]
XThreshold <- 50
for (XThreshold in c(25, 50, 75)) {
  ipyrc[, SubMWX := as.integer(PropToMWX < 1 & TIncomeMonth > 0)]
  ##### PropToMWX <= 1L ==> subMW == 1L
  ####   If DoubleJobRatio  < X & PropToMWX  > 1L ==> subMW == 1L
  ####   If DoubleJobRatio >= X & PropToMWX  > 1L ==> subMW == 0L
  #####     PropToMWX <= 1L ==> subMW == 1L
  #### DoubleJobRatio := round((DJobDurationMonth/TDurationMonth)*100
  ipyrc[DoubleJobRatio >= XThreshold & PropToMWX > 1L, 
    SubMWX := 0L]
  ipyrc[, NumSubMWX := uniqueN(uid[SubMWX==1L]), 
    by = .(Txrf, taxyear)]
  ipyrc[, FAx := round(NumSubMWX/EmployeesX, 6)]
  setnames(ipyrc, c("FAx", "NumSubMWX"), 
    paste0(c("FA", "NumSubMW"), XThreshold))
}
ipyrc[, FinYr2012 := 1L]
ipyrc[DateStart > as.IDate("2012/02/29") | 
  DateEnd > as.IDate("2012/02/29"), FinYr2012 := 0L]
#### Select unique entries for FinYr2012 for each branch
FAdata <- unique(ipyrc[FinYr2012 == 1L, .(
    busprov_geo, 
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno, Txrf, payereferenceno, UID,
    Employees, NumSubMW, PropToMW, SubMW, 
    FA, FA25, FA50, FA75
  )])
setnames(FAdata, c("FA", "FA25", "FA50", "FA75"),
    paste0("FA.", c("orig", 25, 50, 75))
)
summary(FAdata)
FAdata[, Size := "micro"]
FAdata[Employees > 1 & Employees <= 10, Size := "small"]
FAdata[Employees > 10 & Employees <= 50, Size := "medium"]
FAdata[Employees > 50, Size := "large"]
FAdata[is.na(Employees) | Employees == 0, Size := "NA"]
#### Number of affected workers in aggregates (2012)
ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
  , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
  )
  ][, .(
    TotalEmployees, TotalSubMWWorkers, 
    AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
    TotalFirmsWithEmployees, TotalAffectedFirms,
    AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
  )]
ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
  , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
  ), 
  by = .(taxyear, busprov_geo)
  ][, .(
    busprov_geo, TotalEmployees, TotalSubMWWorkers, 
    AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
    TotalFirmsWithEmployees, TotalAffectedFirms,
    AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
  )]
#### Number of employees per branch
(EmpSubMW <- ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  ), 
  by = .(
    taxyear, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, taxrefno, payereferenceno)
  ])
table(EmpSubMW[, TotalSubMWWorkers])
library(ggplot2)
library(cowplot)
g1 <- ggplot(data = EmpSubMW, 
    aes(x=TotalSubMWWorkers)) +
  geom_histogram(binwidth=1)
g2 <- g1 + geom_histogram(bins=20) +
  scale_x_continuous(transform = "log10") 
g1 <- g1 + scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  scale_y_continuous(limits = c(0, 100000))
gs <- plot_grid(
  g1, g2,
  labels = c('natural, up to 10 people', 'log10'),
  align="hv"
)
gs

#### For FA=0, do they employ?
table(EmpSubMW[TotalSubMWWorkers == 0L, TotalEmployees])

qsave(FAdata, paste0(pathdata, "FAdata.qs"), nthreads = 13)
qsave(ipyrc, paste0(pathdata, "ipyrc.qs"), nthreads = 13)
```
```{r plot FAs}
#### possible duplicates with different local municipality names
ii0 <- FAdata[, buslocmuni_geo != "" & busdistmuni_geo != ""
   & busmainplc_geo != ""]
ii1 <- duplicated(FAdata[, .(
   busmainplc_geo, Txrf
 )])
ii2 <- duplicated(FAdata[, .(
   busmainplc_geo, Txrf
 )], fromLast = T)
table(data.table(nonNA=ii0, duplicates=ii1))
FAdata[ii0 & (ii1 | ii2), ]
table(duplicated(FAdata[ii0 & !ii1, ]))
#### plot sensitivity to XThreshold (double job months cutoff)
library(ggplot2)
library(cowplot)
g1 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.orig)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "blue")
g2 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.25)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "red")
g3 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.50)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "green")
g4 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.75)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "orange")
gs <- plot_grid(
  g1, g3, g2, g4,
  labels = c('original', '50', '25', '75'),
  align="hv"
)
ggsave(
  paste0(pathdata, "FAByDoubleJobRatio.jpg"),
  , gs, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
#### Below gives an error: idvar need to specify unique rows
#### FAdataL <- reshape(FAdata[ii0 & !ii1, ], 
####   direction = "long", 
####   idvar = c("busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", 
####     "Txrf", "payereferenceno"), 
####   varying = grepout("^FA", colnames(FAdata))
#### )
gsz <- ggplot(data = FAdata, 
  aes(x = FA.orig)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = c("red", "blue"))
gsz <- gsz + facet_wrap(Size ~ .)
ggsave(
   paste0(pathdata, "FAByFirmSize.jpg"),
  , gsz, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```

# Other

Get observation pattern: ObPattern

```{r create ObPattern, eval = F}
#### icount data is created in chunk "read irp5v5beta"
Icount <- qread(paste0(pathdata, "icount08.qs"), nthreads = 8)
Icount[, Taxyear := 2008]
for (yr in 9:20) {
  if (yr < 10) yr <- paste0("0", yr)
  icount <- qread(paste0(pathdata, "icount", yr, ".qs"), nthreads = 8)
  icount[, Taxyear := 2000+as.numeric(yr)]
  Icount <- rbindlist(list(Icount, icount))
}
qsave(Icount, paste0(pathdata, "Icount.qs"), nthreads = 8)
setkey(Icount, taxrefno, Taxyear)
Icount[, Ob := 1L]
Icount[is.na(Num), Ob := 2L]
Icount <- Icount[grepl("A", natureofperson), ]
Icount <- Icount[busmainplc_geo != "", ]
IcountW <- reshape(Icount, direction = "wide", 
  idvar = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", "taxrefno", "natureofperson"),
  timevar = "Taxyear", v.names= c("Num", "Ob"))

grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]

cols <- grepout("Ob", colnames(IcountW))
IcountW[, (cols) := lapply(.SD, function(x) {x[is.na(x)] <- 0L; x}), .SDcols = cols]
IcountW[, ObPattern := paste(.SD, sep = "", collapse = ""), .SDcols = cols, by = .(taxrefno)]
IcountW[, ObPattern := factor(ObPattern)]
```

Postal codes in stata

```{verbatim}
. use "Z:\Master Data\IRP5\Job level\v5\beta\aux_postal_codes_2008_2023_e6_v2.dta" , replace
. by AddrPostalCode, sort: gen nvals = _n == 1
. count if nvals
  68,140
```

  

