
#  Read IRP5_20yr_cleaned.dta (ver 5 beta) and save by year

```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
pathdataIRP <- "U:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```
Rread irp5 data by year

* read from stata dta
* dropped if
    * natureofperson != "A"
    * taxrefno == ""
#### Note: many empty taxrefno in irp5 (2012) before dropping
#### table(ipyr[, taxrefno == ""])
####   FALSE    TRUE 
#### 8586068 1407896 


```{r read irp5v5beta best geo and create irp5yr files, eval = F}
library(qs)
library(data.table)
library(readstata13)
for (yr in 9:23) {
  if (yr < 10) yr <- paste0("0", yr)
  ## Below code is run only once: Start ##
  # irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, "_cleaned.dta"))
  # ipyr <- data.table(irpyr)
  # rm(irpyr)
  # qsave(ipyr, paste0(pathdata, "irp", yr, ".qs"), nthreads = 13)
  # irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, 
  #   "_cleaned_best_geodata.dta"))
  # ipb <- data.table(irpyr)
  # qsave(ipb, paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  # # merge best geocode to irpYR.qs
  # ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), nthreads = 13)
  # ipb <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  # geonames <- grepout("^bus.*geo$", colnames(ipyr))
  # setnames(ipyr, geonames, paste0("OLD", geonames))
  # ipbyr <- ipyr[ipb, , on = grepout("irp.*aid", colnames(ipb))]
  # nrow(ipbyr[OLDbusmainplc_geo == "" & busmainplc_geo != "", ])
  # qsave(ipbyr, paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## Below code is run only once: End ##
  ipyr <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = 8)
  ## Keep only nature of person is "an individual"
  ipyr <- ipyr[grepl("A", natureofperson), ]
  ## Drop obs with missing taxrefno
  ipyr <- ipyr[taxrefno != "", ]
  ## Note that there are header lines with taxrefno = "NULL", but they can be dropped later
  ## Num: Number of employees in a firm
  ipyr[, Num := as.integer(.N), by = .(taxrefno, 
    busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo)]
  ipyr[, Num := 0L]
  ipyr[, Nationality := "sa"]
  ipyr[passportno != "", Nationality := "non-sa"]
  ipyr[, Nationality := factor(Nationality, levels = c("sa", "non-sa"))]
  irp5yr <- ipyr[, .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, irp5it3aid, taxrefno, UID, Nationality, 
    payereferenceno, gender, dateofbirth, taxyear, 
    periodemployedfrom, periodemployedto,
    #### a3601_income: basic salary
    Num, kerr_income, ptrs_income, a3601_income)]
  qsave(irp5yr, paste0(pathdata, "irp5", yr, ".qs"), nthreads = 13)
}
```

# Copy 2013 location info &rarr; NA in 2012 location info

* use taxrefno-uid match
* this assumes no transfer between establishments in a firm in 2012-2013
* in reality, there are transfers &rarr; measurement errors in location
* but the proportion of transfers must be small 
* benefits of location info outweigh the costs

If there are workers working for multiple establishments ... no problem 

* taxrefno-uid uniquely identifies matches (even if there is a transfer, it does not show up as an error, we will never know)
* base of join = 2012 (fix rows of 2012), and find matches from 2013
* see example shown in codes, 8 entries in 2012, 2013

Note: Firms with anomalous repetetive uids (AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ) are dropped


```{r copy location information from 2013... to 2012}
#### code "copy location"
#### Read data
#### nthreds = 16 at NT-SDF, = 8 with my laptop (less 1, keep 1 for other computations)
ipyr <- qread(paste0(pathdata, "irp512.qs"), nthreads = 13)
ipyr[, uid := gsub(" +$", "", UID)]
rm(ipyrs)
#### uid anomalous entries: AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ
#### these are 0.035% of total number of rows
#### one large employer (with 4632 jobs) contributes 
#### DO NOT use these uids when copying location from future data
dupuid <- ipyr[
    duplicated(uid) | duplicated(uid, fromLast=T), 
    .(taxrefno, uid=uid)
  ]
dupuid[, n := .N, by = .(taxrefno, uid)]
dupuid <- unique(dupuid[order(n, decreasing = T), ])
dupuid[n > 10, ]
print(unique(dupuid[, uid])[1:30], quote = F)
#### repetitive: [A-Z]*8
repetetive <- dupuid[grepl("(.)\\1{7,}", uid), uid]
tbr <- table(repetetive)
tbr <- tbr[order(tbr, decreasing=T)]
repetetive <- unique(repetetive)
#### ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
####   .(taxrefno, uid, busprov_geo, bprov.13, 
####   busmainplc_geo, bmain.13)][uid %in% repetetive, ][
####   order(uid), ]
#### firms with repetetive uids and their counts
tb <- table(ipyr[uid %in% repetetive, taxrefno])
tb <- tb[order(tb, decreasing=T)]
tb[1:20]
#### repetetive entries are concentrated in 3 firms
dupuid[1:20, ]
dupuid[taxrefno %in% names(tb)[1:3], ][1:20, ]
#### firms with 800 same uids: using same uid for all workers
ipyr[uid %in% dupuid[1, uid],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)]
#### same repetetive entries (AAAAA...) are found across firms
ipyr[uid %in% names(tbr)[1:20],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ]
#### Drop firms with repetetive entries if they are above 10
ii <- 10
ipyr[taxrefno %in% names(tb)[tb==ii] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ]
(dropthese <- ipyr[taxrefno %in% names(tb)[tb>10] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ])
table(dropthese[, uid])
table(dropthese[, .(uid, taxrefno)])
qsave(ipyr, paste0(pathdata, "ipyrs.qs"), nthreads = 13)
ipyr <- ipyr[!(taxrefno %in% dropthese[, taxrefno]), ]
#### Below gives non-uique match error for fyr > 13
####  ipyrs <- ipyr2[ipyrs]
#### Below gives memory error for fyr > 13
####    ipyrs <- merge(ipyrs, ipyr, by = c("taxrefno", "uid"), 
####      all.x = T, allow.cartesian = T)
#### So I will just merge 2013
#### (Redundant: ipyrs, an object created by merging 2013, 14, 15)
#### for (fyr in 13:15) {
for (fyr in 13) {
  ipyr2 <- qread(paste0(pathdata, "irp5", fyr, ".qs"), nthreads = 13)
  ipyr2 <- unique(ipyr2[, .(taxrefno, UID, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo)])
  #### future data location variables: bprov, bdist, bloc, bmain
  setnames(ipyr2, c("busprov_geo", "busdistmuni_geo", 
    "buslocmuni_geo", "busmainplc_geo"),
    c("bprov", "bdist", "bloc", "bmain"))
  ipyr2[, uid := gsub(" +$", "", UID)]
  if (fyr == 13) ipyr0 = copy(ipyr) else ipyr0 = copy(ipyrs)
  setkey(ipyr0, taxrefno, uid)
  setkey(ipyr2, taxrefno, uid)
  #### join data.tables ipyr0, ipyr2 keeping rows of ipyr0 (2012 or base)
  #### (keep the rows of ipyr0, attach the matching rows of ipyr2)
  #### "Take ipyr2, and subset it using the rows/keys in ipyr0"
  #### Join by taxrefno-uid as pivots
  ipyrs <- ipyr2[ipyr0]
  #### and copy bmain (2013) ==> busmainplc_geo (2012)
  #### Below gives multiple matches because there can be more than one
  #### establishment per firm and individual can move between them
  #### # ipyrs <- merge(ipyr0, ipyr2, by = c("taxrefno", "uid"), 
  #### #   all.x = T, allow.cartesian = T)
  cat(fyr, "\n")
  #### Empty entries that can be copied from future data
  #### province, district muni, local muni, main place
  ii.prov <- ipyrs[, 
    (busprov_geo == "" | busprov_geo == "EXCEPTION")
    & (bprov != "" & bprov != "EXCEPTION") ]
  ii.dist <- ipyrs[, 
    (busdistmuni_geo == "" | busdistmuni_geo == "EXCEPTION")
    & (bdist != "" & bdist != "EXCEPTION") ]
  ii.loc <- ipyrs[, 
    (buslocmuni_geo == "" | buslocmuni_geo == "EXCEPTION")
    & (bloc != "" & bloc != "EXCEPTION") ]
  ii.main <- ipyrs[, 
    (busmainplc_geo == "" | busmainplc_geo == "EXCEPTION")
    & (bmain != "" & bmain != "EXCEPTION") ]
  cat(fyr, "\n")
  print(table(ii.prov))
  #### dist muni == NA <==> local muni == NA
  print(table(ii.dist))
  print(table(ii.loc))
  print(table(ii.main))
  print(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
    .(taxrefno, uid, busprov_geo, bprov, 
      busdistmuni_geo, bdist,
      buslocmuni_geo, bloc,
      busmainplc_geo, bmain)])
  #### copy from future data
  if (fyr == 13) ipyrs[, c("LocInfoCopiedFrom", 
    "ProvCopied", "DisCopied", "LocCopied", "MaiCopied") := 0L]
  ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
    LocInfoCopiedFrom := as.integer(fyr)]
  ipyrs[ii.prov, c("busprov_geo",     "ProvCopied") := .(bprov, 1L)]
  ipyrs[ii.dist, c("busdistmuni_geo", "DisCopied" ) := .(bdist, 1L)]
  ipyrs[ii.loc,  c("buslocmuni_geo",  "LocCopied" ) := .(bloc,  1L)]
  ipyrs[ii.main, c("busmainplc_geo",  "MaiCopied" ) := .(bmain, 1L)]
  setnames(ipyrs,
      c("bprov", "bdist", "bloc", "bmain"), 
      paste0(c("bprov", "bdist", "bloc", "bmain"), ".", fyr)
    )
}
fyr
ipyrs[, .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)]
dupuid2 <- ipyrs[
  ii.prov | ii.dist | ii.loc | ii.main, 
    .(
      taxrefno, uid, busprov_geo, bprov.13, busmainplc_geo, bmain.13
    )
  ][
    duplicated(uid) | duplicated(uid, fromLast=T), 
  ][, 
    .(taxrefno, uid=uid)
  ]
dupuid2[, n := .N, by = uid]
dupuid2[order(uid), uid]
print(dupuid2[order(uid), uid][1000+1:100], quote = F)
unique(dupuid2[grepl("(.)\1{5,}", uid), uid])
#### Example: 4 entries out of 8 entries are copied
####   2012: All 8 entries are missing location info
####   2013: 7 out of 8 entries have location info, 4 out of 7 match with 2012
####   1 in PE of Eastern Cape, all others are around Capetown in WC
####     ==> Cannot assume this person worked only around Capetown
ii <- 10
#### J
#### X
#### C
#### X
#### B
#### B
#### T
#### G
#### B
#### K
#### C
#### T
#### K
ipyrs[uid %in% unique(dupuid2[n == 4, uid])[ii], .(taxrefno,
  busprov_geo, busmainplc_geo, ProvCopied, MaiCopied,
  uid, dateofbirth, periodemployedfrom, periodemployedto, 
  kerr_income #, ptrs_income, amt3616, source
  )]
ipyrs[uid %in% unique(dupuid2[n == 4, uid])[ii], 
  .(taxrefno, busprov_geo, busmainplc_geo, ProvCopied, MaiCopied, uid)]
ipyr2[uid %in% unique(dupuid2[taxrefno != "" & n == 4, uid])[ii], 
  .(taxrefno, bprov, bmain, uid)]
qsave(ipyrs, paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
```


```{r create long format of irp5, eval = F}
library(qs)
library(data.table)
irp5 <- NULL
for (yr in 9:22) {
  if (yr < 10) yr <- paste0("0", yr)
  if (yr == 12)  
    irp5 <- rbindlist(list(irp5, 
      qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)), use.names = T, fill = T) else 
    irp5 <- rbindlist(list(irp5, 
      qread(paste0(pathdata, "irp5", yr, ".qs"), nthreads = 13)), use.names = T, fill = T)
}
irp5[, Num := as.integer(.N), by = .(busprov_geo,
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno, taxyear)]
#### 1. Use payereferenceno to count the total
####    >>> This is a wrong way to go: payereferenceno has many misreporting <<<
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
irp5[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
irp5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
irp5[, Entity := "private"]
irp5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
irp5[, Entity := factor(Entity, levels = c("private", "gov"))]
irp5Clean <- irp5[!(taxrefno %in% dropthese[, taxrefno]), ]
qsave(irp5, paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = 13)
qsave(irp5Clean, paste0(pathdata, "irp5.qs"), nthreads = 13)

```


