---
title: "MinWage scan summary — IRP5Condense/HHI/MergeData/Impacts (Sessions 16-17 Pale Reed / Cedar Lark)"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-overflow: wrap
---

Review of the four production `.rmd` files (2026-07-02), covering fiscal-year
convention (Session 16), then a full A-D fix pass (Session 17). Unlike the
GUK scan, the underlying IRP5/CIT admin panels are not present on this
machine (server-side data), so evidence chunks below either (a) read the
**live production code** at render time to confirm what actually shipped, or
(b) run a **minimal synthetic reproduction** of the bug mechanism. Row-count
claims (9.0204%, `DESSw.qs` contents, etc.) require the next pipeline run on
the server — flagged explicitly where that applies.

```{r setup}
#| message: false
library(data.table)
pathprogram <- "c:/data/MinWageMarketPower/analysis/program/"
readfile <- function(f) readLines(paste0(pathprogram, f), warn = FALSE)
live <- function(x, from, to) paste(trimws(x[from:to]), collapse = "\n")
```

# Issue index

All 18 findings, open first. Severity: **Major** = wrong results or a
blocked run; **Minor** = performance, display, or comment-only.

## Open

### Issue 1 (Major): Finding 12 — inverted `exists()` guard back in IRP5HHI.rmd

What
:   The Session-15 fix of the inverted `exists()` cleanup guards was reverted
    in the working tree; the wrong-direction guard is live again.

Damage
:   `rm()` on an absent object errors on a standalone run, or memory is never
    freed — the live direction per variable is currently unverified.

Fix
:   Open — re-check the direction of every `exists("irp5[LDP]")` guard in the
    current file, not the inline comment ([Finding 12](#finding-12-open-live-bug-inverted-exists-guard-reintroduced-by-user-revert)).

### Issue 2 (Unknown severity): Finding 13 — `FinYr20yr` computed but never used

What
:   `FinYr20yr` flags rows whose `DateStart`/`DateEnd` contradict the
    `taxyear` label, but no count filters on it and it is excluded from
    `NeededCols`.

Damage
:   Unknown — `NumSubMW`/`Jobs`/`Employees`/`MP` counts trust the raw
    `taxyear` label; the share of mislabeled rows has not been measured.

Fix
:   Open — wire the flag into the counts or remove it
    ([Finding 13](#finding-13-open-dead-code-finyr20yr-computed-but-never-used-to-filter-counts)).

### Issue 3 (Housekeeping): Finding 14 — `eval=` flags of FA/LShare panel chunks

What
:   `{create FA panel}` / `{create LShare panel}` `eval=` flags must track
    whether the main loop currently accumulates `FAD` in RAM.

Damage
:   None yet — but the setting has flipped twice across Sessions 15-17.

Fix
:   Re-verify before every render ([Finding 14](#finding-14-config-not-a-bug-create-fa-panel-create-lshare-panel-eval-flags-need-re-checking)).

## Solved

### Issue 4 (Major): Finding 1 — NA→0L fill loop no-op (`||`) and missing paren

What
:   Scalar `||` inside `which()` collapsed the row-wise `Ob.YYYY` NA fill to
    a no-op; a missing closing paren also left the file unparseable.

Damage
:   The intended fill never fired; the file would not parse as last committed.

Fix
:   Loop replaced by `dcast(..., fill = 0L)`.

### Issue 5 (Major): Finding 2 — `EestabWith1` typo

What
:   `EstabWith1` assigned, `EestabWith1` referenced.

Damage
:   `object not found` abort on any clean run.

Fix
:   Fill rewritten as a keyed update join; no second spelling remains.

### Issue 6 (Major): Finding 3 — `paste(.SD, collapse=)` recycled one string

What
:   `collapse=` deparses all of `.SD` into one string instead of per-row
    concatenation.

Damage
:   `I`/`E`/`FObPattern` all held one recycled string.

Fix
:   `do.call(paste0, .SD)` at all 3 sites.

### Issue 7 (Major): Finding 4 — `DropThisForRev` dropped revisions, kept originals

What
:   The drop filter was evaluated globally (`by=` is a no-op on `i`) with the
    conjunct inverted (`TrueTaxYear != taxyear`).

Damage
:   Revisions dropped, superseded originals kept — panel composition wrong;
    the saved 9.0204% figure is stale.

Fix
:   Per-person update join on `RevYears`; **server pipeline re-run still
    required**.

### Issue 8 (Major): Finding 5 — grouped `.N` count immediately overwritten

What
:   `Num := 0L` ran textually after the grouped `.N` count (pattern repeated
    in ~15 files).

Damage
:   `Num` was always 0.

Fix
:   Dead init commented out; the count itself was correct and kept.

### Issue 9 (Minor): Finding 6 — `all(uid == UID)` not NA-safe

What
:   An NA in either vector makes `all()` return NA, not FALSE.

Damage
:   `if()` errors (or a branch silently misfires) when ids contain NA.

Fix
:   `isTRUE(all(...))`.

### Issue 10 (Major): Finding 7 — `stats::reshape` aborts on R 4.4

What
:   `reshape()`'s `interaction()` over 9 composite id columns
    integer-overflows with ~50M `EUIndID` levels.

Damage
:   "cannot allocate 298.0 Gb" abort — the production reshape could not
    complete at all.

Fix
:   Replaced by `dcast()` (~6s at benchmark scale).

### Issue 11 (Minor): Finding 8 — unused `setkey()` pair before an `on=` join

What
:   Two `setkey()` sorts preceded a join that already specified `on=`.

Damage
:   Two wasted 175M-row sorts.

Fix
:   Removed.

### Issue 12 (Minor): Finding 9 — `FillInLocMuni` grouped-closure fill

What
:   Per-group closure dispatch scales badly past ~1M groups.

Damage
:   Slow fill; no correctness issue.

Fix
:   Rewritten as subset → dedupe → update join.

### Issue 13 (Minor): Finding 10 — `TrueTaxYear` comment said `DateStart`

What
:   Worked-example comment was headed "DateStart" while the code anchors on
    `DateEnd`.

Damage
:   None in code — `DateEnd` is bounded inside the correct fiscal window by
    construction; the comment was misleading.

Fix
:   Comment relabelled to `DateEnd` with the reason (2026-07-03).

### Issue 14 (Minor): Finding 11 — `nHHI`/`nHHIG` NaN in one-worker markets

What
:   The normalised-HHI formula divides by `1 - 1/N`, which is 0 at `N = 1`.

Damage
:   NaN instead of NA (both drop from estimation, but the stored value was
    wrong).

Fix
:   `fifelse(WorkersInMarket > 1, ..., NA_real_)` guard.

### Issue 15 (Minor): Finding 15 — base-year display undercounted establishments

What
:   Establishments tabulated via `num == 1L`, but `num` is inherited from
    LSMa1 numbering rather than being 1 for every establishment.

Damage
:   Display table undercounted establishments per HHI base year.

Fix
:   `unique(LSMa2, by = c("geo","taxrefno"))`.

### Issue 16 (Major): Finding 16 — `ReportEveryYear` grouping omitted `busprov_geo`

What
:   Grouping did not match `EstID`'s granularity.

Damage
:   Flag computed across merged geographies → wrong values.

Fix
:   `busprov_geo` added to `by=`.

### Issue 17 (Major): Finding 17 — winsorized results written to `DESS`, not `DESSw`

What
:   The nested-list assignment targeted `DESS` instead of `DESSw`.

Damage
:   `DESSw.qs` saved as a list of NULLs; `DESS` silently clobbered with
    winsorized values.

Fix
:   Assignment target corrected; **`DESSw.qs` must be regenerated on the next
    server run**.

### Issue 18 (Minor): Finding 18 — winsorization comment stated the wrong multiple

What
:   Comment disagreed with the `> 6` (600%) threshold in code.

Damage
:   Misleading comment only.

Fix
:   Comment corrected.

# IRP5Condense.rmd

## Finding 1 (fixed, confirmed live): NA→0L fill loop retired, along with its `||` bug

What
:   Scalar `||` inside `which()` collapses the whole column to one
    `TRUE`/`FALSE` — the row-by-row `Ob.YYYY` fill never fired; a missing
    closing paren also left the file unparseable as last committed.

Damage
:   NAs never filled to `0L`; the file would not parse.

Fix
:   Session 17 replaced the loop with `dcast(..., fill = 0L)` — no loop, so
    neither defect can recur.

Live code:

```{r finding1}
gk <- readfile("IRP5Condense.rmd")
cat(live(gk, 1029, 1050), "\n")
```

```{r finding1-repro}
#### minimal reproduction of why scalar || silently no-ops a row-wise fill.
#### R >= 4.3 hard-errors on `||`/`&&` given a length>1 argument (used to
#### warn and silently take element 1) - the historical (pre-4.3) behaviour
#### is reproduced explicitly below via `[1]` to show what the fill loop
#### actually evaluated on every call.
x <- c(1L, NA, 3L, NA); y <- c(NA, 2L, NA, NA)
cat("is.na(x)[1] || is.na(y)[1]  [buggy: pre-4.3 `||` behaviour, scalar]:",
    is.na(x)[1] || is.na(y)[1], "-> which() over a length-1 logical picks at most index 1\n")
cat("which(is.na(x) | is.na(y))   [correct, vectorised]:",
    which(is.na(x) | is.na(y)), "\n")
```

Status: **fixed** (loop removed, not just patched).

## Finding 2 (fixed, confirmed live): `EestabWith1` typo — superseded by update-join rewrite

What
:   `EstabWith1` assigned, `EestabWith1` referenced.

Damage
:   `object not found` abort on any clean run.

Fix
:   Session 17 rewrote the whole fill as a keyed update join; no second
    spelling remains to diverge.

Live code:

```{r finding2}
cat(live(gk, 1139, 1144), "\n")
```

Status: **fixed** (superseded by rewrite, not a find-replace patch).

## Finding 3 (fixed, confirmed live): `paste(.SD, collapse=)` recycled one string into the Ob-pattern columns

What
:   `paste(.SD, sep=, collapse=)` collapses the whole `.SD` argument into one
    deparsed string — not a per-row concatenation.

Damage
:   All three pattern columns (`I`/`E`/`FObPattern`) held one recycled
    string.

Fix
:   `do.call(paste0, .SD)` at all 3 sites.

Live code:

```{r finding3}
cat(grep("ObPattern := do.call", gk, value = TRUE), sep = "\n")
```

```{r finding3-repro}
DT <- data.table(a = c("x","y"), b = c("1","2"), c = c("A","B"))
cat("paste(.SD, collapse='') [buggy]: ",
    DT[, paste(.SD, collapse = "")], "\n")
cat("do.call(paste0, .SD)   [fixed]: ",
    DT[, do.call(paste0, .SD)], "\n")
```

Status: **fixed**, all 3 sites (`do.call(paste0, .SD)`).

## Finding 4 (fixed, pipeline re-run required): `DropThisForRev` dropped revisions, kept superseded originals

What
:   `i` was evaluated globally (`by=` is a no-op on `i`, only groups `j`)
    and the conjunct was `TrueTaxYear != taxyear` — the opposite of the
    documented example.

Damage
:   Revisions dropped, superseded originals kept — panel composition wrong.
    The saved 9.0204% figure and the mismatch table (old L666-670/693) are
    stale: they predate this fix **and** the Session 16 end-year
    `TrueTaxYear` revert. Full mechanism, duplicate-person-year consequence,
    and cross-person contamination: `analysis/program/DataTableFixExamples.qmd` §A3.

Fix
:   Per-person update join on `RevYears` (Session 17); regenerate the stale
    figures on the next server run before quoting.

Live code, post-fix:

```{r finding4}
cat(live(gk, 705, 745), "\n")
```

Status: **fixed in code**, **not yet verified against production data**.

## Finding 5 (fixed, confirmed live): grouped `.N` count immediately overwritten

What
:   `Num := 0L` ran textually after the grouped `.N` count — the same
    pattern repeats across roughly 15 files.

Damage
:   The count was erased; `Num` always 0.

Fix
:   Comment out the init, keep the count (user's final call — the count
    itself was correct, only the ordering/dead init was wrong).

Live code:

```{r finding5}
cat(live(gk, 130, 138), "\n")
```

```{r finding5-repro}
DT <- data.table(g = c("a","a","b"))
DT[, Num := as.integer(.N), by = g]
cat("count only:\n"); print(DT)
DT[, Num := 0L] #### the bug: this line, if left in, erases the count above
cat("after the erasing line:\n"); print(DT)
```

Status: **fixed**.

## Finding 6 (fixed, confirmed live): `all(uid == UID)` not NA-safe

What
:   An NA in either id vector makes `all(uid == UID)` return NA, not FALSE.

Damage
:   `if()` errors, or the branch silently misfires, when ids contain NA.

Fix
:   `isTRUE(all(...))`.

Live code:

```{r finding6}
cat(grep("isTRUE\\(all\\(idyr", gk, value = TRUE), "\n")
```

```{r finding6-repro}
uid <- c(1, 1, NA); UID <- c(1, 1, 1)
cat("all(uid==UID)        [buggy, NA propagates]: ", all(uid == UID), "\n")
cat("isTRUE(all(uid==UID)) [fixed, NA-safe]:      ", isTRUE(all(uid == UID)), "\n")
```

Status: **fixed**.

## Finding 7 (fixed, confirmed live): `stats::reshape` aborts on R 4.4 with 9 composite id columns

What
:   `reshape()`'s `interaction()` call indexes the Cartesian product of
    every id column's factor levels — `EUIndID` alone carries ~50M levels in
    production.

Damage
:   Integer overflow → "cannot allocate 298.0 Gb" abort; the production
    reshape (old L975) could not complete on R 4.4 at all —
    correctness-blocking, not style.

Fix
:   Replaced by `dcast()` (~6s at 1.8M-row benchmark scale; repro in
    `DataTableFixExamples.qmd` §A5).

Live code:

```{r finding7}
cat(live(gk, 1045, 1050), "\n")
```

Status: **fixed**. This was a *correctness-blocking* bug, not just a style
issue — the production reshape at the old L975 would not have completed on
R 4.4 at all.

## Finding 8 (fixed, confirmed live): unused `setkey()` pair before an `on=`-specified join

What
:   Two `setkey()` sorts preceded a join that already specified `on=`.

Damage
:   Two wasted 175M-row sorts; the keys were never used.

Fix
:   `setkey()` pair removed.

Live code:

```{r finding8}
cat(live(gk, 1204, 1209), "\n")
```

Status: **fixed** — two 175M-row sorts removed; the join already specified
`on=` so the keys were never used.

## Finding 9 (fixed, confirmed live): `FillInLocMuni` grouped-closure fill rewritten as update join

What
:   `FillInLocMuni` filled via closures dispatched per group — the pattern
    the Session 13 lesson (`feedback_datatable_scale.md`) flags as scaling
    badly past ~1M groups.

Damage
:   Slow fill; no correctness issue.

Fix
:   Rewritten as subset → dedupe → update join.

Live code:

```{r finding9}
cat(live(gk, 555, 562), "\n")
```

Status: **fixed**.

## Finding 10 (resolved 2026-07-03): `TrueTaxYear` anchor — `DateEnd` confirmed correct, comment was stale

What
:   Code anchors on `DateEnd`; the worked-example comment was headed
    "DateStart".

Damage
:   None in code — `periodemployedto` (`DateEnd`) never exceeds end-February
    of the return's own filing year, so it is bounded inside the correct
    fiscal window by construction. `DateStart` is the unbounded one:
    anchoring on it would misclassify records into a past year (matching the
    pre-existing `## wrong` annotation on that line). Misleading comment
    only.

Fix
:   Comment relabelled to `DateEnd`; code untouched.

Live code, post-fix:

```{r finding10}
cat(live(gk, 671, 697), "\n")
```

```{r finding10-repro}
#### why DateStart (not DateEnd) is the one that can misfire: a filing-year
#### return can report on an employee whose employment started years ago
DateStart <- as.Date("2011-06-01")   #### employment began years earlier
DateEnd   <- as.Date("2015-01-20")   #### but this return covers Mar2014-Feb2015
yr <- function(d) year(d) + as.integer(month(d) >= 3)
cat("year(DateStart)-anchor [wrong]: ", yr(DateStart), "-> unrelated past year\n")
cat("year(DateEnd)-anchor   [correct]: ", yr(DateEnd),
    "-> matches the return's own Mar2014-Feb2015 filing year\n")
```

Status: **fixed** (comment-only; the code was already correct).

# IRP5HHI.rmd

## Finding 11 (fixed, confirmed live): `nHHI`/`nHHIG` produced `NaN` for one-worker markets

What
:   The normalised-HHI formula divides by `1 - 1/WorkersInMarket`, which is
    0 in one-worker markets.

Damage
:   `nHHI`/`nHHIG` stored `NaN` instead of `NA`.

Fix
:   `fifelse(WorkersInMarket > 1, ..., NA_real_)` guard.

Live code:

```{r finding11}
hhi <- readfile("IRP5HHI.rmd")
cat(live(hhi, 1183, 1192), "\n")
```

```{r finding11-repro}
WorkersInMarket <- c(1L, 5L); HHI <- c(1, 0.4)
cat("unguarded [buggy, NaN at N=1]: ",
    (HHI - 1/WorkersInMarket) / (1 - 1/WorkersInMarket), "\n")
cat("fifelse guard [fixed, NA]:     ",
    fifelse(WorkersInMarket > 1,
      (HHI - 1/WorkersInMarket) / (1 - 1/WorkersInMarket), NA_real_), "\n")
```

Status: **fixed**.

## Finding 12 (open, live bug): inverted `exists()` guard reintroduced by user revert

What
:   Claude fixed this in Session 15 (commit `0e8602a`); the user then
    reverted the commit in the working tree while dropping the FAD in-RAM
    accumulation, which brought the inverted guard back.

Damage
:   The guard either never frees `irp5L` or errors if the object is absent
    on a standalone run. The comment at L1121 claims the condition was
    already fixed — the claim cannot be trusted after two flips.

Fix
:   Open — re-check which direction is now live, per variable, against the
    current file rather than the comment.

Live code:

```{r finding12}
cat(grep("exists\\(\"irp5[LDP]\"\\)", hhi, value = TRUE), sep = "\n")
```

Status: **open — verify current direction is correct**, not just that a
comment claims it is.

## Finding 13 (open, dead code): `FinYr20yr` computed but never used to filter counts

What
:   `FinYr20yr` flags whether `DateStart`/`DateEnd` match the current
    `taxyear` label, but the `NumSubMW`/`Jobs`/`Employees`/`MP`-level counts
    built earlier in the same chunk never filter on it, and it is excluded
    from `NeededCols` — the flag exists but does nothing.

Damage
:   Unknown — counts trust the raw `taxyear` label uncorrected for
    forward-mislabeled rows; the share of such rows has not been measured.

Fix
:   Open — wire the flag into the counts or remove it.

Live code:

```{r finding13}
cat(live(hhi, 338, 347), "\n")
```

Status: **open**.

## Finding 14 (config, not a bug): `{create FA panel}` / `{create LShare panel}` `eval=` flags need re-checking

What
:   `{create FA panel}` `eval=` status must track whether the main loop is
    currently accumulating `FAD` in RAM or not.

Damage
:   None yet — but the setting has flipped twice across Sessions 15-17
    (user removed accumulation, then restored it), so the flag is a moving
    target.

Fix
:   Re-check before every render rather than trusting the current setting.

Status: **housekeeping — re-verify before next render**.

# IRP5MergeData.rmd

## Finding 15 (fixed, confirmed live): base-year display undercounted establishments

What
:   The display tabulated `HHIBaseYear` over rows with `num == 1L`, but
    `num` is inherited from LSMa1 numbering — `num == 1L` is not "one row
    per establishment".

Damage
:   Display table undercounted establishments per HHI base year.

Fix
:   `unique(LSMa2, by = c("geo","taxrefno"))` — one row per establishment.

Live code:

```{r finding15}
mg <- readfile("IRP5MergeData.rmd")
cat(live(mg, 293, 296), "\n")
```

```{r finding15-repro}
#### num is inherited from LSMa1 numbering, not unique per establishment
LSMa2 <- data.table(
  geo = c("a","a","b","b","b"), taxrefno = c(1,1,2,2,2),
  num = c(1,2,1,2,3), HHIBaseYear = c(2013,2013,2015,2015,2015))
cat("table(LSMa2[num==1L]) [buggy, undercounts if num!=1 rows share a base year]:\n")
print(table(LSMa2[num == 1L, HHIBaseYear]))
cat("table(unique(..., by=geo+taxrefno)) [fixed, one row per establishment]:\n")
print(table(unique(LSMa2, by = c("geo","taxrefno"))[, HHIBaseYear]))
```

Status: **fixed**.

## Finding 16 (fixed, confirmed live): `ReportEveryYear` grouping omitted `busprov_geo`

What
:   `ReportEveryYear` was grouped without `busprov_geo` — coarser than
    `EstID`'s granularity.

Damage
:   Flag computed across merged geographies → wrong values.

Fix
:   `busprov_geo` added to the `by=`.

Live code:

```{r finding16}
cat(grep("by = \\.\\(busprov_geo", mg, value = TRUE)[1:2], sep = "\n")
```

Status: **fixed** — `busprov_geo` now included, matching `EstID`'s
granularity.

# IRP5Impacts.rmd

## Finding 17 (fixed, pipeline re-run required): `DESS` overwritten instead of `DESSw` being filled

What
:   The nested-list assignment targeted `DESS` instead of `DESSw`.

Damage
:   `DESSw.qs` was saved as a list of `NULL`s; `DESS` (a separate, unrelated
    accumulator) was silently overwritten with winsorized values it
    shouldn't hold.

Fix
:   Assignment target corrected; `DESSw.qs` must be regenerated on the next
    server run.

Live code, post-fix:

```{r finding17}
imp <- readfile("IRP5Impacts.rmd")
cat(live(imp, 792, 798), "\n")
```

```{r finding17-repro}
#### minimal reproduction of the nested-list assignment slip
DESS <- list(); DESSw <- list("2010" = list())
dess <- "winsorized value"
DESS[["2010"]][["02"]] <- dess   #### buggy line as originally written
cat("DESSw after buggy assignment (still empty list):\n"); str(DESSw)
DESSw[["2010"]][["02"]] <- dess  #### fixed line
cat("DESSw after fixed assignment:\n"); str(DESSw)
```

Status: **fixed in code, not yet re-run** — `DESSw.qs` needs regenerating.
Also noted but **not fixed** (left to the user): the prose near L266-290
describes winsorization dropping changes "by more than 3000" while the
code's threshold is `abs(dJobsMP) > 2000` — text vs. code disagree and
neither was touched this session.

## Finding 18 (fixed, confirmed live): winsorization comment said the wrong multiple

What
:   The winsorization comment stated the wrong multiple.

Damage
:   Misleading comment only.

Fix
:   Comment now matches the `> 6` (600%) threshold in code.

Live code:

```{r finding18}
cat(grep("abs\\(rJobsMP\\)", imp, value = TRUE)[1:2], sep = "\n")
```

Status: **fixed** — comment now matches the `> 6` (600%) threshold in code.

# Status summary

```{r status-summary}
#| message: false
library(tinytable)
st <- data.frame(
  Finding = 1:18,
  Issue = c(
    "NA-fill / || loop", "EestabWith1 typo", "paste(.SD) ObPatterns",
    "DropThisForRev", "grouped .N overwrite", "all() NA-unsafe",
    "reshape -> dcast", "unused setkey", "FillInLocMuni fill",
    "TrueTaxYear anchor comment", "nHHI/nHHIG NaN",
    "inverted exists() guard", "FinYr20yr dead flag", "eval= flags",
    "base-year undercount", "ReportEveryYear grouping",
    "DESS/DESSw swap", "winsor comment"),
  Status = c(
    "Fixed (loop removed)", "Fixed (superseded by rewrite)", "Fixed",
    "Fixed in code - pipeline re-run required", "Fixed", "Fixed",
    "Fixed (was correctness-blocking on R 4.4)", "Fixed", "Fixed",
    "Fixed (comment-only)", "Fixed",
    "OPEN - live bug, reverted by user", "OPEN - dead flag",
    "Housekeeping", "Fixed", "Fixed",
    "Fixed in code - pipeline re-run required", "Fixed"))
tt(st) |>
  group_tt(i = list(
    "IRP5Condense.rmd"  = 1,
    "IRP5HHI.rmd"       = 11,
    "IRP5MergeData.rmd" = 15,
    "IRP5Impacts.rmd"   = 17))
```

# Next steps

1. **Server pipeline re-run is a prerequisite**, not optional — Finding 4
   (`DropThisForRev`) changes which rows survive into `irp5gir`, so every
   downstream `.qs` and the 9.0204% figure are stale until re-run.
2. Re-verify Finding 12 (`exists()` guard direction) directly in the current
   file rather than trusting the inline comment, since it has flipped twice
   via reverts.
3. Wire or remove Finding 13 (`FinYr20yr`) — currently dead weight.
4. Reconcile the Finding 17 prose/code mismatch (2000 vs "3000") — user's
   call, not touched.
