---
title: "Session 19 Moss Plover — findings explained with mock examples"
subtitle: "Companion to Session19_MinWageScanSummary.html"
date: 2026-07-07
format:
  html:
    toc: true
    toc-location: left
    number-sections: false
    embed-resources: true
engine: knitr
execute:
  eval: true
  echo: true
  warning: true
---

Written by Claude (model: **Fable 5**). Every section links to the exact chunk
in the original notebook, then reproduces the mechanism on mock data small
enough to run anywhere. All chunks below execute at render time — outputs
shown are live, not pasted.

```{r setup, message=FALSE}
library(data.table)
library(qs)
```

# A1 — `use_alt_rep` is a `qread()` argument, `qsave()` has no such thing

Original: [IRP5Condense.rmd](../../analysis/program/IRP5Condense.rmd),
13 calls, e.g. chunk `{r merge industry code}` L378, chunk
`{r create ObPattern}` L1223-1236.

The installed qs 0.27.3 signatures:

```{r}
args(qs::qsave)
args(qs::qread)
```

`use_alt_rep` exists in the package — but only in `qread()`. Passing it to
`qsave()` is an immediate error, reproduced here:

```{r}
tf <- tempfile(fileext = ".qs")
try(qsave(data.table(x = 1:3), tf, nthreads = 2, use_alt_rep = TRUE))
#### the fix: drop the argument from qsave; keep it in qread
qsave(data.table(x = 1:3), tf, nthreads = 2)
qread(tf, use_alt_rep = TRUE)
```

Rule takeaway: "argument exists somewhere in the package" is not
verification — it must belong to *that* function (`feedback_debug_rule.md`).

# B1 — `dropthese` filter narrower than its comment (user-confirmed intentional, 2026-07-07)

**Status: closed.** The user confirms the narrower selection is deliberate;
this section is kept only to record what the scan saw and why it was raised.

Original: [IRP5Condense.rmd](../../analysis/program/IRP5Condense.rmd)
chunk `{r dropping rows}` L973; comment at L956:
"Drop firms with repetetive entries if they are above 10".

Mock: three firms; F1 has 12 repetitive-uid rows, F2 has 10, F3 has 2.
Intent: drop F1 (and any other firm above 10).

```{r}
ipyr <- data.table(
  taxrefno = rep(c("F1", "F2", "F3"), c(12, 10, 2)),
  uid      = c(rep("AAAAAAAAAA", 12), rep("CCCCCCCCCC", 10), rep("ZZZZZZZZZZ", 2))
)
tb <- table(ipyr$taxrefno)   # repetitive rows per firm
ii <- 10
#### current code's row filter:
dupuid <- ipyr[duplicated(uid) | duplicated(uid, fromLast = TRUE), .(taxrefno, uid)]
cur <- ipyr[taxrefno %in% names(tb)[tb > 10] &
            uid %in% dupuid[taxrefno %in% names(tb)[tb == ii], uid][1:5], ]
#### proposed:
prop <- ipyr[taxrefno %in% names(tb)[tb > 10], ]
list(intent_drops = "F1",
     current_selects = unique(cur$taxrefno),
     proposed_selects = unique(prop$taxrefno))
```

The current filter only keeps F1's rows **because** F2 happens to have
exactly 10 rows *and* shares no uid restriction — here F1's uid is not among
F2's first-5 uids, so the current code selects **nothing**: the drop set is
empty and no firm is dropped downstream. And if no firm had exactly 10:

```{r}
tb2 <- tb[c("F1", "F3")]     # pretend F2 does not exist
dupuid[taxrefno %in% names(tb2)[tb2 == ii], uid][1:5]
```

`character(0)[1:5]` = five NAs → `uid %in% NA` is all-FALSE → `dropthese`
empty, silently.

# B2 — `shift(type="lead")` runs in *physical* order, which is not chronological here

Original: [IRP5HHI.rmd](../../analysis/program/IRP5HHI.rmd) chunk
`{r fraction affected for all years}`: `setkey(ipyrc, busmainplc_geo,
taxrefno, UID, DateStart)` (L215) then `shift(..., type = "lead")`
grouped by `(Txrf, taxyear, UID)` (L217-218).

Mock: one worker, one firm, two establishments. Job 1 (Jan-Mar) is at place
"ZZZ-town", job 2 (Feb-Jun) at "AAA-town". Sorted by place first, the "next
row" inside the person-firm group is chronologically the *earlier* job:

```{r}
ipyrc <- data.table(
  busmainplc_geo = c("ZZZ-town", "AAA-town"),
  Txrf = "F1", UID = "W1",
  DateStart = as.POSIXct(c("2013-01-01", "2013-02-01")),
  DateEnd   = as.POSIXct(c("2013-03-31", "2013-06-30"))
)
setkey(ipyrc, busmainplc_geo, Txrf, UID, DateStart)   # as in the original
ipyrc[, DateStart2 := shift(DateStart, type = "lead"), by = .(Txrf, UID)]
ipyrc[, .(busmainplc_geo, DateStart, DateEnd, DateStart2)]
```

Row 1 ("AAA-town", starts Feb) gets `DateStart2` = Jan — the *previous* job.
The overlap conditions (`DateStart20yr < DateStart2 & DateEnd > DateStart2`,
etc.) then mis-fire. Proposed: order by start date within the shift group:

```{r}
setkey(ipyrc, Txrf, UID, DateStart)
ipyrc[, DateStart2 := shift(DateStart, type = "lead"), by = .(Txrf, UID)]
ipyrc[, .(busmainplc_geo, DateStart, DateEnd, DateStart2)]
```

Now the Jan job sees the Feb job as its successor, and the overlap window
(Feb 1 – Mar 31) is computed correctly.

# B3 — `with = FALSE` evaluates `j` *outside* the table

Original: [IRP5HHI.rmd](../../analysis/program/IRP5HHI.rmd) chunk
`{r define irp5M by location granularity}` L999-1004.

```{r}
dt <- data.table(Txrf = c("F1", "F2"), x = 1:2)
#### the original pattern: column masking attempted inside a with=F selector
try(dt[, c(Firm = letters[as.integer(as.factor(Txrf))], "x"), with = FALSE])
```

`Txrf` is looked up in the calling environment, not in `dt` — error. The
`msk` pattern used elsewhere in the file does it in two steps:

```{r}
out <- dt[, c("Txrf", "x"), with = FALSE]
out[, Firm := letters[as.integer(factor(Txrf))]][, Txrf := NULL]
out
```

# B4 — axis limits (1, 100) on a variable that lives in [0, 1]

Original: [IRP5HHI.rmd](../../analysis/program/IRP5HHI.rmd) chunk
`{r plot FA in agri 2012-2020}` L810.

`FA = NumSubMW/Jobs` is a fraction. What survives each pair of limits:

```{r}
FA <- c(0, .1, .25, .5, .8, 1)          # typical FA values
data.table(
  limits    = c("c(1, 100)  (current)", "c(0, 1)  (proposed)"),
  rows_kept = c(sum(FA >= 1 & FA <= 100), sum(FA >= 0 & FA <= 1)),
  rows_total = length(FA)
)
```

With the current limits the density in `FAOfAg2010-2022.jpg` is drawn from
the FA == 1 sliver only. The two sibling plots in the same file (L684, L715)
already use (0, 1).

# B5 — firm-level `by=` under an establishment-level variable name (user-confirmed intentional, 2026-07-07)

**Status: closed.** The user confirms both levels are deliberate — firm level
in the Lf block (L223-), establishment level in the LSMa block (L144-). This
section records the mechanism that was flagged.

Original: [IRP5MergeData.rmd](../../analysis/program/IRP5MergeData.rmd)
L236-237 (`by = taxrefno`) vs the establishment-keyed version of the same
variable at L144-145.

Mock: firm F1 has establishment A since 2010 and opens establishment B
in 2015:

```{r}
Lf <- data.table(
  taxrefno = "F1",
  busmainplc_geo = c("A", "A", "B", "B"),
  taxyear  = c(2010, 2015, 2015, 2016)
)
Lf[, ExistedBefore2013 := as.integer(taxyear < 2013)]
#### current: firm level
Lf[, firmlevel  := ExistedBefore2013[ExistedBefore2013 == 1L][1], by = taxrefno]
#### establishment level
Lf[, establevel := ExistedBefore2013[ExistedBefore2013 == 1L][1],
   by = .(busmainplc_geo, taxrefno)]
Lf
```

Under `by = taxrefno`, the new establishment B inherits "Existed" from A
(`firmlevel` = 1 everywhere); under establishment keys B stays NA (→ 0 after
the NA-fill line). This feeds `AgeSample` and the `LfCE` estimation
subsamples — **decision needed on which definition is intended** before any
edit.

# B6 — `intersect()` keeps an establishment if *any* row is moderate

Original: [IRP5Impacts.rmd](../../analysis/program/IRP5Impacts.rmd) chunk
`{r examining estimation data}` L337-338.

```{r}
lfdata <- data.table(
  EstID  = c(1, 1, 2),
  rJobsMP = c(10, 500, 5),      # EstID 1 has one moderate and one extreme row
  dJobsMP = c(0, -5, 0)
)
cur  <- lfdata[EstID %in% intersect(EstID[abs(rJobsMP) < 200],
                                    EstID[dJobsMP > -1000]), ]
prop <- lfdata[!(EstID %in% EstID[abs(rJobsMP) >= 200 | dJobsMP <= -1000]), ]
list(current_rows = cur, proposed_rows = prop)
```

The current filter keeps EstID 1 — including its rJobsMP = 500 outlier row,
which then stretches the "after dropping outliers" plot. The proposed form is
the same idiom the winsorize block (L696-700) already uses.

# A2 — the archival-file syntax errors, reproduced

Original: e.g. [CITIRP5BranchLevel.rmd](../../analysis/program/CITIRP5BranchLevel.rmd)
chunk `{r hhi}` (7 copies), [ReadCITIRP5Data.rmd](../../analysis/program/ReadCITIRP5Data.rmd)
same chunks, [IRP5Impacts_P.rmd](../../analysis/program/IRP5Impacts_P.rmd) ~L264.
All four files set `eval=F` globally, so these are latent, not live.

```{r}
#### missing comma between a c(...) vector and the .(...) selector:
try(parse(text = 'LShare[Txrf %in% c("a", "b")  .(Txrf, AreaLevel)]'))
#### missing c() around a multi-column on=:
try(parse(text = 'x[y, on = "a", "b", "c"]'))
#### missing commas between column names (Impacts_P/Test .lb/.ub block):
try(parse(text = 'x[, c("lb" "ub") := .(1, 2)]'))
```

The live counterpart, [IRP5Impacts.rmd](../../analysis/program/IRP5Impacts.rmd)
L377-386, has the commas and the `c()` and parses fine:

```{r}
length(parse(text = 'x[, c("lb", "ub") := .(1, 2)]'))
```

# C6 — why `...`-functions escape the signature scanner

`geom_pointrange(width = .01)`
([IRP5Impacts.rmd](../../analysis/program/IRP5Impacts.rmd) L392-394): ggplot
layer functions absorb unknown arguments through `...` and only *warn*
("Ignoring unknown parameters"), so a name check against `formals()` passes
trivially. That is the reason the new rule says: if the signature has `...`,
bad names pass silently — confirm in `?fn`.

```{r}
"..." %in% names(formals(ggplot2::geom_pointrange))
```

# D2 — the qsave→qread round trip is deliberate (do not "fix")

[IRP5HHI.rmd](../../analysis/program/IRP5HHI.rmd) L504-506 saves `Fadata`
and immediately re-reads it with `use_alt_rep = TRUE`. That converts
materialised string columns into ALTREP representations, cutting RAM on the
100M-row tables. It looks like a no-op; it is a memory optimisation:

```{r}
tf <- tempfile(fileext = ".qs")
big <- data.table(s = rep(sprintf("firm%06d", 1:100), 100))
qsave(big, tf)
alt <- qread(tf, use_alt_rep = TRUE)
identical(big$s, alt$s)   # same values; alt's strings are ALTREP-backed
```
