---
title: "Establishment level analysis of CIT-IRP5 files"
date: "r format(Sys.time(), "%Y%m%d %R")"
output: html_document
#### path <- "Z:/Workbenches/epguest/seiro_ito/outfiles/"; rmarkdown::render(paste0(path, "program/ReadCITIRP5Data.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
```
```{r path, eval = T}
#### path <- "Z:/Workbenches/epguest/seiro_ito/"
#### pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
#### pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
livbrary(qs); library(data.table); library(ggplot2)
```



# Read CIT, IRP5 e5v1

## Read `stmfp_2008_2022_e5_v1.txt` &rarr; `ci5_2008-2022_e5v1.qs`

Read tab-separated text file `stmfp_2008_2022_e5_v1.txt` and save as a `qs` format `ci5_2008-2022_e5v1.qs`.

```{r read save ci5 agri, eval = F}
#### stata's export messes something so I used excel to read the tab-separated file and save as .txt file extension
ci5a <- data.table::fread(paste0(pathdata, "ci5_2008-2022_e5v1.txt"), sep = "\t", fill = T, na.strings = ".")
library(qs)
qs::qsave(ci5a, paste0(pathdata, "ci5_2008-2022_e5v1.qs"))
#### ci5 <- data.table::fread(paste0(pathdata, "stmfp_2008_2022_e5_v1.tsv"))
```


## Read `stmfp_2008_2022_e5_v1.dta` &rarr; `stmfp_2008_2022_e5_v1.qs`

Read stata data file `stmfp_2008_2022_e5_v1.dta`

```{r read save ci5 full, eval = F}
library(qs)
library(data.table)
ci5 <- readstata13::read.dta13(paste0(pathdataCITIRP, "stmfp_2008_2022_e5_v1.dta"))
ci5 <- data.table(ci5)
qs::qsave(ci5, paste0(pathdata, "stmfp_2008_2022_e5_v1.qs"))
#### ci5 <- data.table::fread(paste0(pathdata, "stmfp_2008_2022_e5_v1.tsv"))
```
`read.dta13` took 2 hours on the windows-server 2016 with 1.5TB RAM. Warning messages:

1. In readstata13::read.dta13(paste0(pathdataCITIRP, "stmfp_2008_2022_e5_v1.dta")) :
  
   Duplicated factor levels for variables
   irp5_actcode_4d, vat_actcode_4d, imp_mic_sic7_4d
   Unique labels for these variables have been generated.

2. In readstata13::read.dta13(paste0(pathdataCITIRP, "stmfp_2008_2022_e5_v1.dta")) :
  
   Factor codes of type double or float detected in variables

   IT14_c_cfcownershipit80, IT14_c_fnclssres2ad, c_grpturnover,
   c_dividendsincome, c_divdndsinspecie, c_interestrepymnt, c_gvtgrants,
   c_fruitlssexprecvrd, c_impairmntlossrevers, c_royaltyadjstmnt,
   c_foreignxchngeloss, c_deductnsnotallow, c_cryptoincomeamt,
   c_cryptoprofitamt, c_dda25ifrs, c_dda40ifrs, c_dda40allowifrs,
   c_dda85ifrs, c_dda85allowifrs, c_ddatotifrs, c_ddatotallowifrs,
   c_ddaifrs40, c_ddaifrs40allow, c_ifrsapprvdSARSrate, c_ddaifrstot,
   c_ddaifrstotallow, c_apprvdSARSrate, c_ddaaprvdpct, c_dda85aprvdallow,
   c_ddatotal, c_ddatotalallow, c_ddaaprvd, c_ddaapprvdallowifrs,
   c_solarexceedng1MW, c_concentrtdsolar, c_redemptnallow,
   c_divndsinspecie, c_ifrsdoubtfldebt, c_doubtfldebt, c_lessorexpndtreS8G

   No labels have been assigned.
   Set option 'nonint.factors = TRUE' to assign labels anyway.

3. In readstata13::read.dta13(paste0(pathdataCITIRP, "stmfp_2008_2022_e5_v1.dta")) :
  
   Missing factor labels for variables
   irp5_actcode_4d
   No labels have been assigned.
   Set option 'generate.factors=TRUE' to generate labels.


## 


# Read IRP5_20yr_cleaned_best_geodata.dta (ver 5 beta) and save ipbYR.qs files

Read IRP5 v5 data.

Use `readstata13::read.dta13` to read IRP5 stata data files of 2009-2022, save as `qs` format files. This should take one day to complete.  

* Save as irpYR.qs with YR = 09, ..., 22.  
* Extract entries only of natural persons with non-NA `taxrefno`. For each year, each row is a unique individual tax entry, `taxrefno`, the number of employees in the firm, and geographic information. Save as icountYR.qs. 


```{r read irp5v5beta with , eval = F}
library(qs)
library(data.table)
library(readstata13)
for (yr in 9:22) {
  if (yr < 10) yr <- paste0("0", yr)
  ## Below code is run only once: Start
  irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, "_cleaned.dta"))
  ipyr <- data.table(irpyr)
  rm(irpyr)
  qsave(ipyr, paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## Below code is run only once: End
  ipyr <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## Keep only nature of person is "an individual"
  ipyr <- ipyr[grepl("A", natureofperson), ]
  ## Drop obs with kerr_income == 0
  ipyr <- ipyr[kerr_income != 0, ]
  ## Note that there are header lines with taxrefno = "NULL"
  ## taxrefno = "NULL": no business geo info
  ## Num: Number of employees in a firm, after dropping taxrefno = "NULL"
  ipyr[, Num := as.integer(.N), by = .(taxrefno)]
  ipyr[is.na(Num) | grepl("NULL", taxrefno), Num := 0L]
  ## Num2: Number of employees in a firm, treat taxrefno = "NULL" as sole proprietor
  ## To do so, create and assign taxrefno to them
  ipyr[, Tref := taxrefno]
  ipyr[grepl("NULL", taxrefno), Tref := 1:.N]
  ipyr[, Num2 := as.integer(.N), by = .(Tref)]
  ipyr[is.na(Num) | grepl("NULL", taxrefno), Num := 0L]
  icount <- unique(ipyr[, .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, taxrefno, 
    #### need to run below if we want payereferenceno
    #### payereferenceno,
    TRef, natureofperson, kerr_income, Num, Num2)])
  qsave(icount, paste0(pathdata, "icount", yr, ".qs"), nthreads = 8)
}
```

## Read text file derived data `ci5_2008-2022_e5v1.qs`

Read `ci5_2008-2022_e5v1.qs`.

```{r read ci5a, eval = F}
library(qs)
ci5a <- qs::qread(paste0(pathdata, "ci5_2008-2022_e5v1.qs"))
```



# HHI

## Condensed chunks sent to Noreen

### Read IRP5_20yr_cleaned.dta (ver 5 beta) and save icount by year

```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```

```{r read irp5 file, eval = F}
#### irp5.qs is created in IRP5Condense.rmd
library(qs)
library(data.table)
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
```

```{r plot number of employees 2009-2015}
library(qs)
library(data.table)
library(ggplot2)
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
geovars <- grepout("^bus.*geo$", colnames(irp5))
irp5ForPlot <- unique(irp5[taxyear >= 2012 & taxyear <= 2015, 
  c(geovars, taxrefno, Num)])
g1 <- ggplot(data = irp5ForPlot, aes(x = Num, group = taxyear, 
  colour = taxyear, fill = taxyear)) +
  geom_density(alpha = .3) + 
  scale_color_viridis_d("Tax year") +
  scale_fill_viridis_d() +
  theme_bw()
g1 + facet_wrap(taxyear ~ .)
```

## Compute HHI

```{r hhi}
#### code "hhi"
library(qs)
library(data.table)
irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
geovars <- grepout("^bus.*geo$", colnames(irp5))

#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
ipr5[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipr5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
ipr5[, Entity := "private"]
ipr5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ipr5[, Entity := factor(Entity, levels = c("private", "gov"))]

for (yr in 12:22) {
#### for (yr in 12) {
  ipyr <- mee[taxyear == 2000+yr, ]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachRows := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, TotalRows := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalEmp := sum(EachEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalNeEmp := sum(EachNeEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, ShareRows := round(EachRows/TotalRows, 8)]
    ipGeo[, ShareEmp := round(EachEmp/TotalEmp, 8)]
    ipGeo[, ShareNeEmp := round(EachNeEmp/TotalNeEmp, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachEmp, TotalEmp, ShareEmp)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), 
      GTotalNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[grepl("gov", Entity), 
      GTotalEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### copy GTotalEmp to all rows
    ipGeo[, GTotalEmp := GTotalEmp[!is.na(GTotalEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotalNeEmp := GTotalNeEmp[!is.na(GTotalNeEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachEmpG by replacing EachEmp of GovEntity with 0L
    ipGeo[, EachEmpG := EachEmp]
    ipGeo[grepl("gov", Entity), EachEmpG := 0L]
    ipGeo[, TotalEmpG := TotalEmp-GTotalEmp]
    ipGeo[, ShareEmpG := round(EachEmpG/TotalEmpG, 8)]
    ipGeo[, EachNeEmpG := EachNeEmp]
    ipGeo[grepl("gov", Entity), EachNeEmpG := 0L]
    ipGeo[, TotalNeEmpG := TotalNeEmp-GTotalNeEmp]
    ipGeo[, ShareNeEmpG := round(EachNeEmpG/TotalNeEmpG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "irp5_kerr_iweight_b", "Newgt.kerr","wgt.kerr",
      "EachEmp", "EachNeEmp", "TotalEmp", "TotalNeEmp", 
      "ShareEmp", "ShareNeEmp", 
      "EachEmpG", "TotalEmpG", "ShareEmpG", 
      "EachNeEmpG", "TotalNeEmpG", "ShareNeEmpG"
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(ShareNeEmp^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/TotalEmp)/(1-1/TotalEmp)]
    lshare[, HHIG := sum(ShareEmpG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalEmpG)/(1-1/TotalEmpG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"), nthreads = 13)
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareNeEmp, HHI)]

ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)]
print(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
summary(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)])
print(LShare[TotalNeEmp <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
print(LShare[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
library(ggplot2)
ggplot(data = LShare[!is.na(Newgt.kerr), ], 
  aes(x = ShareEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
ggplot(data = LShare[!is.na(Newgt.kerr) & ShareEmp != 0, ], 
  aes(x = ShareNeEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
summary(LShare[!is.na(Newgt.kerr) & ShareNeEmp != 0, ShareNeEmp])
ggplot(data = unique(LShare[!is.na(Newgt.kerr), .(busmainplc_geo, HHI)]), 
  aes(x = HHI)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
```


```{r copy location information from 2013... to 2012}
#### code "copy location"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### Read data
#### nthreds = 16 at NT-SDF, = 8 with my laptop (less 1, keep 1 for other computations)
#### ipyr <- qread(paste0(pathdata, "irp20.qs"), nthreads = 13)
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, uid := gsub(" +$", "", UID)]
rm(ipyrs)
#### Below gives non-uique match error for fyr > 13
####  ipyrs <- ipyr2[ipyrs]
#### Below gives memory error for fyr > 13
####    ipyrs <- merge(ipyrs, ipyr, by = c("taxrefno", "uid"), 
####      all.x = T, allow.cartesian = T)
#### So I will just merge 2013
#### for (fyr in 13:15) {
for (fyr in 13) {
  ipyr2 <- qread(paste0(pathdata, "irp", fyr, ".qs"), nthreads = 13)
  ipyr2 <- ipyr2[, .(taxrefno, UID, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo)]
  setnames(ipyr2, c("busprov_geo", "busdistmuni_geo", 
    "buslocmuni_geo", "busmainplc_geo"),
    c("bprov", "bdist", "bloc", "bmain"))
  ipyr2[, uid := gsub(" +$", "", UID)]
  if (fyr == 13)
    setkey(ipyr, taxrefno, uid) else 
    setkey(ipyrs, taxrefno, uid)
  setkey(ipyr2, taxrefno, uid)
  #### join data.tables ipyr(s), ipyr2 keeping rows of ipyr(s)
  #### (keep the rows of ipyr(s), attach the matching rows of ipyr2)
  #### "Take ipyr2, and subset it using the rows/keys in ipyr(s)"
  if (fyr == 13)
    ipyrs <- ipyr2[ipyr] else 
    ipyrs <- ipyr2[ipyrs]
  # if (fyr == 13)
  #  ipyrs <- merge(ipyr, ipyr2, by = c("taxrefno", "uid"), 
  #    all.x = T, allow.cartesian = T) else 
  #  ipyrs <- merge(ipyrs, ipyr, by = c("taxrefno", "uid"), 
  #    all.x = T, allow.cartesian = T)
  cat(fyr, "\n")
  ii.prov <- ipyrs[, 
    (busprov_geo == "" | busprov_geo == "EXCEPTION")
    & (bprov != "" & bprov != "EXCEPTION") ]
  ii.dist <- ipyrs[, 
    (busdistmuni_geo == "" | busdistmuni_geo == "EXCEPTION")
    & (bdist != "" & bdist != "EXCEPTION") ]
  ii.loc <- ipyrs[, 
    (buslocmuni_geo == "" | buslocmuni_geo == "EXCEPTION")
    & (bloc != "" & bloc != "EXCEPTION") ]
  ii.main <- ipyrs[, 
    (busmainplc_geo == "" | busmainplc_geo == "EXCEPTION")
    & (bmain != "" & bmain != "EXCEPTION") ]
  cat(fyr, "\n")
  print(table(ii.prov))
  print(table(ii.dist))
  print(table(ii.loc))
  print(table(ii.main))
  print(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
    .(taxrefno, uid, busprov_geo, bprov, 
      busdistmuni_geo, bdist,
      buslocmuni_geo, bloc,
      busmainplc_geo, bmain)])
  #### copy from future data
  if (fyr == 13) ipyrs[, LocInfoCopiedFrom := 0L]
  ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
    LocInfoCopiedFrom := as.integer(fyr)]
  ipyrs[ii.prov, busprov_geo := bprov]
  ipyrs[ii.dist, busdistmuni_geo := bdist]
  ipyrs[ii.loc, buslocmuni_geo := bloc]
  ipyrs[ii.main, busmainplc_geo := bmain]
  setnames(ipyrs,
      c("bprov", "bdist", "bloc", "bmain"), 
      paste0(c("bprov", "bdist", "bloc", "bmain"), ".", fyr)
    )
}
fyr
ipyrs[, .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)]
ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][
  duplicated(uid) | duplicated(uid, fromLast=T)][
  order(uid, busmainplc_geo, taxrefno)]
tb <- table(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][
  duplicated(uid) | duplicated(uid, fromLast=T)][, 
  uid])
unique(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][
  duplicated(uid) | duplicated(uid, fromLast=T)][, 
  (uid=uid)])[grepl("(.)\1{5,}", uid)]
#### uid anomalous entries: AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ
#### these are 1.25% of total number of rows
#### one large employer (with 70K+ jobs) contributes 
#### to 1/2 of the total cases
uids <- unique(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][
  duplicated(uid) | duplicated(uid, fromLast=T)][, 
  (Uid=uid)])
#### repetitive: [A-Z]*8
repetetive <- uids[grepl("(.)\\1{7,}", uids)]
ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][uid %in% repetetive, ][
  order(uid), ]
tb <- table(ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
  .(taxrefno, uid, busprov_geo, bprov.13, 
  busmainplc_geo, bmain.13)][uid %in% repetetive, taxrefno])
tb <- tb[order(tb, decreasing=T)]
tb
round((142638/11406422)*100, 5)
#### duplicated TU ("taxrefno-uid")
ipyrs[, TU := .N, by = .(taxrefno, uid)]
tb <- table(ipyrs[, TU])
tb <- tb[order(tb, decreasing=T)]
tb

#### Drop repetitive (anomalous) entries
ipyrs.clean <- ipyrs[!(uid %in% repetetive), ]
qsave(ipyrs, paste0(pathdata, "ipyrs.qs"), nthreads = 13)
qsave(ipyrs.clean, paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
```

`IRP5_2012_cleaned.dta` of T:/IRP5/Job level/v5/beta/ &rarr; `irp12.qs` &rarr; `ipyrsClean.qs` (after dropping anomalous/invalid UID entries such as AAAAAAA, ZZZZZZZ, etc.) &rarr; `BranchWeights2012.qs`

* Individual level cross sectional data with `taxrefno` of employee.
* Compute branch^[A branch is defined by location up to Main Place level. If there are multiple branches in a same Main Place, they are aggregated as one branch. ] weights using `kerr_income`
   * Branch weight = $\frac{\mbox{total income of the branch}}{\mbox{total income of the firm}}$


`stmfp_2008_2022_e5_v1.dta` &rarr; `stmfp_2008_2022_e5_v1.qs`^[`read.dta13` took 2 hours on the windows-server 2016 with 1.5TB RAM.  ]

* Matched employee-employer panel, aka CIT-IRP5 panel. Employees are linked only to HQ, not actual branch they are hired at


`stmfp_2008_2022_e5_v1.qs`+`BranchWeights2012.qs` using Main Place and taxrefno as keys.

* We use branch weights to distribute the number of jobs across branches. 



`CIT.dta` &rarr; `CIT.qs`

* Firm data. Job entries of all branches are attached to a single row (no branch level info. 







```{r mock copy location information codes}
#### code "copy location" mock data codes
library(qs)
library(data.table)
set.seed(100)
ipyr12 <- data.table(
    taxrefno = 
      sample(letters[1:2], 20, replace = T),
    UID = 
      paste0(
        sample(letters[1:10], 20, replace = T),
        sample(letters[1:10], 20, replace = T),
        sample(letters[1:10], 20, replace = T)
      ),
    val = rpois(20, 10)
  )
ipyr12[, uid := gsub(" +$", "", UID)]
for (fyr in 13:15) {
  z <- data.table(
      taxrefno = ipyr12[, taxrefno],
      UID = c(
          sample(as.character(ipyr12[1:(18-(fyr-13)*2), UID]), replace = F),
          paste0(ipyr12[1:(2+(fyr-13)*2), UID], 0)
        ),
      val = rpois(20, 20)
    )
  z[, uid := gsub(" +$", "", UID)]
  setnames(z, "val", paste0("val", fyr))
  setkey(ipyr, taxrefno, uid)
  assiagn(paste0("ipyr", fyr), z)
}
rm(ipyrs)
for (fyr in 13:15) {
  ipyr2 = copy(get(paste0("ipyr", fyr)))
  if (fyr == 13)
    setkey(ipyr12, taxrefno, uid) else 
    setkey(ipyrs, taxrefno, uid)
  setkey(ipyr2, taxrefno, uid)
  #### join data.tables ipyr(s), ipyr2 keeping rows of ipyr(s)
  #### (keep the rows of ipyr(s), attach the matching rows of ipyr2)
  #### "Take ipyr2, and subset it using the rows/keys in ipyr(s)"
  if (fyr == 13)
    ipyrs <- ipyr2[ipyr] else 
    ipyrs <- ipyr2[ipyrs]
  cat(fyr, "\n")
  print(ipyrs[, 
    .(taxrefno, UID, uid, val, val13)])
  #### copy from future data
  if (fyr == 13) ipyrs[, LocInfoCopiedFrom := 0L]
  ipyrs[!is.na(get(paste0("val", fyr))) &
    LocInfoCopiedFrom == 0L, 
    LocInfoCopiedFrom := as.integer(fyr)]
}
ipyrs

z <- data.table(
    geo=sample(LETTERS[1:2], 200, T)
  , uid=paste0(sample(letters[1:2], 200, T), 
      putzeroontop(sample(1:20, 200, T)))
  , val=rpois(200, 20))
tb <- table(z[, uid])
tb <- tb[order(tb, decreasing=T)]


```

### Compute branch weights

* For a given taxrefno, labour information is aggregated over branches and attached to the main branch/HQ
* One needs to distribute the total labour size to each branch
* One way to do so is to use an operation scale variable and compute branch weights that sum to 1

```{r compute branch weights}
#### code "compute branch weights"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### Read data
#### nthreds = 16 at NT-SDF, = 8 with my laptop (less 1, keep 1 for other computations)
#### ipyr <- qread(paste0(pathdata, "irp20.qs"), nthreads = 13)
#### ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)

#### Using irp5-2012 to compute weights
#### ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyrc <- qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
#### total incomes/wages of each branch (up to main place level) of taxrefno
#### Below gives an error: Check if is.data.table() = true, ...
#### wgt12 <- ipyrc[, .(
####   subtot.kerr := sum(kerr_income, na.rm = T),
####   nonempty.kerr := sum(kerr_income[
####       busmainplc_geo != "" | grepl("NULL", busmainplc_geo)
####     ], na.rm = T)
####   ), 
####   by = .(taxrefno, busprov_geo, busdistmuni_geo, 
####     buslocmuni_geo, busmainplc_geo)]
#### Above gives an error, so I take an easy way
wgt12 <- ipyrc[, .(
  subtot.kerr = sum(kerr_income, na.rm = T)
  ), 
  by = .(taxrefno, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo)]
wgt12n <- ipyrc[
      busmainplc_geo != "" | grepl("NULL", busmainplc_geo), 
      .(nonempty.kerr = sum(kerr_income, na.rm = T)
  ), 
  by = .(taxrefno, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo)]
wgt12n[, c("Netot.kerr", "NeBranchNum", "NeTotalBranch") := 
    list(
      sum(nonempty.kerr, na.rm = T),
      1:.N,
      .N
    ), 
  by = taxrefno]
wgt12n[, Newgt.kerr := round((nonempty.kerr/Netot.kerr)*100, 3)]
print(wgt12n[NeTotalBranch == 4, ], topn = 12, nrows = 30)

wgt12[, c("tot.kerr", "BranchNum", "TotalBranch") := 
    list(
      sum(subtot.kerr, na.rm = T),
      1:.N,
      .N
    ), 
  by = taxrefno]
wgt12[, wgt.kerr := round((subtot.kerr/tot.kerr)*100, 3)]
print(wgt12[TotalBranch == 4, ], topn = 12)

setkey(wgt12, taxrefno, busprov_geo, busdistmuni_geo, 
  buslocmuni_geo, busmainplc_geo)
setkey(wgt12n, taxrefno, busprov_geo, busdistmuni_geo, 
  buslocmuni_geo, busmainplc_geo)
wg12 <- wgt12n[wgt12]
dim(wgt12); dim(wgt12n); dim(wg12); 
qsave(wg12, paste0(pathdata, "BranchWeights2012.qs"), nthreads = 13)

print(wg12[TotalBranch == 4, 
  .(taxrefno, busprov_geo, busmainplc_geo, wgt.kerr, Newgt.kerr)], 
  topn = 12)
print(wg12[taxrefno == "", 
  .(taxrefno, busprov_geo, busmainplc_geo, wgt.kerr, Newgt.kerr)], 
  topn = 12)
table(wg12[, .(NeTotalBranch, emptytxrefno = taxrefno == "")])

grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
wg12[, Year := 2012L]
wg12ForL = copy(wg12)
wg12ForL[, grepout("^non|^subt", colnames(wg12ForL)) := NULL]
setnames(wg12LForL, 
  c("tot.kerr", "Netot.kerr", "wgt.kerr", "Newgt.kerr"), 
  c("kerr.tot", "kerr.Netot", "kerr.wgt", "kerr.Newgt"), 
)
wg12L <- reshape(wg12ForL, 
   direction = "long", 
       idvar = c("taxrefno", "busprov_geo", 
         "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"),
     timevar = "Year",
     varying = grepout("^kerr", colnames(wg12ForL))
 )
library(ggplot2)
ggplot(data = wg12L[taxrefno != "" & grepl("wgt", Year), ], 
  aes(x = kerr, group = Year, color = Year)) +
  geom_density(alpha = .1) + xlim(0, 100) + 
  scale_color_manual(values = c("red", "blue"))

meanwgt <- unique(wg12L[taxrefno != "" &  grepl("^Newgt", time), 
    .(meanweight = mean(kerr, na.rm = T),
      medianweight = median(kerr, na.rm = T)), by = .(taxrefno)])
ggplot(data = meanwgt, 
  aes(x = meanweight, group = time, color = time)) +
  geom_density(alpha = .1) + xlim(0, 100) + 
  scale_color_manual(values = c("red", "blue"))
ggplot(data = meanwgt, 
  aes(x = medianweight, group = time, color = time)) +
  geom_density(alpha = .1) + xlim(0, 100) + 
  scale_color_manual(values = c("red", "blue"))



ggplot(data = wg12[taxrefno != "", ], aes(x = Newgt.kerr)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 100)
ggplot(data = wg12[taxrefno != "", ], aes(x = wgt.nekerr)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 100)
ggplot(data = wg12[taxrefno != "", ], aes(x = wgt.kerr)) +
  geom_histogram(colour="red", binwidth = 1)
ggplot(data = wg12[taxrefno != "", ], aes(x = wgt.nekerr)) +
  geom_histogram(colour="red", binwidth = 1)

meanwgt <- unique(wg12[taxrefno != "", 
    .(meanweight = mean(wgt.kerr, na.rm = T),
      meanneweight = mean(wgt.nekerr, na.rm = T)), by = taxrefno])
summary(meanwgt[meanweight < 25, ] )
meanwgt[meanweight < 25, ][order(meanweight, decreasing = F)]
View(meanwgt[meanweight < 10, ][order(meanweight, decreasing = F)][1:50, ])
View(meanwgt[meanneweight < 10, ][order(meanneweight, decreasing = F)][1:50, ])
e#### #### Using cit-irp5-2012 to compute weights
#### ci5p <- qread(paste0(pathdata, "ci5p.qs")) 
#### cip12 <- ci5p[taxyear == 2012, ]
#### #### total incomes/wages of each branch (up to main place level) of taxrefno
#### wgt12 <- cip12[, .(
####   subtot.kerr = sum(kerr_income, na.rm = T), 
####   subtot.xwage = sum(x_wages, na.rm = T), 
####   BranchNum = 1:.N), 
####   by = .(taxrefno, busprov_geo, busdistmuni_geo, 
####     buslocmuni_geo, busmainplc_geo)]
#### #### total incomes/wage payments of taxrefno
#### wgt12[, tot.kerr := sum(subtot.kerr, na.rm = T), by = taxrefno]
#### wgt12[, tot.xwage := sum(subtot.xwage, na.rm = T), by = taxrefno]
#### wgt12[, wgt.kerr := round((subtot.kerr/tot.kerr)*100, 3)]
#### wgt12[, wgt.xwage := round((subtot.xwage/tot.xwage)*100, 3)]
#### library(ggplot2)
#### ggplot(data = wgt12[, .(wgt.kerr, wgt.xwage)]) +
####   geom_point(size=1, colour="blue") + 
####   geom_jitter(width = 0.01, height = 0.01) + 
####   geom_abline(intercept = 0, slope = 1, colour = "orange")
```


### Merge branch weights and create LShare data: HHI at LocalMunicipality, Districts levels in 2012


* Using this and branch weight file, we allocate jobs over branches


```{r merge branch weights with MEE}
#### code "merge branch weights"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### Use the "clean" version (anomalous uid's are dropped)
#### ipr2012 ==> ipyrsClean.qs
####   142638/11406422 = 1.25% of all rows are useless uid's
####   more than half of it is done by taxrefno ZAJJQCTBKC
#### ipyrs <- qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
#### ci5 <- readstata13::read.dta13(paste0(pathdataCITIRP, 
####  "stmfp_2008_2022_e5_v1.dta"))
#### ci5 <- data.table(ci5)
#### qsave(ci5, paste0(pathdata, "stmfp_2008_2022_e5_v1.qs"))
#### stmfp_2008_2022_e5_v1.qs is MEE: CIT-IRP5 panel
ci5 <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1.qs"), nthreads = 13)
#### ci5 <- qread(paste0(pathdata, "ci5_2008-2022_e5v1.qs"), nthreads = 13)
wg12 <- qread(paste0(pathdata, "BranchWeights2012.qs"), nthreads = 13)
#### keep only nonempty taxrefno
ipyrs <- ci5[taxrefno != "", ]
wg12 <- wg12[taxrefno != "", ]
dim(ipyrs)
dim(ipyrs[taxrefno %in% wg12[, taxrefno], ])
dim(wg12[taxrefno %in% ipyrs[, taxrefno], ])
#### multi (taxrefno-taxyear) - multi (taxrefno-branch) merge
#### Set allow.cartesian = T
#### Base = ipyrs and merge whichever matched rows in wg12
iprw12 <- wg12[ipyrs, on = "taxrefno", allow.cartesian = T]
#### mock merge example
set.seed(10)
x11 <- data.table(id = rep(1:2, each = 3), year = rep(11:13, 2), 
  val = rpois(6, 20)) 
x12 <- data.table(id = rep(1:3, each = 3), year = rep(11:13, 3), 
  val = rpois(9, 20)) 
x2 <- data.table(id = rep(1:2, each = 3), branch = rep(1:3, 2), 
  location = letters[1:6]) 
#### This fails. Use allow.cartesian = T.
merge(x11, x2, by = "id")
x122 <- x2[x12, on = "id", allow.cartesian = TRUE]
x12; x2; x122
x212 <- x12[x2, on = "id", allow.cartesian = TRUE]
x12; x2; x212
print(iprw12[TotalBranch == 5, 
  .(taxyear, taxrefno, busmainplc_geo, BranchNum, TotalBranch)], 
  topn = 50)
rbind(MEE=dim(ipyrs), weights=dim(wg12), merged=dim(iprw12))
table(
(NumRows <- iprw12[!grepl("^NULL$", taxrefno), 
  .(NumRows=.N), by = taxrefno])[, NumRows]
)
NumRows
NumRows[grepl("NULL|AAAAAAA", taxrefno), ]
iprw12[, NumRows :=.N, by = taxrefno]
print(iprw12[NumRows == 15, 
  .(taxrefno, NumRows, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, tot.kerr, wgt.kerr)],
  topn = 45)
print(iprw12[NumRows == 12 & TotalBranch == 3 & !is.na(busmainplc_geo), 
  .(taxrefno, NumRows, NumBranch, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, tot.kerr, wgt.kerr)],
  topn = 45)
iprw12[, imputed.kerr := tot.kerr*wgt.kerr/100]
iprw12[, Neimputed.kerr := Netot.kerr*Newgt.kerr/100]
iprw12[, imputed.emp := irp5_kerr_iweight_b*wgt.kerr/100]
iprw12[, Neimputed.emp := irp5_kerr_iweight_b*Newgt.kerr/100]
nrow(iprw12[kerr_income != 0, ]); (nrow(iprw12[kerr_income != 0, ])/nrow(iprw12))*100
nrow(iprw12[Neimputed.kerr != 0, ]); (nrow(iprw12[Neimputed.kerr != 0, ])/nrow(iprw12))*100
qsave(iprw12, paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"), nthreads = 13)
```

### Compute HHI

```{r hhi}
#### code "hhi"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
mee <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"),
  nthreads = 13)
mee[taxrefno == "AACBTBBAJZ", ]
print(mee[,
  .(taxrefno, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr, Neimputed.emp)], topn = 25)
summary(mee[,
  .(taxrefno, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr, Neimputed.emp)])
#### mee[, NumB := .N, by = taxrefno]
#### mee[, EmptyLocInfo := sum(busmainplc_geo == ""), by = taxrefno]
#### table(mee[, .(NumB, EmptyLocInfo)])
#### Not sure if we need this. Test how many obs will be dropped.
nrow(mee[tot.kerr != 0, ]); (nrow(mee[tot.kerr != 0, ])/nrow(mee))*100
nrow(mee[Neimputed.kerr != 0, ]); (nrow(mee[Neimputed.kerr != 0, ])/nrow(mee))*100
#### For 2012, no NA in payereferenceno
#### (n0 <- nrow(mee[is.na(payereferenceno) | payereferenceno == "", ])); (n0)/nrow(mee)*100
#### 14% are gov employees [is.na(taxrefno) | taxrefno == ""]
#### So we cannot simply drop entries with NAs in taxrefno
(n0 <- nrow(mee[grepl("NULL", taxrefno) | is.na(taxrefno) | taxrefno == "", ])); (n0/nrow(mee))*100
#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
mee[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
mee[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
mee[, Entity := "private"]
mee[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
mee[, Entity := factor(Entity, levels = c("private", "gov"))]
for (yr in 12:22) {
#### for (yr in 12) {
  ipyr <- mee[taxyear == 2000+yr, ]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachRows := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, TotalRows := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalEmp := sum(EachEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalNeEmp := sum(EachNeEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, ShareRows := round(EachRows/TotalRows, 8)]
    ipGeo[, ShareEmp := round(EachEmp/TotalEmp, 8)]
    ipGeo[, ShareNeEmp := round(EachNeEmp/TotalNeEmp, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachEmp, TotalEmp, ShareEmp)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), 
      GTotalNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[grepl("gov", Entity), 
      GTotalEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### copy GTotalEmp to all rows
    ipGeo[, GTotalEmp := GTotalEmp[!is.na(GTotalEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotalNeEmp := GTotalNeEmp[!is.na(GTotalNeEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachEmpG by replacing EachEmp of GovEntity with 0L
    ipGeo[, EachEmpG := EachEmp]
    ipGeo[grepl("gov", Entity), EachEmpG := 0L]
    ipGeo[, TotalEmpG := TotalEmp-GTotalEmp]
    ipGeo[, ShareEmpG := round(EachEmpG/TotalEmpG, 8)]
    ipGeo[, EachNeEmpG := EachNeEmp]
    ipGeo[grepl("gov", Entity), EachNeEmpG := 0L]
    ipGeo[, TotalNeEmpG := TotalNeEmp-GTotalNeEmp]
    ipGeo[, ShareNeEmpG := round(EachNeEmpG/TotalNeEmpG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "irp5_kerr_iweight_b", "Newgt.kerr","wgt.kerr",
      "EachEmp", "EachNeEmp", "TotalEmp", "TotalNeEmp", 
      "ShareEmp", "ShareNeEmp", 
      "EachEmpG", "TotalEmpG", "ShareEmpG", 
      "EachNeEmpG", "TotalNeEmpG", "ShareNeEmpG"
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(ShareNeEmp^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/TotalEmp)/(1-1/TotalEmp)]
    lshare[, HHIG := sum(ShareEmpG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalEmpG)/(1-1/TotalEmpG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"), nthreads = 13)
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareNeEmp, HHI)]

ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)]
print(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
summary(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)])
print(LShare[TotalNeEmp <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
print(LShare[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
library(ggplot2)
ggplot(data = LShare[!is.na(Newgt.kerr), ], 
  aes(x = ShareEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
ggplot(data = LShare[!is.na(Newgt.kerr) & ShareEmp != 0, ], 
  aes(x = ShareNeEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
summary(LShare[!is.na(Newgt.kerr) & ShareNeEmp != 0, ShareNeEmp])
ggplot(data = unique(LShare[!is.na(Newgt.kerr), .(busmainplc_geo, HHI)]), 
  aes(x = HHI)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
```

1. Use IRP5 2012 (`ipyrc`) to count total number of workers, number of affected, FA = affected/total
1. Merge branch weight data () with FAdata
    a) Distribute workers and number of affected workers proportionately to branch weights
    a) Compute FA per branch


```{r define fraction affected}
#### code "compute FA"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### ipyrsClean.qs=irp12.qs+dropped anomalous taxrefno
ipyrc <- qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
ipyrc[taxrefno == "AACBTBBAJZ", ]
#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
ipyrc[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipyrc[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
ipyrc[, Entity := "private"]
ipyrc[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ipyrc[, Entity := factor(Entity, levels = c("private", "gov"))]
#### Choose only unique uid-income rows
ipyrc <- ipyrc[!duplicated(
    ipyrc[, .(busmainplc_geo, taxrefno, uid, a3601_income)]
  ), ]
#### Compute job duration and monthly income
####   (DateEnd-DateStart2012)/30.5
####   a3601_income/DurationMonth
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
ipyrc[, DateStart := as.IDate(periodemployedfrom)]
ipyrc[, DateEnd := as.IDate(periodemployedto)]
ipyrc[, DateStart2012 := DateStart]
ipyrc[DateStart < as.IDate("2011/03/01"), 
  DateStart2012 := as.IDate("2011/03/01")]
ipyrc[, DurationMonth := round((DateEnd-DateStart2012)/30.5, 3)]
ipyrc[, IncomeMonth := round(a3601_income/DurationMonth, 0)]

setkey(ipyrc, busmainplc_geo, taxrefno, uid, DateStart)
ipyrc[, DateStart2 := shift(DateStart, n = 1L, type = "lead"), 
  by = .(Txrf, taxyear, uid)]
ipyrc[, DateEnd2 := shift(DateEnd, n = 1L, type = "lead"), 
  by = .(Txrf, taxyear, uid)]
#### Double work duration: DateEnd - DateStart2
####  <----job1---->
####           <----job2---->
ipyrc[, DJobDurationMonth := round((DateEnd - DateStart2)/30.5, 3)]
ipyrc[, TDurationMonth := 
  round((DateEnd2 - DateStart2012)/30.5, 3)]
#### If job2 ends before job1 ends: DateEnd > DateEnd2
####  <----------job1---------->
####           <----job2---->
####   (DateEnd2 - DateStart2)/30.5 (double work duration)
####   (DateEnd-DateStart2012)/30.5 (job1+job2 duration)
ipyrc[DateEnd > DateEnd2, 
  DJobDurationMonth := round((DateEnd2 - DateStart2)/30.5, 3)]
ipyrc[DateEnd > DateEnd2, TDurationMonth := 
  round((DateEnd-DateStart2012)/30.5, 3)]
ipyrc[, TDurationMonth := TDurationMonth[!is.na(TDurationMonth)][1], 
  by = .(Txrf, taxyear, uid)]
ipyrc[, DoubleJobRatio := round((DJobDurationMonth/TDurationMonth)*100, 2)]
ipyrc[, DoubleJobRatio := DoubleJobRatio[!is.na(DoubleJobRatio)][1],
  by = .(Txrf, taxyear, uid)]
ipyrc[, NumJobsPerWorker := .N, by = .(Txrf, taxyear, uid)]
ipyrc[, TIncomeMonth := sum(IncomeMonth, na.rm = T), 
  by = .(Txrf, taxyear, uid)]
#### check data
ipyrc[busmainplc_geo != "" & Txrf != "" & NumJobsPerWorker > 1, 
  .(busmainplc_geo, Txrf, uid, NumJobsPerWorker, 
    DateStart, DateEnd,
    DateStart2, DateEnd2, DurationMonth, 
    DJobDurationMonth, TDurationMonth, DoubleJobRatio,
    a3601_income, IncomeMonth, TIncomeMonth)]
print(ipyrc[NumJobsPerWorker == 5, 
  .(busmainplc_geo, Txrf, uid, NumJobsPerWorker, 
    DateStart, DateEnd,
    DateStart2, DateEnd2, DurationMonth, 
    DJobDurationMonth, TDurationMonth, DoubleJobRatio,
    a3601_income, IncomeMonth, TIncomeMonth)], topn = 30)
ipyrc[busmainplc_geo != "" & Txrf != "", 
  .(busmainplc_geo, taxrefno, DateStart, DateEnd, 
    DurationMonth, a3601_income, IncomeMonth)]
#### Find sub MW workers and define fraction affected
#### We can use IncomeMonth or TIncomeMonth
####  PropToMW: Proportion to MW line
####  SubMW: PropToMW < 1 & IncomeMonth > 0
####  NumSubMW: number of entries with SubMW == 1L
####  Employees: number of entries with IncomeMonth > 0
####  FA: NumSubMW/Employees
ipyrc[, PropToMW := round(IncomeMonth/2274, 6)]
ipyrc[, SubMW := as.integer(PropToMW < 1 & IncomeMonth > 0)]
ipyrc[, NumSubMW := uniqueN(UID[SubMW==1L]), 
  by = .(Txrf, taxyear)]
ipyrc[, Employees := uniqueN(UID[IncomeMonth > 0]), 
  by = .(Txrf, taxyear)]
ipyrc[, FA := round(NumSubMW/Employees, 6)]
ipyrc[grepl("^BCGACCKJZC", Txrf), 
  .(busmainplc_geo, 
    Txrf, uid, UID, DateStart, DurationMonth, 
    IncomeMonth, SubMW, Employees, FA)]
#### Vary sub MW workers by DoubleJobRatio X >= 50
####  SubMWX: PropToMW < 1 & TIncomeMonth > 0 & DoubleJobRatio >= X
####  NumSubMWX: number of entries with SubMW == 1L & DoubleJobRatio >= X
####  EmployeesX: number of entries with IncomeMonth > 0 & DoubleJobRatio >= X
####  FA25: NumSubMWX/EmployeesX with X=25
####  FA50: NumSubMWX/EmployeesX with X=50
####  FA75: NumSubMWX/EmployeesX with X=75
ipyrc[, PropToMWX := round(TIncomeMonth/2274, 6)]
  by = .(Txrf, taxyear)]
ipyrc[, EmployeesX := uniqueN(uid[TIncomeMonth > 0]), 
  by = .(Txrf, taxyear)]
XThreshold <- 50
for (XThreshold in c(25, 50, 75)) {
  ipyrc[, SubMWX := as.integer(PropToMWX < 1 & TIncomeMonth > 0)]
  ##### PropToMWX <= 1L ==> subMW == 1L
  ####   If DoubleJobRatio  < X & PropToMWX  > 1L ==> subMW == 1L
  ####   If DoubleJobRatio >= X & PropToMWX  > 1L ==> subMW == 0L
  #####     PropToMWX <= 1L ==> subMW == 1L
  #### DoubleJobRatio := round((DJobDurationMonth/TDurationMonth)*100
  ipyrc[DoubleJobRatio >= XThreshold & PropToMWX > 1L, 
    SubMWX := 0L]
  ipyrc[, NumSubMWX := uniqueN(uid[SubMWX==1L]), 
    by = .(Txrf, taxyear)]
  ipyrc[, FAx := round(NumSubMWX/EmployeesX, 6)]
  setnames(ipyrc, c("FAx", "NumSubMWX"), 
    paste0(c("FA", "NumSubMW"), XThreshold))
}
ipyrc[, FinYr2012 := 1L]
ipyrc[DateStart > as.IDate("2012/02/29") | 
  DateEnd > as.IDate("2012/02/29"), FinYr2012 := 0L]
#### Select unique entries for FinYr2012 for each branch
FAdata <- unique(ipyrc[FinYr2012 == 1L, .(
    busprov_geo, 
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno, Txrf, payereferenceno, UID,
    Employees, NumSubMW, PropToMW, SubMW, 
    FA, FA25, FA50, FA75
  )])
setnames(FAdata, c("FA", "FA25", "FA50", "FA75"),
    paste0("FA.", c("orig", 25, 50, 75))
)
summary(FAdata)
FAdata[, Size := "micro"]
FAdata[Employees > 1 & Employees <= 10, Size := "small"]
FAdata[Employees > 10 & Employees <= 50, Size := "medium"]
FAdata[Employees > 50, Size := "large"]
FAdata[is.na(Employees) | Employees == 0, Size := "NA"]
#### Number of affected workers in aggregates (2012)
ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
  , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
  )
  ][, .(
    TotalEmployees, TotalSubMWWorkers, 
    AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
    TotalFirmsWithEmployees, TotalAffectedFirms,
    AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
  )]
ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
  , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
  ), 
  by = .(taxyear, busprov_geo)
  ][, .(
    busprov_geo, TotalEmployees, TotalSubMWWorkers, 
    AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
    TotalFirmsWithEmployees, TotalAffectedFirms,
    AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
  )]
#### Number of employees per branch
(EmpSubMW <- ipyrc[, .(
    TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
  , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
  ), 
  by = .(
    taxyear, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, taxrefno, payereferenceno)
  ])
table(EmpSubMW[, TotalSubMWWorkers])
library(ggplot2)
library(cowplot)
g1 <- ggplot(data = EmpSubMW, 
    aes(x=TotalSubMWWorkers)) +
  geom_histogram(binwidth=1)
g2 <- g1 + geom_histogram(bins=20) +
  scale_x_continuous(transform = "log10") 
g1 <- g1 + scale_x_continuous(limits = c(0, 10), breaks = 0:10) +
  scale_y_continuous(limits = c(0, 100000))
gs <- plot_grid(
  g1, g2,
  labels = c('natural, up to 10 people', 'log10'),
  align="hv"
)
gs

#### For FA=0, do they employ?
table(EmpSubMW[TotalSubMWWorkers == 0L, TotalEmployees])

qsave(FAdata, paste0(pathdata, "FAdata.qs"), nthreads = 13)
qsave(ipyrc, paste0(pathdata, "ipyrc.qs"), nthreads = 13)
```
```{r plot FAs}
#### possible duplicates with different local municipality names
ii0 <- FAdata[, buslocmuni_geo != "" & busdistmuni_geo != ""
   & busmainplc_geo != ""]
ii1 <- duplicated(FAdata[, .(
   busmainplc_geo, Txrf
 )])
ii2 <- duplicated(FAdata[, .(
   busmainplc_geo, Txrf
 )], fromLast = T)
table(data.table(nonNA=ii0, duplicates=ii1))
FAdata[ii0 & (ii1 | ii2), ]
table(duplicated(FAdata[ii0 & !ii1, ]))
#### plot sensitivity to XThreshold (double job months cutoff)
library(ggplot2)
library(cowplot)
g1 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.orig)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "blue")
g2 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.25)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "red")
g3 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.50)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "green")
g4 <- ggplot(data = FAdata[ii0 & !ii1, ], 
  aes(x = FA.75)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = "orange")
gs <- plot_grid(
  g1, g3, g2, g4,
  labels = c('original', '50', '25', '75'),
  align="hv"
)
ggsave(
  paste0(pathdata, "FAByDoubleJobRatio.jpg"),
  , gs, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
#### Below gives an error: idvar need to specify unique rows
#### FAdataL <- reshape(FAdata[ii0 & !ii1, ], 
####   direction = "long", 
####   idvar = c("busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", 
####     "Txrf", "payereferenceno"), 
####   varying = grepout("^FA", colnames(FAdata))
#### )
gsz <- ggplot(data = FAdata, 
  aes(x = FA.orig)) +
  geom_density(alpha = .1) + 
  xlim(0, 1) +
  scale_color_manual(values = c("red", "blue"))
gsz <- gsz + facet_wrap(Size ~ .)
ggsave(
   paste0(pathdata, "FAByFirmSize.jpg"),
  , gsz, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r merge branch weight with FAdata and compute FA by branch}
#### code "merge branch weights, compute FA by branch"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### FAdata0 is branch level info
FAdata0 <- qread(paste0(pathdata, "FAdata.qs"), nthreads = 13)
#### Branch weight
#### wg12 <- qread(paste0(pathdata, "BranchWeights2012.qs"), nthreads = 13)
#### stmfp_2008_2022_e5_v1_Branch=stmfp_2008_2022_e5_v1+BranchWeights2012
iprw12 <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"), 
  nthreads = 13)
FAdata <- FAdata0[taxrefno != "", ]
#### multi (taxrefno-taxyear) - multi (taxrefno-branch) merge
#### Set allow.cartesian = T
#### Base = iprw12 and merge whichever matched rows in FAdata
meefa <- FAdata[iprw12, .(taxrefno, payereferenceno, Employees, 
    PropToMW, SubMW, NumSubMW, FA.orig, FA.25, FA.50, FA.75, 
    taxyear, UID, 
    busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, BranchNum, 
    wgt.kerr, Newgt.kerr, Txrf, TotalBranch), 
  on = "busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
     "busmainplc_geo", "taxrefno"), 
  allow.cartesian = T]
rbind(
    FAdata = dim(FAdata), 
    stmfp = dim(iprw12), 
    meefa = dim(meefa)
  )
print(meefa[BranchNum == 5, 
  .(taxyear, taxrefno, busmainplc_geo, UID, BranchNum)], 
  topn = 50)
#### Distributing number of workers, number of subMW workers
#### Assume that number of workers and subMW workers 
#### are proportional to branch incomes
meefa[, NumSubMWBranch := floor(NumSubMW*wgt.kerr/100)]
meefa[, EmployeesBranch := Employees*wgt.kerr/100]
meefa[, FABranch := round(NumSubMWBranch/EmployeesBranch, 6)]
meefa[, NeNumSubMWBranch := floor(NumSubMW*Newgt.kerr/100)]
meefa[, NeEmployeesBranch := Employees*Newgt.kerr/100]
meefa[, NeFABranch := round(NeNumSubMWBranch/NeEmployeesBranch, 6)]
#### Assume that a part of subMW workers are fixed
#### No good way to model it...
qsave(meefa, paste0(pathdata, "stmfp2008-2022e5v1_Branch_FAdata.qs"), 
  nthreads = 13)
meefa <- qread(paste0(pathdata, "stmfp2008-2022e5v1_Branch_FAdata.qs"), 
  nthreads = 13)

LShare0 <- qread(paste0(pathdata, "ShareHHI12.qs"), nthreads = 13)
LShare <- LShare0[, .(busprov_geo, 
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo,
    taxrefno, Txrf, 
    EachEmp, ShareEmp, HHI, ShareEmpG, HHIG, 
    EachNeEmp, ShareNeEmp, nHHI, ShareNeEmpG, nHHIG, 
    AreaLevel)]
meefaBranch <- unique(meefa[, .(taxyear, busprov_geo, 
  busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
  Txrf, taxrefno, BranchNum, TotalBranch, 
  Employees, PropToMW, SubMW, NumSubMW, 
  FA.orig, FA.25, FA.50, FA.75, 
  wgt.kerr, Newgt.kerr)])
#### check duplicates
dup <- function(x, indexonly=F) {
  require(data.table)
  ii <- data.table(
    cbind(fromFirst=duplicated(x), fromLast=duplicated(x, fromLast = T))
    )
  if (indexonly) 
    ii else 
    ii[, as.integer(fromFirst)+as.integer(fromLast) == 1L]
}
ii1 <- dup(LShare[, .(busmainplc_geo, taxrefno)])
if (any(ii1)) LShare[ii1, .(busmainplc_geo, taxrefno)]
ii2 <- dup(meefaBranch[, .(busmainplc_geo, taxrefno)])
if (any(ii2)) meefaBranch[ii2, .(busmainplc_geo, taxrefno)]
LShare[ii1, ][1:10, ]
meefaBranch[ii2, ][1:10, ]

meehhi <- LShare[meefaBranch, 
  on = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
     "busmainplc_geo", "taxrefno")]


LShareW <- reshape(LShare, direction = "wide", 
    idvar = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
      "busmainplc_geo", "taxrefno", "Txrf", "EachEmp", "EachNeEmp"),
    timevar = "AreaLevel",
    v.names = grepout("Sh|HH", colnames(LShare))
  )

print(
  unique(meefa[TotalBranch == 2L & taxyear == 2012 &
    taxrefno %in% FAdata[FA.orig != 0, taxrefno], 
    .(Txrf, taxrefno, payereferenceno, 
      Employees, NumSubMW, EmployeesBranch, NumSubMWBranch, 
      FA.orig, FABranch, wgt.kerr)][order(Txrf), ])
  , topn = 30)


```





```{r count number of UIDs in ipyrc}
#### obsolete code: "count number of UIDs"
ipyrc <- qread(paste0(pathdata, "ipyrc.qs"), nthreads = 13)
#### totalperiodsinyearofassessment
grepout("[^0-9]| ", unique(ipyrc[, totalperiodsinyearofassessment]))
ipyrc[, totpy := totalperiodsinyearofassessment]
ipyrc[, totpy := gsub("\\.| |-", "", totpy)]
ipyrc[, totpy := as.integer(totpy)]
ipyrc[is.na(totpy) & !is.na(totalperiodsinyearofassessment), 
  .(totpy, totalperiodsinyearofassessment)]
ipyrc[, totN := sum(totpy, na.rm = T), by = Txrf]
ipyrc[, .(busmainplc_geo, taxrefno, totpy, totN)]
#### density plot of before and after dropping obs 
#### with missing Txrf and main place info
irpBefore = copy(ipyrc[, .(busmainplc_geo, Txrf, totpy, totN)])
irpBefore[, timing := "before"]
irpAfter = irpBefore[busmainplc_geo != "" & Txrf != "", 
  .(busmainplc_geo, Txrf, totpy, totN)]
irpAfter[, timing := "after"]
irpBA <- rbindlist(list(irpBefore, irpAfter))
library(ggplot2)
g1 <- ggplot(data = irpBA, 
  aes(x = totpy, group = timing, color = timing)) +
  geom_density(alpha = .1) + 
  scale_color_manual(values = c("red", "blue"))
g1 <- g1 + xlim(0, 300)
g1 

for (yr in 12:22) {
#### for (yr in 12) {
  ipyr <- ipyrc[taxyear == 2000+yr, ]
  LShare2 <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachUID := as.integer(length(unique(UID))), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, TotalUID := as.integer(length(unique(UID))), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, ShareUID := round(EachUID/TotalUID, 8)]
    cat(GeoLevel[g], "\n")
    ipGeo[grepl("gov", Entity), 
      GTotalUID := as.integer(length(unique(UID))), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### copy GTotalUID to all rows
    ipGeo[, GTotalUID := GTotalUID[!is.na(GTotalUID)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachEmpG by replacing EachEmp of GovEntity with 0L
    ipGeo[, EachUIDG := EachUID]
    ipGeo[grepl("gov", Entity), EachUIDG := 0L]
    ipGeo[, TotalUIDG := TotalUID-GTotalUID]
    ipGeo[, ShareUIDG := round(EachUIDG/TotalUIDG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "irp5_kerr_iweight_b", "Newgt.kerr","wgt.kerr",
      "EachUID", "TotalUID", "ShareUID", 
      "EachUIDG", "TotalUIDG", "ShareUIDG"
      ), with = F])
    print(summary(lshare))
    lshare[, HHI2 := sum(ShareUID^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI2 := (HHI2-1/TotalEmp)/(1-1/TotalEmp)]
    lshare[, HHI2G := sum(ShareEmpG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI2G := (HHI2G-1/TotalEmpG)/(1-1/TotalEmpG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI2)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalUID", "EachUID", "ShareUID", 
      "HHI2", "nHHI2", "TotalUIDG"), with = F])
  }
  LShare2 <- rbindlist(list(LShare2, lshare), use.names = T, fill = T)
  setkey(LShare2, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare2, paste0(pathdata, "ShareHHI2", yr, ".qs"), nthreads = 13)
}
LShare2 <- qread(paste0(pathdata, "ShareHHI212.qs"))
LShare2[, Nums := .N, by = Txrf]
tb2 <- table(LShare2[, Nums])
numtb2 <- as.numeric(names(tb2))
numtb2 <- numtb2[order(-numtb2)]
numtb2[1:(length(numtb2)-5)]
tb2
LShare2[Nums %in% numtb2[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI)]

ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare2[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI2)]
print(LShare2[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI2)], topn = 25)
summary(LShare2[TotalNeUID <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI2)])
print(LShare2[TotalNeUID <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI2)], topn = 25)
print(LShare2[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachUID, TotalUID, ShareUID, HHI2)], topn = 25)
library(ggplot2)
ggplot(data = LShare2[!is.na(Newgt.kerr) & ShareUID != 0, ], 
  aes(x = ShareUID)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
summary(LShare2[!is.na(Newgt.kerr) & ShareUID != 0, ShareUID])
ggplot(data = unique(LShare2[!is.na(Newgt.kerr), .(busmainplc_geo, HHI2)]), 
  aes(x = HHI2)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
```


```{r hhi based on kerr income}
#### code "hhi based on kerr income"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
mee <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"),
  nthreads = 13)
mee[taxrefno == "AACBTBBAJZ", ]
print(mee[,
  .(taxrefno, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr, Neimputed.emp)], topn = 25)
summary(mee[,
  .(taxrefno, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr, Neimputed.emp)])
#### mee[, NumB := .N, by = taxrefno]
#### mee[, EmptyLocInfo := sum(busmainplc_geo == ""), by = taxrefno]
#### table(mee[, .(NumB, EmptyLocInfo)])
#### Not sure if we need this. Test how many obs will be dropped.
nrow(mee[tot.kerr != 0, ]); (nrow(mee[tot.kerr != 0, ])/nrow(mee))*100
nrow(mee[Neimputed.kerr != 0, ]); (nrow(mee[Neimputed.kerr != 0, ])/nrow(mee))*100
#### For 2012, no NA in payereferenceno
#### (n0 <- nrow(mee[is.na(payereferenceno) | payereferenceno == "", ])); (n0)/nrow(mee)*100
#### 14% are gov employees [is.na(taxrefno) | taxrefno == ""]
#### So we cannot simply drop entries with NAs in taxrefno
(n0 <- nrow(mee[grepl("NULL", taxrefno) | is.na(taxrefno) | taxrefno == "", ])); (n0/nrow(mee))*100
#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
mee[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
mee[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
mee[, Entity := "private"]
mee[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
mee[, Entity := factor(Entity, levels = c("private", "gov"))]
for (yr in 12:22) {
#### for (yr in 12) {
  ipyr <- mee[taxyear == 2000+yr, ]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachRows := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, EachNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, TotalRows := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalEmp := sum(EachEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, TotalNeEmp := sum(EachNeEmp, na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, ShareRows := round(EachRows/TotalRows, 8)]
    ipGeo[, ShareEmp := round(EachEmp/TotalEmp, 8)]
    ipGeo[, ShareNeEmp := round(EachNeEmp/TotalNeEmp, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachEmp, TotalEmp, ShareEmp)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), 
      GTotalNeEmp := round(sum(Neimputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[grepl("gov", Entity), 
      GTotalEmp := round(sum(imputed.emp, na.rm = T), 2), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### copy GTotalEmp to all rows
    ipGeo[, GTotalEmp := GTotalEmp[!is.na(GTotalEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotalNeEmp := GTotalNeEmp[!is.na(GTotalNeEmp)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachEmpG by replacing EachEmp of GovEntity with 0L
    ipGeo[, EachEmpG := EachEmp]
    ipGeo[grepl("gov", Entity), EachEmpG := 0L]
    ipGeo[, TotalEmpG := TotalEmp-GTotalEmp]
    ipGeo[, ShareEmpG := round(EachEmpG/TotalEmpG, 8)]
    ipGeo[, EachNeEmpG := EachNeEmp]
    ipGeo[grepl("gov", Entity), EachNeEmpG := 0L]
    ipGeo[, TotalNeEmpG := TotalNeEmp-GTotalNeEmp]
    ipGeo[, ShareNeEmpG := round(EachNeEmpG/TotalNeEmpG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "irp5_kerr_iweight_b", "Newgt.kerr","wgt.kerr",
      "EachEmp", "EachNeEmp", "TotalEmp", "TotalNeEmp", 
      "ShareEmp", "ShareNeEmp", 
      "EachEmpG", "TotalEmpG", "ShareEmpG", 
      "EachNeEmpG", "TotalNeEmpG", "ShareNeEmpG"
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(ShareNeEmp^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/TotalEmp)/(1-1/TotalEmp)]
    lshare[, HHIG := sum(ShareEmpG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalEmpG)/(1-1/TotalEmpG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", 
      "TotalEmp", "EachEmp", "ShareEmp", 
      "TotalNeEmp", "EachNeEmp", "ShareNeEmp", 
      "HHI", "nHHI", "TotalEmpG", "TotalNeEmpG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"), nthreads = 13)
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareNeEmp, HHI)]

ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)]
print(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
summary(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)])
print(LShare[TotalNeEmp <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
print(LShare[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
library(ggplot2)
ggplot(data = LShare[!is.na(Newgt.kerr), ], 
  aes(x = ShareEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
ggplot(data = LShare[!is.na(Newgt.kerr) & ShareEmp != 0, ], 
  aes(x = ShareNeEmp)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
summary(LShare[!is.na(Newgt.kerr) & ShareNeEmp != 0, ShareNeEmp])
ggplot(data = unique(LShare[!is.na(Newgt.kerr), .(busmainplc_geo, HHI)]), 
  aes(x = HHI)) +
  geom_density(colour="blue", alpha = .1) + 
  xlim(0, 1)
```


```{r hhi using kerr old}
#### code "hhi using kerr"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"

iprw12[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
iprw12[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
for (yr in 12:22) {
#### for (yr in 12) {
  if (yr == 12) 
    ipyr <- qread(paste0(pathdata, "ipyrsClean.qs"), 
      nthreads = 13) else 
    ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), 
      nthreads = 13)
  #### ipyr = copy(ipyr)[grepl("gungun", busdistmuni_geo) &
  ####    grepl("Richmond", buslocmuni_geo), ]
  ipyr[, Txrf := taxrefno]
  #### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
  ipyr[, Entity := "private"]
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
  ipyr[, Entity := factor(Entity, levels = c("private", "gov"))]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachNum := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, Total := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, Share := round(EachNum/Total, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachNum, Total, Share)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), GTotal := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotal := GTotal[!is.na(GTotal)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachNumG by replacing EachNum of GovEntity with 0L
    ipGeo[, EachNumG := EachNum]
    ipGeo[grepl("gov", Entity), EachNumG := 0L]
    ipGeo[, TotalG := Total-GTotal]
    ipGeo[, ShareG := round(EachNumG/TotalG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "EachNum", "Total", "Share", "EachNumG", "TotalG", "ShareG" 
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(Share^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/Total)/(1-1/Total)]
    lshare[, HHIG := sum(ShareG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalG)/(1-1/TotalG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"))
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, AreaLevel, 
    EachNum, Total, HHI)]
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, EachNum, Total, Share, HHI)]
```

```{r hhi this is older}
#### code "hhi"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
#### Use the "clean" version (anomalous uid's are dropped)
#### ipr2012 ==> ipyrsClean.qs
####   142638/11406422 = 1.25% of all rows are useless uid's
####   more than half of it is done by taxrefno ZAJJQCTBKC
#### ipyrs <- qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)
ci5 <- qread(paste0(pathdata, "ci5_2008-2022_e5v1.qs"), nthreads = 13)
wg12 <- qread(paste0(pathdata, "BranchWeights2012.qs"), nthreads = 13)
#### keep only nonempty taxrefno
ipyrs <- ci5[taxyear == 2012 & taxrefno != "", ]
wg12 <- wg12[taxrefno != "", ]
table(ipyrs[!grepl("^NULL$", taxrefno), 
  .(NumBranch = .N), by = taxrefno][, NumBranch])
#### ipyrs <- ipyrs[grepl("A", natureofperson), ]
iprw12 <- merge(ipyrs, wg12, by = "taxrefno", all = T)
table(
(NumRows <- iprw12[!grepl("^NULL$", taxrefno), 
  .(NumRows=.N), by = taxrefno])[, NumRows]
)
NumRows
NumRows[grepl("NULL|AAAAAAA", taxrefno), ]
iprw12[, NumRows :=.N, by = taxrefno]
print(iprw12[NumRows == 5, 
  .(taxrefno, NumRows, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, tot.kerr, wgt.kerr)][1:50, ],
  topn = 50)
iprw12[, imputed.kerr := tot.kerr*wgt.kerr]
iprw12[, Neimputed.kerr := Netot.kerr*Newgt.kerr]
iprw12[taxrefno == "AACBTBBAJZ", ]
#### iprw12[, NumB := .N, by = taxrefno]
#### iprw12[, EmptyLocInfo := sum(busmainplc_geo == ""), by = taxrefno]
#### table(iprw12[, .(NumB, EmptyLocInfo)])
#### Not sure if we need this. Test how many obs will be dropped.
nrow(ipyrs[kerr_income != 0, ]); (nrow(ipyrs[kerr_income != 0, ])/nrow(ipyrs))*100
nrow(ipyrs[Neimputed.kerr != 0, ]); (nrow(ipyrs[Neimputed.kerr != 0, ])/nrow(ipyrs))*100
#### For 2012, no NA in payereferenceno
#### (n0 <- nrow(ipyrs[is.na(payereferenceno) | payereferenceno == "", ])); (n0)/nrow(ipyrs)*100
#### 14% are gov employees [is.na(taxrefno) | taxrefno == ""]
#### So we cannot simply drop entries with NAs in taxrefno
(n0 <- nrow(ipyrs[grepl("NULL", taxrefno) | is.na(taxrefno) | taxrefno == "", ])); (n0/nrow(ipyrs))*100
#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
ipyrs[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipyrs[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
for (yr in 12:22) {
#### for (yr in 12) {
  if (yr == 12) 
    ipyr <- qread(paste0(pathdata, "ipyrsClean.qs"), 
      nthreads = 13) else 
    ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), 
      nthreads = 13)
  #### ipyr = copy(ipyr)[grepl("gungun", busdistmuni_geo) &
  ####    grepl("Richmond", buslocmuni_geo), ]
  ipyr[, Txrf := taxrefno]
  #### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
  ipyr[, Entity := "private"]
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
  ipyr[, Entity := factor(Entity, levels = c("private", "gov"))]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachNum := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, Total := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, Share := round(EachNum/Total, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachNum, Total, Share)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), GTotal := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotal := GTotal[!is.na(GTotal)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachNumG by replacing EachNum of GovEntity with 0L
    ipGeo[, EachNumG := EachNum]
    ipGeo[grepl("gov", Entity), EachNumG := 0L]
    ipGeo[, TotalG := Total-GTotal]
    ipGeo[, ShareG := round(EachNumG/TotalG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "EachNum", "Total", "Share", "EachNumG", "TotalG", "ShareG" 
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(Share^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/Total)/(1-1/Total)]
    lshare[, HHIG := sum(ShareG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalG)/(1-1/TotalG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"))
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, AreaLevel, 
    EachNum, Total, HHI)]
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, EachNum, Total, Share, HHI)]
```

## Merge MEE and HHI

```{r Merge MEE and HHI}
#### code "Merge MEE and HHI"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
mee <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"),
  nthreads = 13)
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"), nthreads = 13)
meeh <- LShare[mee, on = c("taxrefno", 
  "busprov_geo", "busdistmuni_geo", 
  "buslocmuni_geo", "busmainplc_geo"), all = T]
print(meeh[, .(busmainplc_geo, Txrf, taxyear, 
  EachNeEmp, TotalNeEmp, ShareNeEmp, HHI)], topn = 40)
qsave(meeh, paste0(pathdata, "stmfp_2008_2022_e5_v1_hhi.qs"), nthreads = 13)
```



## Checks 

### Checks on counting number of rows

```{r Check0 number of rows}
#### Check number of Government entities at main place level in 2012
#### Note: Names in mult.mnpl appear more than once
#### Need to use district, local municipality to separately count each of them
path <- "Z:/Workbenches/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
livbrary(qs); library(data.table); library(ggplot2)
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"), nthreads = 16)
ipyr[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
ipyr[, Entity := "private"]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ipyr[, Entity := factor(Entity, levels = c("private", "gov"))]
mnpl <- unique(LShare$busmainplc_geo)
mnpl <- mnpl[!is.na(mnpl)]
mnpl <- mnpl[mnpl != ""]
Check0 <- NULL
Check.wrong <- list()
iiwrong <- 1
mult.mnpl <- c("Cape Town", "Heidelberg", "Centurion", "eMalahleni NU", "Vereeniging", 
  "Balfour", "Kanana", "Middelburg", "New Rest", "Rietfontein")
mult.mnpl.collapsed <- paste(mult.mnpl[2:4], sep = "", collapse = "|")
for (ii in 1:length(mnpl)) {
    mnpname <- paste0("^", mnpl[ii], "$")
    ls.num <- LShare[
      grepl(mnpname, busmainplc_geo) & grepl("Gov", Txrf), 
      .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo, EachNum)]
    ip.num <- data.table(table(ipyr[
      grepl(mnpname, busmainplc_geo) & grepl("Gov", Txrf), 
      .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo)] ))
    #### Multiple instance of main place names: nrow(ls.num) > 1
    #### Convert table to data.table, and drop entries with N==0
    if (nrow(ls.num) > 1) ip.num <- ip.num[N!=0, ]
    #### Zero instance of main place names
    if (length(ls.num)==0) {
      ls.num <- data.table(t(rep(as.character(NA), 3)))
      setnames(ls.num, c("busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"))
      ls.num[, EachNum := as.integer(NA)]
    }
    if (length(ip.num)==0) {
      ip.num <- data.table(t(c("this main place", "is not found", mnpl[ii])))
      ip.num[, N := as.integer(NA)]
    }
    if (length(ls.num) > 4 | length(ip.num) > 4) {
      Check.wrong[[iiwrong]] <- list(ls.num, ip.num)
      names(Check.wrong)[iiwrong] <- mnpl[ii]
      iiwrong <- iiwrong + 1
    }
    lsip <- cbind(ls.num, ip.num)
    setnames(lsip, letters[1:ncol(lsip)])
    Check0 <- rbindlist(list(Check0, lsip), use.names = T, fill = T)
}
Check0
ls.num
ip.num
setnames(Check0, c("district.LS", "local.LS", "main.LS", "Gov.LS", 
  "district.ipyr", "local.ipyr", "main.ipyr", "Gov.ipyr"))
Check0
Check0[Gov.LS != Gov.ipyr, ]
qsave(Check0, paste0(pathdata, "CheckGov", yr, ".qs"), nthreads = 16)
```
```{r Check1 number of rows}
#### Check number of non-Government entities at main place level in 2012
#### Check number of Government entities at main place level in 2012
#### Note: Names in mult.mnpl appear more than once
#### Need to use district, local municipality to separately count each of them
path <- "Z:/Workbenches/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
livbrary(qs); library(data.table); library(ggplot2)
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"), nthreads = 16)
ipyr[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
ipyr[, Entity := "private"]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ipyr[, Entity := factor(Entity, levels = c("private", "gov"))]
mnpl <- unique(LShare$busmainplc_geo)
mnpl <- mnpl[!is.na(mnpl)]
mnpl <- mnpl[mnpl != ""]
Check1.ls <- Check1.ip <- NULL
Check.wrong1 <- list()
iiwrong <- 1
mult.mnpl <- c("Cape Town", "Heidelberg", "Centurion", "eMalahleni NU", "Vereeniging", 
  "Balfour", "Kanana", "Middelburg", "New Rest", "Rietfontein")
mult.mnpl.collapsed <- paste(mult.mnpl[2:4], sep = "", collapse = "|")
for (ii in c(1:10, grep(mult.mnpl.collapsed, mnpl))) {
#### for (ii in 1:length(mnpl)) {
    mnpname <- paste0("^", mnpl[ii], "$")
    ls.num <- LShare[
      grepl(mnpname, busmainplc_geo) & !grepl("Gov", Txrf), 
      .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf, EachNum)]
    #### Convert table to data.table
    ip.num <- data.table(table(ipyr[
      grepl(mnpname, busmainplc_geo) & !grepl("Gov", Txrf), 
      .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf)] ))
    #### Multiple instance of main place names: nrow(ls.num) > 1
    #### drop entries with N==0
    if (nrow(ls.num) > 1) ip.num <- ip.num[N!=0, ]
    #### Zero instance of main place names
    if (length(ls.num)==0) {
      ls.num <- data.table(t(rep(as.character(NA), 4)))
      setnames(ls.num, c("busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", "Txrf"))
      ls.num[, EachNum := as.integer(NA)]
    }
    if (length(ip.num)==0) {
      ip.num <- data.table(t(c("this main place", "is not found", mnpl[ii], "none found")))
      ip.num[, N := as.integer(NA)]
    }
    if (length(ls.num) > 5 | length(ip.num) > 5) {
      Check.wrong1[[iiwrong]] <- list(ls.num, ip.num)
      names(Check.wrong1)[iiwrong] <- mnpl[ii]
      iiwrong <- iiwrong + 1
    }
    setnames(ls.num, letters[1:ncol(ls.num)])
    setnames(ip.num, letters[1:ncol(ip.num)])
    Check1.ls <- rbindlist(list(Check1.ls, ls.num), use.names = T, fill = T)
    Check1.ip <- rbindlist(list(Check1.ip, ip.num), use.names = T, fill = T)
}
setkey(Check1.ls, a, b, c, d)
setkey(Check1.ip, a, b, c, d)
Check1 <- merge(Check.ls, Check.ip, by = letters[1:ncol(ls.num)], 
  all = T, suffixes = c(".LS", ".ipyr"))
Check1
ls.num
ip.num
setnames(Check1, c("district.LS", "local.LS", "main.LS", "Txrf.LS", "EachNum.LS", 
  "district.ipyr", "local.ipyr", "main.ipyr", "Txrf.ipyr", "N.ipyr"))
Check1
Check1[Txrf.LS == Txrf.ipyr & EachNum.LS != N.ipyr, ]
Check.wrong1
qsave(Check1, paste0(pathdata, "CheckPvt", yr, ".qs"), nthreads = 16)
```

### Checks and tests using Pietermarizberg, etc.

```{r Check PMT}
path <- "Z:/Workbenches/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
install.package("qs")
install.package("data.table")
install.package("ggplot2")
library(qs); library(data.table); library(ggplot2)
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipPTM0 <- ipyr[grepl("gungun", busdistmuni_geo), ]
ipPTM0 [grepl("Kupholeni", busmainplc_geo) & grepl("Richmond", buslocmuni_geo), 
  .(buslocmuni_geo, taxrefno)]

ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipPTM0 <- ipyr[grepl("gungun", busdistmuni_geo), ]
ipPTM0[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
ipPTM0[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
ipPTM0[, Entity := "private"]
ipPTM0[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ipPTM0[, Entity := factor(Entity, levels = c("private", "gov"))]
LShare <- NULL
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
for (g in 1:length(GeoLevel)) {
  ipGeo = copy(ipPTM0)
  ipGeo[, EachNum := as.integer(.N), by = 
    eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
  ipGeo[, Total := as.integer(.N), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  ipGeo[, Share := round(EachNum/Total, 8)]
  #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
  #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
  ipGeo[grepl("gov", Entity), GTotal := as.integer(.N), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  ipGeo[, GTotal := GTotal[!is.na(GTotal)][1], by =
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  #### We define EachNumG by replacing EachNum of GovEntity with 0L
  ipGeo[, EachNumG := EachNum]
  ipGeo[grepl("gov", Entity), EachNumG := 0L]
  ipGeo[, TotalG := Total-GTotal]
  ipGeo[, ShareG := round(EachNumG/TotalG, 8)]
  ipGeo[, AreaLevel := GeoLevel[g]]
  lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], 
    "Entity", "Txrf",  #"taxrefno",
    "AreaLevel", 
    "EachNum", "Total", "Share", "EachNumG", "TotalG", "ShareG" 
    ), with = F])
  lshare[, HHI := sum(Share^(2), na.rm = T), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  lshare[, nHHI := (HHI-1/Total)/(1-1/Total)]
  lshare[, HHIG := sum(ShareG^(2), na.rm = T), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  lshare[, nHHIG := (HHIG-1/TotalG)/(1-1/TotalG)]
  print(GeoLevel[g])
  print(lshare[1:10, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[g], 
    "Txrf", "Total", "EachNum", "Share", "HHI", "nHHI", "TotalG"), with = F])
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
}

LShare[grepl("Pieter", busmainplc_geo) & grepl("The Msunduzi", buslocmuni_geo) & grepl("Gov", Txrf), 
  .(EachNum, TotalG, Total)]
table(ipPTM0[grepl("Pieter", busmainplc_geo) & grepl("The Msunduzi", buslocmuni_geo) & grepl("Gov", Txrf), 
  .(buslocmuni_geo, busmainplc_geo)])
```

# Merge Data file and HHI file (LShare 2012)

* Data file: `../data/DataFile.tsv`
* HHI file: `../Data/LShare12.qs`
* Both files need to have `taxrefno`


```{r merge Data file and HHI file}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs)
library(data.table)
library(readstata13)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
#### df <- fread(paste0(pathdata, "DataFile.tsv"), sep = "\t", fill = T, 
####  na.strings = "", nThread = 15)
df <- read.dta13(paste0(pathdata, "DataFile.dta"))
df <- data.table(df)
df[, .(buslocmuni_geo, busmainplc_geo, natureofperson, 
  taxrefno, kerr_income, kerr_emp, passportno)]
ls <- qread(paste0(pathdata, "ShareHHI12.qs"), nthreads = 16)
setnames(ls, "Txrf", "taxrefno")
setkey(df, busprov_geo, busdistmuni_geo, buslocmuni_geo, 
  busmainplc_geo)
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
for (g in 1:length(GeoLevel)) {
#### for (g in 1) {
  cat(GeoLevel[g], "\n")
  lsGeo <- unique(ls[grepl(GeoLevel[g], AreaLevel), 
    .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
      busmainplc_geo, HHI)])
  setnames(lsGeo, "HHI", paste0("HHI.", GeoLevel[g]))
  setkey(lsGeo, busprov_geo, busdistmuni_geo, buslocmuni_geo, 
      busmainplc_geo)
  print(dim(df)); print(dim(lsGeo))
  #### df <- merge(df, lsGeo, by = c("busprov_geo", "busdistmuni_geo", 
  ####  "buslocmuni_geo", "busmainplc_geo","taxrefno"), 
  ####  all = T, allow.cartesian = T)
  df2 <- lsGeo[df]
  print(dim(df2))
  print(df2[, c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
    "busmainplc_geo", "taxrefno", grepout("HHI", colnames(df2)))])
}
```




# Other

Get observation pattern: ObPattern

```{r create ObPattern, eval = F}
#### icount data is created in chunk "read irp5v5beta"
Icount <- qread(paste0(pathdata, "icount08.qs"), nthreads = 8)
Icount[, Taxyear := 2008]
for (yr in 9:20) {
  if (yr < 10) yr <- paste0("0", yr)
  icount <- qread(paste0(pathdata, "icount", yr, ".qs"), nthreads = 8)
  icount[, Taxyear := 2000+as.numeric(yr)]
  Icount <- rbindlist(list(Icount, icount))
}
qsave(Icount, paste0(pathdata, "Icount.qs"), nthreads = 8)
setkey(Icount, taxrefno, Taxyear)
Icount[, Ob := 1L]
Icount[is.na(Num), Ob := 2L]
Icount <- Icount[grepl("A", natureofperson), ]
Icount <- Icount[busmainplc_geo != "", ]
IcountW <- reshape(Icount, direction = "wide", 
  idvar = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", "taxrefno", "natureofperson"),
  timevar = "Taxyear", v.names= c("Num", "Ob"))

grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]

cols <- grepout("Ob", colnames(IcountW))
IcountW[, (cols) := lapply(.SD, function(x) {x[is.na(x)] <- 0L; x}), .SDcols = cols]
IcountW[, ObPattern := paste(.SD, sep = "", collapse = ""), .SDcols = cols, by = .(taxrefno)]
IcountW[, ObPattern := factor(ObPattern)]
```
Multiple taxrefno detected. Because there are more than one establishment per firm. 
```{r}
IcountW[, ObPattern := paste(.SD, sep = "", collapse = ""), .SDcols = cols, by = .(taxrefno, busmainplc_geo)]
IcountW[, ObPattern := factor(ObPattern)]
IcountW2[, (cols) := NULL]
qsave(IcountW2, paste0(pathdata, "IcountW2.qs"), nthreads = 8)
IcountL <- reshape(IcountW2, direction = "long", 
  idvar = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", "taxrefno"),
  timevar = "Taxyear", varying = grepout("Num", colnames(IcountW)))
qsave(IcountL, paste0(pathdata, "IcountL.qs"), nthreads = 8)
```

Still multiple establishments per main place. Need to sum them up to be a single establishment. Fine as long as hiring decisions are shared between establishments in a given main place, which is likely.

```{verbatim}
   taxrefno         natureofperson        Num.2012        Ob.2012    
 Length:44509       Length:44509       Min.   :    1   Min.   :0.00  
 Class :character   Class :character   1st Qu.:    2   1st Qu.:1.00  
 Mode  :character   Mode  :character   Median :    7   Median :1.00  
                                       Mean   : 1052   Mean   :0.76  
                                       3rd Qu.:   29   3rd Qu.:1.00  
                                       Max.   :99216   Max.   :1.00  
                                       NA's   :10680                 
    Num.2008        Ob.2008                ObPattern          Est         
 Min.   :    1   Min.   :0.0000   01            :10667   Min.   :  1.000  
 1st Qu.:    2   1st Qu.:0.0000   10            :30901   1st Qu.:  1.000  
 Median :    5   Median :0.0000   11            : 2911   Median :  1.000  
 Mean   :  176   Mean   :0.3054   c(1, 1)c(0, 0):   14   Mean   :  3.287  
 3rd Qu.:   15   3rd Qu.:1.0000   0:11:0        :    2   3rd Qu.:  1.000  
 Max.   :76899   Max.   :1.0000   c(0, 0)c(1, 1):   12   Max.   :280.000  
 NA's   :30917                    c(1, 1)1:0    :    2             
```

### Merge firms22 with CIT-IRP5 (ci5p)

```{r Merge firms22 with CIT-IRP5}
firms22 <- qread(paste0(pathdata, "firms22.qs")) 
firms22[, taxyear := 2022]
setkey(ci5p, taxrefno, taxyear)
setkey(firms22, taxrefno, taxyear)
ci5pf <- firms22[ci5p]
ci5pf[!is.na(busmainplc_geo),  busmainplc_geo]
```

```{verbatim}
> dim(ci5pf)
[1] 705166   1104
> dim(ci5p)
[1] 533396   1098
> dim(firms22)
[1] 13047047        8
```

```{r check distribution of Num}
ip21[, Num := .N, by = .(taxrefno)]
icount21 <- unique(ip21[, .(busprov_geo, busdistmuni_geo, buslocmuni_geo, taxrefno, natureofperson, Num)])
icount21 <- icount21[taxrefno != "", ]
library(ggplot2)
g <- ggplot(data=icount21, aes(Num)) +
  geom_histogram(bins=50) 
g + scale_x_log10()
```
```{r check distribution of Num log 10 for Y}
g + scale_y_log10()
```



```


```{r hhi 21, eval = F}
icount21[, NLocMuni := sum(Num), by = .(buslocmuni_geo)]
icount21[, NDistrict := sum(Num), by = .(busdistmuni_geo)]
#### When oligopsony matters, it shows up in HHI *and* own share in aggregate employment (own dominance).
icount21[, ShareMainPlace := Num/NMainPlace]
icount21[, ShareLocMuni := Num/NLocMuni]
icount21[, ShareDistrict := Num/NDistrict]
icount21[, HHIMainPlace := sum(ShareMainPlace^(2)), by = busmainplc_geo]
icount21[, HHILocMuni := sum(ShareLocMuni^(2)), by = buslocmuni_geo]
icount21[, HHIDistrict := sum(ShareMainPlace^(2)), by = busdistmuni_geo]
print(unique(icount21[, .(busdistmuni_geo, NDistrict)]), nrows=1100L)
print(unique(icount21[, .(buslocmuni_geo, NLocMuni)]), nrows=300L)
```



Postal codes in stata

```{verbatim}
. use "Z:\Master Data\IRP5\Job level\v5\beta\aux_postal_codes_2008_2023_e6_v2.dta" , replace
. by AddrPostalCode, sort: gen nvals = _n == 1
. count if nvals
  68,140
```

Other info:

* Dates are formatted as: YYYY-MM-DD
* `busprop_geo`: Northern Cape, KwaZulu-Natal
* `busidistmuni_geo`: Thabo Mofut...

For an exercise purpose, slim down data to include only a few industries. 

```{r read slimmed down data}
cip5 <- qread(paste0(pathdata, "ci5p.qs"), nthreads = 16)
ci5p[, Num := .N, by = .(taxrefno, taxyear)]
```

# Merge rainfall data (incomplete)

CreateAnalsysDataSet_final.do
line 31 
rainfall data ends at 2019, need until 2022


### obsolete 

```{r hhi}
#### code "hhi"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
mee <- qread(paste0(pathdata, "stmfp_2008_2022_e5_v1_Branch.qs"))
mee[taxrefno == "AACBTBBAJZ", ]
#### mee[, NumB := .N, by = taxrefno]
#### mee[, EmptyLocInfo := sum(busmainplc_geo == ""), by = taxrefno]
#### table(mee[, .(NumB, EmptyLocInfo)])
#### Not sure if we need this. Test how many obs will be dropped.

nrow(mee[kerr_income != 0, ]); (nrow(mee[kerr_income != 0, ])/nrow(mee))*100
nrow(mee[Neimputed.kerr != 0, ]); (nrow(mee[Neimputed.kerr != 0, ])/nrow(mee))*100
#### For 2012, no NA in payereferenceno
#### (n0 <- nrow(mee[is.na(payereferenceno) | payereferenceno == "", ])); (n0)/nrow(mee)*100
#### 14% are gov employees [is.na(taxrefno) | taxrefno == ""]
#### So we cannot simply drop entries with NAs in taxrefno
(n0 <- nrow(mee[grepl("NULL", taxrefno) | is.na(taxrefno) | taxrefno == "", ])); (n0/nrow(mee))*100
#### 1. Use payereferenceno to count the total
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
mee[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
mee[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
#### for (yr in 12:22) {
for (yr in 12) {
  if (yr != 12) 
    ipyr <- mee else 
    ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), 
      nthreads = 13)
  #### ipyr = copy(ipyr)[grepl("gungun", busdistmuni_geo) &
  ####    grepl("Richmond", buslocmuni_geo), ]
  ipyr[, Txrf := taxrefno]
  #### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
  ipyr[, Entity := "private"]
  ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
  ipyr[, Entity := factor(Entity, levels = c("private", "gov"))]
  LShare <- NULL
  GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
  for (g in 1:length(GeoLevel)) {
    ipGeo = copy(ipyr)
    ipGeo[, EachNum := as.integer(.N), by = 
      eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
    ipGeo[, Total := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, Share := round(EachNum/Total, 8)]
    cat(GeoLevel[g], "\n")
    print(summary(ipGeo[, .(EachNum, Total, Share)]))
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
    ipGeo[grepl("gov", Entity), GTotal := as.integer(.N), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    ipGeo[, GTotal := GTotal[!is.na(GTotal)][1], by =
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    #### We define EachNumG by replacing EachNum of GovEntity with 0L
    ipGeo[, EachNumG := EachNum]
    ipGeo[grepl("gov", Entity), EachNumG := 0L]
    ipGeo[, TotalG := Total-GTotal]
    ipGeo[, ShareG := round(EachNumG/TotalG, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
    lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo"), 
      "Entity", "Txrf",  "taxrefno",
      "AreaLevel", 
      "EachNum", "Total", "Share", "EachNumG", "TotalG", "ShareG" 
      ), with = F])
    print(summary(lshare))
    lshare[, HHI := sum(Share^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHI := (HHI-1/Total)/(1-1/Total)]
    lshare[, HHIG := sum(ShareG^(2), na.rm = T), by = 
      eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
    lshare[, nHHIG := (HHIG-1/TotalG)/(1-1/TotalG)]
    print(GeoLevel[g])
    setkey(lshare, Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, HHI)
    print(lshare[, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
    print(lshare[1:30, c("busprov_geo", "busdistmuni_geo", 
      "buslocmuni_geo", "busmainplc_geo", 
      "Txrf", "Total", "EachNum", "Share", 
      "HHI", "nHHI", "TotalG"), with = F])
  }
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  setkey(LShare, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"))
}
#### LShare:  Txrf, EachNum, Share, HHI, ShareG, HHIG, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, AreaLevel, 
    EachNum, Total, HHI)]
ipyr <- qread(paste0(pathdata, "irp12.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, EachNum, Total, Share, HHI)]
```

```{r, eval = F}
x <- data.table(
  A = rep(letters[1], 10),
  B = rep(letters[1+10], 10),
  C = rep(letters[20], 10)
  )
tbx <- table(x)
dtb.tbx <- data.table(tbx)

x <- data.table(
  A = sample(letters[1:3], 10, replace = T),
  B = sample(letters[1:3+10], 10, replace = T),
  C = letters[20]
  )
x[c(2, 7, 9), c("A", "B") := ""]
tbx <- table(x)
dtb.tbx <- data.table(tbx)
dtb.tbx <- dtb.tbx[N!=0, ]



#### old check, start ####
mnpl <- unique(LShare$busmainplc_geo)
mnpl <- mnpl[!is.na(mnpl)]
mnpl <- mnpl[mnpl != ""]
length(mnpl)
Check <- NULL
for (ii in 1:length(mnpl)) {
    mnpname <- mnpl[ii]
    ls.num <- LShare[grepl(mnpl[ii], busmainplc_geo) & grepl("Gov", Txrf), EachNum]
    ip.num <- as.numeric(table(ipyr[grepl(mnpl[ii], busmainplc_geo) & grepl("Gov", Txrf), 
          .(busmainplc_geo)]))
    if (length(ls.num)==0) ls.num <- NA
    if (length(ip.num)==0) ip.num <- NA
    Check <- c(Check, c(mnpname, ls.num, ip.num))
}
Check <- data.table(matrix(Check, byrow = T, ncol = 3))
setnames(Check, c("MainPlace", "LShare", "ipyr"))
qsave(Check, paste0(pathdata, "Check12.qs"), nthreads = 16)
Check
#### old check, end ####


#### check duplicated main places in capetown
ii <- 1
mnpname <- mnpl[ii]
LShare[grepl(mnpl[ii], busmainplc_geo) & grepl("Gov", Txrf), 
  .(busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf, EachNum)]



#### Keep only Pietermaritzberg
#### nrow(ipPTM0 <- ipyr[grepl("gungun", busdistmuni_geo), ])
#### Keep only ORTambo district muni, E Cape
#### nrow(ipORT0 <- ipyr[grepl("Tambo", busdistmuni_geo), ])
#### data.table::fwrite(ipORT0, paste0(pathdata, "IpORT0.prn"), sep = "\t", quote = F, nThread = 16)
ip = copy(ipyr)
#### Drop 2nd GovEntity entries at each geo level <== wrong? I guess.
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
for (g in 1:length(GeoLevel)) {
  ip[, Gov2 := 0L]
  ip[grepl("Gov", Txrf), Gov2 := as.integer(1:.N), by=
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  #### ip <- ip[Gov2 != 2, ]
}
ip[, Entity := "private"]
ip[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
ip[, Entity := factor(Entity, levels = c("private", "gov"))]
#### table(ipry[, Entity], exclude = NULL)
#### Count shares and HHI for the entire country
#### #Keep only Capetown (obsolete)
#### #nrow(ipCPT0 <- ipyr[grepl("Cape", busdistmuni_geo), ])
LShare <- NULL
for (g in 1:length(GeoLevel)) {
  ipGeo = copy(ip)
  ipGeo[, EachNum := as.integer(.N), by = 
    eval(c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "Txrf"))]
  ipGeo[, Total := as.integer(.N), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  ipGeo[, Share := round(EachNum/Total, 8)]
  #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
  #### We compute "total without GovEntity" by subtracting GTotal (nrow of Entity==gov)
  ipGeo[grepl("gov", Entity), GTotal := as.integer(.N), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  ipGeo[, GTotal := GTotal[!is.na(GTotal)][1], by =
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  #### We define EachNumG by replacing EachNum of GovEntity with 0L
  ipGeo[, EachNumG := EachNum]
  ipGeo[grepl("gov", Entity), EachNumG := 0L]
  ipGeo[, TotalG := Total-GTotal]
  ipGeo[, ShareG := round(EachNumG/TotalG, 8)]
  ipGeo[, AreaLevel := GeoLevel[g]]
  lshare <- unique(ipGeo[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], 
    "Entity", "Txrf",  #"taxrefno",
    "AreaLevel", 
    "EachNum", "Total", "Share", "EachNumG", "TotalG", "ShareG" 
    ), with = F])
  lshare[, HHI := sum(Share^(2), na.rm = T), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  lshare[, nHHI := (HHI-1/Total)/(1-1/Total)]
  lshare[, HHIG := sum(ShareG^(2), na.rm = T), by = 
    eval(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  lshare[, nHHIG := (HHIG-1/TotalG)/(1-1/TotalG)]
  print(GeoLevel[g])
  print(lshare[1:10, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[g], 
    "Txrf", "Total", "EachNum", "Share", "HHI", "nHHI", "TotalG"), with = F])
  LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
}
setkey(LShare, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf)
qsave(LShare, paste0(pathdata, "ShareHHI12.qs"), nthreads = 16)



LShare[, TaxRef := as.numeric(as.factor(Txrf))]
LShare[, Gov := 0L]
LShare[grepl("Gov", Txrf), Gov := 1L]
LShare[Gov == 1, TaxRef := "GovEntity"]
for (g in 1:length(GeoLevel)) {
  print(LShare[grepl(GeoLevel[g], AreaLevel) & grepl("Impendle", buslocmuni_geo), 
    .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    AreaLevel, TaxRef, #taxrefno, 
    Total, EachNum, Share, HHI, nHHI, TotalG)])
}



for (g in 1:length(GeoLevel)) {
  print(LShare[grepl(GeoLevel[g], AreaLevel) & grepl("gungun", busdistmuni_geo), 
    .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    AreaLevel, #Txrf, 
    Total, EachNum, Share, HHI, nHHI, TotalG)][1:20, ])
}
#qsave(LShare, paste0(pathdata, "ShareHHI12.qs"), nthreads = 16)

data.table::fwrite(LShare, paste0(pathdata, "ShareHHI12Test.prn"),
  sep = "\t", quote = F, nThread = 16)


#### Number of entries (employees) at province level
ipyr[, NumPrv := as.integer(.N), 
  by = .(busprov_geo, payereferenceno)]
summary(ipyr[, NumPrv])
ipyr[is.na(NumPrv), ][1:2, ]


grepl("^NA$", ipyr[, payereferenceno])
#### ipyr <- ipyr[kerr_income != 0, ]
ipyr[, .(NAInPayeref = is.na(payereferenceno) | payereferenceno == "", 
  , ]


#### Number of entries (employees) in entire SA
ipyr[, NumSA := as.integer(.N), 
  by = .(taxrefno)]
ipyr[is.na(NumSA) | grepl("NULL", taxrefno), NumSA := 0L]
ipyr[grepl("NULL", taxrefno) | taxrefno == "", NumSA := NA]
#### Number of entries (employees) at province level
ipyr[, NumPrv := as.integer(.N), 
  by = .(busprov_geo, taxrefno)]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), NumPrv := NA]
summary(ipyr[, NumPrv])
ipyr[is.na(NumPrv), ][1:2, ]
#### Number of entries (employees) at district level
ipyr[, NumDis := as.integer(.N), 
  by = .(busprov_geo, busdistmuni_geo, taxrefno)]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), NumDis := NA]
#### Number of entries (employees) at local municipality level
ipyr[, NumLoc := as.integer(.N), 
  by = .(busprov_geo, busdistmuni_geo, buslocmuni_geo, taxrefno)]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), NumLoc := NA]
#### Number of entries (employees) at main place level
ipyr[, NumMai := as.integer(.N), 
  by = .(busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, taxrefno)]
ipyr[is.na(NumMai) | grepl("NULL", taxrefno), NumMai := 0L]
ipyr[grepl("NULL", taxrefno) | taxrefno == "", NumMai := NA]
#### Keep only Pretoria and Sandton
IpSP0 <- ipyr[grepl("Sandton|Pretoria", busmainplc_geo), ]
#### Keep only one entry per firm at various levels
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
InSP <- NULL
for (g in 1:length(GeoLevel)) {
  #### Keep unique entries of texrefno, NumGEOLEVEL
  IpSP1 <- unique(IpSP0[, c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], 
    "taxrefno", "natureofperson", 
    paste0("Num", GeoLevel[g])), with = F])
  #### Count number of rows per taxrefno
  IpSP1[, NumRow := as.integer(.N), by = 
    c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g], "taxrefno")]
  tbnum <- table(table(IpSP1[, NumRow]))
  if (length(tbnum) > 1 | tbnum != 1) 
    stop(message = paste("More than 1 entry per taxrefno at", GeoLevel[g], "\n"))
  #### Copy NumGEOLEVEL to Num
  IpSP1[, Num := eval(parse(text=paste0("Num", GeoLevel[g])))]
  #### Sum employees in area
  IpSP1[, Total := as.integer(sum(Num, na.rm = T))
    , by = c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  #### Firm shares of employees, HHI in area
  IpSP1[, Share := round(Num/Total, 6)]
  IpSP1[, HHI := sum(Share^(2), na.rm = T)
    , by = c(c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo")[1:g])]
  IpSP1[, AreaLevel := GeoLevel[g]]
  InSP <- rbindlist(list(InSP, IpSP1), use.names = T, fill = T)
}
#qsave(InSP, paste0(pathdata, "TestData_SandtonPretoria20.qs"), nthreads = 16)
qsave(InSP, paste0(pathdata, "TestData_SandtonPretoria12.qs"), nthreads = 16)
InSP


for (yr in 9:20) {
  if (yr < 10) yr <- paste0("0", yr)
  Ic <- qread(paste0(pathdata, "Icount", yr, ".qs"), nthreads = 8)
  #### Drop from computation
  #### 1. unmatched "sole proprietor" (there are 260900 rows with taxrefno=NULL)
  #### 2. empty taxrefno
  Ic[grepl("NULL", taxrefno) | taxrefno == "", Num := NA]
  #### Num: Number of employees in a firm
  #### Num2: Number of employees in a firm, treat taxrefno = "NULL" as sole proprietors
  #### NMainPlace: Total number of employees in MainPlace
  colns <- c("NMainPlace", "NLocMuni", "NDistrict", 
    "ShareMainPlace", "ShareLocMuni", "ShareDistrict",
    "HHIMainPlace", "HHILocMuni", "HHIDistrict")
  for (added in c(2, "")) {
    NuM <- paste0("Num", added)
    SumNuM <- paste0("sum(", NuM, ", na.rm = T)")
    Ic[, NMainPlace := eval(parse(text=SumNuM)), by = .(busmainplc_geo)]
    Ic[, NLocMuni := eval(parse(text=SumNuM)), by = .(buslocmuni_geo)]
    Ic[, NDistrict := eval(parse(text=SumNuM)), by = .(busdistmuni_geo)]
    #### When oligopsony matters, it shows up in HHI *and* own share in aggregate employment (own dominance).
    Ic[, ShareMainPlace := eval(parse(text=paste0(NuM, "/NMainPlace")))]
    Ic[, ShareLocMuni := eval(parse(text=paste0(NuM, "/NLocMuni")))]
    Ic[, ShareDistrict := eval(parse(text=paste0(NuM, "/NDistrict")))]
    #### HHI at each level
    Ic[, HHIMainPlace := sum(ShareMainPlace^(2), na.rm = T), by = busmainplc_geo]
    Ic[, HHILocMuni := sum(ShareLocMuni^(2), na.rm = T), by = buslocmuni_geo]
    Ic[, HHIDistrict := sum(ShareMainPlace^(2), na.rm = T), by = busdistmuni_geo]
    if (added == 2) setnames(Ic, colns, paste0(colns, added))
  }
  qsave(Ic, paste0(pathdata, "Ic", yr, ".qs"))
}
Ic
summary(Ic)
Ic[grepl("NULL", taxrefno) & grepl("EXCE", busprov_geo), ]
```
Check anomalous entries.
```{r check anomalous entries}
Ic <- qread(paste0(pathdata, "Ic20.qs"))
Ic[grepl("NULL", taxrefno) | taxrefno == "" | busprov_geo == "", .(buslocmuni_geo, Num, Num2, kerr_income)] 
table(Ic[grepl("NULL", taxrefno) | taxrefno == "" | busprov_geo == "", 
  .(locmuni=buslocmuni_geo, kerr=kerr_income>0)],  useNA = "ifany")
Ic <- qread(paste0(pathdata, "Ic20.qs"))
Ic[, .(buslocmuni_geo, busmainplc_geo, taxrefno, Tref, Num, Num2, kerr_income, 
  NMainPlace, NLocMuni, ShareMainPlace, ShareLocMuni, HHIMainPlace, HHILocMuni)] 
Ic[grepl("Sandton", busmainplc_geo), ]

Icnt <- qread(paste0(pathdata, "Icount20.qs"), nthreads = 8)


#### Market shares
NuM <- "Num"
SumNuM <- paste0("sum(", NuM, ", na.rm = T)")
IpSP[, NMainPlace := eval(parse(text=SumNuM)), by = .(busmainplc_geo)]
IpSP[, NMainPlace3 := eval(parse(text=SumNuM)), 
  by = .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo)]
IpSP[, NLocMuni := eval(parse(text=SumNuM)), by = .(buslocmuni_geo)]
IpSP[, NLocMuni3 := eval(parse(text=SumNuM)), by = .(busdistmuni_geo, buslocmuni_geo)]
IpSP[, ShareMainPlace := eval(parse(text=paste0(NuM, "/NMainPlace")))]
IpSP[, ShareMainPlace3 := eval(parse(text=paste0(NuM, "/NMainPlace3")))]
IpSP[, ShareLocMuni := eval(parse(text=paste0(NuM, "/NLocMuni")))]
IpSP[, ShareLocMuni3 := eval(parse(text=paste0(NuM, "/NLocMuni3")))]
IpSP[grepl("Sandton", busmainplc_geo), 
  .(busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    NMainPlace, NMainPlace3, ShareMainPlace, ShareMainPlace3,
    NLocMuni, NLocMuni3, ShareLocMuni, ShareLocMuni3,
    taxrefno, natureofperson, kerr_income, Num)]
IpSP[grepl("Pretoria", busmainplc_geo), 
  .(busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    NMainPlace, NMainPlace3, ShareMainPlace, ShareMainPlace3,
    NLocMuni, NLocMuni3, ShareLocMuni, ShareLocMuni3,
    taxrefno, natureofperson, kerr_income, Num)]


NuM <- "Num"
SumNuM <- paste0("sum(", NuM, ", na.rm = T)")
Ipyr[, NMainPlace := eval(parse(text=SumNuM)), by = .(busmainplc_geo)]
Ipyr[, NMainPlace3 := eval(parse(text=SumNuM)), 
  by = .(busdistmuni_geo, buslocmuni_geo, busmainplc_geo)]
Ipyr[, ShareMainPlace := eval(parse(text=paste0(NuM, "/NMainPlace")))]
Ipyr[, ShareMainPlace3 := eval(parse(text=paste0(NuM, "/NMainPlace3")))]
Ipyr[grepl("Sandton", busmainplc_geo), 
  .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, NMainPlace, NMainPlace3, ShareMainPlace, ShareMainPlace3,
    taxrefno, natureofperson, kerr_income, Num)]
unique(
  Ipyr[grepl("Sandton", busmainplc_geo), 
    .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
      busmainplc_geo, NMainPlace, NMainPlace3, ShareMainPlace, ShareMainPlace3,
      taxrefno, natureofperson, kerr_income, Num)]
  )


colns <- c("NMainPlace", "NLocMuni", "NDistrict", 
  "ShareMainPlace", "ShareLocMuni", "ShareDistrict",
  "HHIMainPlace", "HHILocMuni", "HHIDistrict")

```

Plot time series profiles of HHIs. 

```{r merge IcYR files, eval = F}
IcL <- NULL
for (yr in 9:20) {
  if (yr < 10) yr0 <- paste0("0", yr)
  Ic <- qread(paste0(pathdata, "Ic", yr0, ".qs"), nthreads = 8)
  Ic[, Year := as.integer(2000+yr)]
  Ic[, Year0 := yr0]
  IcL <- rbindlist(list(IcL, Ic), use.names = T, fill = T)
}
qsave(IcL, paste0(pathdata, "IcL.qs"))
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
grepout("^N", colnames(IcL))
IcL[, Year := NULL]
IcW <- reshape(IcL, direction = "wide", 
  idvar = c("busmainplc_geo", "buslocmuni_geo", "busdistmuni_geo", 
    "busprov_geo", "taxrefno"), timevar = "Year0",
  v.names = grepout("^N|^Sha|^HH", colnames(IcL)))
qsave(IcW, paste0(pathdata, "IcW.qs"))
```


