---
title: "data.table pitfalls and fixes — worked examples (Session 17 Cedar Lark)"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    code-overflow: wrap
---

Worked, executable examples behind the Session 17 review findings on
`IRP5Condense.rmd`. Each section shows the failing construct on a small
data.table, the observed (wrong) result, and the fix with its result.

```{r setup}
#| message: false
library(data.table)
```

# A2. `paste(.SD, collapse = "")` cannot build per-row patterns

`.SD` is a *list of column vectors*. `paste()` receives it as **one**
argument, so it deparses each entire column into a single string and
`collapse` glues those column-strings into one scalar, which `:=` then
recycles to every row.

```{r a2-wrong-vs-right}
dt <- data.table(Ob.2008 = c(1L, 0L, 1L), Ob.2009 = c(0L, 1L, 1L))
dt[, wrong := paste(.SD, sep = "", collapse = ""), .SDcols = c("Ob.2008", "Ob.2009")]
dt[, right := do.call(paste0, .SD),                .SDcols = c("Ob.2008", "Ob.2009")]
dt
```

* `wrong` is the same deparsed-columns string in every row — a factor built
  on it has exactly one level, and at IRP5 scale (15 columns x ~50M rows)
  that one string is enormous.
* `right` is the intended per-row observation pattern.

Step by step, what `paste()` saw:

```{r a2-step-by-step}
#### Step 1-2: the single list argument is coerced column-by-column
as.character(list(dt$Ob.2008, dt$Ob.2009))
#### Step 3: collapse joins the two column-strings into ONE scalar
paste(as.character(list(dt$Ob.2008, dt$Ob.2009)), collapse = "")
```

`do.call(paste0, .SD)` instead expands the list into separate parallel
arguments — literally `paste0(Ob.2008, Ob.2009)` — and `paste0` vectorizes
across rows.

# A3. `DropThisForRev`: pooled `i`, inverted target, duplicate person-years

Mini panel. Person A revised their 2011 report in 2012; person B revised
2012 in 2013; person C revised 2021 in 2022. Originals have
`TrueTaxYear == taxyear`; revisions have `TrueTaxYear < taxyear`.

```{r a3-data}
irp <- data.table(
  person      = c("A",  "A",  "B",  "B",  "C",  "C"),
  taxyear     = c(2011, 2012, 2012, 2013, 2021, 2022),
  TrueTaxYear = c(2011, 2011, 2012, 2012, 2021, 2021),
  role        = c("original 2011", "revision of 2011 (filed 2012)",
                  "original 2012", "revision of 2012 (filed 2013)",
                  "original 2021", "revision of 2021 (filed 2022)")
)
irp
```

**Intended outcome** (per the comment block at IRP5Condense.rmd L679-684):
drop each superseded *original* (rows 1, 3, 5), keep the revisions and
relabel them to their true year → one corrected row per person-year.

## What the current code does

```{r a3-current}
cur <- copy(irp)
cur[, DropThisForRev := 0L]
cur[taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear,
  DropThisForRev := 1L,
  by = person]
cur[]
```

The pooled set `TrueTaxYear[TrueTaxYear != taxyear]` is evaluated on the
**whole table** (in `DT[i, j := v, by=g]`, `i` never sees the groups):

```{r a3-pooled-set}
sort(unique(irp[TrueTaxYear != taxyear, TrueTaxYear]))
```

Row by row:

* Rows 1, 3, 5 (the originals that *should* be dropped): the conjunct
  `TrueTaxYear != taxyear` is FALSE → **never flagged**. Target inverted.
* Row 2 (A's correction, stated 2012): `2012 %in% {2011,2012,2021}` is TRUE
  — but 2012 entered the set through **person B's** revision → dropped
  because of someone else's filing.
* Row 4 (B's correction, stated 2013): `2013` is *not* in the set → kept.
* Row 6 (C's correction, stated 2022): `2022` is not in the set (a
  revision's TrueTaxYear is always earlier than its stated year, so the
  set can never contain the last stated year) → kept.

In the real IRP5 data the pooled set is `{2003, ..., 2021}` (the mismatch
table saved at L666-669), while stated taxyears run 2008-2022. Hence every
revision stated 2008-2021 is dropped (~9.02% of rows ≈ all 18.5M revision
entries) and only 2022-filed revisions escape — the real-data analogue of
row 6.

## Consequence: duplicate person-years

After the drop and the `TrueTaxYear -> taxyear` rename, C's surviving
revision lands on 2021 — next to C's *kept* outdated original:

```{r a3-duplicates}
curkept <- cur[DropThisForRev == 0L]
setnames(curkept, c("TrueTaxYear", "taxyear"), c("taxyear", "Old.taxyear"))
curkept[order(person, taxyear), .(person, taxyear, Old.taxyear, role)]
table(curkept[, .(person, taxyear)])
```

* Person A lost the corrected 2011 record entirely (only the outdated
  original remains).
* Persons B and C now have **two rows** for 2012 and 2021 respectively —
  the stale original *and* the relabeled revision — inflating any
  count aggregated on `taxyear`.

## Cross-person contamination, isolated

Delete person B and rerun: A's correction (identical row!) now survives,
because 2012 left the pooled set. A row's fate depends on other people's
filings:

```{r a3-contamination}
cur2 <- copy(irp[person != "B"])
cur2[, DropThisForRev := 0L]
cur2[taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear,
  DropThisForRev := 1L, by = person]
cur2[person == "A"]
```

## The fix: per-person update join

Build a lookup of (person, superseded year) pairs from the revision rows,
then flag — within the join — only rows that are originals of that year:

```{r a3-fix}
fix <- copy(irp)
fix[, DropThisForRev := 0L]
rev <- unique(fix[TrueTaxYear != taxyear, .(person, taxyear = TrueTaxYear)])
rev
fix[rev, DropThisForRev := as.integer(TrueTaxYear == taxyear),
  on = c("person", "taxyear")]
fix[]
```

Rows 1, 3, 5 flagged — exactly the superseded originals. After the drop
and rename, one corrected row per person-year:

```{r a3-fix-result}
fixkept <- fix[DropThisForRev == 0L]
setnames(fixkept, c("TrueTaxYear", "taxyear"), c("taxyear", "Old.taxyear"))
fixkept[order(person, taxyear), .(person, taxyear, Old.taxyear, role)]
table(fixkept[, .(person, taxyear)])
```

In `IRP5Condense.rmd` the person key is
`c(geovars, "taxrefno", "UID")` and the lookup carries `taxyear =
TrueTaxYear`; the join is one keyed pass over the 175M-row table.

# A5. `stats::reshape` vs `dcast()` — same result, different cost

The problem with `reshape()` is the **composite id**: with several
`idvar` columns (9 in `idyr`) it builds the row key with
`interaction(data[, idvar], drop = TRUE)`, which indexes the full
Cartesian product of all id-column levels. With high-cardinality ids that
code space exceeds `.Machine$integer.max`, the codes overflow to NA, and
the fallback allocation is absurd — `reshape` **aborts** rather than runs
slowly. The example reproduces the real structure: 200,000 units carrying
4 geo strings + 5 id columns, each observed in a random subset of
2008-2022.

```{r a5-data}
set.seed(1)
n_id <- 2e5
#### 9 id columns, mimicking idyr (4 geo strings + 5 establishment/firm ids)
units <- data.table(
  busprov_geo     = sample(paste0("PR", 1:9),     n_id, replace = TRUE),
  busdistmuni_geo = sample(paste0("DC", 1:52),    n_id, replace = TRUE),
  buslocmuni_geo  = sample(paste0("LM", 1:205),   n_id, replace = TRUE),
  busmainplc_geo  = sample(paste0("MP", 1:14000), n_id, replace = TRUE),
  FirmIDTx  = seq_len(n_id),
  FirmID    = seq_len(n_id),
  EstabIDTx = seq_len(n_id),
  EstabID   = seq_len(n_id),
  EUIndID   = paste0(seq_len(n_id), "-", seq_len(n_id))
)
#### each unit observed in a random subset of 2008-2022 (no duplicates)
ny   <- sample(3:15, n_id, replace = TRUE)
long <- units[rep(seq_len(n_id), ny)]
long[, taxyear := unlist(lapply(ny, sample, x = 2008:2022))]
long[, Ob := 1L]
idcols <- setdiff(names(long), c("taxyear", "Ob"))
dim(long)
```

## `reshape()` on the 9-column composite id: it does not run at all

```{r a5-reshape-fails}
#| warning: true
res <- try(
  w1 <- reshape(as.data.frame(long), direction = "wide",
                idvar = idcols, timevar = "taxyear", v.names = "Ob"),
  silent = TRUE
)
cat("reshape():", attr(res, "condition")$message, "\n")
```

On this machine the failure is
`cannot allocate vector of size 298.0 Gb`, preceded by
`NAs produced by integer overflow` warnings from `interaction()` — the
level-product of the 9 id columns (9 x 52 x 205 x 14000 x 2e5 x ...) has
overflowed integer indexing. The production `idyr` at IRP5Condense.rmd
L975 sits in the same regime (`EUIndID` alone has tens of millions of
levels), so the current `reshape()` call is not merely slow — on
R >= 4.4 it is exposed to exactly this abort.

## `dcast()` on the same task

```{r a5-dcast-succeeds}
t_dcast <- system.time(
  w2 <- dcast(long,
    busprov_geo + busdistmuni_geo + buslocmuni_geo + busmainplc_geo +
    FirmIDTx + FirmID + EstabIDTx + EstabID + EUIndID ~ taxyear,
    value.var = "Ob", fill = 0L, fun.aggregate = max)
)
t_dcast
dim(w2)
w2[1:3, 9:14, with = FALSE]  #### unobserved years already 0L via fill = 0L
```

`dcast` never materializes the level product — it radix-orders the
observed id combinations — so cardinality is harmless, and `fill = 0L`
already delivers what the separate NA→0 fill loop (the loop carrying the
L985 paren regression) was for.

## Honest toy-scale comparison, and equivalence of results

Where `reshape` *can* run — a single low-cardinality id — it is
competitive at toy scale (`dcast` carries fixed setup overhead), so raw
speed on small examples is **not** the argument; feasibility at scale and
the built-in `fill` are:

```{r a5-single-id}
small <- long[, .(id = FirmID, taxyear, Ob)]
t_r <- system.time(
  w1s <- reshape(as.data.frame(small), direction = "wide",
                 idvar = "id", timevar = "taxyear", v.names = "Ob"))
t_d <- system.time(
  w2s <- dcast(small, id ~ taxyear, value.var = "Ob",
               fill = 0L, fun.aggregate = max))
rbind(reshape = t_r, dcast = t_d)
w1s[1:3, 1:6]                #### reshape: unobserved years are NA
w2s[1:3, 1:6, with = FALSE]  #### dcast:   already 0L
```

Rename dcast's year columns to the `Ob.YYYY` convention and confirm the
two routes agree once reshape's NAs are filled:

```{r a5-equivalence}
setnames(w2s, as.character(2008:2022), paste0("Ob.", 2008:2022))
w1sdt <- setDT(w1s)
cols <- grep("^Ob\\.", names(w1sdt), value = TRUE)
w1sdt[, (cols) := lapply(.SD, function(x) {x[is.na(x)] <- 0L; x}), .SDcols = cols]
setkey(w1sdt, id); setkey(w2s, id)
all.equal(w1sdt, w2s[, names(w1sdt), with = FALSE], check.attributes = FALSE)
```

At `idyr` scale (9 id columns, ~200M input rows) the gap widens further;
the data.table
[reshape vignette](https://cran.r-project.org/web/packages/data.table/vignettes/datatable-reshape.html)
notes `dcast`/`melt` are implemented in C with fast radix ordering and are
designed for large (10Gb+) in-memory data. `fun.aggregate = max` must be
stated explicitly: it defines what happens when a person-establishment-year
appears twice (`reshape` silently keeps the first row; `dcast` errors
without it — safer).

# The `fifelse()` update join, and how `i`, `by`, `j` scope

## What `fifelse(taxyear <= 2015, i.v, FillInLocMuni)` does

Toy version of the A9 loc-muni fill. `lk` holds one looked-up value per
person; the join writes only matched rows, and only their `taxyear <= 2015`
entries:

```{r fifelse-join}
main <- data.table(
  FirmUInd      = c(1L, 1L, 1L, 2L, 3L),
  taxyear       = c(2014, 2015, 2016, 2014, 2014),
  FillInLocMuni = c("", "", "", "", "Cape Town")
)
lk <- data.table(FirmUInd = c(1L, 2L), v = c("Stellenbosch", "eThekwini"))
main[lk, on = "FirmUInd",
  FillInLocMuni := fifelse(taxyear <= 2015, i.v, FillInLocMuni)]
main[]
```

* `lk` occupies the `i` slot → `j` runs **only on matched rows**
  (FirmUInd 1 and 2). FirmUInd 3 is never touched.
* Inside `j`, bare names (`taxyear`, `FillInLocMuni`) are the matched
  rows' own columns; `i.v` is the joined-in lookup value.
* `fifelse` (fast, type-strict vectorized if-else) writes `i.v` where
  `taxyear <= 2015` and re-assigns the current value elsewhere (a no-op) —
  needed because the `i` slot is already taken by the join, so the year
  restriction cannot go there without materializing a subset copy.

## `by=` scopes `j` over the `i`-selected rows — never `i` itself

Evaluation order in `dt[i, j, by]`: `i` selects rows (evaluated **once, on
the full table's columns**) → `by` forms groups *within that subset* → `j`
runs per group on that subset.

```{r scoping-demo}
dd <- data.table(g = c("a", "a", "b"), x = 1:3)
#### j IS grouped: max(x) per group
dd[, gmax := max(x), by = g]
#### i is NOT grouped: max(x) inside i is the GLOBAL max (3), by= cannot change it
dd[x == max(x), flag := 1L, by = g]
dd[]
```

`gmax` differs by group (2, 3) — `j` saw the groups. `flag` marks only
`x == 3` — the `i` expression used the global maximum even though `by = g`
was supplied; with `j` a constant, that `by=` did nothing at all. This is
precisely the A3 mechanism: the pooled revision-year set lived in `i`,
out of `by=`'s reach. Conversely, in the A9 lookup

```r
lk <- irp5gi[CanFill == 1L & taxyear <= 2015 & FirmUInd %in% need,
  .(v = Bus_adr_Geo_Munic[1]), by = FirmUInd]
```

`by = FirmUInd` groups **only the rows that passed `i`** (the small
CanFill subset), which is why the grouped `[1]` is cheap there.
