---
title: "Compute HHI and FA"
date: "`{r} format(Sys.time(), "%Y %m %d %R")`"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 23/IRP5HHI.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
```
```{r path, eval = T}
#### path <- "Z:/Workbenches/epguest/seiro_ito/"
#### pathdataCITIRP <- "Z:/Master Data/CIT-IRP5 Panel/"
#### pathdataIRP <- "Z:/Master Data/IRP5/Job level/v5/beta/"
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathresults <- paste0(path, "results/")
pathdata <- paste0(path, "data/")
pathsaveddata <- paste0(path, "saveddata/")
pathdataCITIRP <- "U:/CIT-IRP5 Panel/"
pathdataIRP <- "U:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```


* FA does not require location info
   * We use FA data that have jobs and employment as the base
* HHI requires location info, so we use `irp5M` that have information of Main Place
   * Whenever location information is updated, we need to compute HHI
   * To update location information, need to update irp5.qs &rarr; irp5M.qs &rarr; LShareHHI.qs



# Fraction affected

1. Use IRP5 to count total number of workers, number of affected, FA = affected/total for all years
1. Need to use unfiltered data like `irp5.qs` (except for NatureOfPer == A and empty taxrefno) 
   * Entries of irp5 look naturally balanced across years
   * Many NAs in location info in early years &rArr; loss of sample in early years

```{r read irp5 file, eval = F}
if (exists(irp5gi))    rm(irp5gi)
if (exists(irp5gir))   rm(irp5gir)
if (exists(irp5Clean)) rm(irp5Clean)
#### irp5.qs is created in IRP5Condense.rmd
library(qs)
library(data.table)
irp5 <- qread(paste0(pathsaveddata, "irp5.qs"), nthreads = 13)
```

```{r fraction affected for all years, eval = F}
#### code "compute FA"
library(qs)
library(data.table)
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathsaveddata <- paste0(path, "saveddata/")
pathdataCITIRP <- "U:/CIT-IRP5 Panel/"
pathdataIRP <- "U:/IRP5/Job level/v5/beta/"
#### Takes 20-30 mins: Start  ####
#### irp5M <- qread(paste0(pathdata, "irp5M.qs"), nthreads = 13)
#### ipyrc0 <- irp5M[
#### !duplicated(irp5M[, .(busmainplc_geo, taxrefno, UID, 
#### DateBirth, DateStart, DateEnd,
#### kerr_income, ptrs_income, a3601_income)]), 
#### ]
#### Takes 20-30 mins: End  ####
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
aggsummary <- NULL
for (yr in 8:22) { 
  ipyrc <- irp5[taxyear == 2000+yr, ]#[1:100000, ]
  #### Compute job duration and monthly income
  ####   (DateEnd-DateStart2012)/30.5
  ####   a3601_income/DurationMonth
  ipyrc[, DateStart20yr := DateStart]
  #### If DateStart is before start of taxyear, March 1 
  ####  ==> DateStart20yr = March 1 of taxyear-1
  ipyrc[DateStart < as.IDate(paste0(2000+yr-1, "/03/01")), 
          DateStart20yr := as.IDate(paste0(2000+yr-1,"/03/01"))]
  #### IncomeMonth = monthly income, duration is measured as the same taxyear
  ipyrc[, DurationMonth := round(as.numeric((DateEnd-DateStart)/30.5), 3)]
  ipyrc[, DurMonthMarch := round(as.numeric((DateEnd-DateStart20yr)/30.5), 3)]
  #### DurationMonth can be zero if work is only for 1 day
  ipyrc[DateStart == DateEnd, c("DurationMonth", "DurMonthMarch") := round(1/30.5, 3)]
  lapply(ipyrc[, .(DurationMonth, DurMonthMarch)], 
    quantile, c(seq(0, 1, .25)[-5], .985, .995, 1), na.rm = T)
  #### DateStart too old for some rows (typo for previous year or actual starting dates)  
  #### DurationMonth == DurMonthMarch for these entries
  #### Threshold is set to 20 because typo for prev Ma 1 will give DurationMonth==24
  print(ipyrc[DurationMonth > 20, 
      .(Txrf, DateStart, DateStart20yr, DateEnd, DurationMonth, DurMonthMarch)])
  ipyrc[DurationMonth > 20, DurationMonth := DurMonthMarch]
  #### typos (because we corrected for revision reporting for previous years)
  print(ipyrc[DurMonthMarch > 100, 
    .(Txrf, DateStart, DateStart20yr, DateEnd, DurationMonth, DurMonthMarch)])
  ipyrc[, IncomeMonth   := round(a3601_income/DurationMonth, 0)]
  setkey(ipyrc, busmainplc_geo, taxrefno, UID, DateStart)
  #### DateStart2, DateEnd2: Dates for 2nd job in the same firm
  ipyrc[, DateStart2 := shift(DateStart, n = 1L, type = "lead"), by = .(Txrf, taxyear, UID)]
  ipyrc[,   DateEnd2 := shift(DateEnd, n = 1L, type = "lead"), by = .(Txrf, taxyear, UID)]
  ipyrc[, c("TDurationMonth", "DJobDurationMonth") := 0]
  #### If job2 after job1 ends: 
  #### Double work duration: DateEnd  - DateStart2
  ####  Total work duration: DateEnd2 - DateStart20yr
  ####  <----job1---->
  ####           <----job2---->
  ####  DateStart20yr < DateStart2
  ####  DateEnd < DateEnd2
  ####  DateEnd > DateStart2
  ipyrc[DateStart20yr < DateStart2 & DateEnd < DateEnd2 & DateEnd > DateStart2, 
    DJobDurationMonth := round(as.numeric((DateEnd - DateStart2)/30.5), 3)]
    #### Negative TDurationMonth for 35 entries for 2012 
    #### because their reported DateEnd2 is before 2011/03/01
  ipyrc[DateStart20yr < DateStart2 & DateEnd < DateEnd2 & DateEnd > DateStart2, 
    TDurationMonth := round(as.numeric((DateEnd2 - DateStart20yr)/30.5), 3)]
  #### If job2 ends before job1 ends: 
  #### Double work duration: DateEnd2 - DateStart2
  ####  Total work duration: DateEnd  - DateStart20yr
  ####  <----------job1---------->
  ####           <----job2---->
  ####  DateStart20yr < DateStart2
  ####  DateEnd > DateEnd2
  ####     (DateEnd2 - DateStart2   )/30.5 (double work duration)
  ####     (DateEnd  - DateStart20yr)/30.5 (job1+job2 duration)
  ipyrc[DateStart20yr < DateStart2 & DateEnd > DateEnd2, 
         DJobDurationMonth := round(as.numeric((DateEnd2 - DateStart2)/30.5), 3)]
  ipyrc[DateStart20yr < DateStart2 & DateEnd > DateEnd2, 
         TDurationMonth    := round(as.numeric((DateEnd - DateStart20yr)/30.5), 3)]
  #### TDurationMonth = Total duration of jobs
  ipyrc[, TDurationMonth   := TDurationMonth[TDurationMonth == max(TDurationMonth, na.rm = T)][1],
    by = .(Txrf, taxyear, UID)]
  ipyrc[, DoubleJobRatio   := round((DJobDurationMonth/TDurationMonth)*100, 2)]
  ipyrc[, DoubleJobRatio   := DoubleJobRatio[!is.na(DoubleJobRatio)][1],by = .(Txrf, taxyear, UID)]
  ipyrc[, NumJobsPerWorker := as.integer(.N), by = .(Txrf, taxyear, UID)]
  #### TIncomeMonth = sum of monthly income across jobs
  ipyrc[, TIncomeMonth     := sum(IncomeMonth, na.rm = T), by = .(Txrf, taxyear, UID)]
  #### check data
  summary(ipyrc[busmainplc_geo != "" & Txrf != "" & NumJobsPerWorker > 1, 
    .(busmainplc_geo, Txrf, UID, uid, NumJobsPerWorker, 
      DateStart, DateEnd,
      DateStart2, DateEnd2, DurationMonth,DurMonthMarch, 
      DJobDurationMonth, TDurationMonth, DoubleJobRatio,
      a3601_income, 
      kerr_income, IncomeMonth, TIncomeMonth)])
  print(ipyrc[NumJobsPerWorker > 10, 
    .(busmainplc_geo, Txrf, UID, uid, NumJobsPerWorker, 
      DateStart, DateEnd,
      DateStart2, DateEnd2, DurationMonth, 
      DJobDurationMonth, TDurationMonth, DoubleJobRatio,
      a3601_income, IncomeMonth, TIncomeMonth)], topn = 30)
  ipyrc[busmainplc_geo != "" & Txrf != "", 
    .(busmainplc_geo, taxrefno, DateStart, DateEnd, 
      DurationMonth, a3601_income, IncomeMonth)]
  #### Find sub MW workers and define fraction affected
  #### Sums (Jobs, NumSubMW, etc.) counts total in two levels
  ####   a. Firm (Txrf):                good: more obs, bad: not granular
  ####   b. Establishment (Main Place): good: granular, bad: fewer obs
  #### We can use IncomeMonth or TIncomeMonth
  ####   PropToMW: Proportion to MW line = IncomeMonth/2274 ==> SubMW is wage per job
  ####      SubMW: PropToMW < 1 & IncomeMonth > 0
  ####  NumSubMW : number of entries with SubMW == 1L
  ####  NumSubMWe: number of unique UID entries with SubMW == 1L
  ####       Jobs: number of entries with IncomeMonth > 0
  ####  Employees: number of unique UID with IncomeMonth > 0
  ####        FA : NumSubMW/Jobs (FA in jobs)
  ####        FAe: NumSubMWe/Employees (FA in employment)
  #### Note: uid is NA in some entries, UID is not 
  ipyrc[, PropToMW  := round(IncomeMonth/2274, 6)]
  ipyrc[, SubMW     := as.integer(PropToMW < 1 & IncomeMonth > 0)]
  #### a. Firm level
  ipyrc[, NumSubMW  := as.integer(sum(SubMW==1L)), 
    by = .(Txrf, taxyear)]
  ipyrc[, Jobs      := as.integer(sum(IncomeMonth > 0)), 
    by = .(Txrf, taxyear)]
  ipyrc[, NumSubMWe := as.integer(uniqueN(UID[SubMW==1L])), 
    by = .(Txrf, taxyear)]
  ipyrc[, Employees := as.integer(uniqueN(UID[IncomeMonth > 0])), 
    by = .(Txrf, taxyear)]
  #### b. Establishment (Main Place) level
  ipyrc[, NumSubMWMP  := as.integer(sum(SubMW==1L)), 
    by = .(Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, taxyear)]
  ipyrc[, JobsMP      := as.integer(sum(IncomeMonth > 0)), 
    by = .(Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, taxyear)]
  ipyrc[, NumSubMWeMP := as.integer(uniqueN(UID[SubMW==1L])), 
    by = .(Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, taxyear)]
  ipyrc[, EmployeesMP := as.integer(uniqueN(UID[IncomeMonth > 0])), 
    by = .(Txrf, busprov_geo, busdistmuni_geo, 
      buslocmuni_geo, busmainplc_geo, taxyear)]
  ipyrc[, FA     := round(NumSubMW/Jobs, 6)]
  ipyrc[, FAe    := round(NumSubMWe/Employees, 6)]
  ipyrc[, FAMP   := round(NumSubMWMP/JobsMP, 6)]
  ipyrc[, FAeMP  := round(NumSubMWeMP/EmployeesMP, 6)]
  ####ipyrc[grepl("^BCJXBCC", Txrf), 
  ####  .(busmainplc_geo, taxyear, 
  ####    Txrf, uid, UID, DateStart, DurationMonth, 
  ####    IncomeMonth, SubMW, NumSubMW, Jobs, FA,
  ####    NumSubMWe, Employees, FAe)]
  #### Fin year 2013 = 2012 Mar 01 - 2013 Feb 29
  ####  <==> 0L for DateEnd < 2012 Mar 01 or DateStart >= 2013 Mar 01 
  #### FinYr20yr := 1L ==> DateStart & DateEnd are correct for current taxyear
  ipyrc[, FinYr20yr := 1L]
  ipyrc[
    DateEnd    < as.IDate(paste0(2000+(yr-1), "/03/01")) |
    DateStart >= as.IDate(paste0(2000+yr,     "/03/01")), 
    FinYr20yr := 0L]
  #### Select variables to keep
  FAdata <- ipyrc[, .(
      busprov_geo, 
      busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
      taxrefno, Txrf, UID, uid,
      DurationMonth, DurMonthMarch,
      a3601_income, IncomeMonth, NumJobsPerWorker,
      PropToMW, SubMW, 
      Jobs, NumSubMW, Employees, NumSubMWe,  
      JobsMP, NumSubMWMP, EmployeesMP, NumSubMWeMP,  
      FA, FAe, FAMP, FAeMP
    )]
  #### summary(FAdata)
  #### summary(FAdata[is.na(FA), .(FA, Jobs, a3601_income, IncomeMonth, DurationMonth)])
  FAdata[, Size := "micro"]
  FAdata[Employees >= 10 & Employees <= 49, Size := "small"]
  FAdata[Employees >= 50 & Employees <= 249, Size := "medium"]
  FAdata[Employees >= 250, Size := "large"]
  FAdata[is.na(Employees) | Employees == 0, Size := NA]
  FAdata[, Size := factor(Size, levels = c("micro", "small", "medium", "large"))]
  #### Number of affected workers in aggregates
  ag1 <- ipyrc[, .(
      TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
    , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
    , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
    , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
    )
    ][, .(
      TotalEmployees, TotalSubMWWorkers, 
      AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
      TotalFirmsWithEmployees, TotalAffectedFirms,
      AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
    )]
  ag1[, taxyear := 2000+yr]
  aggsummary <- rbindlist(list(aggsummary, ag1), use.names = T, fill = T)
  ag2 <- ipyrc[, .(
      TotalEmployees = uniqueN(UID[IncomeMonth > 0]) 
    , TotalSubMWWorkers = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
    , TotalFirmsWithEmployees = uniqueN(taxrefno[IncomeMonth > 0]) 
    , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
    ), 
    by = .(taxyear, busprov_geo)
    ][, .(
      busprov_geo, TotalEmployees, TotalSubMWWorkers, 
      AggFA = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
      TotalFirmsWithEmployees, TotalAffectedFirms,
      AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmployees)*100, 2)
    )]
  ag2[, taxyear := 2000+yr]
  aggsummary <- rbindlist(list(aggsummary, ag2), use.names = T, fill = T)
  FAdata[, taxyear := 2000+yr]
  if (yr < 10) yr <- paste0("0", yr)
  qsave(FAdata, paste0(pathdata, "FA", yr, ".qs"), nthreads = 13)
} 
qsave(aggsummary, paste0(pathdata, "AggregateSummary.qs"), nthreads = 13)
aggsummary <- qread(paste0(pathdata, "AggregateSummary.qs"), nthreads = 13)
```

1. For each taxyear, `FAdata` has 4 indicators of "Fraction Affected":
   * `FA`: At firm level, computed with jobs
   * `FAe`: At firm level, computed with individuals
   * `FAMP`: At establishment (Main Place) level, computed with jobs
   * `FAeMP`: At establishment (Main Place) level, computed with individuals

    Accordingly, we have created for `FA`
   * Jobs: Number of jobs
   * NumSubMW: Number of sub MW jobs
   * Employees: Number of employees
   * NumSubMWe: Number of sub MW employees

    Also for `FAe`, `FAMP`, `FAeMP`, we have created these four variables 
1. We append `FAdata` of each taxyear &rarr; `FAD` (FA panel at individual level)
1. We select rows of `FAD` as
   *  `Fadata`: FAFirstRow.qs, establishment level FA panel (1st rows of FAD by location-taxrefno-taxyear)
   * `fadatai`: FAAndJobs.qs, duration within 1 year of Fadata + industry
   *     `faa`: FAOfAgri.qs, fadatai[Crop==1L|Animal==1L|Forest==1L, ]
   *   `estfa`: fadataUniqueEntries.qs, unique entries by location-taxrefno-taxyear in fadatai (almost same as fadatai except uniqueness is explicitly imposed)


```{r create FA panel, eval = F}
FAD <- NULL
for (yr in 8:22) {
  if (yr < 10)  yr <- paste0("0", yr)
  FAdata <- qread(paste0(pathdata, "FA", yr, ".qs"))
  FAdata[, taxyear := 2000+as.numeric(yr)]
  FAD <- rbindlist(list(FAD, FAdata))
}
qsave(FAD, paste0(pathdata, "FAD.qs"), nthreads = 13)

table(FAD[,taxyear])
#### 2009     2010     2011     2012     2013     2014     2015     2016     2017     2018     2019     2020     2021     2022 
####  200625  5289890   535385  4379788  8910574 10078717 10249875 10412189 11013133 11183320 11163246 10901136  9882070  9683715
```

```{r industry code for ag, eval = F}
#### DO NOT RUN THIS TWICE. It has been run and let us keep it that way.
#### Merge industry code, taking fadata as base
ind <- qread(paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = 13)
setkey(irp5, taxrefno); setkey(ind, taxrefno, taxyear)
ind[, n := 1:.N, by = taxrefno]
inda <- ind[n == 1 & grepl("^Anim|^Plant pro|crops|Logging|forest", imp_mic_sic7_3d), ]
inda[, n := NULL]
inda[, c("Crop", "Animal", "Forest") := 0L]
inda[grepl("Plant|cr", imp_mic_sic7_3d), Crop := 1L]
inda[grepl("Anim", imp_mic_sic7_3d),   Animal := 1L]
inda[grepl("Lo|for", imp_mic_sic7_3d), Forest := 1L]
inda[, grepout("mp_mic", colnames(inda)) := NULL]
setkey(inda, taxrefno)
setnames(inda, "taxyear", "FirstYearOnSIC7")
qsave(inda, paste0(pathdata, "TaxrefnoOfAgri.qs"), nthreads = 13)
```

```{r Fadata is first lines of FAD in each year, eval = F}
FAD <- qread(paste0(pathsaveddata, "FAD.qs"), nthreads = 13)
FAD[, num := 1:.N, by = .(busprov_geo, busdistmuni_geo, 
  buslocmuni_geo, busmainplc_geo, taxrefno, taxyear)]
Fadata <- FAD[num == 1L, ]
table(Fadata[, taxyear])
####  2008   2009   2010   2011   2012   2013   2014   2015   2016   2017   2018   2019   2020   2021   2022 
#### 23791  15641 168258  15521 243798 283160 299154 311918 323084 331015 341569 352993 364654 329671 309533 
qsave(Fadata, paste0(pathdata, "FAFirstRow.qs"), nthreads = 13)
Fadata <- qread(paste0(pathdata, "FAFirstRow.qs"), nthreads = 13)
#### Drop date errors (Negative and over duration)
fadata <- Fadata[0 <= DurMonthMarch & DurMonthMarch <= 12, ]
nrow(Fadata); nrow(fadata) # reduced 1.2M rows...
qsave(fadata, 
  paste0(pathdata, "FAFirstRowWithoutNegativeDurations.qs"), 
  nthreads = 13)
#### Merge industry code, taking fadata as base
inda <- qread(paste0(pathdata, "TaxrefnoOfAgri.qs"), nthreads = 13)
fadatai <- inda[fadata, on = "taxrefno"]
qsave(fadatai, paste0(pathdata, "FAAndJobs.qs"), nthreads = 13)
fadatain <- fadatai[busmainplc_geo != "", ]
faa <- fadatain[Crop == 1L | Animal == 1L | Forest == 1L, ]
setkey(faa, taxrefno)
faa[, SubDiv := "Crop"]
faa[Animal == 1L, SubDiv := "Animal"]
faa[Forest == 1L, SubDiv := "Forest"]
faa[, SubDiv := factor(SubDiv, levels = c("Crop", "Animal", "Forest"))]
faa[, .(TotalJobs = sum(Jobs, na.rm = T)), by = .(taxyear, SubDiv)][order(SubDiv, taxyear)]
faa[, FApositive := 0L]
faa[taxyear < 2014, FApositive:= as.integer(FA > 0)]
faa[, FApositive := FApositive[FApositive == 1L][1], by = taxrefno]
faa[is.na(FApositive), FApositive := 0L]
faa[, Sample := "Unexposed"]
faa[FApositive == 1L, Sample := "Exposed"]
faa[, Sample := factor(Sample)]
#### faa is almost unique at establishment level
table(duplicated(faa[, .(taxrefno, taxyear, busmainplc_geo)]))
#### FALSE  TRUE 
#### 59627     5 
qsave(faa, paste0(pathdata, "FAOfAgri.qs"), nthreads = 13)
#### estfa: establishment level FAs with location, taxrefno, taxyear
estfa <- unique(fadatai[, 
  grepout("^bus|^tax[ry]|^Tx|Num|Jo|Emp|FA", colnames(fadatai)), with = F])
qsave(estfa, paste0(pathdata, "fadataUniqueEntries.qs"), nthreads = 13)
```

```{r copy FAD and derived files to saveddata, eval = F}
FilesToCopy <- paste0(c(
  "FAD",        # row-bound FAdata over taxyears
  "FAFirstRow", # Fadata : 1st rows of FAD by location-taxrefno-taxyear
  "FAAndJobs",  # fadatai: Duration within 1 year of Fadata + industry
  "FAOfAgri",   # faa    : fadatain[Crop==1L|Animal==1L|Forest==1L, ]
  "fadataUniqueEntries" ,
                # estfa  : unique entries by location-taxrefno-taxyear
  "AggregateSummary.qs"
                # aggsummary : summary stats, busprov_geo, TotalEmployees, TotalSubMWWorkers, AggFA, TotalFirmsWithEmployees, TotalAffectedFirms,      AggFirmFA
  ), ".qs")
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
  )
```



```{r descriptive stats FA and jobs, eval = F}
library(tinytable)
aggsummary <- 
  qread(paste0(pathsaveddata, "AggregateSummary.qs"), 
  nthreads = 13)
#### total jobs, total sub MW jobs, aggregate FA
aggsum <- 
  aggsummary[, .(
      TotalJobs               = sum(TotalEmployees)
    , TotalSubMWJobs          = sum(TotalSubMWWorkers)
    , MeanFAJobs              = mean(TotalSubMWWorkers/TotalEmployees)
    , TotalFirmsWithEmployees = sum(TotalFirmsWithEmployees)
    , TotalAffectedFirms      = sum(TotalAffectedFirms)
  ), by = taxyear][, .(
        taxyear, TotalJobs, TotalSubMWJobs, 
        TotalFirmsWithEmployees, TotalAffectedFirms,
        MeanFAJobs
      , AggregateFA = round((TotalSubMWJobs/TotalJobs)*100, 2)
    )
  ]
tb <- tt(aggsum, width = 0.8, caption = "Jobs and sub-MW jobs, all industries")
options(tinytable_format_digits = 2)
tb.agg <- format_tt(tb
    j = 1:4,
    num_mark_big = " "
  )
tb.agg
#### Agriculture
faa <- qread(paste0(pathsaveddata, "FAOfAgri.qs""), nthreads = 13)
agsum <- faa[, .(
      TotalJobs          = sum(Jobs[IncomeMonth > 0]) 
    , TotalSubMWJobs     = uniqueN(Jobs[IncomeMonth > 0 & SubMW==1L]) 
    , TotalEmployees     = uniqueN(UID[IncomeMonth > 0]) 
    , TotalSubMWWorkers  = uniqueN(UID[IncomeMonth > 0 & SubMW==1L]) 
    , TotalFirmsWithEmp  = uniqueN(taxrefno[IncomeMonth > 0]) 
    , TotalAffectedFirms = uniqueN(taxrefno[IncomeMonth > 0 & SubMW==1L]) 
    ), 
    by = .(taxyear)
    ][, .(
      taxyear, 
      TotalJobs, TotalSubMWJobs, TotalEmployees, TotalSubMWWorkers, 
      AggFA = round((TotalSubMWJobs/TotalJobs)*100, 2),
      AggFAe = round((TotalSubMWWorkers/TotalEmployees)*100, 2),
      TotalFirmsWithEmp, TotalAffectedFirms,
      AggFirmFA = round((TotalAffectedFirms/TotalFirmsWithEmp)*100, 2)
    )]
tb.ag <- tt(agsum, width = 0.8, caption = "Jobs and sub-MW jobs, agriculture")
options(tinytable_format_digits = 2)
tb.ag <- format_tt(tb.ag,
    j = c(2:5, 8:9),
    num_mark_big = " "
  )
tb.ag
#### estfa <- 
####  qread(paste0(pathdata, "fadataUniqueEntries.qs"), nthreads = 13)
```


# Plot employment, FAs


```{r plot FAs by size, eval = F}
fadatai <- qread(paste0(pathsaveddata, "FAAndJobs.qs"), 
  nthreads = 13)
#### possible duplicates with different local municipality names
ii0 <- fadatai[, buslocmuni_geo != "" & busdistmuni_geo != ""
   & busmainplc_geo != ""]
ii1 <- duplicated(fadatai[, .(
   busmainplc_geo, Txrf
 )])
ii2 <- duplicated(fadatai[, .(
   busmainplc_geo, Txrf
 )], fromLast = T)
table(data.table(nonNA=ii0, duplicates=ii1))
fadatai[ii0 & (ii1 | ii2), ]
#### ii0 & !ii1 = nonempty & non-duplicated
table(duplicated(fadatai[ii0 & !ii1, ]))
#### plot sensitivity to XThreshold (double job months cutoff)
#### FA.orig : FA of jobs
#### FAe.orig: FA of employment/individuals
library(ggplot2)
library(cowplot)
#### Below gives an error: idvar need to specify unique rows
#### fadataiL <- reshape(fadatai[ii0 & !ii1, ], 
####   direction = "long", 
####   idvar = c("busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo", 
####     "Txrf", "payereferenceno"), 
####   varying = grepout("^FA", colnames(fadatai))
#### )
gsz <- ggplot(data = fadatai[ii0 & !ii1 & !is.na(Size), ], 
  aes(x = FA.orig)) +
  geom_density(alpha = .1, fill = "blue") + 
  xlim(0, 1) +
  scale_color_manual(values = c("red", "blue"))
gsz <- gsz + facet_wrap(Size ~ .)
gsz
ggsave(
   paste0(pathresults, "FAByFirmSize.jpg")
  , gsz, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r plot FAs by jobs employment, eval = F}
#### possible duplicates with different local municipality names
ii0 <- fadatai[, buslocmuni_geo != "" & busdistmuni_geo != ""
   & busmainplc_geo != ""]
ii1 <- duplicated(fadatai[, .(
   busmainplc_geo, Txrf
 )])
ii2 <- duplicated(fadatai[, .(
   busmainplc_geo, Txrf
 )], fromLast = T)
table(data.table(nonNA=ii0, duplicates=ii1))
table(duplicated(fadatai[ii0 & !ii1, ]))
#### plot sensitivity to XThreshold (double job months cutoff)
library(ggplot2)
library(cowplot)
g1 <- ggplot(data = fadatai[ii0 & !ii1, ], 
  aes(x = FA.orig)) +
  geom_density(alpha = .1, fill = "blue") + 
  ggtitle("Fraction of affected jobs")+
  scale_x_continuous(name = "fraction per firm", limits=c(0, 1)) +
  scale_color_manual(values = "blue")
g2 <- ggplot(data = fadatai[ii0 & !ii1, ], 
  aes(x = FAe.orig)) +
  geom_density(alpha = .1, fill = "blue") + 
  ggtitle("Fraction of affected workers")+
  scale_x_continuous(name = "fraction per firm", limits=c(0, 1)) +
  scale_color_manual(values = "red")
gs <- plot_grid(
  g1, g2,
  labels = c('a.', 'b.'),
  align="h"
)
gs
ggsave(
  paste0(pathresults, "FAInJobsAndEmployment.jpg")
  , gs, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r plot number of agri jobs taxyear 2012-2020, eval = F}
library(qs)
library(data.table)
library(ggplot2)
#### irp5Ma is created in 
#### Location information granularity checks (next section)
#### irp5Ma <- qread(paste0(pathdata, "irp5Ma.qs"), nthreads = 13)
faa <- qread(paste0(pathsaveddata, "FAOfAgri.qs""), nthreads = 13)
#### Below is resolved. Negative duration reflects 
#### "revision reporting to SARS" by firms.
#### To Michael, 2026 Jan 21, negative duration entries
#### FAD[, .(Num = .N, Neg = sum(DurMonthMarch<0)), by = (yr)][, 
####  .(yr, Num, NegRate = round(Neg/Num, 4))]
geovars <- grepout("^bus.*geo$", colnames(faa))
faa[, Num := as.integer(.N), by = c(geovars, "taxrefno", "taxyear")]
irp5ForPlot <- unique(faa[taxyear >= 2010 & taxyear <= 2020, 
  c(geovars, "taxyear", "taxrefno", "Num", "FA.orig"), with = F])
irp5ForPlot[, taxyear := as.character(taxyear)]
g1 <- ggplot(data = irp5ForPlot[busmainplc_geo != "", ], 
    aes(x = Num, group = taxyear, colour = taxyear, fill = taxyear)) +
  geom_histogram(bins = 50) + 
  scale_color_viridis_d("Tax year") +
  scale_fill_viridis_d() +
  scale_x_continuous(trans = "log10") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
  facet_wrap(taxyear ~ ., ncol = 2)
ggsave(
  paste0(pathresults, "NumberOfAgEmployees2010-2020.jpg")
  , g1, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r agri Jobs by firm age, eval = F}
#### irp5M, agriculture
#### Jobs by firm age (existed prior to 2013 vs. new entrants after 2013)
irp5Ma <- qread(paste0(pathdata, "irp5Ma.qs"), nthreads = 13)
irp5Ma[, .(num = 1:.N, Num), by = .(taxyear, taxrefno)][num == 1, 
  .(TotalJobs = sum(Num, na.rm = T)), by = taxyear]
irp5Ma[, .(num = 1:.N, Num = .N), by = .(taxyear, taxrefno)][num == 1, 
  .(TotalJobs = sum(Num, na.rm = T)), by = taxyear]
irp5Ma[, ExistedBefore2013 := as.integer(NA)]
irp5Ma[taxyear < 2013, ExistedBefore2013 := 1L]
irp5Ma[, ExistedBefore2013 := ExistedBefore2013[!is.na(ExistedBefore2013)][1], by = taxrefno]
irp5Ma[is.na(ExistedBefore2013), ExistedBefore2013 := 0L]
JobsFirms <- irp5Ma[, .(num = 1:.N, Num = .N), 
  by = .(ExistedBefore2013, taxyear, taxrefno)][
  num == 1, .(TotalJobs = sum(Num, na.rm = T)), by = .(ExistedBefore2013, taxyear)]
JobsFirms[, Sample := "Entrants"]
JobsFirms[ExistedBefore2013 == 1L, Sample := "Incumbents"]
g1 <- ggplot(data = JobsFirms, aes(x = taxyear, y = TotalJobs)) +
  geom_col() +
  scale_x_continuous(breaks = 2009:2022) +
  facet_grid(~ Sample) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
ggsave(
  paste0(pathresults, "NumberOfAgEmployees.jpg")
  , g1, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r plot FA in agri 2012-2020, eval = F}
g2 <- ggplot(data = irp5ForPlot[busmainplc_geo != "", ], 
    aes(x = FA, group = taxyear, colour = taxyear, fill = taxyear)) +
  geom_density(alpha = .3) + 
  scale_color_viridis_d("Tax year") +
  scale_fill_viridis_d() +
  scale_x_continuous(limits = c(1, 100)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
  facet_wrap(taxyear ~ ., ncol = 2)
ggsave(
  paste0(pathresults, "FAOfAg2010-2022.jpg")
  , g2, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```
```{r plot FA of agri, eval = F}
#### FA of agri
g2 <- ggplot(data = faa, aes(x = as.factor(taxyear), y = FA)) +
#  geom_point(size = .1, position = position_jitterdodge()) +
#  geom_point(data = faama, size = 1, fill = "blue")
  geom_boxplot(fill = "white", colour = "blue", outlier.alpha = 0) + 
#  geom_jitter(width = .2) +
  scale_x_discrete(name = "Tax year") +
  facet_grid(~ Sample) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
ggsave(
  paste0(pathresults, "FAOfAgri.jpg")
  , g2, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )

```


# Location information granularity checks


We need to know how granular the location information is. Based on granularity, we decide at what level one can use the sampled firm.

```{r define irp5M by location granularity, eval = F}
rm(FAD); rm(Fadata); rm(fadata); rm(inda); 
rm(fadata2); rm(fadata3); rm(faa)
irp5 <- qread(paste0(pathsaveddata, "irp5.qs"), nthreads = 13)
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
geovars <- grepout("^bus.*geo$", colnames(irp5))
#### Location granularity indicator
irp5[, LocGranular := "none"]
for (yr in 8:22) {
  irp5[Txrf %in% Txrf
    #### Granular down to Main Place
    [
      busprov_geo     != "" & 
      busdistmuni_geo != "" & 
      buslocmuni_geo  != "" & 
      busmainplc_geo  != "" & 
      taxyear == 2000 + yr
    ] &
    taxyear == 2000 + yr, 
    LocGranular := GeoLevel[4]]
  irp5[Txrf %in% Txrf
    #### Granular down to Loc muni
    [
      busprov_geo     != "" & 
      busdistmuni_geo != "" & 
      buslocmuni_geo  != "" & 
      busmainplc_geo  == "" &
      taxyear == 2000 + yr
    ] &
    taxyear == 2000 + yr, 
    LocGranular := GeoLevel[3]]
  irp5[Txrf %in% Txrf
    #### Granular down to Dist muni
    [
      busprov_geo     != "" & 
      busdistmuni_geo != "" & 
      buslocmuni_geo  == "" & 
      busmainplc_geo  == "" &
      taxyear == 2000 + yr
    ] &
    taxyear == 2000 + yr, 
    LocGranular := GeoLevel[2]]
  irp5[Txrf %in% Txrf
    #### Granular down to Province
    [
      busprov_geo     != "" & 
      busdistmuni_geo == "" & 
      buslocmuni_geo  == "" & 
      busmainplc_geo  == "" &
      taxyear == 2000 + yr
    ] &
    taxyear == 2000 + yr, 
    LocGranular := GeoLevel[1]]
}
irp5[, LocGranular := factor(LocGranular, levels = c(GeoLevel, "none"))]
#### LocGranularFirm: Most granular location info (order sensitive)
#### LocGranularJob: Most granular location info (order insensitive)
setnames(irp5, "LocGranular", "LocGranularFirm")
irp5[, LocGranularJob := "none"]
irp5[busprov_geo != "", LocGranularJob := GeoLevel[1]]
irp5[busdistmuni_geo != "", LocGranularJob := GeoLevel[2]]
irp5[buslocmuni_geo != "", LocGranularJob := GeoLevel[3]]
irp5[busmainplc_geo != "", LocGranularJob := GeoLevel[4]]

#### Location granularity due to establishment clustering
#### When location info is missing in some rows, one can allocate workers
#### only up to location level where all establishments are clustered
for (mm in 1:length(GeoLevel)) {
  irp5[, (GeoLevel[mm]) := get(geovars[mm])]
  irp5[get(geovars[mm]) == "", (GeoLevel[mm]) := NA]
}
#### Number of unique entries at each level 
#### Subtract 1 because NA is counted as 1 (all NA entries are MainPlNum == 1L)
irp5[, MainPlNum := as.integer(length(unique(Mai)))-1L, by = .(Txrf, taxyear)]
irp5[, LocMunNum := as.integer(length(unique(Loc)))-1L, by = .(Txrf, taxyear)]
irp5[, DisMunNum := as.integer(length(unique(Dis)))-1L, by = .(Txrf, taxyear)]
irp5[, ProvNum   := as.integer(length(unique(Prv)))-1L, by = .(Txrf, taxyear)]
irp5[, MainPlNA := as.integer(all(is.na(Mai))), by = .(Txrf, taxyear)]
irp5[, LocMunNA := as.integer(all(is.na(Loc))), by = .(Txrf, taxyear)]
irp5[, DisMunNA := as.integer(all(is.na(Dis))), by = .(Txrf, taxyear)]
irp5[, ProvNA := as.integer(all(is.na(Prv))), by = .(Txrf, taxyear)]

irp5[, CommonLocality := "none"]
irp5[
      ProvNum == 1L & 
    DisMunNum == 1L & 
    LocMunNum == 1L & 
    MainPlNum == 1L &
    MainPlNA  == 0L, 
  CommonLocality := GeoLevel[4]]
irp5[
    ProvNum   == 1L & 
    DisMunNum == 1L & 
    LocMunNum == 1L & 
    MainPlNum != 1L & 
    LocMunNA  == 0L, 
  CommonLocality := GeoLevel[3]]
irp5[
    ProvNum   == 1L & 
    DisMunNum == 1L & 
    LocMunNum != 1L & 
    MainPlNum != 1L &
    DisMunNA  == 0L, 
  CommonLocality := GeoLevel[2]]
irp5[
    ProvNum   == 1L & 
    DisMunNum != 1L & 
    LocMunNum != 1L & 
    MainPlNum != 1L & 
    ProvNA  == 0L, 
  CommonLocality := GeoLevel[1]]
irp5[, CommonLocality := factor(CommonLocality, levels = c(GeoLevel, "none"))]
irp5[, NARowsMai := as.integer(sum(is.na(Mai))), by = .(Txrf, taxyear)]
irp5[, NARowsLoc := as.integer(sum(is.na(Loc))), by = .(Txrf, taxyear)]
irp5[, NARowsDis := as.integer(sum(is.na(Dis))), by = .(Txrf, taxyear)]
irp5[, NARowsPrv := as.integer(sum(is.na(Prv))), by = .(Txrf, taxyear)]
irp5[, TotalRowsInTxrf := as.integer(.N), by = .(Txrf, taxyear)]
qsave(irp5, paste0(pathdata, "irp5_LocationGranularity.qs"), nthreads = 13)

unique(irp5[Txrf %in% Txrf[grepl("^M", LocGranularFirm) | grepl("^M", CommonLocality)][1]
  & taxyear == 2009, 
  c("Txrf", "UID", geovars, grepout("Num|Rows|Gran|Commo", colnames(irp5))), with = F])

#### LocGranularFirm: the most granular location info at firm level
#### LocGranularJob : the most granular location info at each job level
#### MainPlNum      : the number of unique entries at main place level for firm-taxyear
#### NARowsMai      : the number of NA entries at main place level for firm-taxyear
#### CommonLocality : Lowest area level common among all the jobs for firm-taxyear
####   E.g., if all jobs are in Stellenbosch (main place), it is main place
####   If there are jobs attached to a branch in Capetown, then it is province
####   If location info is missing in some rows, then "none"

#### We know that rows with no location info can be allocated to the area level by:
####   MainPlNum == 1L                    ==> main place
####   MainPlNum >  1L & LocMunNum == 1L  ==> local muni
####   LocMunNum >  1L & DisMunNum == 1L  ==> district muni
####   DisMunNum >  1L & ProvNum == 1L    ==> province 

#### Panel of various levels
#### Main place panel: 
####  a. location entries are complete & location is granular down to main place level, or,
####        NARowsMai == 0L & LocGranularFirm == Mai
####  b. location == NA for some rows & uniquely common locality level is main place
####        NARowsMai > 0L  & MainPlNum == 1L
#### Local municipality panel: (addition to main place panel)
####  a. location entries are complete down to local municipality & location is granular down to local municipality, or,
####        NARowsLoc == 0L & LocGranularFirm == Loc
####  b. location == NA for some rows & uniquely common locality level is local municipality
####        NARowsLoc > 0L  & LocMunNum == 1L
#### District municipality panel: (addition to local municipality panel)
####  a. location entries are complete down to district municipality & location is granular down to district municipality, or,
####        NARowsDis == 0L & LocGranularFirm == Dis
####  b. location == NA for some rows & uniquely common locality level is district municipality
####        NARowsDis > 0L  & DisMunNum == 1L
#### Province panel: (addition to district municipality panel)
####  a. location entries are complete down to province & location is granular down to province, or,
####        NARowsPrv == 0L & LocGranularFirm == Prv
####  b. location == NA for some rows & uniquely common locality level is province
####        NARowsPrv > 0L  & ProvNum == 1L
iiM <- irp5[,
  NARowsMai == 0L & grepl("^M", LocGranularFirm) | 
  NARowsMai >  0L & MainPlNum == 1L]
iiL <- !iiM & irp5[, 
  NARowsLoc == 0L & grepl("^L", LocGranularFirm) | 
  NARowsLoc >  0L & LocMunNum == 1L]
iiD <- !iiM & !iiL & irp5[, 
  NARowsDis == 0L & grepl("^D", LocGranularFirm) | 
  NARowsDis >  0L & DisMunNum == 1L]
iiP <- !iiM & !iiL & !iiD & irp5[, 
  NARowsPrv == 0L & grepl("^P", LocGranularFirm) | 
  NARowsPrv >  0L & ProvNum   == 1L] 
nrow(irp5M <- irp5[iiM, ])
nrow(irp5L <- irp5[iiL, ])
nrow(irp5D <- irp5[iiD, ])
nrow(irp5P <- irp5[iiP, ])
qsave(irp5M, paste0(pathdata, "irp5M.qs"), nthreads = 13)
qsave(irp5L, paste0(pathdata, "irp5L.qs"), nthreads = 13)
qsave(irp5D, paste0(pathdata, "irp5D.qs"), nthreads = 13)
qsave(irp5P, paste0(pathdata, "irp5P.qs"), nthreads = 13)

irp5L[Txrf %in% unique(Txrf)[4]] #### 5 rows missing location, 2 branches in same loc muni
irp5D[Txrf %in% unique(Txrf)[5]] #### 1 row missing location, 2 branches in same dist muni
irp5P[Txrf %in% unique(Txrf)[5]] #### 1 row missing location, 3 branches in same prov

#### Missingness in location information
#### 1. Maximum sample size data sets
####   a. Main Place data = firms with information down to Main Place (irp5M) 
####   b. Loc Muni data   = firms with information down to Loc Muni   (irp5L) + a
####   c. Dis Muni data   = firms with information up to Dist Muni    (irp5D) + a + b
####   d. Prov data       = firms with information up to Prov         (irp5P) + a + b + c
#### 2. Same sample data sets (irp5M) 
####   a. Main Place data = firms with information up to Main Place
####   b. Loc Muni data   = a
####   c. Dis Muni data   = a
####   d. Prov data       = a

#### Agricultural panel
irp5Ma <- irp5M[grepl("^Anim|^Plant pro|crops|Logging|forest", imp_mic_sic7_3d), ]
qsave(irp5Ma, paste0(pathdata, "irp5Ma.qs"), nthreads = 13)
```

```{r copy irp5M and derived files to saveddata, eval = F}
FilesToCopy <- c(paste0("irp5", c("M", "L", "D", "P"), ".qs"), "irp5Ma.qs")
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
    )
```


# Compute HHI


We use main place level sample `irp5M` by following the "same sample data sets"

`irp5M` data only holds entries with location information of Main Place

1. location entries are complete & location is granular down to main place level, or, NARowsMai == 0L & LocGranularFirm == Mai
1. location == NA for some rows & uniquely common locality level is main place

For Jan 23, 2026, we use `irp5[busmainplc_geo != "", ]`, which is close to `irp5M`, because I do not have time to run to get `irp5M`


```{r hhi, eval = F}
#### code "hhi"
library(qs)
library(data.table)
#### irp5M <- qread(paste0(pathdata, "irp5M.qs"), nthreads = 13)
irp5 <- qread(paste0(pathsaveddata, "irp5.qs"), nthreads = 13)
geovars <- grepout("^bus.*geo$", colnames(irp5))
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
for (yr in 8:22) {
####  ipyr <- irp5M[taxyear == 2000+yr, ]
  ipyr <- irp5[taxyear == 2000+yr, ]
  LShare <- NULL
  for (g in 1:length(GeoLevel)) {
    print(GeoLevel[g])
    byvar <- 	c("busprov_geo", "busdistmuni_geo", 
                "buslocmuni_geo", "busmainplc_geo")[1:g]
    ipGeo = copy(ipyr)
    #### Drop rows with no location info and location is "EXCEPTION"
    ipGeo <- ipGeo[get(byvar[g]) != "" & !is.na(get(byvar[g])) & get(byvar[g]) != "EXCEPTION", ]
    #### ipGeo <- ipGeo[!grepl("none", LocGranularJob) & get(byvar[g]) != "EXCEPTION", ]
    print(summary(ipGeo[, grepout("NARows", colnames(ipGeo)), with = F]))
    ipGeo[, WorkersAtEstab := as.integer(.N), by = c(byvar, "Txrf")]
    #### ggplot(ipGeo[Txrf %in% Txrf[WorkersAtEstab > 50000][2], .(Txrf, a3601_income)], aes(x=a3601_income)) +
    #### geom_density() + scale_x_continuous(trans="log10") ==> # of workers > 50000 at main place seems correct
    #### E.g., Toyota in Umhlanga, Durban had over 80K workers    
	  #### Some firms have missing location information in some (but not all) rows
	  #### Cannot allocate 8 workers with missing location info between branches
    if (yr == 12) print(
      ipGeo[Txrf %in% unique(Txrf[busmainplc_geo != ""])[2], 
        c(geovars, "Txrf", "payereferenceno", "UID",  "dateofbirth", 
        "periodemployedfrom", "WorkersAtEstab", "kerr_income"), with = F])
    ipGeo[, WorkersInMarket := as.integer(.N), by = byvar]
    ipGeo[, Share := round(WorkersAtEstab/WorkersInMarket, 8)]
    #### ShareG, HHIG: Share and HHI after dropping GovEntity ####
    ipGeo[, WorkersAtEstabNoGov := 0L]
    ipGeo[!grepl("gov", Entity), WorkersAtEstabNoGov := WorkersAtEstab]
	  #### Below gives private firm totals for rows of private firms
    ipGeo[, WorkersInMarketNoGov := as.integer(.N), by = c(byvar, "Entity")]
    ipGeo[, ShareNoGov := round(WorkersAtEstabNoGov/WorkersInMarketNoGov, 8)]
    ipGeo[, AreaLevel := GeoLevel[g]]
	  #### employment shares
    lshare <- unique(ipGeo[, c(
        byvar, "Entity", "Txrf",  "taxrefno",
        "AreaLevel", grepout("^Wo|^Sha", colnames(ipGeo))
      ), with = F])
    print(summary(lshare))
    lshare[WorkersAtEstab > 50000, ] 
    #### Shoprite-Checkers has HQ in Brackenfell, 160K employes
    #### MultiChoice (DSTV, etc.) has HW in Randburg, 8K employees
    #### PicknPay has HQ in Kinilworth (Main Place = Capetown), 90K employees
    #### Woolworth has HQ in Capetown, 33K employees
    lshare[, HHI := sum(Share^(2), na.rm = T), by = byvar]
    lshare[, nHHI := (HHI-1/WorkersInMarket)/(1-1/WorkersInMarket)]
    lshare[, HHIG := sum(ShareNoGov^(2), na.rm = T), by = byvar]
    lshare[, nHHIG := (HHIG-1/WorkersInMarketNoGov)/(1-1/WorkersInMarketNoGov)]
    LShare <- rbindlist(list(LShare, lshare), use.names = T, fill = T)
  }
  LShare[, AreaLevel := factor(AreaLevel, levels = GeoLevel)]
  LShare[, (geovars) := lapply(.SD, factor), .SDcols = geovars]
  setkey(LShare, AreaLevel, Txrf, busprov_geo, busdistmuni_geo, 
    buslocmuni_geo, busmainplc_geo, HHI)
  setcolorder(LShare, c("Txrf", "taxrefno", 
    "busprov_geo", "busdistmuni_geo", "buslocmuni_geo", "busmainplc_geo",
    "Entity", "AreaLevel", 
    "WorkersAtEstab", "WorkersInMarket", 
    "WorkersAtEstabNoGov", "WorkersInMarketNoGov"))
  if (yr < 10) yr <- paste0("0", yr)
  qsave(LShare, paste0(pathdata, "ShareHHI", yr, ".qs"), nthreads = 13)
}
#### LShare:  Txrf, Share, ShareNoGov, HHI, HHINoGov, AreaLevel for all entities 
#### in all area levels (main place/local municipality/district municipality/province)
#### This will be merged to ipyr data using Txrf
#### Before the merge, we check integrity of LShare.

library(ggplot2)
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[is.na(busdistmuni_geo) & grepl('D', AreaLevel), ]
LShare[grepl('M', AreaLevel) & WorkersInMarket < 500, ]
g12 <- ggplot(data = LShare, aes(x = HHI, colour = AreaLevel, fill = AreaLevel)) +
  geom_density(alpha = .2) + 
  scale_fill_manual(values = rep("blue", 4)) +
  scale_color_manual(values = rep("blue", 4)) +
  scale_x_continuous(trans = "log10") +
  theme(legend.position = "none") +
  facet_grid(AreaLevel ~ .)
LShare <- qread(paste0(pathdata, "ShareHHI13.qs"))
g13 <- ggplot(data = LShare, aes(x = HHI, colour = AreaLevel, fill = AreaLevel)) +
  geom_density(alpha = .2) + 
  scale_fill_manual(values = rep("blue", 4)) +
  scale_color_manual(values = rep("blue", 4)) +
  scale_x_continuous(trans = "log10") +
  theme(legend.position = "none") +
  facet_grid(AreaLevel ~ .)
ggsave(
  paste0(pathresults, "HHI2012byAreaLevel.jpg")
  , g12, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
ggsave(
  paste0(pathresults, "HHI2013byAreaLevel.jpg")
  , g13, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
#### HHI in 2012 and 2013 are similar 
#### HHI at Dist muni and Loc muni are similar
```

```{r create LShare panel, eval = F}
LS <- NULL
for (yr in 8:22) {
  if (yr < 10) yr <- paste0("0", yr)
  ls <- qread(paste0(pathdata, "ShareHHI", yr, ".qs"))
  ls[, taxyear := 2000+as.numeric(yr)]
  LS <- rbindlist(list(LS, ls))
}
qsave(LS, paste0(pathdata, "LShareHHI.qs"), nthreads = 13)
```


```{r LSa is LS of agri only}
inda <- qread(paste0(pathsaveddata, "TaxrefnoOfAgri.qs"), nthreads = 13)
LS <- qread(paste0(pathdata, "LShareHHI.qs"), nthreads = 13)
LSia <- inda[LS, on = "taxrefno"]
LSa <- LSia[Crop == 1L | Animal == 1L | Forest == 1L, ]
#### LSa is rows of agri only, but HHI is measured at all industries
qsave(LSa, paste0(pathdata, "HHIAgriRows.qs"), nthreads = 13)
LSa <- qread(paste0(pathdata, "HHIAgriRows.qs"), nthreads = 13)
table(LSa[grepl("M", AreaLevel), .(Nums = 1:.N), 
  by = c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
  "busmainplc_geo", "taxrefno")][Nums == 1L, taxyear])
```

```{r copy LShareHHI.qs HHIAgriRows.qs to saveddata, eval = F}
FilesToCopy <- c("LShareHHI.qs", "HHIAgriRows.qs")
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
    )
```

```{r further checks on shares, eval = F}
#### further checks
LShare <- qread(paste0(pathdata, "ShareHHI12.qs"))
LShare[, Nums := .N, by = Txrf]
tb <- table(LShare[, Nums])
numtb <- as.numeric(names(tb))
numtb <- numtb[order(-numtb)]
numtb[1:(length(numtb)-5)]
tb
LShare[Nums %in% numtb[1:10], 
  .(Txrf, busprov_geo, busmainplc_geo, AreaLevel, 
    WorkersAtEstab, WorkersAtEstabNoGov, 
  	WorkersInMarket, WorkersInMarketNoGov, Share, 
	  HHI)]
ipyr <- qread(paste0(pathdata, "irp512.qs"), nthreads = 13)
ipyr <- ipyr[grepl("A", natureofperson), ]
ipyr[, Txrf := taxrefno]
ipyr[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
LShare[Txrf %in% c(ipyr[1:5, Txrf], 
  "AACAACCBBZ", "AACAACGZCZ", "AACAACZZTZ", "AACAAKCAQZ", 
  "AACABACJJZ", "AACABCAXTZ", "AACABCBCGZ", "AACABCQTQZ")
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)]
print(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
summary(LShare[TotalNeEmp <= 1000,
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)])
print(LShare[TotalNeEmp <= 1000 & !is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
print(LShare[!is.na(Newgt.kerr),
  .(Txrf, AreaLevel, busprov_geo, busmainplc_geo, 
    irp5_kerr_iweight_b, Newgt.kerr,
    EachNeEmp, TotalNeEmp, ShareEmp, HHI)], topn = 25)
```

# Plot HHI

``{r plot HHI by year, eval = F}
library(ggplot2)
LS <- qread(paste0(pathsaveddata, "LShareHHI.qs"), nthreads = 13)
gsh <- ggplot(data = LS, 
    aes(x = HHI)) +
  geom_density(alpha = .2) + 
  scale_fill_manual(values = "blue") +
  scale_color_manual(values = "blue") +
  scale_x_continuous(trans = "log10") +
  theme(legend.position = "none") +
  facet_wrap( ~ taxyear, dir = "tl")
ggsave(
  paste0(pathresults, "HHI2010-2022.jpg")
  , gsh, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```


