---
title: "Construct IRP5 panel"
date: "`{r} format(Sys.time(), "%Y %b %d %R")`"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 23/IRP5Condense.rmd"))
---

#  Read IRP5_20yr_cleaned.dta (ver 5 beta) and save by year

```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
pathdataIRP <- "U:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
geovars <- c("busprov_geo", "busdistmuni_geo", 
  "buslocmuni_geo", "busmainplc_geo")
```
Rread irp5 data by year

* read from stata dta
* dropped if
    * natureofperson != "A"
    * taxrefno == ""
#### Note: many empty taxrefno in irp5 (2012) before dropping
#### table(ipyr[, taxrefno == ""])
####   FALSE    TRUE 
#### 8586068 1407896 


```{r read irp5v5beta best geo and create irp5yr files, eval = F}
library(qs)
library(data.table)
library(readstata13)
for (yr in 8:22) {
#### for (yr in 8) {
  if (yr < 10) yr <- paste0("0", yr)
  ## Below code is run only once: Start ##
  ## irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, "_cleaned.dta"))
  ## ipyr <- data.table(irpyr)
  ## rm(irpyr)
  ## qsave(ipyr, paste0(pathdata, "irp", yr, ".qs"), nthreads = 13)
  ## irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, 
  ##   "_cleaned_best_geodata.dta"))
  ## ipb <- data.table(irpyr)
  ## qsave(ipb, paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## # merge best geocode to irpYR.qs
  ## ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), nthreads = 13)
  ## ipb <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## geonames <- grepout("^bus.*geo$", colnames(ipyr))
  ## setnames(ipyr, geonames, paste0("OLD", geonames))
  ## ipbyr <- ipyr[ipb, , on = grepout("irp.*aid", colnames(ipb))]
  ## nrow(ipbyr[OLDbusmainplc_geo == "" & busmainplc_geo != "", ])
  ## qsave(ipbyr, paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## Below code is run only once: End ##
  ipyr <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = 13)
  ## Keep only nature of person is "an individual"
  ### ipyr <- ipyr[grepl("A", natureofperson), ]
  ## Drop obs with missing taxrefno
  ipyr <- ipyr[taxrefno != "", ]
  ## Note that there are header lines with taxrefno = "NULL", but they can be dropped later
  ## Num: Number of employees in a firm
  ipyr[, Num := as.integer(.N), by = .(taxrefno, 
    busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo)]
  ipyr[, Num := 0L]
  ipyr[, Nationality := "sa"]
  ipyr[passportno != "", Nationality := "non-sa"]
  ipyr[, Nationality := factor(Nationality, levels = c("sa", "non-sa"))]
  irp5yr <- ipyr[, .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, irp5it3aid, taxrefno, UID, idno,
    Nationality, natureofperson,
    payereferenceno, gender, dateofbirth, taxyear, 
    periodemployedfrom, periodemployedto,
    totalperiodsworked, totalperiodsinyearofassessment,
    #### a3601_income: basic salary
    Num, kerr_income, ptrs_income, a3601_income)]
  qsave(irp5yr, paste0(pathdata, "irp5", yr, ".qs"), nthreads = 13)
}
```



# Read industry code


```{r read industry code data, eval = F}
library(readstata13)
ind <- read.dta13(paste0(pathdata, "IndustryCodeFromCITv5.dta"))
ind <- data.table(ind)
setkey(ind, taxrefno)
indNoNA <- ind[!(is.na(imp_mic_sic7_1d) & is.na(imp_mic_sic7_2d) & 
  is.na(imp_mic_sic7_3d) & is.na(imp_mic_sic7_5d)), ]
for (ii in c(1:3, 5))
  indNoNA[, (paste0("change", ii, "d")) := 
    as.integer(uniqueN(get(paste0("imp_mic_sic7_", ii, "d")))), 
    by = taxrefno]
table(indNoNA[, change3d])
qsave(ind, paste0(pathdata, "IndustryCodeFromCIT.qs"), nthreads = 13)
qsave(indNoNA, paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = 13)
```

# Create irp5 panel

Appending all annual files

```{r create long format of irp5, eval = F}
library(qs)
library(data.table)
irp5 <- NULL
for (yr in 8:22) {
  if (yr < 10) yr <- paste0("0", yr)
  if (yr == 12)  
    irp5 <- rbindlist(list(irp5, 
      qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = 13)), use.names = T, fill = T) else 
    irp5 <- rbindlist(list(irp5, 
      qread(paste0(pathdata, "irp5", yr, ".qs"), nthreads = 13)), use.names = T, fill = T)
}
irp5[, Num := as.integer(.N), by = .(busprov_geo,
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno, taxyear)]
#### 1. Use payereferenceno to count the total
####    >>> This is a wrong way to go: payereferenceno has many misreporting <<<
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
irp5[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
irp5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
irp5[, Entity := "private"]
irp5[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
irp5[, Entity := factor(Entity, levels = c("private", "gov"))]
geovars <- grepout("^bus.*geo$", colnames(irp5))
#### Below takes about 2 hours!!!: Start  ###
irp5[, DateStart := as.IDate(periodemployedfrom)]
irp5[, DateEnd   := as.IDate(periodemployedto)]
irp5[, DateBirth := as.IDate(dateofbirth)]
#### Below takes about 2 hours!!!: End  ###
```

```{r checking 2009, eval = F}
irp09 <- qread(paste0(pathdata, "irp509.qs"), nthreads = 13)
irp09[, c("Num", "num") := list(.N, 1:.N), by = c(geovars, "taxrefno", "taxyear")]
irp09[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
irp09[num == 1L & grepl("A", NatureOfPer), .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####     1381105
irp10 <- qread(paste0(pathdata, "irp510.qs"), nthreads = 13)
irp10[, c("Num", "num") := list(.N, 1:.N), by = c(geovars, "taxrefno", "taxyear")]
irp10[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
irp10[num == 1L & grepl("A", NatureOfPer), .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####    12303557
irp14 <- qread(paste0(pathdata, "irp514.qs"), nthreads = 13)
irp14[, c("Num", "num") := list(.N, 1:.N), by = c(geovars, "taxrefno", "taxyear")]
irp14[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
irp14[num == 1L & grepl("A", NatureOfPer), .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####    12303557
irp5[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
irp5[, num := 1:.N, by = c(geovars, "taxrefno", "taxyear")]
irp5[num == 1L & grepl("A", NatureOfPer), .(JobTotal = sum(Num, na.rm = T)), by = taxyear]
#### Looks fine in irp5!
####     taxyear JobTotal
####       <int>    <int>
####  1:    2008 12323161
####  2:    2009 12083707
####  3:    2010 12135165
####  4:    2011 13034038
####  5:    2012 13496238
####  6:    2013 13930394
####  7:    2014 14188697
####  8:    2015 15586123
####  9:    2016 15240490
#### 10:    2017 15267221
#### 11:    2018 15566584
#### 12:    2019 15635489
#### 13:    2020 15654027
#### 14:    2021 13989801
#### 15:    2022 13930391
irp5a <- qread(paste0(pathdata, "irp5a.qs"), nthreads = 13)
irp5a[, c("Num", "num") := list(.N, 1:.N), by = c(geovars, "taxrefno", "taxyear")]
irp5a[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
irp5a[num == 1L & grepl("A", NatureOfPer), .(JobTotal = sum(Num, na.rm = T)), by = taxyear]
#### Looks fine in irp5a!
####    taxyear JobTotal
####       <int>    <int>
####  1:    2008   193888
####  2:    2009   233919
####  3:    2010   251261
####  4:    2011   326655
####  5:    2012   379034
####  6:    2013   420265
####  7:    2014   407802
####  8:    2015   461713
####  9:    2016   494848
#### 10:    2017   513754
#### 11:    2018   531506
#### 12:    2019   559162
#### 13:    2020   569491
#### 14:    2021   652310
#### 15:    2022   566829
irp5a[, .(total = sum(busmainplc_geo != "")), by = taxyear]
#### Looks horrible once conditioning on location info in irp5a!
####     taxyear  total
####       <int>  <int>
####  1:    2008   4338
####  2:    2009   4357
####  3:    2010 164006
####  4:    2011   6085
####  5:    2012 230056
####  6:    2013 396541
####  7:    2014 405210
####  8:    2015 457141
####  9:    2016 490655
#### 10:    2017 512344
#### 11:    2018 529679
#### 12:    2019 556736
#### 13:    2020 565447
#### 14:    2021 646384
#### 15:    2022 561929
```

```{r correcting information of IRP panel, eval = F}
####  addmargins(table(irp5[is.na(uid) | is.na(UID), .(uid=is.na(uid), UID=is.na(UID))]))
####        UID
#### uid         FALSE      TRUE       Sum
####   FALSE         0   4409392   4409392
####   TRUE  199028655         0 199028655
####   Sum   199028655   4409392 203438047
#### #### So sub
irp5[ is.na(uid) & !is.na(UID), uid := UID]
irp5[!is.na(uid) &  is.na(UID), UID := uid]
#### Copy natureofperson across years for same uid
irp5[, NatureOfPer := "ZZ"]
irp5[uid %in% uid[grepl("A", natureofperson)], NatureOfPer := "A"]
qsave(irp5, paste0(pathdata, "irp5_BeforeDroppingRepetitiveUIDs.qs"), nthreads = 13)
```

```{r merge industry code, eval = F}
#### Merge industry code, taking irp5 as base
irp5 <- qread(paste0(pathdata, "irp5_BeforeDroppingRepetitiveUIDs.qs"), 
  nthreads = 13)
ind <- qread(paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = 13)
setkey(irp5, taxrefno); setkey(ind, taxrefno, taxyear)
ind[, n := 1:.N, by = taxrefno]
indFirstYear <- ind[n == 1, ]
indFirstYear[, n := NULL]
setkey(indFirstYear, taxrefno)
setnames(indFirstYear, "taxyear", "FirstYearOnSIC7")
irp5i <- indFirstYear[irp5, on = "taxrefno"]
setcolorder(irp5i, c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
  "busmainplc_geo", "taxrefno", "Txrf", "uid", "UID", "gender",
  grepout("Date", colnames(irp5i))))  
qsave(irp5i, paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = 13)
```

# Copy location information from other source: IRP5_Geo

```{r Open the geolocation file and produce geoloc file, eval = F}
library(readstata13)
geodat <- read.dta13(paste0(path, "IRP5_Geo.dta"))
geodat <- data.table(geodat)
setkey(geodat, IRP5IT3aID)
setnames(geodat, 
  c("IRP5IT3aID", "TaxYear", "PAYERefno"),
  c("irp5it3aid", "taxyear", "payereferenceno")
)
geodat[, taxyear := as.integer(taxyear)]
geoloc <- geodat[taxyear==2015, ]
qsave(geoloc, paste0(pathdata, "GeoLocFile.qs"), nthreads = 13)
geodat[, HaveInfo := 0L]
geodat[(Bus_adr_Geo_Munic != "" & !is.na(Bus_adr_Geo_Munic)), 
  HaveInfo := 1L]
table(geodat[, .(HaveInfo, taxyear)], useNA = "ifany")
qsave(geodat, paste0(pathdata, "GeoLocationFile.qs"), nthreads = 13)
```

```{r smallirp5, eval = F}
irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = 13)
smallirp5 <- irp5i[grepl("A", NatureOfPer), .(taxrefno, payereferenceno)]
smallirp5 <- smallirp5[!duplicated(smallirp5), ]
qsave(smallirp5, paste0(pathdata, "smallIRP5.qs"), nthreads = 13)
```

```{r GeoLF, eval = F}
geoloc <- qread(paste0(pathdata, "GeoLocFile.qs"), nthreads = 13)
smallirp5 <- qread(paste0(pathdata, "smallIRP5.qs"), nthreads = 13)
setkey(geoloc, payereferenceno)
setkey(smallirp5, payereferenceno)
GeoLF<- smallirp5[geoloc, on = "payereferenceno"]
qsave(GeoLF,paste0(pathdata, "GeoLF.qs"), nthreads = 13)
rm(geoloc)
```

```{r merge irp5 GeoLF, eval = T}
irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = 13)
GeoLF <- qread(paste0(pathdata, "GeoLF.qs"), nthreads = 13)
GeoLF[, dup := 1:.N, by = irp5it3aid]
nrow(GeoLF[dup > 1, ])
GeoLF <- GeoLF[dup == 1L, ]
GeoLF[, dup := NULL]
irp5gi <- GeoLF[irp5i, on = c("irp5it3aid", "payereferenceno", "taxrefno")]
colnames(GeoLF)
rbind(GeoLF=dim(GeoLF), irp5i=dim(irp5i), irp5ig=dim(irp5ig))
colnames(irp5ig)
table(irp5gi[, taxyear], useNA = "ifany")
table(irp5gi[, i.taxyear], useNA = "ifany")
table(GeoLF[, taxyear], useNA = "ifany")
table(irp5i[, taxyear], useNA = "ifany")
irp5gi[, taxyear := NULL]
if (any(grepl("i.taxyear", colnames(irp5gi)))) 
  setnames(irp5gi, "i.taxyear", "taxyear")
qsave(irp5gi, paste0(pathdata, "irp5_MergedGeoLF.qs"), 
  nthreads = 13))
```

Copy from GeoLF if 

* `UID` and `taxrefno` are the keys

```{r copy 2015 location info, eval = F, echo = F}
irp5gi <- qread(paste0(pathdata, "irp5_MergedGeoLF.qs"), 
  nthreads = 13))
irp5gi[, uidtaxref := 0L]
irp5gi[, uidtaxref := as.integer(factor(paste(taxrefno, uid, sep = "-")))]
irp5gi[, c("NoLocMuni", "CanFill", "ShouldFill") := 0L]
irp5gi[buslocmuni_geo == "" | is.na(buslocmuni_geo),
  NoLocMuni := 1L]
irp5gi[Bus_adr_Geo_Munic != "" & !is.na(Bus_adr_Geo_Munic), 
  CanFill := 1L]
irp5gi[NoLocMuni == 1L & CanFill == 1L, ShouldFill := 1L]
#### How much is loc muni information populated?
table(irp5gi[, .(NoLocMuni == 0L, taxyear)], useNA = "ifany")
NAsInLocMuniBefore <- 
  irp5gi[, .(BeforeFill = sum(NoLocMuni)), by = taxyear]
print(NAsInLocMuniBefore)
#### to be safe, create a variable for "fill in"
irp5gi[, FillInLocMuni := buslocmuni_geo]
#### For each uidtaxref that ShouldFill == 1L, 
#### fill in loc muni info from the 1st CanFill row
irp5gi[taxyear < 2015 & ShouldFill == 1L, 
  FillInLocMuni := Bus_adr_Geo_Munic[CanFill == 1L][1], 
  by = uidtaxref] 
#### How much is loc muni information populated after this?
#### Number of NAs in loc muni info
NAsInLocMuni <- irp5gi[, .(
  BeforeFill = sum(NoLocMuni), 
  AfterFill  = sum(FillInLocMuni == "" | is.na(FillInLocMuni))), 
  by = taxyear]
print(NAsInLocMuni)
#### table(irp5gi[, .(
####   LocMuniNotNA = buslocmuni_geo != "" | !is.na(buslocmuni_geo), 
####   taxyear)], 
####   useNA = "ifany")
#### How useful is GeoLocFile in fillling information?
#### table(irp5gi[, .(ShouldFill, taxyear)], useNA = "ifany")
qsave(irp5gi, paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"), 
  nthreads = 13))
```



Copy location info

* use taxrefno-uid match
* this assumes no transfer between establishments in a firm in 2012-2013
* in reality, there are transfers &rarr; measurement errors in location
* but the proportion of transfers must be small 
* benefits of location info outweigh the costs

If there are workers working for multiple establishments ... no problem 

* taxrefno-uid uniquely identifies matches (even if there is a transfer, it does not show up as an error, we will never know)
* base of join = 2012 (fix rows of 2012), and find matches from 2013
* see example shown in codes, 8 entries in 2012, 2013


```{r correcting information of IRP panel, eval = T}
#### case: taxyear = 2015, DateStart = Dec 01, 2013, DateEnd = Jan 10, 2014
#### taxyear 2015: Mar 2014 - Feb 2015
#### This is a revision report of taxyear 2014, so drop this person's report of 2013, 
#### use this row for 2014
#### paste0 operation is not slow
#### library(data.table)
#### n <- 1e5
#### dt <- data.table(
####   x1 = sample(1000:2000, n, replace= T),
####   x2 = sample(1000:2000, n, replace= T)
#### )
#### dt[, date1 := as.Date(x1, origin = "1990/01/01")]
#### dt[, date2 := as.Date(x2, origin = "1990/01/01")]
#### dt[, taxyear := year(date1)]
#### system.time(dt[, posix1 := as.POSIXct(paste0(taxyear, "/03/01"))])
####    user  system elapsed 
####    1.24    0.01    1.25 
#### system.time(dt[, idate1 := as.IDate(paste0(taxyear, "/03/01"))])
####    user  system elapsed 
####    0.42    0.00    0.44 
#### system.time(paste0(as.character(dt[, taxyear]), "/03/01"))
####    user  system elapsed 
####    0.02    0.00    0.02 
irp5gi <- qread(paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"), 
  nthreads = 13))
#### irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = 13)
irp5gi[, TYStart := as.IDate(paste0(taxyear-1, "/03/01"))]
irp5gi[, TYEnd   := as.IDate(paste0(taxyear, "/03/01"))]
irp5gi[, TrueTaxYear := taxyear]
#### Entries of earlier jobs
irp5gi[DateStart < TYStart & DateEnd < TYEnd & (taxyear - year(DateStart) <= 5) , 
  TrueTaxYear := year(DateStart)]
apply(table(irp5gi[TrueTaxYear != taxyear, .(TrueTaxYear, taxyear)]), 1, sum)
####    2003    2004    2005    2006    2007    2008    2009    2010    2011    2012 
####     376    1370    5448   16307   56706   65947 1219105 1205674 1541903 1483088 
####    2013    2014    2015    2016    2017    2018    2019    2020    2021 
#### 1482391 1356213 1427595 1561429 1599350 1437092 1496579 1436222 1173744 
#### Total is  18566539
####  TrueTaxYear taxyear
####         2010    2010
####         2011    2011 * drop this
####         2011    2012 * taxyear := TrueTaxYear
####         2012    2012
####         2013    2013
#### TrueTaxYear[TrueTaxYear != taxyear] == 2011 is year that needs to be dropped
#### taxyear == TrueTaxYear[TrueTaxYear != taxyear] is flagged as DropThisForRev
####  TrueTaxYear taxyear
####         2010    2010
####         2011    2011 ** drop this 1
####         2011    2012 *  taxyear := TrueTaxYear
####         2012    2012 ** drop this 2
####         2012    2013 *  taxyear := TrueTaxYear
#### TrueTaxYear[TrueTaxYear != taxyear] == 2011, 2012 
####   taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear 
#### is flagged as DropThisForRev
irp5gi[, DropThisForRev := 0L]
irp5gi[taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear, 
  DropThisForRev := 1L, 
  by = c(geovars, "taxrefno", "UID")]
round((sum(irp5gi[, DropThisForRev])/nrow(irp5gi))*100, 4)
####  9.0204
#### Rename 
####  taxyear ==> Old.taxyear
####  TrueTaxYear ==> taxyear, becomes
####  taxyear Old.taxyear
####     2010        2010
####     2011        2012 *  
####     2012        2013 *  
#### Revision rows dropped
irp5gir <- irp5gi[DropThisForRev == 0L, ]
c(nrow(irp5gi), nrow(irp5gir))
setnames(irp5gir, c("TrueTaxYear", "taxyear"), c("taxyear", "Old.taxyear"))
irp5gir[, DropThisForRev := NULL]
#### Copy location information from 2013 ==> 2008-2012
#### Use taxrefno-uid match ==> saved 10.9M rows
#### Keep original variables
irp5gir[, orig.bprov := busprov_geo]
irp5gir[, orig.bdist := busdistmuni_geo]
irp5gir[, orig.bloc  := buslocmuni_geo]
irp5gir[, orig.bmain := busmainplc_geo]
#### If we do not condition on first row "[1]", we wind up 
#### having multiple cases ==> error
#### If there are multiple matches, we have no idea which to pick, so pick first
irp5gir[, bprov13 := busprov_geo[taxyear == 2013][1], by = .(taxrefno, uid)]
irp5gir[, bdist13 := busdistmuni_geo[taxyear == 2013][1], by = .(taxrefno, uid)]
irp5gir[, bloc13  := buslocmuni_geo[taxyear == 2013][1], by = .(taxrefno, uid)]
irp5gir[, bmain13 := busmainplc_geo[taxyear == 2013][1], by = .(taxrefno, uid)]
irp5gir[orig.bloc  == "" & bloc13  != "", .(taxrefno, uid, taxyear, bloc13, buslocmuni_geo)]
irp5gir[orig.bloc  == "" & bloc13  != "", .(taxrefno, uid, taxyear, 
  bloc13, buslocmuni_geo, bmain13, busmainplc_geo)]
irp5gir[, c("LocInfoCopiedFrom", "ProvCopied", "DisCopied", "LocCopied", "MaiCopied") := 0L]
#### Copy from bXXXX13 
irp5gir[
  taxyear <= 2012 &
  (busprov_geo == "" | is.na(busprov_geo) | busprov_geo == "EXCEPTION") & 
  (bprov13 != "" & !is.na(bprov13) & bprov13 != "EXCEPTION"), 
   c("busprov_geo", "ProvCopied", "LocInfoCopiedFrom") := .(bprov13, 1L, 2013L)
]
irp5gir[
  taxyear <= 2012 &
  (busdistmuni_geo == "" | is.na(busdistmuni_geo) | busdistmuni_geo == "EXCEPTION") & 
  (bdist13 != "" & !is.na(bdist13) & bdist13 != "EXCEPTION"),
  c("busdistmuni_geo", "DisCopied", "LocInfoCopiedFrom") := .(bdist13, 1L, 2013L)
]
irp5gir[
  taxyear <= 2012 &
  (buslocmuni_geo == "" | is.na(buslocmuni_geo) | buslocmuni_geo == "EXCEPTION") &
  (bloc13 != "" & !is.na(bloc13) & bloc13 != "EXCEPTION"),
  c("buslocmuni_geo",  "LocCopied", "LocInfoCopiedFrom") := .(bloc13,  1L, 2013L)
]
irp5gir[
  taxyear <= 2012 &
  (busmainplc_geo == "" | is.na(busmainplc_geo) | busmainplc_geo == "EXCEPTION") &
  (bmain13 != "" & !is.na(bmain13) & bmain13 != "EXCEPTION"),
  c("busmainplc_geo",  "MaiCopied", "LocInfoCopiedFrom") := .(bmain13, 1L, 2013L)
]
qsave(irp5gir, paste0(pathdata, "irp5_RevReports.qs"), nthreads = 13)
```
```{r dropping rows, eval = T}
#### Note: Firms with anomalous repetetive uids (AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ) are dropped
#### code "dropthese"
#### Read data
#### nthreds = 16 at NT-SDF, = 8 with my laptop (less 1, keep 1 for other computations)
ipyr <- qread(paste0(pathdata, "irp512.qs"), nthreads = 13)
ipyr[, uid := gsub(" +$", "", UID)]
rm(ipyrs)
#### uid anomalous entries: AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ
#### these are 0.035% of total number of rows
#### one large employer (with 4632 jobs) contributes 
#### DO NOT use these uids when copying location from future data
dupuid <- ipyr[
    duplicated(uid) | duplicated(uid, fromLast=T), 
    .(taxrefno, uid=uid)
  ]
dupuid[, n := .N, by = .(taxrefno, uid)]
dupuid <- unique(dupuid[order(n, decreasing = T), ])
dupuid[n > 10, ]
print(unique(dupuid[, uid])[1:30], quote = F)
#### repetitive: [A-Z]*8
repetetive <- dupuid[grepl("(.)\\1{7,}", uid), uid]
tbr <- table(repetetive)
tbr <- tbr[order(tbr, decreasing=T)]
repetetive <- unique(repetetive)
#### ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
####   .(taxrefno, uid, busprov_geo, bprov.13, 
####   busmainplc_geo, bmain.13)][uid %in% repetetive, ][
####   order(uid), ]
#### firms with repetetive uids and their counts
tb <- table(ipyr[uid %in% repetetive, taxrefno])
tb <- tb[order(tb, decreasing=T)]
tb[1:20]
#### repetetive entries are concentrated in 3 firms
dupuid[1:20, ]
dupuid[taxrefno %in% names(tb)[1:3], ][1:20, ]
#### firms with 800 same uids: using same uid for all workers
ipyr[uid %in% dupuid[1, uid],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)]
#### same repetetive entries (AAAAA...) are found across firms
ipyr[uid %in% names(tbr)[1:20],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ]
#### Drop firms with repetetive entries if they are above 10
ii <- 10
ipyr[taxrefno %in% names(tb)[tb==ii] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ]
(dropthese <- ipyr[taxrefno %in% names(tb)[tb>10] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],  
  .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
  order(uid, taxrefno), ])
table(dropthese[, .(uid, taxrefno)])
table(dropthese[, uid])
####      CCCCCCCC CCCCCCCCCCCCC 
####           191          2341 
qsave(dropthese, paste0(pathdata, "DroppedTheseRepetetiveIDs.qs"), nthreads = 13)

#### Keep only nature of person is "an individual"
#### we did this: irp5[uid %in% uid[grepl("A", natureofperson)], NatureOfPer := "A"]
#### irp5ir <- qread(paste0(pathdata, "irp5_RevReports.qs"), nthreads = 13)
addmargins(table(irp5gir[, .(natureof = grepl("A", natureofperson), 
  NatureOf = grepl("A", NatureOfPer))]))
####         NatureOf
#### natureof      TRUE       Sum
####    FALSE  42731611  42731611  * 42M saved!!!
####    TRUE  132692587 132692587
####    Sum   175424198 175424198
irp5gir <- irp5gir[grepl("A", NatureOfPer), ]
#### repetetive UID
irp5Clean <- irp5gir[!(taxrefno %in% dropthese[, taxrefno]), ]
c(nrow(irp5gir), nrow(irp5Clean))
#### [1] 175424198 175176433
#### if (any(iiEmpty <- grepl("", colnames(irp5Clean)))) 
#### Agriculture panel
irp5a <- irp5Clean[grepl("^Anim|^Plant pro|crops|Logging|forest", imp_mic_sic7_3d), ]
#### grepl("^Crop|^Fore", imp_mic_sic7_2d), ]
qsave(irp5Clean, paste0(pathdata, "irp5.qs"), nthreads = 13)
qsave(irp5a,     paste0(pathdata, "irp5a.qs"), nthreads = 13)
```

```{r MeanStdN of employment, eval = F}
#### code: MeanStdN of employment
#### For some reasons, it gives "cannot allocate a vector of size 36Gb..." error
library(data.table)
library(qs)
irp5Clean <- qread(paste0(pathdata, "irp5.qs"), nthreads = 13)
irp5Clean[, c("num", "Num") := list(as.integer(1:.N), as.integer(.N)), 
  by = .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
  busmainplc_geo, taxrefno, taxyear)]
MeanStdN <- irp5Clean[num == 1L, .(
  mean = round(mean(num, na.rm = T), 3),
  std = round(var(num, na.rm = T)^(.5), 3),
  N = .N
  ), by = taxyear]
print(MeanStdN)
fwrite(MeanStdN, paste0(pathdata, "MeanStdNOfEmployment.tsv"), sep = "\t")
```

