---
title: "Construct IRP5 panel"
date: "`{r} format(Sys.time(), "%Y %b %d %R")`"
output: 
  litedown::html_format:
    meta:
      css: ["MyCSS.css", "@default", "@tabsets", "@copy-button"]
      js: ["@sidenotes", "@appendix", "@tabsets", "@copy-button"]
#### output: html_document
#### path <- "W:/epguest/seiro_ito/"; RunSep <- F; litedown::fuse(paste0(path, "outfiles/2026 01 23/IRP5Condense.rmd"))
---

#  Read IRP5_20yr_cleaned.dta (ver 5 beta) and save by year

```{r, child='setup.Rmd', message = F, warning = F}
#### setup.rmd sets eval, echo, time to T
```
```{r read functions from child file, child='destatfactor.Rmd', message = F, warning = F}
```

```{r setup, eval = F, echo = F}
library(qs); library(data.table); library(ggplot2)
library(rmarkdown); library(fixest)
library(knitr); library(tufte)
library(litedown)
litedown::reactor(
  echo = TRUE,
  eval = TRUE, 
  time = TRUE
)
options(htmltools.dir.version = FALSE, width = 100)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
#### mask_dots: replace every char except positions in `keep` with ".".
#### Used to anonymise taxrefno / uid in display-only prints (not in keys or joins).
mask_dots <- function(x, keep = c(3, 6, 7, 8, 11)) {
  x <- as.character(x)
  n <- nchar(x)
  out <- strrep(".", n)
  for (k in keep) {
    m <- !is.na(x) & n >= k
    substr(out[m], k, k) <- substr(x[m], k, k)
  }
  out
}
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
geovars <- c("busprov_geo", "busdistmuni_geo", 
  "buslocmuni_geo", "busmainplc_geo")
```

<!--
Git workflow:
----
git checkout -b work/EstImpacts2024Apr
git push -u origin work/EstImpacts2024Apr

#### many small commits
git add .
git commit -m "..."

git pull --rebase   # 鐃緒申 add this before pushing (important)
git push

#### squash & merge on GitHub

git checkout main
git pull --rebase
git branch -d work/EstImpacts2024Apr
----
-->



Read irp5 data by year

* read from stata dta
* dropped if
    * natureofperson != "A"
    * taxrefno == ""
#### Note: many empty taxrefno in irp5 (2012) before dropping
#### table(ipyr[, taxrefno == ""])
####   FALSE    TRUE 
#### 8586068 1407896 


```{r read irp5v5beta best geo and create irp5yr files, eval = F}
#| attr.source = '.r .line-numbers',
library(qs)
library(data.table)
library(readstata13)
for (yr in 8:22) {
#### for (yr in 8) {
  if (yr < 10) yr <- paste0("0", yr)
  ## Below code is run only once: Start ##
  ## irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, "_cleaned.dta"))
  ## ipyr <- data.table(irpyr)
  ## rm(irpyr)
  ## qsave(ipyr, paste0(pathdata, "irp", yr, ".qs"), nthreads = DTThreads)
  ## irpyr <- read.dta13(paste0(pathdataIRP, "IRP5_20", yr, 
  ##   "_cleaned_best_geodata.dta"))
  ## ipb <- data.table(irpyr)
  ## qsave(ipb, paste0(pathdata, "irpb", yr, ".qs"), nthreads = DTThreads)
  ## # merge best geocode to irpYR.qs
  ## ipyr <- qread(paste0(pathdata, "irp", yr, ".qs"), nthreads = DTThreads)
  ## ipb <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = DTThreads)
  ## geonames <- grepout("^bus.*geo$", colnames(ipyr))
  ## setnames(ipyr, geonames, paste0("OLD", geonames))
  ## ipbyr <- ipyr[ipb, , on = grepout("irp.*aid", colnames(ipb))]
  ## nrow(ipbyr[OLDbusmainplc_geo == "" & busmainplc_geo != "", ])
  ## qsave(ipbyr, paste0(pathdata, "irpb", yr, ".qs"), nthreads = DTThreads)
  ## Below code is run only once: End ##
  ipyr <- qread(paste0(pathdata, "irpb", yr, ".qs"), nthreads = DTThreads,
    use_alt_rep = TRUE)
  ## Keep only nature of person is "an individual"
  ### ipyr <- ipyr[grepl("A", natureofperson), ]
  ## Drop obs with missing taxrefno
  ipyr <- ipyr[taxrefno != "", ]
  ## Note that there are header lines with taxrefno = "NULL", but they can be dropped later
  ## Num: Number of employees in a firm
  #### CLAUDE tpo: 2026-07-02
  ####   Bug: "Num := 0L" ran AFTER the grouped count and erased it, so
  ####   every saved irp5YY.qs carried Num = 0.
  ####   Fix: keep only the count. No initialisation is needed because
  ####   ":=" with "by=" already writes a value into every row.
  ipyr[, Num := as.integer(.N), by = .(taxrefno,
    busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo)]
  #### ipyr[, Num := 0L]
  ipyr[, Nationality := "sa"]
  ipyr[passportno != "", Nationality := "non-sa"]
  ipyr[, Nationality := factor(Nationality, levels = c("sa", "non-sa"))]
  irp5yr <- ipyr[, .(busprov_geo, busdistmuni_geo, buslocmuni_geo, 
    busmainplc_geo, irp5it3aid, taxrefno, UID, idno,
    Nationality, natureofperson,
    payereferenceno, gender, dateofbirth, taxyear, 
    periodemployedfrom, periodemployedto,
    totalperiodsworked, totalperiodsinyearofassessment,
    #### a3601_income: basic salary
    Num, kerr_income, ptrs_income, a3601_income)]
  #### CLAUDE tpo: 2026-07-07 use_alt_rep is qread-only (qs 0.27.3);
  ####   qsave() has no such arg -> unused argument error
  #### qsave(irp5yr, paste0(pathdata, "irp5", yr, ".qs"), nthreads = DTThreads,
  ####   use_alt_rep = TRUE)
  qsave(irp5yr, paste0(pathdata, "irp5", yr, ".qs"), nthreads = DTThreads)
}
```



# Read industry code


```{r read industry code data}
#| attr.source = '.r .line-numbers',
library(readstata13)
ind <- read.dta13(paste0(pathdata, "IndustryCodeFromCITv5.dta"))
ind <- data.table(ind)
setkey(ind, taxrefno)
indNoNA <- ind[!(is.na(imp_mic_sic7_1d) & is.na(imp_mic_sic7_2d) & 
  is.na(imp_mic_sic7_3d) & is.na(imp_mic_sic7_5d)), ]
for (ii in c(1:3, 5))
  indNoNA[, (paste0("change", ii, "d")) := 
    as.integer(uniqueN(get(paste0("imp_mic_sic7_", ii, "d")))), 
    by = taxrefno]
table(indNoNA[, change3d])
#### CLAUDE tpo:
#### qsave(ind, paste0(pathdata, "IndustryCodeFromCIT.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
#### qsave(indNoNA, paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(ind, paste0(pathdata, "IndustryCodeFromCIT.qs"), nthreads = DTThreads)
qsave(indNoNA, paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = DTThreads)
```

# Create irp5 panel

Appending all annual files

```{r create long format of irp5}
#| attr.source = '.r .line-numbers',
library(qs)
library(data.table)
library(fasttime)
irp5pnl <- NULL
for (yr in 8:22) {
  if (yr < 10) yr <- paste0("0", yr)
  if (yr == 12)  
    irp5pnl <- rbindlist(list(irp5pnl, 
      qread(paste0(pathdata, "ipyrsClean.qs"), nthreads = DTThreads,
        use_alt_rep = TRUE)), 
      use.names = T, fill = T) else 
    irp5pnl <- rbindlist(list(irp5pnl, 
      qread(paste0(pathdata, "irp5", yr, ".qs"), nthreads = DTThreads,
        use_alt_rep = TRUE)), 
      use.names = T, fill = T)
}
irp5pnl[, Num := as.integer(.N), by = .(busprov_geo,
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno, taxyear)]
#### 1. Use payereferenceno to count the total
####    >>> This is a wrong way to go: payereferenceno has many misreporting <<<
#### 2. Compute the shares of each firms
####   a. In doing so, create a hypothetical "gov entity" to aggregate the entries with NAs in taxrefno
####   b. Compute the shares of each firms including "gov entity" thence HHI
irp5pnl[, Txrf := taxrefno]
#### Note: GovEntity has taxrefno == "" or NULL, so omit from unique operation below
irp5pnl[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Txrf := "GovEntity"]
irp5pnl[, Entity := "private"]
irp5pnl[grepl("NULL", taxrefno) | taxrefno == "" | is.na(taxrefno), Entity := "gov"]
irp5pnl[, Entity := factor(Entity, levels = c("private", "gov"))]
geovars <- grepout("^bus.*geo$", colnames(irp5pnl))
#### Below takes about 2 hours (with as.IDate)!!!: Start  ###
library(fasttime) 
#### library(anytime)  #### for DOB, fasttime works only for year > 1970
irp5pnl[, DateStart := fastPOSIXct(periodemployedfrom, NULL, 3L)]
irp5pnl[, DateEnd   := fastPOSIXct(periodemployedto,   NULL, 3L)]
irp5pnl[, DateBirth := as.IDate(dateofbirth)]
irp5pnl[!is.na(DateStart), 
  .(DateStart, periodemployedfrom, DateEnd, periodemployedto, 
  DateBirth, dateofbirth)][1:20, ]
#### Below takes about 2 hours!!!: End  ###
```

```{r checking 2009 2010 2014, eval = F}
#| attr.source = '.r .line-numbers',
irp09 <- qread(paste0(pathdata, "irp509.qs"), nthreads = DTThreads,
  use_alt_rep = TRUE)
irp09[, c("Num", "num") := list(.N, 1:.N), 
  by = c(geovars, "taxrefno", "taxyear")]
irp09[UID %in% UID[grepl("^A$", natureofperson)], 
  NatureOfPer := "A"]
irp09[num == 1L & grepl("A", NatureOfPer), 
  .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####     1381105
irp10 <- qread(paste0(pathdata, "irp510.qs"), nthreads = DTThreads,
  use_alt_rep = TRUE)
irp10[, c("Num", "num") := list(.N, 1:.N), 
  by = c(geovars, "taxrefno", "taxyear")]
irp10[UID %in% UID[grepl("^A$", natureofperson)], 
  NatureOfPer := "A"]
irp10[num == 1L & grepl("A", NatureOfPer), 
  .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####    12303557
irp14 <- qread(paste0(pathdata, "irp514.qs"), nthreads = DTThreads,
  use_alt_rep = TRUE)
irp14[, c("Num", "num") := list(.N, 1:.N), 
  by = c(geovars, "taxrefno", "taxyear")]
irp14[UID %in% UID[grepl("^A$", natureofperson)], 
  NatureOfPer := "A"]
irp14[num == 1L & grepl("A", NatureOfPer), 
  .(JobTotal = sum(Num, na.rm = T))]
####    JobTotal
####       <int>
####    12303557
#### irp5a <- qread(paste0(pathdata, "irp5a.qs"), nthreads = DTThreads)
#### irp5a[, c("Num", "num") := list(.N, 1:.N), 
####   by = c(geovars, "taxrefno", "taxyear")]
#### irp5a[UID %in% UID[grepl("^A$", natureofperson)], NatureOfPer := "A"]
#### irp5a[num == 1L & grepl("A", NatureOfPer), 
####   .(JobTotal = sum(Num, na.rm = T)), by = taxyear]
#### #### Looks fine in irp5a!
#### ####    taxyear JobTotal
#### ####       <int>    <int>
#### ####  1:    2008   193888
#### ####  2:    2009   233919
#### ####  3:    2010   251261
#### ####  4:    2011   326655
#### ####  5:    2012   379034
#### ####  6:    2013   420265
#### ####  7:    2014   407802
#### ####  8:    2015   461713
#### ####  9:    2016   494848
#### #### 10:    2017   513754
#### #### 11:    2018   531506
#### #### 12:    2019   559162
#### #### 13:    2020   569491
#### #### 14:    2021   652310
#### #### 15:    2022   566829
#### irp5a[, .(total = sum(busmainplc_geo != "")), by = taxyear]
#### #### Looks horrible once conditioning on location info in irp5a!
#### ####     taxyear  total
#### ####       <int>  <int>
#### ####  1:    2008   4338
#### ####  2:    2009   4357
#### ####  3:    2010 164006
#### ####  4:    2011   6085
#### ####  5:    2012 230056
#### ####  6:    2013 396541
#### ####  7:    2014 405210
#### ####  8:    2015 457141
#### ####  9:    2016 490655
#### #### 10:    2017 512344
#### #### 11:    2018 529679
#### #### 12:    2019 556736
#### #### 13:    2020 565447
#### #### 14:    2021 646384
#### #### 15:    2022 561929
```

```{r correcting information of IRP panel}
#| attr.source = '.r .line-numbers',
#### CLAUDE com: 2026-07-13 (S19 item "316 vs 350 regex inconsistency"):
####   unified to the LOOSE grepl("A", ...) that has always governed the
####   pipeline filter (the "^A$" here fed only the JobTotal display below;
####   NatureOfPer is rebuilt from scratch at the "ZZ" block). Behavior-
####   preserving. The diagnostic print surfaces the actual value set on the
####   next server run -- if values are exactly single letters, loose and
####   exact are identical; if not, revisit BOTH sites together.
print(table0(irp5pnl$natureofperson))
#### irp5pnl[UID %in% UID[grepl("^A$", natureofperson)],
####   NatureOfPer := "A"]
irp5pnl[UID %in% UID[grepl("A", natureofperson)],
  NatureOfPer := "A"]
irp5pnl[, num := 1:.N, by = c(geovars, "taxrefno", "taxyear")]
irp5pnl[num == 1L & grepl("A", NatureOfPer), 
  .(JobTotal = sum(Num, na.rm = T)), by = taxyear]
#### Looks fine in irp5pnl!
####     taxyear JobTotal
####       <int>    <int>
####  1:    2008 12323161
####  2:    2009 12083707
####  3:    2010 12135165
####  4:    2011 13034038
####  5:    2012 13496238
####  6:    2013 13930394
####  7:    2014 14188697
####  8:    2015 15586123
####  9:    2016 15240490
#### 10:    2017 15267221
#### 11:    2018 15566584
#### 12:    2019 15635489
#### 13:    2020 15654027
#### 14:    2021 13989801
#### 15:    2022 13930391
####  addmargins(table(irp5pnl[is.na(uid) | is.na(UID), .(uid=is.na(uid), UID=is.na(UID))]))
####        UID
#### uid         FALSE      TRUE       Sum
####   FALSE         0   4409392   4409392
####   TRUE  199028655         0 199028655
####   Sum   199028655   4409392 203438047
#### #### So sub
irp5pnl[ is.na(uid) & !is.na(UID), uid := UID]
irp5pnl[!is.na(uid) &  is.na(UID), UID := uid]
#### Copy natureofperson across years for same uid
irp5pnl[, NatureOfPer := "ZZ"]
irp5pnl[uid %in% uid[grepl("A", natureofperson)], NatureOfPer := "A"]
#### CLAUDE tpo:
#### qsave(irp5pnl,
####   paste0(pathdata, "irp5_BeforeDroppingRepetitiveUIDs.qs"),
####   nthreads = DTThreads,
####     use_alt_rep = TRUE)
qsave(irp5pnl,
  paste0(pathdata, "irp5_BeforeDroppingRepetitiveUIDs.qs"),
  nthreads = DTThreads)
```

```{r merge industry code}
#| attr.source = '.r .line-numbers',
#### Merge industry code, taking irp5 as base
irp5 <- qread(paste0(pathdata, "irp5_BeforeDroppingRepetitiveUIDs.qs"), 
  nthreads = DTThreads, use_alt_rep = TRUE)
ind <- qread(paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), 
  nthreads = DTThreads, use_alt_rep = TRUE)
setkey(irp5, taxrefno); setkey(ind, taxrefno, taxyear)
ind[, n := 1:.N, by = taxrefno]
indFirstYear <- ind[n == 1, ]
indFirstYear[, n := NULL]
setkey(indFirstYear, taxrefno)
setnames(indFirstYear, "taxyear", "FirstYearOnSIC7")
irp5i <- indFirstYear[irp5, on = "taxrefno"]
setcolorder(irp5i, c("busprov_geo", "busdistmuni_geo", "buslocmuni_geo", 
  "busmainplc_geo", "taxrefno", "Txrf", "uid", "UID", "gender",
  grepout("Date", colnames(irp5i))))  
#### CLAUDE tpo:
#### qsave(irp5i, paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(irp5i, paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = DTThreads)
```

```{r copy ipr5pnl irp5i to saveddata, eval = F}
FilesToCopy <- c(
  "irp5_BeforeDroppingRepetitiveUIDs.qs", 
  "irp5_WithRepetetiveUIDs.qs"
)
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
    )
```


# Copy location information from other source: IRP5_Geo

Geo location file has following variables for each jobs:

* `taxyear`
* `irp5it3aid`: Job id (with location-taxyear-taxrefno)
* `IRP5IT3aPersonalDetailID`: personal id (ID number for SA nationals or passport number for foreign nationals)
* `payereferenceno`: useless
* `Bus_adr_Geo_province`, `Str_adr_Geo_province`, `Bus_adr_Geo_Munic`, `Str_adr_Geo_Munic`
   * Most granular: Local municipality
* `HaveInfo`: ==1L if Bus_adr_Geo_Munic != "" & !is.na(Bus_adr_Geo_Munic)

We can uniquely (one-to-one match) merge the geo location file with IRP5 data on `irp5it3aid`

```{r Open the geolocation file and produce geoloc file, eval = F}
#| attr.source = '.r .line-numbers',
library(readstata13)
geodat <- read.dta13(paste0(path, "IRP5_Geo.dta"))
geodat <- data.table(geodat)
setkey(geodat, IRP5IT3aID)
setnames(geodat, 
  c("IRP5IT3aID", "TaxYear", "PAYERefno"),
  c("irp5it3aid", "taxyear", "payereferenceno")
)
geodat[, taxyear := as.integer(taxyear)]
geoloc <- geodat[taxyear==2015, ]
#### CLAUDE tpo:
#### qsave(geoloc, paste0(pathdata, "GeoLocFile.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(geoloc, paste0(pathdata, "GeoLocFile.qs"), nthreads = DTThreads)
geodat[, HaveInfo := 0L]
geodat[(Bus_adr_Geo_Munic != "" & !is.na(Bus_adr_Geo_Munic)), 
  HaveInfo := 1L]
table(geodat[, .(HaveInfo, taxyear)], useNA = "ifany")
#### CLAUDE tpo:
#### qsave(geodat, paste0(pathdata, "GeoLocationFile.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(geodat, paste0(pathdata, "GeoLocationFile.qs"), nthreads = DTThreads)
```

```{r smallirp5}
#| attr.source = '.r .line-numbers',
#### CLAUDE: exists() for exact names, ls(pattern=) for partial/regex.
####   && / || for scalar short-circuit. Double negation -> positive logic.
if (exists("geodat")) rm(geodat)
if (!exists("RunSep")) RunSep <- T
if (RunSep && exists("geoloc")) rm(geoloc)
gc()
#### Read from disk if running separately OR object not in memory
if (RunSep || length(ls(pattern="^irp5i"))==0)
  irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
#### smallirp5 <- irp5i[grepl("A", NatureOfPer), .(taxrefno, payereferenceno)]
smallirp5 <- unique(irp5i[grepl("A", NatureOfPer),
  c(geovars, "taxrefno", "irp5it3aid"), with = F])
#### CLAUDE dea: 2026-07-13 (S19 item): unique() above already deduplicates;
####   this second pass was a no-op.
#### smallirp5 <- smallirp5[!duplicated(smallirp5), ]
#### CLAUDE tpo:
#### qsave(smallirp5, paste0(pathdata, "smallIRP5.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(smallirp5, paste0(pathdata, "smallIRP5.qs"), nthreads = DTThreads)
```

```{r GeoLF}
#| attr.source = '.r .line-numbers',
#### CLAUDE exi, sca, pos:
if (!exists("RunSep")) RunSep <- T
if (RunSep || !exists("geoloc"))
  geoloc <- qread(paste0(pathdata, "GeoLocFile.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
if (RunSep || length(ls(pattern="smalli"))==0)
  smallirp5 <- qread(paste0(pathdata, "smallIRP5.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
#### setkey(geoloc, payereferenceno)
#### setkey(smallirp5, payereferenceno)
#### GeoLF <- smallirp5[geoloc, on = "payereferenceno"]
setkey(geoloc, irp5it3aid)
setkey(smallirp5, irp5it3aid)
GeoLF <- smallirp5[geoloc, on = "irp5it3aid"]
#### CLAUDE tpo:
#### qsave(GeoLF, paste0(pathdata, "GeoLF.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(GeoLF, paste0(pathdata, "GeoLF.qs"), nthreads = DTThreads)
```

```{r copy GeoXX and derived files to saveddata, eval = F}
FilesToCopy <- c(
  "GeoLocationFile.qs", 
  "GeoLocFile.qs", 
  "smallIRP5.qs", 
  "GeoLF.qs"
)
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
    )
```

Copy from GeoLF if 

* `UID` and `taxrefno` are the keys

```{r merge irp5_WithRepetetiveUIDs GeoLF}
#| attr.source = '.r .line-numbers',
#### irp5_WithRepetetiveUIDs.qs is created in "merge industry code"
#### CLAUDE exi, sca, pos:
if (!RunSep && exists("geoloc")) rm(geoloc)
gc()
if (RunSep || length(ls(pattern="irp5i"))==0)
  irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
if (RunSep || length(ls(pattern="GeoLF"))==0)
  GeoLF <- qread(paste0(pathdata, "GeoLF.qs"), nthreads = DTThreads,
    use_alt_rep = TRUE)
GeoLF[, dup := 1:.N, by = irp5it3aid]
nrow(GeoLF[dup > 1, ])
GeoLF.nodup <- GeoLF[dup == 1L, ]
GeoLF.nodup[, dup := NULL]
#### irp5gi <- GeoLF.nodup[irp5i, on = c("irp5it3aid", "payereferenceno", "taxrefno")]
irp5gi <- GeoLF.nodup[irp5i, on = c(geovars, "irp5it3aid", "taxrefno")]
colnames(GeoLF.nodup)
rbind(GeoLF=dim(GeoLF.nodup), irp5i=dim(irp5i), irp5gi=dim(irp5gi))
colnames(irp5gi)
table(irp5gi[, taxyear], useNA = "ifany")
table(irp5gi[, i.taxyear], useNA = "ifany")
table(GeoLF.nodup[, taxyear], useNA = "ifany")
table(irp5i[, taxyear], useNA = "ifany")
irp5gi[, taxyear := NULL]
if (any(grepl("i.taxyear", colnames(irp5gi)))) 
  setnames(irp5gi, "i.taxyear", "taxyear")
#### qsave(irp5gi, paste0(pathdata, "irp5_MergedGeoLF.qs"), 
####  nthreads = DTThreads)
#### irp5gi <- qread(paste0(pathdata, "irp5_MergedGeoLF.qs"), 
####  nthreads = DTThreads)
#### r copy 2015 location info
#### CLAUDE msk: define anonymous ID columns; Corp/UInd used in all html display; taxrefno/uid retained for joins/filters only.
#### CLAUDE frg: 2026-07-13 (S19 item): separator added -- separator-less
####   paste0 lets different (geo, id) tuples collide ("AB"+"C" == "A"+"BC").
####   Integer factor labels change, but both IDs are regenerated and joined
####   only WITHIN a pipeline run (ObPattern join below uses them from the
####   same lineage), so this requires the usual full re-run from Condense,
####   nothing more. Matches the EstID fix in IRP5MergeData.rmd.
#### irp5gi[, EstabIDTx := as.integer(factor(paste0(
####     busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf)))]
#### irp5gi[, EstabID   := as.integer(factor(paste0(
####     busdistmuni_geo, buslocmuni_geo, busmainplc_geo, taxrefno)))]
irp5gi[, EstabIDTx := as.integer(factor(paste(
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, Txrf, sep = "|")))]
irp5gi[, EstabID   := as.integer(factor(paste(
    busdistmuni_geo, buslocmuni_geo, busmainplc_geo, taxrefno, sep = "|")))]
irp5gi[, FirmIDTx  := as.integer(factor(Txrf))]
irp5gi[, FirmID    := as.integer(factor(taxrefno))]
irp5gi[, UIndID    := as.integer(factor(uid))]
irp5gi[, EUIndID   := paste(EstabIDTx, UIndID, sep = "-")]
irp5gi[, Corp      := FirmID]
irp5gi[, UInd      := UIndID]
#### CLAUDE dea: 2026-07-13 (S19 item): dead initialisation, overwritten on
####   the next line.
#### irp5gi[, FirmUInd := 0L]
irp5gi[, FirmUInd := as.integer(factor(paste(taxrefno, uid, sep = "-")))]
irp5gi[, c("NoLocMuni", "CanFill", "ShouldFill") := 0L]
irp5gi[buslocmuni_geo == "" | is.na(buslocmuni_geo),
  NoLocMuni := 1L]
irp5gi[Bus_adr_Geo_Munic != "" & !is.na(Bus_adr_Geo_Munic), 
  CanFill := 1L]
irp5gi[NoLocMuni == 1L & CanFill == 1L, ShouldFill := 1L]
#### How much is loc muni information populated?
table(irp5gi[NoLocMuni == 0L, .(CanFill, taxyear)], useNA = "ifany")
NAsInLocMuniBefore <- 
  irp5gi[, .(BeforeFill = sum(NoLocMuni)), by = taxyear]
print(NAsInLocMuniBefore)
#### to be safe, create a variable for "fill in"
irp5gi[, FillInLocMuni := buslocmuni_geo]

#### For each FirmUInd that ShouldFill == 1L, 
#### fill in loc muni info from the 1st CanFill row
#### CLAUDE eff: code correct but slow; faster native equivalent exists.
#### 2026-07-02
####   Old code evaluated "Bus_adr_Geo_Munic[CanFill == 1L][1]" once per
####   FirmUInd group - millions of groups over the 100M-row table.
####   New code, three steps:
####     1. need   = persons that have at least one row needing a fill
####     2. lkfill = one municipality per such person, computed on the
####        small subset of rows that actually carry the 2015 geo info
####     3. update join writes it back; fifelse() restricts the write to
####        taxyear <= 2015 rows (others keep their current value), and
####        persons absent from lkfill are never touched
#### irp5gi[taxyear <= 2015 & FirmUInd %in% FirmUInd[ShouldFill == 1L],
####   FillInLocMuni := Bus_adr_Geo_Munic[CanFill == 1L][1],
####   by = FirmUInd]
need <- irp5gi[ShouldFill == 1L, unique(FirmUInd)]
lkfill <- irp5gi[CanFill == 1L & taxyear <= 2015 & FirmUInd %in% need,
  .(v = Bus_adr_Geo_Munic[1]), by = FirmUInd]
irp5gi[lkfill, on = "FirmUInd",
  FillInLocMuni := fifelse(taxyear <= 2015, i.v, FillInLocMuni)]
summary(irp5gi[taxyear <= 2015 & FirmUInd %in% FirmUInd[ShouldFill == 1L], 
  .(taxyear, NoLocMuni, CanFill, Bus_adr_Geo_Munic, 
    naFillInLocMuni=is.na(FillInLocMuni), nabuslocmuni_geo=is.na(buslocmuni_geo))])
#### CLAUDE tpo: original used `=` instead of `:=` in data.table j (invalid syntax); removed entirely, masking now done inline in the select.
#### irp5gi[, (c("taxrefno.", "uid.")) = lapply(.SD,
####   function(x) gsub("....$", "....", gsub("^....", "....", x))),
####   .SDcols = c("taxrefno", "uid")][
####   taxyear <= 2015 & FirmUInd %in% FirmUInd[ShouldFill == 1L][1:10],
####   .(taxrefno., uid., taxyear, NoLocMuni, CanFill, Bus_adr_Geo_Munic,
####     FillInLocMuni, buslocmuni_geo)]
irp5gi[taxyear <= 2015 & FirmUInd %in% FirmUInd[ShouldFill == 1L][1:10],
  .(Corp, UInd, taxyear, NoLocMuni, CanFill, Bus_adr_Geo_Munic,
    FillInLocMuni, buslocmuni_geo)]
#### How much is loc muni information populated after this?
#### Number of NAs in loc muni info
NAsInLocMuni <- irp5gi[, .(
  BeforeFill = sum(NoLocMuni), 
  AfterFill  = sum(FillInLocMuni == "" | is.na(FillInLocMuni))), 
  by = taxyear]
print(NAsInLocMuni)
#### table(irp5gi[, .(
####   LocMuniNotNA = buslocmuni_geo != "" | !is.na(buslocmuni_geo), 
####   taxyear)], 
####   useNA = "ifany")
#### How useful is GeoLocFile in fillling information?
#### table(irp5gi[, .(ShouldFill, taxyear)], useNA = "ifany")
LocMuniCanBeFilled <- irp5gi[, .(
  CanBeFilled  = sum(
    (FillInLocMuni  != "" & !is.na(FillInLocMuni)) &
    (buslocmuni_geo == "" | is.na(buslocmuni_geo)) &
     FillInLocMuni != buslocmuni_geo
    )), 
  by = taxyear]
print(LocMuniCanBeFilled)
setnames(irp5gi, 
    c("buslocmuni_geo", "FillInLocMuni")
  , c("old_buslocmuni_geo", "buslocmuni_geo")) 
#### CLAUDE tpo:
#### qsave(irp5gi, paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
####   nthreads = DTThreads,
####     use_alt_rep = TRUE)
qsave(irp5gi, paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
  nthreads = DTThreads)
```


```{r copy Bus_adr_Geo_Munic, eval = F, echo = F}
#### CLAUDE dea: 2026-07-13 (S19 item): the whole chunk read and rewrote the
####   same file unchanged -- unlike the deliberate ALTREP round trips in
####   IRP5HHI.rmd, nothing uses the re-read object. Body commented out.
#### irp5gi <- qread(paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
####   nthreads = DTThreads, use_alt_rep = TRUE)
#### CLAUDE tpo:
#### qsave(irp5gi, paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
####   nthreads = DTThreads, use_alt_rep = TRUE)
#### qsave(irp5gi, paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
####   nthreads = DTThreads)
```

Copy location info

* use taxrefno-uid match
* this assumes no transfer between establishments in a firm in 2012-2013
* in reality, there are transfers &rarr; measurement errors in location
* but the proportion of transfers must be small 
* benefits of location info outweigh the costs

If there are workers working for multiple establishments ... no problem 

* taxrefno-uid uniquely identifies matches (even if there is a transfer, it does not show up as an error, we will never know)
* base of join = 2012 (fix rows of 2012), and find matches from 2013
* see example shown in codes, 8 entries in 2012, 2013


```{r correcting information of IRP panel irp5gir}
#| attr.source = '.r .line-numbers',
#### case: taxyear = 2015, DateStart = Dec 01, 2013, DateEnd = Jan 10, 2014
#### taxyear 2015: Mar 2014 - Feb 2015
#### This is a revision report of taxyear 2014, so drop this person's report of 2013, 
#### use this row for 2014
#### paste0 operation is not slow
#### library(data.table)
#### library(fasttime)
#### library(anytime)
#### n <- 1e7
#### dt <- data.table(
####   x1 = sample(1000:2000, n, replace= T),
####   x2 = sample(1000:2000, n, replace= T)
#### )
#### dt[, date1 := as.Date(x1, origin = "1990/01/01")]
#### dt[, date2 := as.Date(x2, origin = "1990/01/01")]
#### dt[, taxyear := year(date1)]
#### system.time(dt[, date := paste0(taxyear, "/03/01")])
####   user system elapsed 
####   1.97   0.01    1.97 
#### system.time(dt[, posix := as.POSIXct(date)])
####   user system elapsed 
#### 118.61   1.13  121.34 
#### system.time(dt[, idate := as.IDate(date)])
####  user system elapsed 
#### 36.91   0.22   38.19 
#### system.time(dt[, fdate := fastPOSIXct(date)])
####  user system elapsed
####  0.08   0.01    0.11 
#### system.time(dt[, adate := anytime(date)])
####  user system elapsed 
####  6.89   0.10    7.05 
#### CLAUDE exi, sca, pos:
if (exists("geoloc"))      rm(geoloc)
if (exists("irp5i"))       rm(irp5i)
if (exists("GeoLF"))       rm(GeoLF)
if (exists("GeoLF.nodup")) rm(GeoLF.nodup)
gc()
if (RunSep || length(ls(pattern="^irp5gi$"))==0)
  irp5gi <- qread(paste0(pathdata, "irp5_CopiedLocMuniFromGeoLoc.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
#### irp5i <- qread(paste0(pathdata, "irp5_WithRepetetiveUIDs.qs"), nthreads = DTThreads)
library(fasttime)
irp5gi[, TYStart := fastPOSIXct(paste0(taxyear-1,   "/03/01"))]
irp5gi[, TYEnd   := fastPOSIXct(paste0(taxyear, "/03/01"))]
irp5gi[, TrueTaxYear := taxyear]
#### CLAUDE com: 2026-07-03
####   Anchor = DateEnd (periodemployedto), not DateStart. periodemployedto
####   can never exceed end-Feb of the return's own filing year (annual
####   return, Mar Y-1 - Feb Y coverage) - it is bounded within the correct
####   fiscal window by construction. periodemployedfrom (DateStart) is not:
####   for long-tenured employees "Entries of earlier jobs" (below) allows
####   DateStart several years before the filing year, so applying this
####   formula to DateStart would misclassify the record into an unrelated
####   past year. DateEnd is therefore the only anchor that recovers which
####   annual return a record actually belongs to.
#### Correct: DateEnd (periodemployedto)
####   Wrong: taxyear (for revision entries that are filed in later years)
#### Examples: DateEnd
####             taxyear,  year() month()>=3
####   Jan 2014:  2014  ,   2014      0      ==> 2014+0 = 2014
####   Mar 2014:  2015  ,   2014      1      ==> 2014+1 = 2015
####   Dec 2014:  2015  ,   2014      1      ==> 2014+1 = 2015
####   Feb 2015:  2015  ,   2015      0      ==> 2015+0 = 2015
#### Entries of earlier jobs: DateStart < TYStart & DateEnd < TYEnd
irp5gi[DateStart < TYStart & DateEnd < TYEnd & (taxyear - year(DateStart) <= 5),
####   TrueTaxYear := year(DateStart)] ## wrong: DateStart can be years
####   before the filing year for long-tenured employees (see note above)
  TrueTaxYear := year(DateEnd) + as.integer(month(DateEnd) >= 3)]
apply(table(irp5gi[TrueTaxYear != taxyear, .(TrueTaxYear, taxyear)]), 1, sum)
####    2003    2004    2005    2006    2007    2008    2009    2010    2011    2012 
####     376    1370    5448   16307   56706   65947 1219105 1205674 1541903 1483088 
####    2013    2014    2015    2016    2017    2018    2019    2020    2021 
#### 1482391 1356213 1427595 1561429 1599350 1437092 1496579 1436222 1173744 
#### Total is  18566539
####  TrueTaxYear taxyear
####         2010    2010
####         2011    2011 * drop this
####         2011    2012 * taxyear := TrueTaxYear
####         2012    2012
####         2013    2013
#### TrueTaxYear[TrueTaxYear != taxyear] == 2011 is year that needs to be dropped
#### taxyear == TrueTaxYear[TrueTaxYear != taxyear] is flagged as DropThisForRev
####  TrueTaxYear taxyear
####         2010    2010
####         2011    2011 ** drop this 1
####         2011    2012 *  taxyear := TrueTaxYear
####         2012    2012 ** drop this 2
####         2012    2013 *  taxyear := TrueTaxYear
#### TrueTaxYear[TrueTaxYear != taxyear] == 2011, 2012 
####   taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear 
#### is flagged as DropThisForRev
irp5gi[, DropThisForRev := 0L]
#### CLAUDE tpo: 2026-07-02
####   Two bugs in the old flagging line:
####   1. In DT[i, j := v, by=g], i is evaluated once on the WHOLE table,
####      so "TrueTaxYear[TrueTaxYear != taxyear]" pooled the revision
####      years of ALL taxpayers. by= only groups j (a constant 1L here),
####      so a row could be dropped because a stranger revised that year.
####   2. "& TrueTaxYear != taxyear" selects the revision rows themselves.
####      The worked example above wants the opposite: drop the
####      superseded ORIGINAL row and keep the revision.
####   Net effect of the old code: corrections deleted, outdated
####   originals kept.
####   Fix: list each person's revised years (RevYears), then flag, via
####   update join, only that person's original row of that year
####   (original = row with TrueTaxYear == taxyear).
####   Worked example: DataTableFixExamples.html
#### irp5gi[taxyear %in% TrueTaxYear[TrueTaxYear != taxyear] & TrueTaxYear != taxyear,
####   DropThisForRev := 1L,
####   by = c(geovars, "taxrefno", "UID")]
RevYears <- unique(irp5gi[TrueTaxYear != taxyear,
  c(geovars, "taxrefno", "UID", "TrueTaxYear"), with = F])
setnames(RevYears, "TrueTaxYear", "taxyear")
irp5gi[RevYears, DropThisForRev := as.integer(TrueTaxYear == taxyear),
  on = c(geovars, "taxrefno", "UID", "taxyear")]
round((sum(irp5gi[, DropThisForRev])/nrow(irp5gi))*100, 4)
####  9.0204
#### CLAUDE com: reminder note. 2026-07-02 the 9.0204 figure comes from
####   the old flagging rule; re-run before quoting it.
#### Rename 
####  taxyear ==> Old.taxyear
####  TrueTaxYear ==> taxyear, becomes
####  taxyear Old.taxyear
####     2010        2010
####     2011        2012 *  
####     2012        2013 *  
#### Revision rows dropped
irp5gir <- irp5gi[DropThisForRev == 0L, ]
c(nrow(irp5gi), nrow(irp5gir))
setnames(irp5gir, c("TrueTaxYear", "taxyear"), c("taxyear", "Old.taxyear"))
irp5gir[, DropThisForRev := NULL]

#### Copy location information from 2015 ==> 2008-2014
#### Use taxrefno-uid match ==> saved 10.9M rows
#### Keep original variables
irp5gir[, orig.bprov := busprov_geo]
irp5gir[, orig.bdist := busdistmuni_geo]
irp5gir[, orig.bloc  := buslocmuni_geo]
irp5gir[, orig.bmain := busmainplc_geo]
geo15vars <- c("bprov15", "bdist15", "bloc15", "bmain15")
#### Copy location entries only if unique in an establishment
#### 
#### I. Old way (risks of misscopying)
#### If we do not condition on first row "[1]", we wind up
#### having multiple cases ==> error
#### If there are multiple matches, we have no idea which to pick, so pick first
#### irp5gir[, bprov15 := busprov_geo[taxyear == 2015][1], by = .(taxrefno, uid)]
#### irp5gir[, bdist15 := busdistmuni_geo[taxyear == 2015][1], by = .(taxrefno, uid)]
#### irp5gir[, bloc15  := buslocmuni_geo[taxyear == 2015][1], by = .(taxrefno, uid)]
#### irp5gir[, bmain15 := busmainplc_geo[taxyear == 2015][1], by = .(taxrefno, uid)]
#### CLAUDE tpo: [1] could select ""/NA before filtering; conflicting non-missing
####   values now set to NA (not copied). Loop uses geovars from setup.rmd.
#### 
#### II. Filter on non-NA and non-"", while 
####     AVOIDING elegant expressions and stick to memory efficiency
####  * direct vector subset by creating a working data tmp_2015
####  * no .SD
####  * no grouped evaluation using "by="
####  * no get()
####  * minimal abstraction layers (At 100M-row scale, abstractions 
####    that are elegant or convenient can introduce: temporary objects, 
####    method dispatch, expression evaluation, copying, hidden loops)
#### 1. Initialize output columns to NA_character_ in the main table.
####  This correctly covers groups with no valid data, or >1 unique 
####  value.
irp5gir[, (geo15vars) := NA_character_]
#### 2. Pre-subset the 2015 data ONCE. 
####  Extracting only the necessary columns drastically speeds up 
####  subsetting inside the loop.
ColsToKeep <- c("taxrefno", "uid", geovars)
tmp_2015 <- irp5gir[taxyear == 2015, .SD, .SDcols = ColsToKeep]
for (k in seq_along(geovars)) {
  gv <- geovars[k]
  outv <- geo15vars[k]
  #### a) Filter to non-NA and non-empty records for this variable
  ####    Using tmp_2015[[gv]] prevents 'get()' overhead and 
  ####    evaluates at vector speeds
  valid_idx <- which(!is.na(tmp_2015[[gv]]) & tmp_2015[[gv]] != "")
  #### Below is ugly but copies selected rows of each variables
  #### and does not use get() which is less efficient than data.table
  #### dt_valid <- tmp_2015[valid_idx, .(taxrefno, uid, v = get(gv))]
  #### gemini 3.1 says: 
  ####   By extracting the vectors with $ and [[ and subsetting them 
  ####   directly with [valid_idx], you bypass all of data.table's 
  ####   parsing overhead. You are dropping straight down to 
  ####   R's primitive C functions. 
  dt_valid <- data.table(
    taxrefno = tmp_2015$taxrefno[valid_idx],
    uid      = tmp_2015$uid[valid_idx],
    v        = tmp_2015[[gv]][valid_idx]
  )
  #### b) Get unique combinations of[taxrefno, uid, value]
  ####    (Highly optimized in C via data.table's internal radix sort)
  #### GPT4 says no to unique(dt_valid, by = c("taxrefno", "uid", "v"))
  #### because it may interfere with data.table's internal key/sort-ing
  dt_uniq <- unique(dt_valid)
  #### c) Identify groups that have exactly ONE unique value
  ####    If a person has >1 unique location, their (taxrefno, uid) 
  ####    will be duplicated here. We flag all rows that belong to 
  ####    duplicated groups to drop them.
  is_dup <- duplicated(dt_uniq, by = c("taxrefno", "uid")) | 
            duplicated(dt_uniq, by = c("taxrefno", "uid"), fromLast = TRUE)
  #### Keep only the "singleton" groups
  dt_final <- dt_uniq[!is_dup]
  #### d) Perform a fast update join back to the main 100M+ row dataset
  irp5gir[dt_final, (outv) := i.v, on = .(taxrefno, uid)]
}
#### CLAUDE msk:
#### irp5gir[orig.bloc  == "" & bloc15  != "", .(taxrefno, uid, taxyear, bloc15, buslocmuni_geo)]
irp5gir[orig.bloc  == "" & bloc15  != "",
  .(Corp, UInd, taxyear, bloc15, buslocmuni_geo)]
#### CLAUDE msk:
#### irp5gir[orig.bloc  == "" & bloc15  != "", .(taxrefno, uid, taxyear,
####   bloc15, buslocmuni_geo, bmain15, busmainplc_geo)]
irp5gir[orig.bloc  == "" & bloc15  != "",
  .(Corp, UInd, taxyear, bloc15, buslocmuni_geo, bmain15, busmainplc_geo)]
irp5gir[, c("LocInfoCopiedFrom", "ProvCopied", "DisCopied", "LocCopied", "MaiCopied") := 0L]
#### Copy from bXXXX15 
irp5gir[
  taxyear <= 2014 &
  (busprov_geo == "" | is.na(busprov_geo) | busprov_geo == "EXCEPTION") & 
  (bprov15 != "" & !is.na(bprov15) & bprov15 != "EXCEPTION"), 
   c("busprov_geo", "ProvCopied", "LocInfoCopiedFrom") := .(bprov15, 1L, 2015L)
]
irp5gir[
  taxyear <= 2014 &
  (busdistmuni_geo == "" | is.na(busdistmuni_geo) | busdistmuni_geo == "EXCEPTION") & 
  (bdist15 != "" & !is.na(bdist15) & bdist15 != "EXCEPTION"),
  c("busdistmuni_geo", "DisCopied", "LocInfoCopiedFrom") := .(bdist15, 1L, 2015L)
]
irp5gir[
  taxyear <= 2014 &
  (buslocmuni_geo == "" | is.na(buslocmuni_geo) | buslocmuni_geo == "EXCEPTION") &
  (bloc15 != "" & !is.na(bloc15) & bloc15 != "EXCEPTION"),
  c("buslocmuni_geo",  "LocCopied", "LocInfoCopiedFrom") := .(bloc15,  1L, 2015L)
]
irp5gir[
  taxyear <= 2014 &
  (busmainplc_geo == "" | is.na(busmainplc_geo) | busmainplc_geo == "EXCEPTION") &
  (bmain15 != "" & !is.na(bmain15) & bmain15 != "EXCEPTION"),
  c("busmainplc_geo",  "MaiCopied", "LocInfoCopiedFrom") := .(bmain15, 1L, 2015L)
]
#### CLAUDE tpo:
#### qsave(irp5gir, paste0(pathdata, "irp5_RevReports.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(irp5gir, paste0(pathdata, "irp5_RevReports.qs"), nthreads = DTThreads)
rm(irp5gi)
gc()
```
```{r dropping rows}
#| attr.source = '.r .line-numbers',
#### Note: Firms with anomalous repetetive uids (AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ) are dropped
#### code "dropthese"
#### Read data
#### nthreds = 16 at NT-SDF, = 8 with my laptop (less 1, keep 1 for other computations)
ipyr <- qread(paste0(pathdata, "irp512.qs"), nthreads = DTThreads,
  use_alt_rep = TRUE)
ipyr[, uid := gsub(" +$", "", UID)]
#### CLAUDE msk: define Corp/UInd on ipyr (separate dataset; no Txrf-based IDs needed here).
ipyr[, FirmID := as.integer(factor(taxrefno))]
ipyr[, UIndID := as.integer(factor(uid))]
ipyr[, Corp   := FirmID]
ipyr[, UInd   := UIndID]
#### rm(ipyrs)
#### CLAUDE: ipyrs never defined; only ipyr exists here.
####   Commented out to avoid "object not found" error.
#### uid anomalous entries: AAAAAAAAAA, CCCCCCCCC, ZZZZZZZZ
#### these are 0.035% of total number of rows
#### one large employer (with 4632 jobs) contributes 
#### DO NOT use these uids when copying location from future data
dupuid <- ipyr[
    duplicated(uid) | duplicated(uid, fromLast=T),
    .(taxrefno, uid, Corp, UInd)
  ]
dupuid[, n := .N, by = .(taxrefno, uid)]
dupuid <- unique(dupuid[order(n, decreasing = T), ])
#### CLAUDE msk:
#### dupuid[n > 10, ]
#### print(unique(dupuid[, uid])[1:30], quote = F)
dupuid[n > 10, .(Corp, UInd, n)]
print(unique(dupuid[, UInd])[1:30])
#### repetitive: [A-Z]*8
repetetive <- dupuid[grepl("(.)\\1{7,}", uid), uid]
tbr <- table(repetetive)
tbr <- tbr[order(tbr, decreasing=T)]
repetetive <- unique(repetetive)
#### ipyrs[ii.prov | ii.dist | ii.loc | ii.main, 
####   .(taxrefno, uid, busprov_geo, bprov.13, 
####   busmainplc_geo, bmain.13)][uid %in% repetetive, ][
####   order(uid), ]
#### firms with repetetive uids and their counts
tb <- table(ipyr[uid %in% repetetive, taxrefno])
tb <- tb[order(tb, decreasing=T)]
#### CLAUDE msk: taxrefno should not appear in a table; wrap names() before printing.
#### tb[1:20]
{ t <- tb[1:20]; names(t) <- ipyr[match(names(t), ipyr[["taxrefno"]]), Corp]; t }
#### repetetive entries are concentrated in 3 firms
#### CLAUDE msk:
#### dupuid[1:20, ]
dupuid[1:20, .(Corp, UInd, n)]
#### CLAUDE msk:
#### dupuid[taxrefno %in% names(tb)[1:3], ][1:20, ]
dupuid[taxrefno %in% names(tb)[1:3], ][1:20, .(Corp, UInd, n)]
#### firms with 800 same uids: using same uid for all workers
#### CLAUDE msk:
#### ipyr[uid %in% dupuid[1, uid],
####   .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)]
ipyr[uid %in% dupuid[1, uid],
  .(busprov_geo, busmainplc_geo, Corp, UInd,
    periodemployedfrom, periodemployedto, kerr_income)]
#### same repetetive entries (AAAAA...) are found across firms
#### CLAUDE msk:
#### ipyr[uid %in% names(tbr)[1:20],
####   .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
####   order(uid, taxrefno), ]
ipyr[uid %in% names(tbr)[1:20],
  .(busprov_geo, busmainplc_geo, Corp, UInd,
    periodemployedfrom, periodemployedto, kerr_income)][
  order(UInd, Corp), ]
#### Drop firms with repetetive entries if they are above 10
ii <- 10
#### CLAUDE msk:
#### ipyr[taxrefno %in% names(tb)[tb==ii] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],
####   .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
####   order(uid, taxrefno), ]
ipyr[taxrefno %in% names(tb)[tb==ii] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],
  .(busprov_geo, busmainplc_geo, Corp, UInd,
    periodemployedfrom, periodemployedto, kerr_income)][
  order(UInd, Corp), ]
#### CLAUDE msk: original wrapped the assignment in `(...)` to print as a side-effect;
####   `dropthese` is needed raw at the qsave below, so split into raw assign + masked display.
#### (dropthese <- ipyr[taxrefno %in% names(tb)[tb>10] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],
####   .(busprov_geo, busmainplc_geo, taxrefno, uid, periodemployedfrom, periodemployedto, kerr_income)][
####   order(uid, taxrefno), ])
#### table(dropthese[, .(uid, taxrefno)])
#### table(dropthese[, uid])
dropthese <- ipyr[taxrefno %in% names(tb)[tb>10] & uid %in% dupuid[taxrefno %in% names(tb)[tb==ii], uid][1:5],
  .(busprov_geo, busmainplc_geo, taxrefno, uid, Corp, UInd,
    periodemployedfrom, periodemployedto, kerr_income)][
  order(UInd, Corp), ]
dropthese[, .(busprov_geo, busmainplc_geo, Corp, UInd,
    periodemployedfrom, periodemployedto, kerr_income)]
table(dropthese[, .(UInd, Corp)])
table(dropthese[, UInd])
####      CCCCCCCC CCCCCCCCCCCCC 
####           191          2341 
#### CLAUDE tpo:
#### qsave(dropthese,
####   paste0(pathdata, "DroppedTheseRepetetiveIDs.qs"), nthreads = DTThreads,
####     use_alt_rep = TRUE)
qsave(dropthese,
  paste0(pathdata, "DroppedTheseRepetetiveIDs.qs"), nthreads = DTThreads)

#### Keep only nature of person is "an individual"
#### we did this: irp5[uid %in% uid[grepl("A", natureofperson)], NatureOfPer := "A"]
#### irp5ir <- qread(paste0(pathdata, "irp5_RevReports.qs"), nthreads = DTThreads)
addmargins(table(irp5gir[, .(natureof = grepl("A", natureofperson), 
  NatureOf = grepl("A", NatureOfPer))]))
####         NatureOf
#### natureof      TRUE       Sum
####    FALSE  42731611  42731611  * 42M saved!!!
####    TRUE  132692587 132692587
####    Sum   175424198 175424198
irp5gir <- irp5gir[grepl("A", NatureOfPer), ]
#### repetetive UID
irp5Clean <- irp5gir[!(taxrefno %in% dropthese[, taxrefno]), ]
c(nrow(irp5gir), nrow(irp5Clean))
#### [1] 175424198 175176433
#### if (any(iiEmpty <- grepl("", colnames(irp5Clean)))) 
#### grepl("^Crop|^Fore", imp_mic_sic7_2d), ]
#### CLAUDE tpo:
#### qsave(irp5Clean, paste0(pathdata, "irp5Clean.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(irp5Clean, paste0(pathdata, "irp5Clean.qs"), nthreads = DTThreads)
```

Get observation pattern for individuals, establishments, firms: IObPattern, EObPattern, FObPattern

Also create 

* `EstabIDTx` (busdistmuni_geo)(buslocmuni_geo)(busmainplc_geo)(Txrf)
* `EstabID` (busdistmuni_geo)(buslocmuni_geo)(busmainplc_geo)(taxrefno)
* `FirmIDTx` (Txrf)
* `FirmID` (taxrefno)
* `UIndID` (uid)
* `EUIndID` (EstabIDTx)(UIndID)
* `Corp` (= FirmID; display alias)
* `UInd` (= UIndID; display alias)

```{r create ObPattern}
#| attr.source = '.r .line-numbers',
#### CLAUDE exi, sca, pos:
if (exists("irp5gir")) rm(irp5gir)
gc()
if (RunSep || !exists("irp5Clean"))
  irp5Clean <- qread(paste0(pathdata, "irp5Clean.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
idyr <- irp5Clean[, .(taxyear, Txrf, busprov_geo, busdistmuni_geo,
  buslocmuni_geo, busmainplc_geo, taxrefno, UID, uid,
  EstabIDTx, EstabID, FirmIDTx, FirmID, EUIndID, UIndID, Corp, UInd)]
#### CLAUDE frg: code correct but breaks if the data changes. 2026-07-02
####   If any uid/UID pair is NA, all() returns NA and "if (NA)" stops
####   with an error. isTRUE() turns that NA into FALSE, so the check
####   cannot crash; when NAs exist, UID is simply kept.
#### if (all(idyr[, uid == UID])) idyr[, UID := NULL]
if (isTRUE(all(idyr[, uid == UID]))) idyr[, UID := NULL]
#### Individual observation pattern in a given establishment
#### Reshape wide gives Ob2008, Ob2009, ..., Ob2022
idyr[, Ob := 1L]
#### CLAUDE frg: code correct but breaks when scale grows. 2026-07-02
####   With 9 idvar columns, reshape() builds the row key with
####   interaction(), which indexes every POSSIBLE combination of the
####   9 columns' levels, not just the observed ones. EUIndID alone has
####   millions of levels, the combination count overflows, and R stops
####   with "cannot allocate ... Gb" (reproduced with 1.8M rows in
####   scratch_reshape_test.R). dcast() orders only the OBSERVED
####   combinations, so it cannot overflow. Its two options:
####     fill = 0L           unobserved years become 0 directly, so the
####                         NA -> 0 loop below is no longer needed
####     fun.aggregate = max if the same id-year appears twice it becomes
####                         one 1L (reshape silently kept the first row)
####   Columns dcast does not carry over (Txrf, taxrefno, uid, UIndID,
####   Corp, UInd) are not used downstream of this point.
#### idyrW <- reshape(idyr, direction = "wide",
####   idvar = c(
####     "busprov_geo", "busdistmuni_geo",
####     "buslocmuni_geo", "busmainplc_geo",
####     "FirmIDTx", "FirmID",
####     "EstabIDTx", "EstabID",
####     "EUIndID"),
####   timevar = "taxyear", v.names= c("Ob"))
idyrW <- dcast(idyr,
  busprov_geo + busdistmuni_geo + buslocmuni_geo + busmainplc_geo +
  FirmIDTx + FirmID + EstabIDTx + EstabID + EUIndID ~ taxyear,
  value.var = "Ob", fill = 0L, fun.aggregate = max)
#### dcast names the year columns "2008", "2009", ...; rename to Ob.2008,
#### Ob.2009, ... The grep keeps this robust if the year range changes.
yrcols <- grep("^[0-9]{4}$", colnames(idyrW), value = TRUE)
setnames(idyrW, yrcols, paste0("Ob.", yrcols))
#### ObYYYY = 0L if NA (if obs does not exist in long format data)
#### CLAUDE tpo: 2026-07-02 loop no longer needed: fill = 0L above
####   already writes 0 for unobserved years.
#### for (jj in grep("^Ob", colnames(idyrW)))
####   for (ii in which(is.na(idyrW[, jj, with = F] | idyrW[, jj, with = F] != "")))
####     set(idyrW, i = ii, j = jj, value = 0L)
cols <- grepout("^Ob", colnames(idyrW))
#### idyrW[, (cols) := lapply(.SD, function(x) {x[is.na(x)] <- 0L; x}),
####   .SDcols = cols]
#### CLAUDE tpo: 2026-07-02
####   paste(.SD, collapse=) turns each WHOLE COLUMN into one long string
####   and recycles that single string to every row, so all rows got the
####   same nonsense pattern (and IObPattern had one factor level).
####   do.call() hands the columns to paste0 side by side, which pastes
####   them row by row: the true "0100...1" observation pattern.
####   Worked example: DataTableFixExamples.html
#### idyrW[, IObPattern := paste(.SD, sep = "", collapse = ""),
####   .SDcols = cols]
idyrW[, IObPattern := do.call(paste0, .SD), .SDcols = cols]
idyrW[, IObPattern := factor(IObPattern)]
#### Establishment observation pattern
#### Method: Find any individual with 1L in taxyear-establishment
####         and copy within an establishment
#### EstabObYYYY: Establishment observed in YYYY 
#### FirmObYYYY: Establishment observed in YYYY 
#### Goal: EstabOb.YYYY = 1L if any worker in the establishment
####   has Ob.YYYY == 1L, for each year YYYY.
####   Marks an establishment as "active" in a given year
####   regardless of which specific workers were present.
####   EstabOb.YYYY = NA for rows of uid[is.na(uid)] in YYYY
####   But EstabOb.YYYY = 0L/1L for rows of uid[!is.na(uid)] in YYYY
####   This is fine because there is no generic corporate income tax (CIT)
####   variable in irp5 and the only indication of operation is
####   variables like a3601_income or uid.
####   Used in downstream to define complete-panel establishments.
#### 
#### CLAUDE: Original double loop broken in two ways:
####   1. is.na() always FALSE cols filled with 0L at L806-808.
####   2. Logic inverted targeted NA rows (unobserved workers)
####      to set EstabOb=1L, but intent is to target Ob==1L rows.
####   estabid/firmid no longer needed after replacement.
####   Replaced with data.table by= (correct and faster).
####   Ref: github.com/Rdatatable/data.table/wiki/Benchmarks-:-Grouping
#### 
#### estabid <- unique(idyrW[, EstabID])
#### firmid  <- unique(idyrW[, FirmID])
#### for (jj in grepout("^Ob", colnames(idyrW))) {
####   jjj <- grep(paste0("Estab", jj), colnames(idyrW))
####   for (ii in 1:length(estabid))
####     set(idyrW,
####           i = which(
####                 idyrW[, EstabID] == estabid[ii] &
####                 is.na(idyrW[, jj, with = F])
####               ),
####           j = jjj,
####       value = 1L)
#### }
#### Below is not efficient ("by=" and "get")
#### for (jj in grepout("^Ob", colnames(idyrW))) {
####   jjj <- paste0("Estab", jj)
####   idyrW[, (jjj) := as.integer(any(get(jj) == 1L)), by = EstabID]
#### }
ColsOb <- grepout("^Ob", colnames(idyrW))
estabcols <- paste0("Estab", ColsOb)
#### 1. Initialize the new columns as exact copies
idyrW[, (estabcols) := .SD, .SDcols = ColsOb]
#### 2. Single Loop: Find estab with 1's, update the missing rows
#### CLAUDE eff: 2026-07-02
####   The "%in%" version scanned the whole table twice per year column.
####   A keyed update join finds the rows of the flagged establishments
####   directly. The NA/"" handling is gone because fill = 0L above
####   leaves only 0/1 in the Ob columns, and writing 1L where a 1L
####   already sits changes nothing.
#### for (k in seq_along(ColsOb)) {
####   ov <- ColsOb[k]
####   jjj <- estabcols[k]
####   #### a) Get estab with 1s
####   EstabWith1 <- unique(idyrW[get(ov) %in% c("1", 1, 1L), EstabID])
####   #### b) Filter non-NA and non-empty records for this variable
####   col_data <- idyrW[[ov]]
####   ii.NA0 <- is.na(col_data) | col_data %in% c("", "0", 0)
####   #### c) Update ONLY the missing rows in idyrW directly
####   idyrW[ii.NA0 & (EstabID %in% EstabWith1), (jjj) := 1L]
#### }
for (k in seq_along(ColsOb)) {
  ov <- ColsOb[k]
  jjj <- estabcols[k]
  #### a) establishments with at least one worker observed in this year
  EstabWith1 <- unique(idyrW[get(ov) == 1L, .(EstabID)])
  #### b) update join: every row of those establishments gets 1L
  idyrW[EstabWith1, on = "EstabID", (jjj) := 1L]
}
#### CLAUDE tpo: 2026-07-02 same whole-column paste bug as IObPattern
####   above; do.call() pastes row by row.
#### idyrW[, EObPattern := paste(.SD, sep = "", collapse = ""),
####   .SDcols = estabcols]
idyrW[, EObPattern := do.call(paste0, .SD),
  .SDcols = estabcols]
idyrW[, EObPattern := factor(EObPattern)]

#### Firm observation pattern
#### Find any individual with 1L in a given firm
firmcols <- paste0("Firm", ColsOb)
idyrW[, (firmcols) := .SD, .SDcols = ColsOb]
#### CLAUDE eff: 2026-07-02 same update-join rewrite as the
####   establishment loop above.
#### for (k in seq_along(ColsOb)) {
####   ov <- ColsOb[k]
####   jjj <- firmcols[k]
####   #### a) Get firms with 1s
####   FirmWith1 <- unique(idyrW[get(ov) %in% c("1", 1, 1L), FirmID])
####   #### b) Filter non-NA and non-empty records for this variable
####   col_data <- idyrW[[ov]]
####   ii.NA0 <- is.na(col_data) | col_data %in% c("", "0", 0)
####   #### c) Update ONLY the missing rows in idyrW directly
####   idyrW[ii.NA0 & (FirmID %in% FirmWith1), (jjj) := 1L]
#### }
for (k in seq_along(ColsOb)) {
  ov <- ColsOb[k]
  jjj <- firmcols[k]
  #### a) firms with at least one worker observed in this year
  FirmWith1 <- unique(idyrW[get(ov) == 1L, .(FirmID)])
  #### b) update join: every row of those firms gets 1L
  idyrW[FirmWith1, on = "FirmID", (jjj) := 1L]
}
#### CLAUDE tpo: 2026-07-02 same whole-column paste bug as IObPattern
####   above; do.call() pastes row by row.
#### idyrW[, FObPattern := paste(.SD, sep = "", collapse = ""),
####   .SDcols = firmcols]
idyrW[, FObPattern := do.call(paste0, .SD),
  .SDcols = firmcols]
idyrW[, FObPattern := factor(FObPattern)]
#### for (jj in grepout("^Ob", colnames(idyrW))) {
####   jjj <- paste0("Firm", jj)
####   idyrW[, (jjj) := as.integer(any(get(jj) == 1L)), by = FirmID]
#### }
#### for (yy in 2008:2022) {
####   jjname <- paste0("FirmOb.", yy)
####   idyrW[, (jjname) := get(jjname)[!is.na(get(jjname))][1],
####     by = .(FirmID)]
#### }
#### ObPattern: FirmIDTx, FirmID, EstabIDTx, EstabID, EUIndID, 
####            FObPattern, EObPattern, IObPattern
ObPattern <- idyrW[, c(
    geovars,
    "FirmIDTx", "FirmID", 
    "EstabIDTx", "EstabID", 
    "EUIndID", 
    "FObPattern", "EObPattern", "IObPattern"), with = F]
#### CLAUDE tpo:
#### qsave(ObPattern, paste0(pathdata, "ObPattern.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(ObPattern, paste0(pathdata, "ObPattern.qs"), nthreads = DTThreads)
#### ObPattern <- qread(paste0(pathdata, "ObPattern.qs"), nthreads = DTThreads)
#### CLAUDE eff: 2026-07-02 the two setkey() sorts were never used: the
####   join below picks its own columns via on=. Sorting 175M rows twice
####   cost minutes for nothing.
#### setkey(irp5Clean, EUIndID); setkey(ObPattern, EUIndID)
irp5 <- ObPattern[irp5Clean, on = c(geovars, "EstabIDTx", "EUIndID")]
#### CLAUDE tpo:
#### qsave(irp5, paste0(pathdata, "irp5.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(irp5, paste0(pathdata, "irp5.qs"), nthreads = DTThreads)
#### Agriculture panel
irp5a <- irp5[grepl("^Anim|^Plant pro|crops|Logging|forest", imp_mic_sic7_3d), ]
#### CLAUDE tpo:
#### qsave(irp5a, paste0(pathdata, "irp5a.qs"), nthreads = DTThreads,
####   use_alt_rep = TRUE)
qsave(irp5a, paste0(pathdata, "irp5a.qs"), nthreads = DTThreads)
litedown::timing_data() 
```


```{r MeanStdN of employment}
#| attr.source = '.r .line-numbers',
#### code: MeanStdN of employment
#### For some reasons, it gives "cannot allocate a vector of size 36Gb..." error
library(data.table)
library(qs)
if (length(ls(pattern="irp5C"))==1) rm(irp5Clean) 
if (length(ls(pattern="ObP"))==1) rm(ObPattern)
gc()
if (length(ls(pattern="irp5$"))==0) # do not read if in memory
  irp5 <- qread(paste0(pathdata, "irp5.qs"), nthreads = DTThreads,
    use_alt_rep = TRUE)
#### CLAUDE tpo: 2026-07-02
####   Three problems in the old code:
####   1. After filtering num == 1L, mean(num) is always 1 and var(num)
####      is always 0 - the intended variable was Num (establishment
####      size).
####   2. by = EstabIDTx ignored taxyear, so Num pooled all years of an
####      establishment into one count.
####   3. Adding two columns to the full 175M-row table is what caused
####      the "cannot allocate 36Gb" error noted at the chunk top.
####   Fix: first collapse to one row per establishment-year (a small
####   table), then take the statistics by year from that.
#### irp5[, c("num", "Num") := list(as.integer(1:.N), as.integer(.N)),
####   by = EstabIDTx]
#### MeanStdN <- irp5[num == 1L, .(
####   mean = round(mean(num, na.rm = T), 3),
####   std = round(var(num, na.rm = T)^(.5), 3),
####   N = .N
####   ), by = taxyear]
EstabN <- irp5[, .(Num = .N), by = .(EstabIDTx, taxyear)]
MeanStdN <- EstabN[, .(
  mean = round(mean(Num), 3),
  std  = round(sd(Num), 3),
  N = .N
  ), by = taxyear]
print(MeanStdN)
fwrite(MeanStdN, paste0(pathdata, "MeanStdNOfEmployment.tsv"), sep = "\t")
```

```{r copy ipr5i irp5gi irp5gir irp5Clean to saveddata, eval = F}
#| attr.source = '.r .line-numbers',
FilesToCopy <- c(
####  "irp5_MergedGeoLF.qs", 
  "irp5_CopiedLocMuniFromGeoLoc.qs", 
  "irp5_RevReports.qs", 
  "ObPattern.qs",
  "irp5Clean.qs", 
  "irp5.qs",
  "irp5a.qs",
  "MeanStdNOfEmployment.tsv"
)
for (ff in FilesToCopy)
  file.copy(
           from = paste0(pathdata,      ff)
    ,        to = paste0(pathsaveddata, ff)
    , overwrite = T
    )
```


