---
title: "Minimum wage impacts on employment using IRP5 files"
date: "r format(Sys.time(), "%Y%m%d %R")"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 03 03/IRP5Impacts.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```


# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathsaveddata <- paste0(path, "saveddata/")
pathresults <- paste0(pathprogram, "results/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2); library(rmarkdown); library(fixest)
```





# Estimation

We use the TWFE event sudy specification. 



``{r select sample and estimate, eval = F}
Lf <- qread(paste0(pathsaveddata, "EstSample_Ag.qs"), nthreads = 13)
setkey(Lf, taxrefno, taxyear)
Lf[, ReportEveryYear := 0L]
Lf[, DurationTY := max(taxyear, na.rm = T)-min(taxyear, na.rm = T)+1, 
  by = (
    busprov_geo, busdistmuni_geo, buslocmuni_geo, busmainplc_geo, 
    taxrefno
  )]
Lf[, num := .N, by = .(taxrefno)]
Lf[DurationTY == num, ReportEveryYear := 1L]
#### Complete panel sample: 
####   Existed before 2013 Mar - 2014 Feb: FirstYear2013 = "Before"
####   Jobs0 is defined in chunk "merge faa plus LSMa1 gives Lf"
####   "Jobs in first taxyear" >= 10 <==> "small" or larger
####   "Job duration in current taxyear" > 0
####     DurMonthMarch := round(as.numeric((DateEnd-DateStart20yr)/30.5)
####   Observed duration = latest TY - earliest TY + 1
Lf[taxyear == taxyear[num == 1L], 
  .(
         Micro = sum(Jobs0>0 & Jobs0<10, na.rm = T)
    , NotMicro = sum(Jobs0>=10, na.rm = T)
  ), by = taxyear]
LfC <- Lf[
    grepl("^B", FirstYear2013) & 
                   Jobs0 >= 10 & 
             DurMonthMarch > 0 & 
         ReportEveryYear == 1L , 
  ]
LfC[HHI <= median(HHI[taxyear == 2012], na.rm = T), HHILevel := "Below"]
LfC[HHI > median(HHI[taxyear == 2012], na.rm = T), HHILevel := "Above"]
LfC[, HHILevel := factor(HHILevel)]
LfCE <- LfC[ExistedBefore2013 == 1L, ]
LfCEb <- LfCE[grepl("Be", HHILevel), ])
LfCEa <- LfCE[grepl("Ab", HHILevel), ])
rbind(
    Lf    = table(Lf[, taxyear])
  , LfC   = table(LfC[, taxyear])
  , LfCE  = table(LfCE[, taxyear])
  , LfCEb = table(LfCEb[, taxyear])
  , LfCEa = table(LfCEa[, taxyear])
)
library(fixest)
esLf <- feols(Jobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf)
esLfC <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfC)
esLfCE <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCE)
#### Subsample: High and low concentrated markets
esLfCEa <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCEa)
esLfCEb <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCEb)

library(ggfixest)
eslist <-  list(
    "All sample" = esLf, 
    "Complete panel" = esLfC, 
    "Existed firms, all sample" = esLfCE, 
    "Existed firms, high concentrated markets" = esLfCEa, 
    "Existed firms, low concentrated markets" = esLfCEb 
  )
gFE <- ggiplot(eslist,
  ref.line = 2012, 
  main = "Job impacts: Low and high concentration",
  xlab = "Tax year (March - Februray)",
  multi_style = "facet",
  geom_style = "ribbon",
  theme = theme(
    axis.text.x = element_text(angle = 90, vjust = .5),
    legend.position = "none"
    )
 
) + 
scale_x_continuous(limits = c(2008, 2022), breaks = 2008:2022) +
scale_y_continuous(limits = c(15, -10)) 
ggsave(
  paste0(pathresults, "Results2026Mar03.jpg")
  , gFE, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
gFE
lapply(eslist, summary)
```

```{r estimation with various Job0 and start year}
for (Jb in c(2, 10)) {
  LfC <- Lf[
      grepl("^B", FirstYear2013) & 
                     Jobs0 >= Jb & 
               DurMonthMarch > 0 & 
           ReportEveryYear == 1L , 
    ]
  LfC[HHI <= median(HHI[taxyear == 2012], na.rm = T), HHILevel := "Below"]
  LfC[HHI > median(HHI[taxyear == 2012], na.rm = T), HHILevel := "Above"]
  LfC[, HHILevel := factor(HHILevel)]
  LfCE <- LfC[ExistedBefore2013 == 1L, ]
  LfCEb <- LfCE[grepl("Be", HHILevel), ])
  LfCEa <- LfCE[grepl("Ab", HHILevel), ])
  print(rbind(
      Lf    = table(Lf[, taxyear])
    , LfC   = table(LfC[, taxyear])
    , LfCE  = table(LfCE[, taxyear])
    , LfCEb = table(LfCEb[, taxyear])
    , LfCEa = table(LfCEa[, taxyear])
  ))
  for (Yr in c(2010, 2012)) {
    LfYr    <- Lf[taxyear >= YR, ]
    LfCYr   <- LfC[taxyear >= YR, ]
    LfCEYr  <- LfCE[taxyear >= YR, ]
    LfCEaYr <- LfCEa[taxyear >= UR, ]
    LfCEbYr <- LfCEb[taxyear >= YR, ]
    library(fixest)
    esLfYr <- feols(Jobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
      vcov = ~ taxyear + busmainplc_geo,
      data = LfYr)
    esLfCYr <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
      vcov = ~ taxyear + busmainplc_geo,
      data = LfCYr)
    esLfCEYr <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
      vcov = ~ taxyear + busmainplc_geo,
      data = LfCEYr)
    #### Subsample: High and low concentrated markets
    esLfCEaYr <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
      vcov = ~ taxyear + busmainplc_geo,
      data = LfCEaYr)
    esLfCEbYr <- feols(rJobs ~ i(taxyear, FA0, ref = 2013)|taxyear + taxrefno + busmainplc_geo, 
      vcov = ~ taxyear + busmainplc_geo,
      data = LfCEbYr)
    library(ggfixest)
    eslistYr <-  list(
        "All sample" = esLfYr, 
        "Complete panel" = esLfCYr, 
        "Existed firms, all sample" = esLfCEYr, 
        "Existed firms, high concentrated markets" = esLfCEaYr, 
        "Existed firms, low concentrated markets" = esLfCEbYr 
      )
    gFEYr <- ggiplot(eslistYr,
      ref.line = 2012, 
      main = "Job impacts: Low and high concentration",
      xlab = "Tax year (March - Februray)",
      multi_style = "facet",
      geom_style = "ribbon",
      theme = theme(
        axis.text.x = element_text(angle = 90, vjust = .5),
        legend.position = "none"
        )
     
    ) + 
    scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022) +
    scale_y_continuous(limits = c(15, -10)) 
    if (Jb == 2) Jb <- paste0("0", Jb)
    ggsave(
      paste0(pathresults, "ResultsJob", Jb, Yr, "-2022.jpg")
      , gFE, width = 8*2, height = 4*2, units = "cm",
      dpi = 300
     )
    gFEYr
    lapply(eslistYr, summary)
    assign(paste0("eslist", Jb, Yr), eslistYr)
    assign(paste0("gFE", Jb, Yr), gFEYr)
    qsave(eslistYr, paste0(pathresults, "eslist", Jb, Yr, ".qs"))
  }
}
```

