---
title: "Minimum wage impacts on employment using IRP5 files"
date: "`r format(Sys.time(), '%Y%m%d %R')`"
output: 
  litedown::html_format:
    meta:
      css: ["MyCSS.css", "@default", "@tabsets", "@copy-button"]
      js: ["@sidenotes", "@appendix", "@tabsets", "@copy-button"]
#### output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(pathprogram, "IRP5Impacts.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
/* CLAUDE tpo: 2026-07-12 was a '####' line — not CSS; the parser folded it
   into the selector and silently dropped this whole pre rule (same fix as
   IRP5HHI.rmd S20-4) */
/* Default height of a block */
pre {
  max-height: 700px;
  overflow-y: auto;
}
```


# Setup

```{r read setup from child file, child='setup.Rmd', message = F, warning = F}
#### setup.rmd sets eval, echo, time to T
```
```{r read functions from child file, child='destatfactor.Rmd', message = F, warning = F}
```

```{r setup, eval = F, echo = F}
library(qs); library(data.table); library(ggplot2)
library(rmarkdown); library(fixest)
library(knitr); library(tufte)
library(litedown)
litedown::reactor(
  echo = TRUE,
  eval = TRUE, 
  time = TRUE
)
options(htmltools.dir.version = FALSE, width = 100)
ThisTheme <- ggplot2::theme(
   axis.text.x = element_text(angle = 90, vjust = .5, hjust = .5), 
   axis.text.y = element_text(size = 12), 
   axis.title = element_text(size = 10), 
   strip.text.x = element_text(color = "blue", size = 8, 
     margin = margin(0, 1.25, 0, 1.25, "cm")), 
   strip.text.y = element_text(color = "blue", size = 8, 
     margin = margin(1.5, 0, 1.5, 0, "cm")),
   panel.spacing.x = unit(c(.1), units = "cm"),
   panel.spacing.y = unit(.1, units = "cm")
)
ThisThemeEnd <- ThisTheme + ggplot2::theme(legend.position="bottom")
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
read.tablev <- function(fn, quotemark = "", headertrue = TRUE, 
  blskip = TRUE, mat = F, ...) 
{
  x <- read.table(fn, sep = "\t", header = headertrue, blank.lines.skip = blskip, 
        quote = quotemark, na.strings = "", ...)
  if (is.null(dim(x)) | nrow(x) == 1) 
  {
    cn <- names(x)
    dim(x) <- c(1, length(x))
    colnames(x) <- cn
  }
  if (mat) x <- as.matrix(x)
  return(x)
}
write.tablev <- function(x, fn, colnamestrue = TRUE, rownamestrue = FALSE, 
nastrings = "", ...) 
  write.table(x, fn, sep = "\t", quote = FALSE, 
        row.names = rownamestrue, col.names = colnamestrue, na = nastrings, ...)
GeoLevel <- c("Prv", "Dis", "Loc", "Mai")
geovars <- c("busprov_geo", "busdistmuni_geo", 
  "buslocmuni_geo", "busmainplc_geo")
```



# Estimation

We use the TWFE event sudy specification. 


## Straight estimation

```{r select estimation sample}
#| attr.source = '.r .line-numbers',
Lf <- qread(paste0(pathsaveddata, "EstSample_Ag.qs"), 
  nthreads = DTThreads, use_alt_rep = TRUE)
#### Complete panel
#### definition of complete: observed everyear in start year - 2020
Lf[, c("Num", "DiffTY", "MaxDiffTY") := 
  .(1:.N, c(0L, diff(taxyear)), max(taxyear)-min(taxyear)+1), 
  by = EstID]
Lf[, GapInTY := 0L]
Lf[EstID %in% EstID[DiffTY > 1 & taxyear <= 2020], GapInTY := 1L]
print(Lf[Num == 1L, .(num = .N), by = .(GapInTY, taxyear)][
  order(GapInTY, taxyear), ])
Lf[, num := 1:.N, by = .(EstID, taxyear)]
print(Lf[num == 1L, ][, 
  .(
         Micro = sum(JobsMP0>0 & JobsMP0<10, na.rm = T)
    , NotMicro = sum(JobsMP0>=10, na.rm = T)
  ), by = taxyear])
#### CLAUDE fix: 2026-07-13 (Session 9 item "threshold uses base-year HHI
####   constant"): median(HHI0[taxyear == 2012]) was the median of base-year
####   HHI constants evaluated on 2012 cross-section rows (mostly 2013 HHI),
####   not of actual 2012 market HHI. True 2012 threshold from the HHI panel.
LSMa <- qread(paste0(pathsaveddata, "HHIAgriRowsMainPlaceLevel.qs"),
  nthreads = DTThreads, use_alt_rep = TRUE)
thr2012 <- median(LSMa[taxyear == 2012, HHI], na.rm = TRUE)
rm(LSMa); gc()
#### HHI grouping (first year)
Lf[, HHI0 := HHI[!is.na(HHI)][1], by = EstID]
#### CLAUDE dea: 2026-07-13 (Session 9 item): HHILevel0 was never used in a
####   feols call, and inside the regression sample (ExistedBefore2013==1L)
####   it equals HHILevel identically. Removed; HHI0 kept for descriptives.
#### Lf[HHI0 <= median(HHI0[taxyear == 2012], na.rm = T),
####   HHILevel0 := "Below"]
#### Lf[HHI0 > median(HHI0[taxyear == 2012], na.rm = T),
####   HHILevel0 := "Above"]
#### Lf[, HHILevel0 := factor(HHILevel0)]
#### HHI grouping (pre-2013)
Lf[, Pre2013HHI := HHI[!is.na(HHI) & taxyear < 2013][1], by = EstID]
Lf[Pre2013HHI <= thr2012, HHILevel := "Below"]
Lf[Pre2013HHI >  thr2012, HHILevel := "Above"]
Lf[, HHILevel := factor(HHILevel)]
#### FAclass
#### FAMP0 is first year FAMP of each establishment
#### not necessarily pre 2013. Defined in IRP5MergeData.rmd
#### faa[, c("FA0", "Jobs0", "FAMP0", "JobsMP0") := 
####   list(
####       FA[!is.na(FA) & FA != 0][1]             # FA0
####     , Jobs[!is.na(Jobs) & Jobs != 0][1]       # Jobs0
####     , FAMP[!is.na(FAMP) & FAMP != 0][1]      # FAMP0
####     , JobsMP[!is.na(JobsMP) & JobsMP != 0][1] # JobsMP0
####   ), by = 
####   .(
####     busprov_geo, busdistmuni_geo, 
####     buslocmuni_geo, busmainplc_geo, taxrefno
####   )]
#### FAclass0: Initial year FA class (not necessarily pre-2013)
Lf[, FAclass0 := "low"]
Lf[.2 < FAMP0 & FAMP0 <= .5, FAclass0 := "mid"]
Lf[FAMP0 > .5, FAclass0 := "high"]
Lf[, FAclass0 := factor(FAclass0, levels = c("low", "mid", "high"))]
#### FAclass: Initial year FA class (pre-2013)
Lf[, Pre2013FAMP0 := FAMP[taxyear < 2013 & !is.na(FAMP)][1], by = EstID]
Lf[, FAclass := "low"]
Lf[.2 < Pre2013FAMP0 & Pre2013FAMP0 <= .5, FAclass := "mid"]
Lf[Pre2013FAMP0 > .5, FAclass := "high"]
Lf[, FAclass := factor(FAclass, levels = c("low", "mid", "high"))]
#### Pre2013Size: Initial year Size (pre-2013)
Lf[, Pre2013Size := Size[taxyear < 2013 & !is.na(Size)][1], by = EstID]
Lf[, Pre2013Size := factor(Pre2013Size, 
  levels = c("micro", "small", "medium", "large"))]
DESS <- vector(length = 2, mode = "list")
names(DESS) <- c("02", "10")
DESS <- list("2010" = DESS, "2012" = DESS)
Jbs <- c(2, 10)
Yrs <- c(2010, 2012)
for (jj in 1:length(Jbs)) {
####for (jj in 1) {
  if (Jbs[jj] < 10) 
    Jb <- paste0(0, Jbs[jj]) else 
    Jb <- Jbs[jj]
  LfC <- Lf[
####  grepl("^B", FirstYear2013) &  #### first year < 2013
              JobsMP0 >= Jbs[jj] &  #### baseline jobs > Jbs[jj]
               DurMonthMarch > 0 &  #### job duration > 0 month
                   GapInTY == 0L ,  #### no gap in taxyear for all observed periods
      c(geovars, "EstID", 
        grepout("tax|Exist|Job|Dur|Cr|An|Fo|Are|Work|Sha|HHI|nc|Emp|FA|Siz|Sub", 
        colnames(Lf))), with = F
    ]
  LfCE <- LfC[ExistedBefore2013 == 1L, ]
  LfCEb <- LfCE[grepl("Be", HHILevel), ]
  LfCEa <- LfCE[grepl("Ab", HHILevel), ]
  print(rbind(
      Lf    = table(Lf[, taxyear])
    , LfC   = table(LfC[, taxyear])
    , LfCE  = table(LfCE[, taxyear])
    , LfCEb = table(LfCEb[, taxyear])
    , LfCEa = table(LfCEa[, taxyear])
  ))
  for (yy in 1:length(Yrs)) {
####  for (yy in 1) {
    Yr <- Yrs[yy]
    LfYr    <- Lf[taxyear >= Yr, ]
    LfCYr   <- LfC[taxyear >= Yr, ]
    LfCEYr  <- LfCE[taxyear >= Yr, ]
    LfCEbYr <- LfCEb[taxyear >= Yr, ]
    LfCEaYr <- LfCEa[taxyear >= Yr, ]
    assign(paste0("Lf",    Yr, Jb), LfYr)
    assign(paste0("LfC",   Yr, Jb), LfCYr)
    qsave(LfYr,    paste0(pathdata, "Lf",    Yr, Jb, ".qs"))
    qsave(LfCYr,   paste0(pathdata, "LfC",   Yr, Jb, ".qs"))
    qsave(LfCEYr,  paste0(pathdata, "LfCE",  Yr, Jb, ".qs"))
    qsave(LfCEbYr, paste0(pathdata, "LfCEb", Yr, Jb, ".qs"))
    qsave(LfCEaYr, paste0(pathdata, "LfCEa", Yr, Jb, ".qs"))
    rm(LfCEaYr, LfCEbYr, LfCEYr)
    #### destat for each data
    #### To be specified using destat and destatfactor
    iifac <- sapply(LfCYr, is.factor)
    iich  <- sapply(LfCYr, is.character)
    des1  <- destat(LfCYr[, which(!(iifac|iich)), with = F])
    des2  <- destatfactor(LfCYr, IgnoreIfAbove = 8)
####des2  <- NULL
    des1 <- data.table(rn=rownames(des1), des1)
    des2 <- data.table(rn=rownames(des2), des2)
    dess  <- rbind(des1, des2)
    DESS[[yy]][[jj]] <- dess
    #### LaTeX tables
    ltb1 <- latextab.simple(as.matrix(des1), 
      hcenter = c(8, rep(.8, 5), rep(.6, 2), .5, .3, .5), 
      unit = "cm")
    ltb2 <- latextab.simple(as.matrix(des2), 
      hcenter = c(8, rep(.8, 5), rep(.6, 2), .5, .3, .5), 
      unit = "cm")
    ltb <- matrix(c(
        ltb1[1:2]
      , paste0("\\multicolumn{", ncol(des1)+1, 
        "}{l}{\\textit{\\footnotesize A. Continuous variables}}\\\\") 
      , ltb1[-c(1:2, length(ltb1))]
      , paste0("\\multicolumn{", ncol(des1)+1, 
        "}{l}{\\textit{\\footnotesize B. Categorical variables}}\\\\") 
      , ltb2[-(1:2)]
    ))
    write.tablev(ltb1, paste0(pathresults, "Destat", Yr, Jb, "_1.tex"))
    write.tablev(ltb2, paste0(pathresults, "Destat", Yr, Jb, "_2.tex"))
    write.tablev(ltb , paste0(pathresults, "Destat", Yr, Jb,   ".tex"))
    #### HTML table
    library(kableExtra)
    ktb <- data.table(dess)
    kt <- kable(ktb, align = 
     paste0("r", paste0(rep("c", ncol(ktb)-1), collapse = "")), 
     caption = paste0("Descriptive statistics ", Yr, "-2022, Job0 cutoff", Jb))
    kt <- column_spec(kt, 1, width = "9cm; min-width:9cm;")
    kt <- column_spec(kt, 2:6, width = "2.0cm; min-width:2.0cm;")
    kt <- kableExtra::footnote(kt, general = "Sample of agricultural establishments.")
    qsave(kt, paste0(pathresults, "Destat", Yr, Jb, ".qs"))
    assign(paste0("Destat", Yr, Jb), kt)
  }
}
qsave(DESS, paste0(pathresults, "DESS.qs"))
Destat201002
```

```{r LfCEb and LfCEa contents, echo = F}
LfCEb[order(EstID, taxyear), 
  .(rk=EstID, taxyear, y=JobsMP, ry=rJobsMP, dy=dJobsMP, FAMP )][1:30, ]
LfCEa[order(EstID, taxyear), 
  .(rk=EstID, taxyear, y=JobsMP, ry=rJobsMP, dy=dJobsMP, FAMP )][1:30, ]
Destat201002
```

```{r show estimation data, echo = F}
LfCYr <- qread(paste0(pathdata, "LfC201002.qs"),
  use_alt_rep = TRUE)
LfCYr[EstID %in% EstID[dJobsMP < -1000], ][order(EstID, taxyear),
  .(busdistmuni_geo, buslocmuni_geo, rk=EstID, 
  taxyear, y=JobsMP, dy=dJobsMP, Share, FAMP)][1:50, ]
```
```{r examining estimation data}
lfdata <- unique(LfCYr[, .(busdistmuni_geo, buslocmuni_geo, 
  busmainplc_geo, taxrefno, EstID, taxyear, 
  Crop, Animal, Forest,
  rJobsMP, dJobsMP, JobsMP, JobsMP0, 
  HHI, Share, FAMP, FAMP0, FAclass)])
rbind(Lf = dim(Lf), LfCYr = dim(LfCYr), lfdata = dim(lfdata))
destat(lfdata)
destatfactor(lfdata, IgnoreIfAbove = 8)
desrJobsByYear <- lapply(2010:2022, function(yr) 
  data.table(t(c(taxyear = yr,  
    destat(lfdata[taxyear == yr, rJobsMP], prob = "odd"))))
)
desrJobsByYear <- rbindlist(desrJobsByYear)
#### destat of HHI by year
desHHIByYear <- lapply(2010:2022, function(yr) 
  data.table(t(c(taxyear = yr,  
    destat(lfdata[taxyear == yr, HHI], prob = "odd"))))
)
desHHIByYear <- rbindlist(desHHIByYear)
#### destat of FAclass by year
desFAclassByYear <- lapply(2010:2022, function(yr) 
  data.table(taxyear = yr,  
    destatfactor(lfdata[taxyear == yr, FAclass], prob = "odd"))
)
desFAclassByYear <- rbindlist(desFAclassByYear)
desFAclassByYear <- desFAclassByYear[c(
      seq(1, nrow(desFAclassByYear), 3)
    , seq(2, nrow(desFAclassByYear), 3)
    , seq(3, nrow(desFAclassByYear), 3)
  ), ]
DesDTBy <- rbind(desrJobsByYear, desHHIByYear, desFAclassByYear, 
  use.names = F)
setnames(DesDTBy, colnames(desFAclassByYear))
hcentre.long <- c(3, rep(.8, 7), rep(.6, 2), .5, .3, .5)
destblby <- latextab.simple(as.matrix(DesDTBy), 
   hcenter = hcentre.long, unit = "cm")
DesTblBy <- c(
    destblby[1:2]
  , paste0("\\multicolumn{", ncol(DesDTBy), 
        "}{l}{\\textit{\\footnotesize Rate of job change}}\\\\")
  , destblby[3:15]
  , paste0("\\multicolumn{", ncol(DesDTBy), 
    "}{l}{\\textit{\\footnotesize HHI}}\\\\") 
  , destblby[13+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTBy), 
    "}{l}{\\textit{\\footnotesize FAclass: low}}\\\\") 
  , destblby[13*2+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTBy), 
    "}{l}{\\textit{\\footnotesize FAclass: mid}}\\\\") 
  , destblby[13*3+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTBy), 
    "}{l}{\\textit{\\footnotesize FAclass: high}}\\\\") 
  , destblby[(13*4+3):length(destblby)]
)
write.tablev(DesTblBy, paste0(pathresults, "rJobsHHIFAclassByYear.tex"))
qsave(DesDTBy, paste0(pathresults, "rJobsHHIFAclassByYear.qs"))
#### plot rJobsMP after dropping outliers
#### CLAUDE tpo: intersect() kept any establishment with at least one
####   moderate row, so its extreme rows were still plotted; use the
####   exclusion idiom of the winsorize block below
#### lfdata.dropped <- lfdata[
####   EstID %in% intersect(EstID[abs(rJobsMP) < 200], EstID[dJobsMP > -1000]), ]
lfdata.dropped <- lfdata[
  !(EstID %in% EstID[abs(rJobsMP) >= 200 | dJobsMP <= -1000]), ]
rbind(
  lfdata = dim(lfdata),
  dropped = dim(lfdata.dropped)
)
ggplot(data=lfdata.dropped, 
  aes(x = taxyear, y = rJobsMP, group = EstID, colour = FAclass)) +
geom_line() +
scale_y_continuous(transform = "log10") + 
guides(
  colour = guide_legend(title = "FA classes", nrow = 1)
  ) +
theme(
  legend.text=element_text(size=9),
  legend.position = "bottom"
)
#### mean plot by FAclass
cihalf <- function(n, sigma) qt(.975, n-1)*(sigma/n)^(.5)
lfdata[, c("MeanJobsMP",
        "MeanrJobsMP",
        "Sigma2JobsMP",
        "Sigma2rJobsMP",
        "nJobsMP",
        "nrJobsMP") 
       :=
      .(mean(JobsMP,  na.rm = T), 
        mean(rJobsMP, na.rm = T), 
        var(JobsMP,   na.rm = T), 
        var(rJobsMP,  na.rm = T), 
        sum(!is.na(JobsMP)), 
        sum(!is.na(rJobsMP)))
      , by = .(FAclass, taxyear)
]
lfdata[, c("MeanJobsMP.ci",
        "MeanrJobsMP.ci")
      :=
      .(cihalf(n=nJobsMP , sigma=Sigma2JobsMP), 
        cihalf(n=nrJobsMP, sigma=Sigma2rJobsMP))
]
lfdata[, c("MeanJobsMP.lb"
        , "MeanJobsMP.ub"
        , "MeanrJobsMP.lb"
        , "MeanrJobsMP.ub")
      :=
      .(MeanJobsMP  - MeanJobsMP.ci,
        MeanJobsMP  + MeanJobsMP.ci,
        MeanrJobsMP - MeanrJobsMP.ci,
        MeanrJobsMP + MeanrJobsMP.ci)
]
gMean <- ggplot(data=
  unique(lfdata[, 
    .(taxyear, FAclass, MeanrJobsMP, MeanrJobsMP.lb, MeanrJobsMP.ub)]), 
  aes(x = taxyear, y = MeanrJobsMP, group = FAclass, shape = FAclass,
    fill = FAclass, colour = FAclass)) +
  #### CLAUDE com: 2026-07-13 width is not a geom_pointrange parameter
  ####   (silently swallowed by ... with a runtime warning); removed.
  geom_pointrange(aes(ymin = MeanrJobsMP.lb, ymax = MeanrJobsMP.ub),
    position = position_dodge(width = .3),
    stat = "identity") +
  scale_y_continuous() + 
  scale_x_continuous(breaks = 2010:2022) + 
  labs(title = "Mean job change rates by FA class") + 
  guides(
    colour = guide_legend(title = "FA classes", nrow = 1)
    ) +
  theme(
    legend.text=element_text(size=9),
    legend.position = "bottom"
  )
ggsave(
  paste0(pathresults, "rJobsByFAClass.jpg")
  , gMean, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
)
#### Sub-industry (imp_mic_sic7_3d) wise mean and CIs
####  Crop: ^Plant pro|crops
####  Animal: ^Anim
####  Forest: Logging|forest
lfdata.crop   <- lfdata[EstID %in% EstID[Crop == 1L], ]
lfdata.animal <- lfdata[EstID %in% EstID[Animal == 1L], ]
lfdata.forest <- lfdata[EstID %in% EstID[Forest == 1L], ]
lflist <- list(
  lfdata = lfdata,
  crop   = lfdata.crop,
  animal = lfdata.animal,
  forest = lfdata.forest
)
lapply(lflist, dim)
jbdat <- lapply(lflist, function(x) 
  x[, .(taxyear, JobsMP, rJobsMP, FAclass)])
jcols <- c("JobsMP", "rJobsMP")
meanjbdat <- lapply(jbdat, function(x) 
  x[, c("MeanJobsMP",
        "MeanrJobsMP",
        "Sigma2JobsMP",
        "Sigma2rJobsMP",
        "nJobsMP",
        "nrJobsMP") 
       :=
      .(mean(JobsMP,  na.rm = T), 
        mean(rJobsMP, na.rm = T), 
        var(JobsMP,   na.rm = T), 
        var(rJobsMP,  na.rm = T), 
        sum(!is.na(JobsMP)), 
        sum(!is.na(rJobsMP)))
      , by = .(FAclass, taxyear)])
meanjbdat <- lapply(meanjbdat, function(x) 
  x[, c("MeanJobsMP.ci",
        "MeanrJobsMP.ci")
      :=
      .(cihalf(n=nJobsMP , sigma=Sigma2JobsMP), 
        cihalf(n=nrJobsMP, sigma=Sigma2rJobsMP))
   ])
meanjbdat <- lapply(meanjbdat, function(x) 
  x[, c( "MeanJobsMP.lb"
        , "MeanJobsMP.ub"
        , "MeanrJobsMP.lb"
        , "MeanrJobsMP.ub")
      :=
      .(MeanJobsMP  - MeanJobsMP.ci,
        MeanJobsMP  + MeanJobsMP.ci,
        MeanrJobsMP - MeanrJobsMP.ci,
        MeanrJobsMP + MeanrJobsMP.ci)
   ])
meanjbdat[[1]][, ind := "all"]
meanjbdat[[2]][, ind := "crop"]
meanjbdat[[3]][, ind := "animal"]
meanjbdat[[4]][, ind := "forest"]
mjb <- rbindlist(meanjbdat)
gmjb <- ggplot(data=mjb, 
    aes(x = taxyear, y = MeanrJobsMP, group = FAclass, shape = FAclass,
      fill = FAclass, colour = FAclass)) +
  #### CLAUDE com: 2026-07-13 width is not a geom_pointrange parameter
  ####   (silently swallowed by ... with a runtime warning); removed.
  geom_pointrange(aes(ymin = MeanrJobsMP.lb, ymax = MeanrJobsMP.ub),
    position = position_dodge(width = .3),
    stat = "identity") +
  scale_y_continuous() + 
  scale_x_continuous(breaks = 2010:2022) + 
  labs(title = "Mean job change rates by FA class and by industry") + 
  facet_grid(ind ~ ., scales = "free_y") + 
  guides(
    colour = guide_legend(title = "FA classes", nrow = 1)
    ) +
  theme(
    legend.text=element_text(size=9),
    legend.position = "bottom"
  )
ggsave(
  paste0(pathresults, "rJobsBySubindustry.jpg")
  , gmjb, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
)
```

::: {.fullwidth}
```{r rjobs graphs by FAclass and by Subindustry}
DesDTBy
DesTblBy
gMean
gmjb
```
:::


```{r estimation with various Job0 and start year}
for (Jb in c(2, 10)) {
#### for (Jb in 2) {
  if (Jb < 10) Jb <- paste0(0, Jb)
  for (Yr in c(2010, 2012)) {
#### for (Yr in 2010) {
    LfYr    <- qread(paste0(pathdata, "Lf",    Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCYr   <- qread(paste0(pathdata, "LfC",   Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEYr  <- qread(paste0(pathdata, "LfCE",  Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEbYr <- qread(paste0(pathdata, "LfCEb", Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEaYr <- qread(paste0(pathdata, "LfCEa", Yr, Jb, ".qs"),
      use_alt_rep = TRUE)


    library(fixest)
    #### CLAUDE spl: 2026-07-13 (Session 19 D1): 8 near-identical feols()
    ####   calls folded into one loop over a named list of data subsets --
    ####   same formula/vcov/data, list order unchanged (downstream ggiplot/
    ####   etable index by position). Removes the copy-paste channel that
    ####   produced the S19-6 label swaps. Originals in git history.
    subsetsYr <- list(
        "All sample"                               = LfYr
      , "Complete panel"                           = LfCYr
      , "Existed firms, all sample"                = LfCEYr
      , "Existed firms, high concentrated markets" = LfCEaYr
      , "Existed firms, low concentrated markets"  = LfCEbYr
      , "Existed firms, low FA"                    = LfCEYr[grepl("low", FAclass), ]
      , "Existed firms, mid FA"                    = LfCEYr[grepl("mid", FAclass), ]
      , "Existed firms, high FA"                   = LfCEYr[grepl("hi", FAclass), ]
    )
    eslistYr <- vector("list", length(subsetsYr))
    names(eslistYr) <- names(subsetsYr)
    for (k in seq_along(subsetsYr))
      eslistYr[[k]] <- feols(
        rJobsMP ~ i(taxyear, Pre2013FAMP0, ref = 2012) |
          taxyear + taxrefno + busmainplc_geo,
        vcov = ~ taxyear + busmainplc_geo,
        data = subsetsYr[[k]])
    rm(subsetsYr)
    library(ggfixest)
    dictvec <- c(
        "JobsMP"         = "Number of jobs at the establishment"
      , "FAMP0"          = "Fraction affected in the taxyear 2012"
      , "taxrefno"       = "Firm ID"
      , "busmainplc_geo" = "Main Place"
    )
    setFixest_dict(dict = dictvec, reset = FALSE)
    estdf <- etable(eslistYr, 
      style.df = style.df(
        depvar.title = "Number of jobs", 
        fixef.title = "Fixed effects", 
        fixef.suffix = " FEs", 
        yesNo = "yes")
    )
    estltx <- etable(eslistYr, 
      style.tex = style.tex("aer"), 
      fitstat = ~ r2 + n, 
      tex = TRUE
    )
    gFEYr <- ggiplot(eslistYr,
      ref.line = 2012, 
      main = "Job impacts: Low and high concentration",
      xlab = "Tax year (March - Februray)",
      multi_style = "facet",
      geom_style = "ribbon",
      theme = theme(legend.position = "none")
    ) + 
    ThisTheme + 
    scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
    gFEYrLH <- ggiplot(eslistYr[2:5],
      ref.line = 2012, 
      main = "Job impacts: Low and high concentration",
      xlab = "Tax year (March - Februray)",
      multi_style = "facet",
      geom_style = "ribbon",
      theme = theme(
        legend.position = "none"
        )
    ) + 
    ThisTheme + 
    scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
    gFEYrFAclass <- ggiplot(eslistYr[c(3, 6:8)],
      ref.line = 2012, 
      main = "Job impacts: FA class in pre-2013",
      xlab = "Tax year (March - Februray)",
      multi_style = "facet",
      geom_style = "ribbon",
      theme = theme(
        legend.position = "none"
        )
    ) + 
    ThisTheme + 
    scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
    #### CLAUDE tpo: dead code — Jb already "02" (character) from paste0(0,Jb) at L500.
    #### if (Jb == 2) Jb <- paste0("0", Jb)
    ggsave(
      paste0(pathresults, "Results", Yr, "-2022Job", Jb, ".jpg")
      , gFEYr, width = 8*2, height = 4*2, units = "cm",
      dpi = 300
     )
    ggsave(
      paste0(pathresults, "Results", Yr, "-2022Job", Jb, "_LowHigh.jpg")
      , gFEYrLH, width = 8*2, height = 4*2, units = "cm",
      dpi = 300
     )
    ggsave(
      paste0(pathresults, "Results", Yr, "-2022Job", Jb, "_FAclass.jpg")
      , gFEYrFAclass, width = 8*2, height = 4*2, units = "cm",
      dpi = 300
     )
    lapply(eslistYr, summary)
    #### drop data from estimated objects
    #### eslistShort <- lapply(eslistYr
    #### scores, residuals, fitted.values, 
    assign(paste0("eslist", Yr, Jb),            eslistYr)
    assign(paste0("gFE",    Yr, Jb),            gFEYr)
    assign(paste0("gFE",    Yr, Jb, "LH"),      gFEYrLH)
    assign(paste0("gFE",    Yr, Jb, "FAclass"), gFEYrFAclass)
    qsave(estdf, 
      paste0(pathresults, "est", Yr, Jb, ".qs"))
    write.tablev(estltx, 
      paste0(pathresults, "est", Yr, Jb, ".tex"))
    qsave(eslistYr, paste0(pathresults, "eslist", Yr, Jb, ".qs"))
  }
}
```

::: {.fullwidth}
```{r 201002 results}
gFE201002
gFE201010LH
gFE201010FAclass
```
:::

## Winsorizing

Some entries show wild swings in the number of jobs between contiguous years at the order of a few thousands. This may be due to revision reporting that have not materialised. Therefore, we winsorize establishments that match at least one condition that they annually increase the jobs: 

* For more than 600% (mean of raw rate of change) with a condition that baseline job size is more than 500
   * Number of observations does not change when baseline job size is reduced (so the scope is expanded) to more than 200
* By more than 2000 <!-- CLAUDE com: 2026-07-13 text said 3000; code uses abs(dJobsMP) > 2000 (S19 note) -->



This results in the sample size reduction from 70888 to 70171 in the 13 year period.


```{r select winsorized estimation sample}
Lf <- qread(paste0(pathsaveddata, "EstSample_Ag.qs"), 
  nthreads = DTThreads, use_alt_rep = TRUE)
#### Complete panel
#### definition of complete: observed everyear in start year - 2020
Lf[, c("Num", "DiffTY", "MaxDiffTY") := 
  .(1:.N, c(0L, diff(taxyear)), max(taxyear)-min(taxyear)+1), 
  by = EstID]
Lf[, GapInTY := 0L]
Lf[EstID %in% EstID[DiffTY > 1 & taxyear <= 2020], GapInTY := 1L]
print(Lf[Num == 1L, .(num = .N), by = .(GapInTY, taxyear)][
  order(GapInTY, taxyear), ])
Lf[, num := 1:.N, by = .(EstID, taxyear)]
#### FAclass0: Initial year FA class (not necessarily pre-2013)
Lf[, FAclass0 := "low"]
Lf[.2 < FAMP0 & FAMP0 <= .5, FAclass0 := "mid"]
#### CLAUDE tpo: duplicate line removed.
Lf[FAMP0 > .5, FAclass0 := "high"]
Lf[, FAclass0 := factor(FAclass0, levels = c("low", "mid", "high"))]
#### FAclass: Initial year FA class (pre-2013)
Lf[, Pre2013FAMP0 := FAMP[taxyear < 2013 & !is.na(FAMP)][1], by = EstID]
Lf[, FAclass := "low"]
Lf[.2 < Pre2013FAMP0 & Pre2013FAMP0 <= .5, FAclass := "mid"]
Lf[Pre2013FAMP0 > .5, FAclass := "high"]
Lf[, FAclass := factor(FAclass, levels = c("low", "mid", "high"))]
#### Pre2013Size: Initial year Size (pre-2013)
Lf[, Pre2013Size := Size[taxyear < 2013 & !is.na(Size)][1], by = EstID]
Lf[, Pre2013Size := factor(Pre2013Size, 
  levels = c("micro", "small", "medium", "large"))]
#### Winsorization
#### Establishments with abs(dJobsMP) > 2000 or 
#### abs(rJobsMP) > 6 & JobsMP0 > 500 are dropped
#### CLAUDE com: 2026-07-02 comment said "> 2"; code uses abs(rJobsMP) > 6 (= 600%).
Lfw <- Lf[
 num == 1L & 
 !(EstID %in% EstID[
      abs(dJobsMP) > 2000 |
     (abs(rJobsMP) > 6 & JobsMP0 > 500)])  # keep only small changes
     #### setting JobsMP0 > 200, JobsMP0 > 400 
     #### gives the same selection as JobsMP0 > 500
 , ]
rbind(Lf=dim(Lf), Lfw=dim(Lfw))
Lfw[, 
  .(
         Micro = sum(JobsMP0>0 & JobsMP0<10, na.rm = T)
    , NotMicro = sum(JobsMP0>=10, na.rm = T)
  ), by = taxyear]
#### CLAUDE fix: 2026-07-13 same true-2012 threshold as the unwinsorized
####   chunk; thr2012 is computed there (this chunk always runs after it).
if (!exists("thr2012")) {
  LSMa <- qread(paste0(pathsaveddata, "HHIAgriRowsMainPlaceLevel.qs"),
    nthreads = DTThreads, use_alt_rep = TRUE)
  thr2012 <- median(LSMa[taxyear == 2012, HHI], na.rm = TRUE)
  rm(LSMa); gc()
}
#### HHI grouping (first year)
Lfw[, HHI0 := HHI[!is.na(HHI)][1], by = EstID]
#### CLAUDE dea: 2026-07-13 HHILevel0 removed, as in the unwinsorized chunk.
#### Lfw[HHI0 <= median(HHI0[taxyear == 2012], na.rm = T),
####   HHILevel0 := "Below"]
#### Lfw[HHI0 > median(HHI0[taxyear == 2012], na.rm = T),
####   HHILevel0 := "Above"]
#### Lfw[, HHILevel0 := factor(HHILevel0)]
#### HHI grouping (pre-2013)
Lfw[, Pre2013HHI := HHI[!is.na(HHI) & taxyear < 2013][1], by = EstID]
Lfw[Pre2013HHI <= thr2012, HHILevel := "Below"]
Lfw[Pre2013HHI >  thr2012, HHILevel := "Above"]
Lfw[, HHILevel := factor(HHILevel)]
DESSw <- vector(length = 2, mode = "list")
names(DESSw) <- c("02", "10")
DESSw <- list("2010" = DESSw, "2012" = DESSw)
Jbs <- c(2, 10)
Yrs <- c(2010, 2012)
for (jj in 1:length(Jbs)) {
####for (jj in 1) {
  if (Jbs[jj] < 10) 
    Jb <- paste0(0, Jbs[jj]) else 
    Jb <- Jbs[jj]
  LfwC <- Lfw[
####  grepl("^B", FirstYear2013) &  #### first year < 2013
              JobsMP0 >= Jbs[jj] &  #### baseline jobs > Jbs[jj]
               DurMonthMarch > 0 &  #### job duration > 0 month
                   GapInTY == 0L ,  #### no gap in taxyear for all observed periods
      c(geovars, "EstID", 
        grepout("tax|Exist|Job|Dur|Cr|An|Fo|Are|Work|Sha|HHI|nc|Emp|FA|Siz|Sub", 
        colnames(Lfw))), with = F
    ]
  LfwCE <- LfwC[ExistedBefore2013 == 1L, ]
  LfwCEb <- LfwCE[grepl("Be", HHILevel), ]
  LfwCEa <- LfwCE[grepl("Ab", HHILevel), ]
  LfwCEmi <- LfwCE[grepl("mic", Pre2013Size), ]
  LfwCEsm <- LfwCE[grepl("sma", Pre2013Size), ]
  LfwCEme <- LfwCE[grepl("med", Pre2013Size), ]
  LfwCEla <- LfwCE[grepl("lar", Pre2013Size), ]
  print(rbind(
      Lfw    = table(Lfw[, taxyear])
    , LfwC   = table(LfwC[, taxyear])
    , LfwCE  = table(LfwCE[, taxyear])
    , LfwCEb = table(LfwCEb[, taxyear])
    , LfwCEa = table(LfwCEa[, taxyear])
    , LfwCEmi = table(LfwCEmi[, taxyear])
    , LfwCEsm = table(LfwCEsm[, taxyear])
    , LfwCEme = table(LfwCEme[, taxyear])
    , LfwCEla = table(LfwCEla[, taxyear])
  ))
  for (yy in 1:length(Yrs)) {
    Yr <- Yrs[yy]
    LfYr     <- Lfw[taxyear >= Yr, ]
    LfCYr    <- LfwC[taxyear >= Yr, ]
    LfCEYr   <- LfwCE[taxyear >= Yr, ]
    LfCEbYr  <- LfwCEb[taxyear >= Yr, ]
    LfCEaYr  <- LfwCEa[taxyear >= Yr, ]
    LfCEmiYr <- LfwCEmi[taxyear >= Yr, ]
    LfCEsmYr <- LfwCEsm[taxyear >= Yr, ]
    LfCEmeYr <- LfwCEme[taxyear >= Yr, ]
    LfCElaYr <- LfwCEla[taxyear >= Yr, ]
    assign(paste0("Lfw",    Yr, Jb), LfYr)
    qsave(LfYr,     paste0(pathdata, "Lfw",     Yr, Jb, ".qs"))
    qsave(LfCYr,    paste0(pathdata, "LfwC",    Yr, Jb, ".qs"))
    qsave(LfCEYr,   paste0(pathdata, "LfwCE",   Yr, Jb, ".qs"))
    qsave(LfCEbYr,  paste0(pathdata, "LfwCEb",  Yr, Jb, ".qs"))
    qsave(LfCEaYr,  paste0(pathdata, "LfwCEa",  Yr, Jb, ".qs"))
    qsave(LfCEmiYr, paste0(pathdata, "LfwCEmi", Yr, Jb, ".qs"))
    qsave(LfCEsmYr, paste0(pathdata, "LfwCEsm", Yr, Jb, ".qs"))
    qsave(LfCEmeYr, paste0(pathdata, "LfwCEme", Yr, Jb, ".qs"))
    qsave(LfCElaYr, paste0(pathdata, "LfwCEla", Yr, Jb, ".qs"))
    rm(LfCEaYr, LfCEbYr, LfCEYr)
    #### destat for each data
    #### To be specified using destat and destatfactor
    iifac <- sapply(LfCYr, is.factor)
    iich  <- sapply(LfCYr, is.character)
    des1  <- destat(LfCYr[, which(!(iifac|iich)), with = F])
    des2  <- destatfactor(LfCYr, IgnoreIfAbove = 8)
####des2  <- NULL
    des1 <- data.table(rn=rownames(des1), des1)
    des2 <- data.table(rn=rownames(des2), des2)
    dess  <- rbind(des1, des2)
    #### CLAUDE tpo: 2026-07-02 dess landed in DESS, the UNwinsorized
    ####   object, so DESSw.qs was saved as a list of NULLs and DESS was
    ####   silently overwritten in memory. This chunk must fill DESSw.
    #### DESS[[yy]][[jj]] <- dess
    DESSw[[yy]][[jj]] <- dess
    #### LaTeX tables
    ltb1 <- latextab.simple(as.matrix(des1), 
      hcenter = c(8, rep(.8, 5), rep(.6, 2), .5, .3, .5), 
      unit = "cm")
    ltb2 <- latextab.simple(as.matrix(des2), 
      hcenter = c(8, rep(.8, 5), rep(.6, 2), .5, .3, .5), 
      unit = "cm")
    ltb <- matrix(c(
        ltb1[1:2]
      , paste0("\\multicolumn{", ncol(des1)+1, 
        "}{l}{\\textit{\\footnotesize A. Continuous variables}}\\\\") 
      , ltb1[-c(1:2, length(ltb1))]
      , paste0("\\multicolumn{", ncol(des1)+1, 
        "}{l}{\\textit{\\footnotesize B. Categorical variables}}\\\\") 
      , ltb2[-(1:2)]
    ))
    write.tablev(ltb1, paste0(pathresults, "Destat", Yr, Jb, "w_1.tex"))
    write.tablev(ltb2, paste0(pathresults, "Destat", Yr, Jb, "w_2.tex"))
    write.tablev(ltb , paste0(pathresults, "Destat", Yr, Jb,   "w.tex"))
    #### HTML table
    library(kableExtra)
    ktb <- data.table(dess)
    kt <- kable(ktb, align = 
     paste0("r", paste0(rep("c", ncol(ktb)-1), collapse = "")), 
     caption = paste0("Descriptive statistics ", Yr, "-2022, Job0 cutoff", Jb))
    kt <- column_spec(kt, 1, width = "9cm; min-width:9cm;")
    kt <- column_spec(kt, 2:6, width = "2.0cm; min-width:2.0cm;")
    kt <- kableExtra::footnote(kt, general = "Sample of agricultural establishments.")
    qsave(kt, paste0(pathresults, "Destatw", Yr, Jb, ".qs"))
    assign(paste0("Destatw", Yr, Jb), kt)
  }
}
qsave(DESSw, paste0(pathresults, "DESSw.qs"))
Destatw201002
```

```{r examining winsorized estimation data}
#### CLAUDE com: 2026-07-13 LfwC in memory at this point is the LAST loop
####   iteration (Jb = 10 cutoff), while the unwinsorized twin chunk
####   {show estimation data} explicitly re-reads the Jb = 02 sample
####   (LfC201002.qs). Mirror that here so the two descriptive/plot
####   sections compare the same cutoff.
LfwC <- qread(paste0(pathdata, "LfwC201002.qs"), use_alt_rep = TRUE)
lfwdata <- unique(LfwC[taxyear >= 2010,
 .(busdistmuni_geo, buslocmuni_geo, 
   busmainplc_geo, taxrefno, EstID, taxyear, 
   Crop, Animal, Forest,
   rJobsMP, dJobsMP, JobsMP, JobsMP0, 
   HHI, Pre2013HHI, Pre2013FAMP0, Share, FAMP, FAMP0, FAclass)])
rbind(Lf = dim(Lfw), LfCYr = dim(LfwC), lfdata = dim(lfwdata))
destat(lfwdata)
destatfactor(lfwdata, IgnoreIfAbove = 8)
#### desbytime <- lapply(1:10, function(yr) 
####   data.table(t(c("time" = yr,  
####     destat(x[time == yr, x1], prob = "odd"))))
#### )
#### desbyime <- rbindlist(desbytime)
#### destat of rJobsMP by year
desrJobsByYear <- lapply(2010:2022, function(yr) 
  data.table(t(c(taxyear = yr,  
    destat(lfwdata[taxyear == yr, rJobsMP], prob = "odd"))))
)
desrJobsByYear <- rbindlist(desrJobsByYear)
#### destat of HHI by year
desHHIByYear <- lapply(2010:2022, function(yr) 
  data.table(t(c(taxyear = yr,  
    destat(lfwdata[taxyear == yr, HHI], prob = "odd"))))
)
desHHIByYear <- rbindlist(desHHIByYear)
#### destat of FAclass by year
desFAclassByYear <- lapply(2010:2022, function(yr) 
  data.table(taxyear = yr,  
    destatfactor(lfwdata[taxyear == yr, FAclass], prob = "odd"))
)
desFAclassByYear <- rbindlist(desFAclassByYear)
desFAclassByYear <- desFAclassByYear[c(
      seq(1, nrow(desFAclassByYear), 3)
    , seq(2, nrow(desFAclassByYear), 3)
    , seq(3, nrow(desFAclassByYear), 3)
  ), ]
DesDTwBy <- rbind(desrJobsByYear, desHHIByYear, desFAclassByYear, 
  use.names = F)
setnames(DesDTwBy, colnames(desFAclassByYear))
DesDTwBy
hcentre.long <- c(3, rep(.8, 7), rep(.6, 2), .5, .3, .5)
destblwby <- latextab.simple(as.matrix(DesDTwBy), 
   hcenter = hcentre.long, unit = "cm")
DesTblwBy <- c(
    destblwby[1:2]
  , paste0("\\multicolumn{", ncol(DesDTwBy), 
        "}{l}{\\textit{\\footnotesize Rate of job change}}\\\\")
  , destblwby[3:15]
  , paste0("\\multicolumn{", ncol(DesDTwBy), 
    "}{l}{\\textit{\\footnotesize HHI}}\\\\") 
  , destblwby[13+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTwBy), 
    "}{l}{\\textit{\\footnotesize FAclass: low}}\\\\") 
  , destblwby[13*2+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTwBy), 
    "}{l}{\\textit{\\footnotesize FAclass: mid}}\\\\") 
  , destblwby[13*3+(3:15)]
  , paste0("\\multicolumn{", ncol(DesDTwBy), 
    "}{l}{\\textit{\\footnotesize FAclass: high}}\\\\") 
  , destblwby[(13*4+3):length(destblwby)]
)
write.tablev(DesTblwBy, paste0(pathresults, "rJobsHHIFAclassByYear_w.tex"))
qsave(DesDTwBy, paste0(pathresults, "rJobsHHIFAclassByYear_w.qs"))
#### mean plot by FAclass
cihalf <- function(n, sigma) qt(.975, n-1)*(sigma/n)^(.5)
lfwdata[, c("MeanJobsMP",
        "MeanrJobsMP",
        "Sigma2JobsMP",
        "Sigma2rJobsMP",
        "nJobsMP",
        "nrJobsMP") 
       :=
      .(mean(JobsMP,  na.rm = T), 
        mean(rJobsMP, na.rm = T), 
        var(JobsMP,   na.rm = T), 
        var(rJobsMP,  na.rm = T), 
        sum(!is.na(JobsMP)), 
        sum(!is.na(rJobsMP)))
      , by = .(FAclass, taxyear)
]
lfwdata[, c("MeanJobsMP.ci",
        "MeanrJobsMP.ci")
      :=
      .(cihalf(n=nJobsMP , sigma=Sigma2JobsMP), 
        cihalf(n=nrJobsMP, sigma=Sigma2rJobsMP))
]
lfwdata[, c("MeanJobsMP.lb"
        , "MeanJobsMP.ub"
        , "MeanrJobsMP.lb"
        , "MeanrJobsMP.ub")
      :=
      .(MeanJobsMP  - MeanJobsMP.ci,
        MeanJobsMP  + MeanJobsMP.ci,
        MeanrJobsMP - MeanrJobsMP.ci,
        MeanrJobsMP + MeanrJobsMP.ci)
]
gMeanw <- ggplot(data=
  unique(lfwdata[, 
    .(taxyear, FAclass, MeanrJobsMP, MeanrJobsMP.lb, MeanrJobsMP.ub)]), 
  aes(x = taxyear, y = MeanrJobsMP, group = FAclass, shape = FAclass,
    fill = FAclass, colour = FAclass)) +
  #### CLAUDE com: 2026-07-13 width is not a geom_pointrange parameter
  ####   (silently swallowed by ... with a runtime warning); removed.
  geom_pointrange(aes(ymin = MeanrJobsMP.lb, ymax = MeanrJobsMP.ub),
    position = position_dodge(width = .3),
    stat = "identity") +
  scale_y_continuous() + 
  scale_x_continuous(breaks = 2010:2022) + 
  labs(title = "Mean job change rates by FA class") + 
  guides(
    colour = guide_legend(title = "FA classes", nrow = 1)
    ) +
  theme(
    legend.text=element_text(size=9),
    legend.position = "bottom"
  )
ggsave(
  paste0(pathresults, "rJobsByFAClass_w.jpg")
  , gMeanw, width = 8*2, height = 8*2, units = "cm",
  dpi = 300
)
#### Sub-industry (imp_mic_sic7_3d) wise mean and CIs
####  Crop: ^Plant pro|crops
####  Animal: ^Anim
####  Forest: Logging|forest
lfwdata.crop   <- lfwdata[EstID %in% EstID[Crop == 1L], ]
lfwdata.animal <- lfwdata[EstID %in% EstID[Animal == 1L], ]
lfwdata.forest <- lfwdata[EstID %in% EstID[Forest == 1L], ]
lfwlist <- list(
  lfwdata = lfwdata,
  crop   = lfwdata.crop,
  animal = lfwdata.animal,
  forest = lfwdata.forest
)
lapply(lfwlist, dim)
jbwdat <- lapply(lfwlist, function(x) 
  x[, .(taxyear, JobsMP, rJobsMP, FAclass)])
jcols <- c("JobsMP", "rJobsMP")
meanjbwdat <- lapply(jbwdat, function(x) 
  x[, c("MeanJobsMP",
        "MeanrJobsMP",
        "Sigma2JobsMP",
        "Sigma2rJobsMP",
        "nJobsMP",
        "nrJobsMP") 
       :=
      .(mean(JobsMP,  na.rm = T), 
        mean(rJobsMP, na.rm = T), 
        var(JobsMP,   na.rm = T), 
        var(rJobsMP,  na.rm = T), 
        sum(!is.na(JobsMP)), 
        sum(!is.na(rJobsMP)))
      , by = .(FAclass, taxyear)])
meanjbwdat <- lapply(meanjbwdat, function(x) 
  x[, c("MeanJobsMP.ci",
        "MeanrJobsMP.ci")
      :=
      .(cihalf(n=nJobsMP , sigma=Sigma2JobsMP), 
        cihalf(n=nrJobsMP, sigma=Sigma2rJobsMP))
   ])
meanjbwdat <- lapply(meanjbwdat, function(x) 
  x[, c( "MeanJobsMP.lb"
        , "MeanJobsMP.ub"
        , "MeanrJobsMP.lb"
        , "MeanrJobsMP.ub")
      :=
      .(MeanJobsMP  - MeanJobsMP.ci,
        MeanJobsMP  + MeanJobsMP.ci,
        MeanrJobsMP - MeanrJobsMP.ci,
        MeanrJobsMP + MeanrJobsMP.ci)
   ])
meanjbwdat[[1]][, ind := "all"]
meanjbwdat[[2]][, ind := "crop"]
meanjbwdat[[3]][, ind := "animal"]
meanjbwdat[[4]][, ind := "forest"]
mjbw <- rbindlist(meanjbwdat)
gmjbw <- ggplot(data=mjbw, 
    aes(x = taxyear, y = MeanrJobsMP, group = FAclass, shape = FAclass,
      fill = FAclass, colour = FAclass)) +
  #### CLAUDE com: 2026-07-13 width is not a geom_pointrange parameter
  ####   (silently swallowed by ... with a runtime warning); removed.
  geom_pointrange(aes(ymin = MeanrJobsMP.lb, ymax = MeanrJobsMP.ub),
    position = position_dodge(width = .3),
    stat = "identity") +
  scale_y_continuous() + 
  scale_x_continuous(breaks = 2010:2022) + 
  labs(title = "Mean job change rates by FA class and by industry") + 
  facet_grid(ind ~ ., scales = "free_y") + 
  guides(
    colour = guide_legend(title = "FA classes", nrow = 1)
    ) +
  theme(
    legend.text=element_text(size=9),
    legend.position = "bottom"
  )
ggsave(
  paste0(pathresults, "rJobsBySubindustry_w.jpg")
  , gmjbw, width = 8*2, height = 8*2, units = "cm",
  dpi = 300
)
#### HHI at main place level
gMeanh <- ggplot(data=
  unique(lfwdata[, .(taxyear, Pre2013HHI, num = 1:.N), by = EstID][
    num == 1L & !is.na(Pre2013HHI), .(EstID, Pre2013HHI)]), 
  aes(x = Pre2013HHI)) +
  geom_histogram(bins = 50) + 
  scale_x_continuous() + 
  labs(title = "HHI in pre-2013") + 
  theme(
    legend.text=element_text(size=9),
    legend.position = "none"
  )
ggsave(
  paste0(pathresults, "Pre2013HHI_w.jpg")
  , gMeanh, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
)
```

```{r show job trends in winsorized data}
gmjbw
gMeanh
```

```{r estimation with winsorized data}
for (Jb in c(2, 10)) {
  if (Jb < 10) Jb <- paste0(0, Jb)
  for (Yr in c(2010, 2012)) {
    LfYr     <- qread(paste0(pathdata, "Lfw",     Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCYr    <- qread(paste0(pathdata, "LfwC",    Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEYr   <- qread(paste0(pathdata, "LfwCE",   Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEbYr  <- qread(paste0(pathdata, "LfwCEb",  Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    LfCEaYr  <- qread(paste0(pathdata, "LfwCEa",  Yr, Jb, ".qs"),
      use_alt_rep = TRUE)
    #### CLAUDE dea: 2026-07-13 these four were read but never used -- the
    ####   size-subsample estimations below subset LfCEYr with grepl() on
    ####   Pre2013Size (equivalent rows) instead of the saved LfwCE{mi,sm,
    ####   me,la} files. Reads dropped.
    #### LfCEmiYr <- qread(paste0(pathdata, "LfwCEmi", Yr, Jb, ".qs"),
    ####   use_alt_rep = TRUE)
    #### LfCEsmYr <- qread(paste0(pathdata, "LfwCEsm", Yr, Jb, ".qs"),
    ####   use_alt_rep = TRUE)
    #### LfCEmeYr <- qread(paste0(pathdata, "LfwCEme", Yr, Jb, ".qs"),
    ####   use_alt_rep = TRUE)
    #### LfCElaYr <- qread(paste0(pathdata, "LfwCEla", Yr, Jb, ".qs"),
    ####   use_alt_rep = TRUE)
    library(fixest)
    #### CLAUDE spl: 2026-07-13 (Session 19 D1): 20 near-identical feols()
    ####   calls folded into one loop over a named list of data subsets --
    ####   same formula/vcov/data, list order unchanged (EstSpecs below
    ####   indexes by position: All 1:8, HHI 3:5, FAclass c(3,6:8),
    ####   Size c(3,9:12), SizeHHI c(3,13:20)). CEa = above-median HHI
    ####   (high concentration), CEb = below-median. Originals in git
    ####   history.
    subsetsYrw <- list(
        "All sample"                               = LfYr                              #  1
      , "Complete panel"                           = LfCYr                             #  2
      , "Existed firms, all sample"                = LfCEYr                            #  3
      , "Existed firms, high concentrated markets" = LfCEaYr                           #  4
      , "Existed firms, low concentrated markets"  = LfCEbYr                           #  5
      , "Existed firms, low FA"                    = LfCEYr[grepl("low", FAclass), ]   #  6
      , "Existed firms, mid FA"                    = LfCEYr[grepl("mid", FAclass), ]   #  7
      , "Existed firms, high FA"                   = LfCEYr[grepl("hi", FAclass), ]    #  8
      , "Existed firms, micro size"                = LfCEYr[grepl("mi", Pre2013Size), ]  #  9
      , "Existed firms, small size"                = LfCEYr[grepl("sm", Pre2013Size), ]  # 10
      , "Existed firms, medium size"               = LfCEYr[grepl("me", Pre2013Size), ]  # 11
      , "Existed firms, large size"                = LfCEYr[grepl("la", Pre2013Size), ]  # 12
      , "Existed firms, high HHI and micro size"   = LfCEaYr[grepl("mi", Pre2013Size), ] # 13
      , "Existed firms, high HHI and small size"   = LfCEaYr[grepl("sm", Pre2013Size), ] # 14
      , "Existed firms, high HHI and medium size"  = LfCEaYr[grepl("me", Pre2013Size), ] # 15
      , "Existed firms, high HHI and large size"   = LfCEaYr[grepl("la", Pre2013Size), ] # 16
      , "Existed firms, low HHI and micro size"    = LfCEbYr[grepl("mi", Pre2013Size), ] # 17
      , "Existed firms, low HHI and small size"    = LfCEbYr[grepl("sm", Pre2013Size), ] # 18
      , "Existed firms, low HHI and medium size"   = LfCEbYr[grepl("me", Pre2013Size), ] # 19
      , "Existed firms, low HHI and large size"    = LfCEbYr[grepl("la", Pre2013Size), ] # 20
    )
    eslistYrw <- vector("list", length(subsetsYrw))
    names(eslistYrw) <- names(subsetsYrw)
    for (k in seq_along(subsetsYrw))
      eslistYrw[[k]] <- feols(
        rJobsMP ~ i(taxyear, Pre2013FAMP0, ref = 2012) |
          taxyear + taxrefno + busmainplc_geo,
        vcov = ~ taxyear + busmainplc_geo,
        data = subsetsYrw[[k]])
    rm(subsetsYrw)
    library(ggfixest)
    dictvec <- c(
        "JobsMP"         = "Number of jobs at the establishment"
      , "FAMP0"          = "Fraction affected in the taxyear 2012"
      , "taxrefno"       = "Firm ID"
      , "busmainplc_geo" = "Main Place"
    )
    setFixest_dict(dict = dictvec, reset = FALSE)
    #### estimation
    EstSpecs <- list(
        All     = 1:8
      , HHI     = 3:5
      , FAclass = c(3, 6:8)
      , Size    = c(3, 9:12)
      , SizeHHI = c(3, 13:20)
    )
    TitleSpecs <- c(
        "All"
      , "Low and high concentration"
      , "FA class in pre-2013"
      , "Size in pre-2013"
      , "Size and concentration in pre-2013"
    )
    names(TitleSpecs) <- names(EstSpecs)
    for (ii in 1:length(EstSpecs)) {
      ests <- eslistYrw[EstSpecs[[ii]]]
      estdfw <- etable(ests, 
        style.df = style.df(
          depvar.title = "Number of jobs", 
          fixef.title = "Fixed effects", 
          fixef.suffix = " FEs", 
          yesNo = "yes")
      )
      estltxw <- etable(ests, 
        style.tex = style.tex("aer"), 
        fitstat = ~ r2 + n, 
        tex = TRUE
      )
      gw <- ggiplot(ests,
        ref.line = 2012, 
        main = paste0("Job impacts: ", TitleSpecs[ii], ", winsorized data"),
        xlab = "Tax year (March - Februray)",
        multi_style = "facet",
        geom_style = "ribbon",
        theme = theme(
          legend.position = "none"
          )
      ) + 
      ThisTheme +
      scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
      #### CLAUDE tpo: dead code — Jb already "02" (character) from paste0(0,Jb) at L1044.
      #### if (Jb == 2) Jb <- paste0("0", Jb)
      ggsave(
        paste0(pathresults, 
          "Results", names(EstSpecs)[ii], Yr, "-2022Job", Jb, "w.jpg")
          , gw, width = 8*2, height = 4*2, units = "cm",
          dpi = 300
      )
      assign(paste0("est", names(EstSpecs)[ii], Yr, Jb, "w"), estdfw)
      assign(paste0("g",   names(EstSpecs)[ii], Yr, Jb, "w"), gw)
      qsave(estdfw, 
        paste0(pathresults, "est", names(EstSpecs)[ii], Yr, Jb, "w.qs"))
      write.tablev(estltxw, 
        paste0(pathresults, "est", names(EstSpecs)[ii], Yr, Jb, "w.tex"))
    }
    lapply(eslistYrw, summary)
    qsave(eslistYrw, paste0(pathresults, "eslist", Yr, Jb, "w.qs"))

####     estdfw <- etable(eslistYrw[1:8], 
####       style.df = style.df(
####         depvar.title = "Number of jobs", 
####         fixef.title = "Fixed effects", 
####         fixef.suffix = " FEs", 
####         yesNo = "yes")
####     )
####     #### by Size
####     estdfwSize <- etable(eslistYrw[c(3, 9:12)], 
####       style.df = style.df(
####         depvar.title = "Number of jobs", 
####         fixef.title = "Fixed effects", 
####         fixef.suffix = " FEs", 
####         yesNo = "yes")
####     )
####     estltxw <- etable(eslistYrw[1:8], 
####       style.tex = style.tex("aer"), 
####       fitstat = ~ r2 + n, 
####       tex = TRUE
####     )
####     estltxwSize <- etable(eslistYrw[c(3, 9:12)], 
####       style.tex = style.tex("aer"), 
####       fitstat = ~ r2 + n, 
####       tex = TRUE
####     )
####     gFEYrw <- ggiplot(eslistYrw[1:8],
####       ref.line = 2012, 
####       main = "Job impacts: Low and high concentration, winsorized data",
####       xlab = "Tax year (March - Februray)",
####       multi_style = "facet",
####       geom_style = "ribbon",
####       theme = theme(
####         legend.position = "none"
####         )
####     ) + 
####     ThisTheme +
####     scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
####     gFEYrwLH <- ggiplot(eslistYrw[2:5],
####       ref.line = 2012, 
####       main = "Job impacts: Low and high concentration, winsorized data",
####       xlab = "Tax year (March - Februray)",
####       multi_style = "facet",
####       geom_style = "ribbon",
####       theme = theme(
####         legend.position = "none"
####         )
####     ) + 
####     ThisTheme + 
####     scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
####     gFEYrwFAclass <- ggiplot(eslistYrw[c(3, 6:8)],
####       ref.line = 2012, 
####       main = "Job impacts: FA class in pre-2013, winsorized data",
####       xlab = "Tax year (March - Februray)",
####       multi_style = "facet",
####       geom_style = "ribbon",
####       theme = theme(
####         legend.position = "none"
####         )
####     ) + 
####     ThisTheme + 
####     scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
####     gFEYrwSize <- ggiplot(eslistYrw[c(3, 9:12)],
####       ref.line = 2012, 
####       main = "Job impacts: Size in pre-2013, winsorized data",
####       xlab = "Tax year (March - Februray)",
####       multi_style = "facet",
####       geom_style = "ribbon",
####       theme = theme(
####         legend.position = "none"
####         )
####     ) + 
####     ThisTheme + 
####     scale_x_continuous(limits = c(Yr, 2022), breaks = Yr:2022)
####     if (Jb == 2) Jb <- paste0("0", Jb)
####     ggsave(
####     paste0(pathresults, "Results", Yr, "-2022Job", Jb, "w.jpg")
####       , gFEYrw, width = 8*2, height = 4*2, units = "cm",
####       dpi = 300
####      )
####     ggsave(
####       paste0(pathresults, "Results", Yr, "-2022Job", Jb, "wLowHigh.jpg")
####       , gFEYrwLH, width = 8*2, height = 4*2, units = "cm",
####       dpi = 300
####      )
####     ggsave(
####       paste0(pathresults, "Results", Yr, "-2022Job", Jb, "wFAclass.jpg")
####       , gFEYrwFAclass, width = 8*2, height = 4*2, units = "cm",
####       dpi = 300
####      )
####     ggsave(
####       paste0(pathresults, "Results", Yr, "-2022Job", Jb, "wSize.jpg")
####       , gFEYrwSize, width = 8*2, height = 4*2, units = "cm",
####       dpi = 300
####      )
####     lapply(eslistYrw, summary)
####     assign(paste0("eslist", Yr, Jb, "w"),        eslistYrw)
####     assign(paste0("gFE",    Yr, Jb, "w"),        gFEYrw)
####     assign(paste0("gFE",    Yr, Jb, "wLH"),      gFEYrwLH)
####     assign(paste0("gFE",    Yr, Jb, "wFAclass"), gFEYrwFAclass)
####     assign(paste0("gFE",    Yr, Jb, "wSize"),    gFEYrwSize)
####     qsave(estdfw, 
####       paste0(pathresults, "est", Yr, Jb, "w.qs"))
####     qsave(estdfwSize, 
####       paste0(pathresults, "est", Yr, Jb, "wSize.qs"))
####     write.tablev(estltxw, 
####       paste0(pathresults, "est", Yr, Jb, "w.tex"))
####     write.tablev(estltxwSize, 
####       paste0(pathresults, "est", Yr, Jb, "wSize.tex"))
####     qsave(eslistYrw, paste0(pathresults, "eslist", Yr, Jb, "w.qs"))
  }
}
litedown::timing_data() 
```
::: {.fullwidth}
```{r 201002w results}
gAll201002w
gHHI201010w
gFAclass201010w
gSize201010w
gSizeHHI201010w
```
:::

<!-- files to be exported
latex table files for destat 
"Destat201002w_1.tex"
"Destat201002w_2.tex"
  "Destat201002w.tex"
"Destat201202w_1.tex"
"Destat201202w_2.tex"
  "Destat201202w.tex"
"Destat201010w_1.tex"
"Destat201010w_2.tex"
  "Destat201010w.tex"
"Destat201210w_1.tex"
"Destat201210w_2.tex"
  "Destat201210w.tex"

R table files for destat
"DESS.qs"
"DESSw.qs"

estimation result image files
"Results2010-2022Job02.jpg"
"Results2010-2022Job10.jpg"
"Results2012-2022Job02.jpg"
"Results2012-2022Job10.jpg"
"Results2010-2022Job02w.jpg"
"Results2010-2022Job10w.jpg"
"Results2012-2022Job02w.jpg"
"Results2012-2022Job10w.jpg"
-->

