---
title: "Minimum wage impacts on employment using IRP5 files"
date: "r format(Sys.time(), "%Y%m%d %R")"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 23/IRP5Impacts.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```


# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathsaveddata <- paste0(path, "saveddata/")
pathresults <- paste0(pathprogram, "results/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2); library(rmarkdown); library(fixest)
```





# Simple estimation



``{r select sample and estimate, eval = F}
Lf <- qread(paste0(pathsaveddata, "EstSample_Ag.qs"), nthreads = 13)
setkey(Lf, taxrefno, taxyear)
Lf[, ReportEveryYear := 0L]
Lf[, MaxDiffTY := max(taxyear, na.rm = T)-min(taxyear, na.rm = T)+1, 
  by = (taxrefno)]
Lf[, num := .N, by = (taxrefno)]
Lf[MaxDiffTY == num, ReportEveryYear := 1L]
#### Complete panel sample
LfC <- Lf[Jobs0 > 10 & DurMonthMarch > 0 & ReportEveryYear == 1L, ]
LfC[HHI <= median(HHI[taxyear == 2013], na.rm = T), HHILevel := "Below"]
LfC[HHI > median(HHI[taxyear == 2013], na.rm = T), HHILevel := "Above"]
LfC[, HHILevel := factor(HHILevel)]
LfCE <- LfC[ExistedBefore2013 == 1L, ]
LfCEb <- LfCE[grepl("Be", HHILevel), ])
LfCEa <- LfCE[grepl("Ab", HHILevel), ])
library(fixest)
esLf <- feols(Jobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf)
esLfC <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfC)
esLfCE <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCE)
#### Subsample: High and low concentrated markets
esLfCEa <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCEa)
esLfCEb <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = LfCEb)
library(ggfixest)
eslist <-  list(
    "All sample" = esLf, 
    "Complete panel" = esLfC, 
    "Existed firms, all sample" = esLfCE, 
    "Existed firms, high concentrated markets" = esLfCEa, 
    "Existed firms, low concentrated markets" = esLfCEb 
  )
gFE <- ggiplot(eslist,
  ref.line = 2012, 
  main = "Job impacts: Low and high concentration",
  xlab = "Tax year (March - Februray)",
  multi_style = "facet",
  geom_style = "ribbon",
  theme = theme(
    axis.text.x = element_text(angle = 90, vjust = .5),
    legend.position = "none"
    )
 
) + 
scale_x_continuous(limits = c(2010, 2022), breaks = 2010:2022) +
scale_y_continuous(limits = c(15, -10)) 
ggsave(
  paste0(pathresults, "Results2026Mar03.jpg")
  , gFE, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
gFE
lapply(eslist, summary)
```

