---
title: "Test litedown"
date: "`{r} format(Sys.time(), "%Y %b %d %R")`"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 21/TestRMD.rmd"))
#### setwd("c:/data/MinWageMarketPower/analysis/program/"); litedown::fuse("TestRMD.rmd")
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = T,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```

```{r }
set.seed(111)
x <- data.table(a = letters[1:10], b = 1:10, c = 1e+4+rpois(10, 10))
library(tinytable)
tb <- tt(x, caption = "test")
options(tinytable_format_digits = 2)
tb <- format_tt(tb,
    j = 3,
    num_mark_big = " "
  )
tb
```

```{r }
x[, d := T]
```

```{r export to itable}
library(fixest)
data(airquality)
#### Setting a dictionary 
setFixest_dict(c(
    Ozone = "Ozone (ppb)", 
  Solar.R = "Solar Radiation (Langleys)",
     Wind = "Wind Speed (mph)", 
     Temp = "Temperature"))
#### On multiple estimations: see the dedicated vignette
#### sw0: step wise addition Solar.R, Solar.R+Wind+Temp
#### csw: cumulative step wise addition Month, Month+Day
#### gives 2*2=4 estimations
est = feols(Ozone ~ Solar.R + sw0(Wind + Temp) | csw(Month, Day), 
  airquality, cluster = ~Day)                 
etable(est)
est_slopes = feols(Ozone ~ Solar.R + Wind | Day + Month[Temp], airquality)
etable(est, est_slopes, tex = TRUE)
etable(est, est_slopes, style.tex = style.tex("aer"), 
       signif.code = NA, fitstat = ~ r2 + n, tex = TRUE)
etable(est, 
  style.df = style.df(depvar.title = "", fixef.title = "", 
    fixef.suffix = " fixed effect", yesNo = "yes"))
#### library(pander)
```
```{r, eval = T}
#| results = "asis"
etable(est, postprocess.df = pandoc.table.return, style = "rmarkdown")
my_style = style.df(depvar.title = "", fixef.title = "", 
                    fixef.suffix = " fixed effect", yesNo = "yes")
setFixest_etable(style.df = my_style, 
  postprocess.df = pandoc.table.return)
etable(est[rhs = 2], style = "rmarkdown", 
  caption = "New default values")
```

```{r try out of fixest, eval = F, echo = T, warning = F}
library(fixest)
library(data.table)
library(Matrix)
set.seed(200)
x = rep(letters[1:4], 3)[1:20]
y = sample(1:4, 20, replace = T)
z = rep(0:1, each = 10)
i(x, y, reference = 2)
data(base_did)
base_did <- data.table(base_did)
dt <- data.table(base_did, base_did[, i(period, treat, 5)])
dt[, y := round(y, 2)]
dt[, x1 := round(x1, 2)]
Matrix(as.matrix(dt)[1:50, ], sparse = T)
#### printSpMatrix2(Matrix(as.matrix(dt), sparse = T))

dt[, Gr := 1L]
dt[id > 50, Gr := 2L]
Matrix(as.matrix(dt[, i(period, treat, 5)][1:50, ]), sparse = T)

summary(lm(y ~ i(treat)*i(period, 2)*Gr, data = dt))

data(airquality)
aq = airquality
aq$week = aq$Day %/% 7 + 1
aq <- data.table(aq)
aq[1:11, ]
#### Interacting Month and week:
res_2 = feols(Ozone ~ Solar.R + i(Month, week), aq) # week as numbers (time trend)
res_2F = feols(Ozone ~ Solar.R + i(Month, i.week), aq) # week as factor (dummy)


#### Same but dropping the 5th Month and 1st week
res_2F_bis = feols(Ozone ~ Solar.R + i(Month, i.week, ref = 5, ref2 = 1), aq)

etable(res_2F, res_2F_bis)

```
