---
title: "Test litedown"
date: "`{r} format(Sys.time(), "%Y %b %d %R")`"
output: html_document
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 21/TestRMD.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = T,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```



# Merge data

* FAD (FA, Jobs, Employment) + industry
* LS (HHI, WorkersAtEstablishment) + industry
* irp5Ma

```{r merge data}
#### LS + industry
inda <- qread(paste0(pathdata, "TaxrefnoOfAgri.qs"), nthreads = 13)
LSa <- qread(paste0(pathdata, "HHIAgriRows.qs"), nthreads = 13)
faa <- qread(paste0(pathdata, "FAOfAgri.qs"), nthreads = 13)

#### jobs are recorded poorly prior to 2013
LSa[, .(TotalJobs = sum(WorkersAtEstab, na.rm = T)), by = taxyear]
#### Need to set the population == 2012 firms
LSa[, num := 1:.N, by = .(taxyear, taxrefno)]
LSa[, ExistedBefore2013 := 0L]
LSa[taxrefno %in% taxrefno[taxyear[num==1L]<2013], ExistedBefore2013 := 1L]
lsa <- LSa[, .(TotalJobs = sum(WorkersAtEstab, na.rm = T)), 
  by = .(taxyear, ExistedBefore2013)]
lsa

```



# Employment



```{r plot number of jobs taxyear 2012-2020}
library(qs)
library(data.table)
library(ggplot2)
lsa[, ExistedBefore2013 := factor(ExistedBefore2013)]
g1 <- ggplot(data = lsa, aes(x = taxyear, y = TotalJobs, group = ExistedBefore2013, 
  colour = ExistedBefore2013, fill = ExistedBefore2013)) +
  geom_col() + 
  scale_color_viridis_d("Existing firms") +
  scale_fill_viridis_d() +
  scale_x_discrete("Tax year") + facet_grid(~ ExistedBefore2013) +
  theme_bw()
 g1
```


# Merge data

* FAD (FA, Jobs, Employment) + industry
   * FA is measured at the establishment level
   * `FAD`: Worker level data with establishment level aggregates
   * For estimation, need to aggregate at establishment level, only agriculture &rarr; 
   * `faa`: Collection of first rows of establishment-taxyear
* LS (HHI, WorkersAtEstab) + industry
   * HHI is measured at local market level
   * `LS`: Establishment level data with area level aggregates
   * For estimation, need to aggregate the worker characteristics at establishment level 
* irp5Ma (IRP5), only agriculture
   * All background information at worker level + establishment information

We use:

`faa` (establishment level FAA, Jobs, Employment) + `LSMa` (Main Place level HHI)



```{r merge data}
#### ind <- qread(paste0(pathdata, "IndustryCodeFromCIT_NoNAs.qs"), nthreads = 13)
#### setkey(irp5, taxrefno); setkey(ind, taxrefno, taxyear)
#### ind[, n := 1:.N, by = taxrefno]
#### inda <- ind[n == 1 & grepl("^Anim|^Plant pro|crops|Logging|forest", imp_mic_sic7_3d), ]
#### inda[, n := NULL]
#### inda[, c("Crop", "Animal", "Forest") := 0L]
#### inda[grepl("Plant|cr", imp_mic_sic7_3d), Crop := 1L]
#### inda[grepl("Anim", imp_mic_sic7_3d),   Animal := 1L]
#### inda[grepl("Lo|for", imp_mic_sic7_3d), Forest := 1L]
#### inda[, grepout("mp_mic", colnames(inda)) := NULL]
#### setkey(inda, taxrefno)
#### setnames(inda, "taxyear", "FirstYearOnSIC7")
#### qsave(inda, paste0(pathdata, "TaxrefnoOfAgri.qs"), nthreads = 13)
#### 
#### LSMa = LS + industry
inda <- qread(paste0(pathdata, "TaxrefnoOfAgri.qs"), nthreads = 13)
#### LS <- qread(paste0(pathdata, "LShareHHI.qs"), nthreads = 13)
#### LSia <- inda[LS, on = "taxrefno"]
#### LSa <- LSia[Crop ++ 1L | Animal == 1L | Forest == 1L, ]
#### #### LSa is rows of agri only, but HHI is measured at all industries
#### qsave(LSa, paste0(pathdata, "HHIAgriRows.qs"), nthreads = 13)
#### LSa <- qread(paste0(pathdata, "HHIAgriRows.qs"), nthreads = 13)
#### LSMa <- LSa[grepl("Mai", AreaLevel), ]
#### LSMa[, dJob := c(NA, diff(WorkersAtEstab)), by = .(busprov_geo, busdistmuni_geo, 
####   buslocmuni_geo, busmainplc_geo, taxrefno)]
#### #### jobs are recorded poorly prior to 2013
#### LSMa[, .(TotalJobs = sum(WorkersAtEstab, na.rm = T)), by = taxyear]
#### #### Need to set the population == 2012 firms
#### setorder(LSMa, taxyear, taxrefno)
#### LSMa[, num := 1:.N, by = .(taxyear, taxrefno)]
#### LSMa[, ExistedBefore2013 := as.integer(NA)]
#### LSMa[taxyear < 2013, ExistedBefore2013 := 1L]
#### LSMa[, ExistedBefore2013 := ExistedBefore2013[!is.na(ExistedBefore2013)][1], by = taxrefno]
#### LSMa[is.na(ExistedBefore2013), ExistedBefore2013 := 0L]
#### LSMa[, .(TotalJobs = sum(WorkersAtEstab, na.rm = T)), 
####   by = .(ExistedBefore2013, taxyear)]
#### qsave(LSMa, paste0(pathdata, "HHIAgriRowsMainPlaceLevel.qs"), nthreads = 13)


#### FAD + industry
#### Extract establishment level entries of FA
#### FAD <- qread(paste0(pathdata, "FAD.qs"), nthreads = 13)
#### table(FAD[, taxyear])
#### #### number by geo location, taxrefno, taxyear
#### FAD[, num := 1:.N, by = .(busprov_geo, busdistmuni_geo, 
####   buslocmuni_geo, busmainplc_geo, taxrefno, taxyear)]
#### fadata <- FAD[num == 1L, 
####   grepout("^bus|^tax[ry]|^Tx|Mon|[nN]um|Jo|Emp|FA", colnames(FAD)), with = F]
#### table(fadata[, taxyear])
####   2008   2009   2010   2011   2012   2013   2014   2015   2016   2017   2018   2019   2020   2021   2022 
####  23791  15641 168258  15521 243798 283160 299154 311918 323084 331015 341569 352993 364654 329671 309533 
#### #### Drop date errors (Negative and over duration)
#### fadata2 <- fadata[0 <= DurMonthMarch & DurMonthMarch <= 12, ]
#### nrow(fadata); nrow(fadata2) # 30K rows
#### #### Merge industry code, taking fadata as base
#### fadata3 <- inda[fadata2, on = "taxrefno"]
#### fadata3 <- fadata3[busmainplc_geo != "", ]
#### faa <- fadata3[Crop == 1L | Animal == 1L | Forest == 1L, ]
#### setkey(faa, taxrefno)
#### faa[, SubDiv := "Crop"]
#### faa[Animal == 1L, SubDiv := "Animal"]
#### faa[Forest == 1L, SubDiv := "Forest"]
#### faa[, SubDiv := factor(SubDiv, levels = c("Crop", "Animal", "Forest"))]
#### faa[, .(TotalJobs = sum(Jobs, na.rm = T)), by = .(taxyear, SubDiv)]
#### faa[, FApositive := 0L]
#### faa[taxyear < 2014, FApositive:= as.integer(FA > 0)]
#### faa[, FApositive := FApositive[FApositive == 1L][1], by = taxrefno]
#### faa[is.na(FApositive), FApositive := 0L]
#### faa[, Sample := "Unexposed"]
#### faa[FApositive == 1L, Sample := "Exposed"]
#### faa[, Sample := factor(Sample)]
#### #### #### faa is unique at establishment level
#### table(duplicated(faa[, .(taxrefno, taxyear, busmainplc_geo)]))
####  FALSE  TRUE 
####  59627     5 
#### qsave(faa, paste0(pathdata, "FAOfAgri.qs"), nthreads = 13)
faa <- qread(paste0(pathdata, "FAOfAgri.qs"), nthreads = 13)
faa[, .(mean = mean(FA, na.rm = T), median = median(FA, na.rm = T)), 
  by = .(Sample, taxyear)]
g2 <- ggplot(data = faa, aes(x = as.factor(taxyear), y = FA)) +
#  geom_point(size = .1, position = position_jitterdodge()) +
#  geom_point(data = faama, size = 1, fill = "blue")
  geom_boxplot(fill = "white", colour = "blue", outlier.alpha = 0) + 
#  geom_jitter(width = .2) +
  scale_x_discrete(name = "Tax year") +
  facet_grid(~ Sample) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
ggsave(
  paste0(path, "outfiles/2026 01 21/FA.jpg")
  , g2, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )

#### irp5M, agriculture
irp5Ma <- qread(paste0(pathdata, "irp5Ma.qs"), nthreads = 13)
irp5Ma[, .(num = 1:.N, Num), by = .(taxyear, taxrefno)][num == 1, 
  .(TotalJobs = sum(Num, na.rm = T)), by = taxyear]
irp5Ma[, .(num = 1:.N, Num = .N), by = .(taxyear, taxrefno)][num == 1, 
  .(TotalJobs = sum(Num, na.rm = T)), by = taxyear]
irp5Ma[, ExistedBefore2013 := as.integer(NA)]
irp5Ma[taxyear < 2013, ExistedBefore2013 := 1L]
irp5Ma[, ExistedBefore2013 := ExistedBefore2013[!is.na(ExistedBefore2013)][1], by = taxrefno]
irp5Ma[is.na(ExistedBefore2013), ExistedBefore2013 := 0L]
JobsFirms <- irp5Ma[, .(num = 1:.N, Num = .N), 
  by = .(ExistedBefore2013, taxyear, taxrefno)][
  num == 1, .(TotalJobs = sum(Num, na.rm = T)), by = .(ExistedBefore2013, taxyear)]
JobsFirms[, Sample := "Entrants"]
JobsFirms[ExistedBefore2013 == 1L, Sample := "Incumbents"]
g1 <- ggplot(data = JobsFirms, aes(x = taxyear, y = TotalJobs)) +
  geom_col() +
  scale_x_continuous(breaks = 2009:2022) +
  facet_grid(~ Sample) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5))
ggsave(
  paste0(pathdata, "Employment.jpg")
  , g1, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )


#### faa (base) + LSMa
faa <- qread(paste0(pathdata, "FAOfAgri.qs"), nthreads = 13)
faa[, dJobs := Jobs - mean(Jobs, na.rm = T), by = taxrefno]
faa[, FA0 := FA[1], by = taxrefno]
faa[, Jobs0 := Jobs[1], by = taxrefno]
faa[, rJobs := dJobs/Jobs0]
LSMa <- qread(paste0(pathdata, "HHIAgriRowsMainPlaceLevel.qs"), nthreads = 13)
LSMa1 <- LSMa[busmainplc_geo != "", ]
LSMa1[, num := 1:.N, by = .(busprov_geo, busdistmuni_geo, 
  buslocmuni_geo, busmainplc_geo, taxrefno)]
LSMa1 <- LSMa1[num == 1L, ]
Lf <- LSMa1[faa, on = c("taxrefno", grepout("bus", colnames(LSMa1)))]
Lf[, num := 1:.N, by = .(taxrefno, taxyear)]
Lf <- Lf[num == 1L & grepl("Mai", AreaLevel), ]
Lf[, ExistedBefore2013 := 0L]
Lf[taxyear < 2013, ExistedBefore2013 := 1L]
Lf[, ExistedBefore2013 := ExistedBefore2013[ExistedBefore2013 == 1L][1], by = taxrefno]
Lf[is.na(ExistedBefore2013), ExistedBefore2013 := 0L]
Lf[, FirstYear2013 := "Before"]
Lf[ExistedBefore2013 == 0L, FirstYear2013 := "Not before"]
Lf[, FirstYear2013 := factor(FirstYear2013)]
Lf[, Sample := "Exposed"]
Lf[FA0 == 0L, Sample := "Unexposed"]
Lf[, Sample := factor(Sample)]
Lf[, AgeSample := "Existed, exposed"]
Lf[grepl("Be", FirstYear2013) & grepl("Un", Sample),
  AgeSample := "Existed, unexposed"]
Lf[grepl("N", FirstYear2013) & grepl("Ex", Sample), 
  AgeSample := "New, exposed"]
Lf[grepl("N", FirstYear2013) & grepl("Un", Sample),
  AgeSample := "New, unexposed"]
Lf[, AgeSample := factor(AgeSample, levels = 
  c("Existed, exposed", "Existed, unexposed", "New, exposed", "New, unexposed"))]

  rbind(
    ExistedBefore2013 = table(Lf[ExistedBefore2013 == 1L, taxyear]),
    All = table(Lf[, taxyear])
  )


#### Jobs by firm age (existed prior to 2013 vs. new entrants after 2013)
g3 <- ggplot(data = Lf[, .(Jobs = sum(Jobs, na.rm= T)), by = .(FirstYear2013, taxyear)], 
   aes(x = taxyear, y = Jobs, 
     group = FirstYear2013, fill = FirstYear2013, colour = FirstYear2013)) +
  geom_col() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_x_continuous(breaks = 2009:2022) +
  facet_grid(FirstYear2013 ~ .) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5),
    legend.position = "none"
    )
ggsave(
  paste0(pathdata, "JobsByExistedBefore2013.jpg")
  , g3, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
#### Jobs by firm age * FA == 0
g4 <- 
ggplot(data = Lf[, .(Jobs = sum(Jobs, na.rm= T), AgeSample), by = .(FirstYear2013, Sample, taxyear)], 
   aes(x = taxyear, y = Jobs, group = AgeSample, fill = AgeSample, colour = AgeSample)) +
  geom_col() +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_x_continuous(limits = c(2009, 2022), breaks = 2009:2022) +
  facet_grid(FirstYear2013 ~ Sample) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5),
    legend.position = "none"
    )
ggsave(
  paste0(pathdata, "JobsByExistedBefore2013AndFA.jpg")
  , g4, width = 8*2, height = 4*2, units = "cm",
  dpi = 300
 )
```

```{r select nonanomalous sample}
Lf <- qread(paste0(pathdata, "EstSample_AllSectorHHI.qs"), nthreads = 13)
setkey(Lf, taxrefno, taxyear)
Lf[, ReportEveryYear := 0L]
Lf[, MaxDiffTY := max(taxyear, na.rm = T)-min(taxyear, na.rm = T)+1, by = (taxrefno)]
Lf[, num := .N, by = (taxrefno)]
Lf[MaxDiffTY == num, ReportEveryYear := 1L]

Lf1 <- Lf[Jobs0 > 10 & DurMonthMarch > 0 & ReportEveryYear == 1L, ]

Lf1[HHI <= median(HHI[taxyear == 2013], na.rm = T), HHILevel := "Below"]
Lf1[HHI > median(HHI[taxyear == 2013], na.rm = T), HHILevel := "Above"]
Lf1[, HHILevel := factor(HHILevel)]

library(fixest)
es1 <- feols(Jobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[ExistedBefore2013 == 1L, ])
es2 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[ExistedBefore2013 == 1L, ])
es3 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1)
#### Subsample: High and low concentrated markets
es4 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[ExistedBefore2013 == 1L & grepl("Be", HHILevel), ])
es5 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[ExistedBefore2013 == 1L & grepl("Ab", HHILevel), ])
es6 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[grepl("Be", HHILevel), ])
es7 <- feols(rJobs ~ i(taxyear, FA0, ref = 2012)|taxyear + taxrefno + busmainplc_geo, 
  vcov = ~ taxyear + busmainplc_geo,
  data = Lf1[grepl("Ab", HHILevel), ])
library(ggfixest)
eslist <-  list(
    "Existed firms, all sample" = es2, 
    "Existed firms, low concentrated markets" = es4, 
    "Existed firms, high concentrated markets" = es5 
  )
gFE <- ggiplot(eslist,
  ref.line = 2012, 
  main = "Job impacts: Low and high concentration",
  xlab = "Tax year (March - Februray)",
  multi_style = "facet",
  geom_style = "ribbon",
  theme = theme(
    axis.text.x = element_text(angle = 90, vjust = .5),
    legend.position = "none"
    )
 
) + 
scale_x_continuous(limits = c(2010, 2022), breaks = 2010:2022) 
gFE
lapply(eslist, summary)
```