---
title: "Test litedown"
date: "`{r} format(Sys.time(), "%Y %b %d %R")`"
output: 
  litedown::html_format:
    meta:
      css: ["@default", "@tabsets", "@copy-button"]
      js: ["@sidenotes", "@appendix", "@tabsets", "@copy-button"]
#### path <- "W:/epguest/seiro_ito/"; litedown::fuse(paste0(path, "outfiles/2026 01 21/TestRMD.rmd"))
#### setwd("c:/data/MinWageMarketPower/analysis/program/"); litedown::fuse("TestRMD.rmd")
---
```{css, echo = F}
body {
  background-color: #f5f5f5;
}
.body {
  background-color: transparent;
}
```
<style>
/* background color for the page */
body {
  background-color: #f0f8ff; 
}
/* 2. Background color ONLY for the R Source Code blocks */
pre code[class*="language-"] {
  background-color: #FAEDEA; /* A light brown box */
  color: black;            /* Light text */
  display: block;
  padding: 1px;
  border-radius: 1px;
}
/* 3. Background color ONLY for the R Output blocks */
pre code:not([class*="language-"]) {
  background-color: #ffffff; /* White background for the output */
  color: #333333;            /* Dark text */
  display: block;
  padding: 10px;
  border-left: 1px solid #007acc; /* Optional: A nice blue accent line on the left */
  border-radius: 4px;
}
/* 4. Force tables and all rows/cells to be white and remove transparency */
table {
  background-color: white;
}
</style>

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

::: {.side .side-left}
**Anything** on the left.
:::


# Setup

```{r, child='setup.Rmd', message = F, warning = F}
```


```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "T:/CIT-IRP5 Panel/"
pathdataIRP <- "T:/IRP5/Job level/v5/beta/"
```

```{r }
set.seed(111)
x <- data.table(a = letters[1:10], b = 1:10, c = 1e+4+rpois(10, 10))
library(tinytable)
tb <- tt(x, caption = "test")
options(tinytable_format_digits = 2)
tb <- format_tt(tb,
    j = 3,
    num_mark_big = " "
  )
tb <- style_tt(tb, background = "white") 
tb <- style_tt(tb, i = seq(2, nrow(x), by = 2), background = "#f2f2f2")
tb
```

```{r }
x[, d := T]
x
```

```{r time series plot}
library(ggplot2)
library(reshape2)
library(dplyr)
covid1 = read.csv(file="EUCOVIDdeaths.csv",header=TRUE)
covid_perCapita <- covid1[,-(1:2)] / covid1$PopulationM
covid_perCapita$Country <- covid1$Country
covid_perCapita_deaths <- melt(covid_perCapita,id.vars=c("Country"),
 value.name="value", variable.name="Day")
ggplot(data=covid_perCapita_deaths,
 aes(x=Day, y=value, group = Country, colour = Country)) + 
geom_line() +
labs(y= "Deaths per Capita", x = "Day")  + 
theme_bw(base_size = 16) + 
theme(axis.text.x=element_text(angle=60,hjust=1))+ 
ggtitle("Day-wise COVID-deaths per Capita in Europe in 2020")
```

```{r export to itable}
library(fixest)
data(airquality)
#### Setting a dictionary 
setFixest_dict(c(
    Ozone = "Ozone (ppb)", 
  Solar.R = "Solar Radiation (Langleys)",
     Wind = "Wind Speed (mph)", 
     Temp = "Temperature"))
#### On multiple estimations: see the dedicated vignette
#### sw0: step wise addition Solar.R, Solar.R+Wind+Temp
#### csw: cumulative step wise addition Month, Month+Day
#### gives 2*2=4 estimations
est = feols(Ozone ~ Solar.R + sw0(Wind + Temp) | csw(Month, Day), 
  airquality, cluster = ~Day)                 
etable(est)
est_slopes = feols(Ozone ~ Solar.R + Wind | Day + Month[Temp], airquality)
etable(est, est_slopes, tex = TRUE)
etable(est, est_slopes, style.tex = style.tex("aer"), 
       signif.code = NA, fitstat = ~ r2 + n, tex = TRUE)
etable(est, 
  style.df = style.df(depvar.title = "", fixef.title = "", 
    fixef.suffix = " fixed effect", yesNo = "yes"))
library(pander)
```
```{r, eval = T}
#| results = "asis"
etable(est, postprocess.df = pandoc.table.return, style = "rmarkdown")
my_style = style.df(depvar.title = "", fixef.title = "", 
                    fixef.suffix = " fixed effect", yesNo = "yes")
setFixest_etable(style.df = my_style, 
  postprocess.df = pandoc.table.return)
etable(est[rhs = 2], style = "rmarkdown", 
  caption = "New default values")
```

```{r tabset test}
xfun::tabset(airquality)
```

```{r try out of fixest, eval = T, echo = T, warning = F}
library(fixest)
library(data.table)
library(Matrix)
set.seed(200)
x = rep(letters[1:4], 3)[1:20]
y = sample(1:4, 20, replace = T)
z = rep(0:1, each = 10)
i(x, y, reference = 2)
data(base_did)
base_did <- data.table(base_did)
dt <- data.table(base_did, base_did[, i(period, treat, 5)])
dt[, y := round(y, 2)]
dt[, x1 := round(x1, 2)]
Matrix(as.matrix(dt)[1:50, ], sparse = T)
#### printSpMatrix2(Matrix(as.matrix(dt), sparse = T))

dt[, Gr := 1L]
dt[id > 50, Gr := 2L]
Matrix(as.matrix(dt[, i(period, treat, 5)][1:50, ]), sparse = T)

summary(lm(y ~ i(treat)*i(period, 2)*Gr, data = dt))

data(airquality)
aq = airquality
aq$week = aq$Day %/% 7 + 1
aq <- data.table(aq)
aq[1:11, ]
#### Interacting Month and week:
res_2 = feols(Ozone ~ Solar.R + i(Month, week), aq) # week as numbers (time trend)
res_2F = feols(Ozone ~ Solar.R + i(Month, i.week), aq) # week as factor (dummy)


#### Same but dropping the 5th Month and 1st week
res_2F_bis = feols(Ozone ~ Solar.R + i(Month, i.week, ref = 5, ref2 = 1), aq)

etable(res_2F, res_2F_bis)

```
