---
title: "Test litedown"
date: "`{r} format(Sys.time(), "%Y%m%d %R")`"
output: html_document
#### path <- "W:/epguest/seiro_ito/outfiles/"; litedown::fuse(paste0(path, "TestLitedown.rmd"))
#### path <- "C:/data/MinWageMarketPower/analysis/program/"; litedown::fuse(paste0(path, "TestLitedown.rmd"))
---
```{css, echo=F}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Lucida Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
#### Default height of a block
pre {
  max-height: 700px;
  overflow-y: auto;
}
```

# What we do in the project

1. Create merged panel using stata: "Data file" 
   * Use the same definitions as Michael's codes
   * But include `imp_mic_sic7_3d == 16` industry
   * Done in other do file
   * Save Data file in tsv format to read in R
1. Compute HHI using CIT-IRP5 at various levels: "HHI file"
   * Use CIT-IRP5_2012 HHI in 2012
   * Verify if counts are not erroneous 
1. "Analysis file" = Data file (panel 2009-2022) + HHI file
   * Merge at various levels 
   * Need main place level analysis for larger statistical power
1. Descriptive statistic
1. Hypothesis testing
   * % Employment change = a + b1*FA + b2*HHI + b3*(FA*HHI) + FEs + e
   * Test b3 > 0 labor market power mitigates disemployment impacts
   * Expect b1<0 or b1 = 0
   * No prior for the sign of b2 

# Setup

```{r setup}
library(knitr)
library(tufte)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  #### remove leading hashes in html output
  comment = "", 
  echo = T, cache = F, eval = F, message = F,
  class.source = "SeiroBenign", class.output = "SeiroLightGreen"
  )
options(htmltools.dir.version = FALSE, width = 100)
grepout <- function(str, x)
  # returns element of match (not numbers)
  x[grep(str, x, perl = T)]
table0 <- function(x) table(x, useNA = "ifany")
```
```{r path, eval = T}
path <- "W:/epguest/seiro_ito/"
pathprogram <- paste0(path, "outfiles/")
pathdata <- paste0(path, "data/")
pathdataCITIRP <- "U:/CIT-IRP5 Panel/"
pathdataIRP <- "U:/IRP5/Job level/v5/beta/"
library(qs); library(data.table); library(ggplot2)
```





# Employment



```{r plot number of jobs taxyear 2012-2020}
library(qs)
library(data.table)
library(ggplot2)
x <- 1:10
x
#### g1 <- ggplot(data = lsa, aes(x = TotalJobs, group = ExistedBefore2013, 
####   colour = ExistedBefore2013, fill = ExistedBefore2013)) +
####   geom_col() + 
####   scale_color_viridis_d("Existing firms") +
####   scale_fill_viridis_d() +
####   scale_x_discrete("Tax year") + facet_grid(~ ExistedBefore2013) +
####   theme_bw()
####  g1
```
