---
title: "GNR"
subtitle: ""
author: "Seiro Ito"
date: "`r format(Sys.time(), '%Y年%m月%d日 %R')`"
header_includes:
    - \usepackage{tikz}
    - \usepackage{adjustbox}
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, default-fonts, xaringan-themer.css]
    nature:
      highlightStyle: github
      countIncrementalSlides: false
      ratio: "16:9"
    citation_package: natbib
# path <- "c:/seiro/data/MinWageMarketPower/"; rmarkdown::render("docs/GNR_slides.Rmd")
# renderthis::to_pdf("docs/GNR_slides.Rmd")
# git add /docs
# git commit -am "initial commit"
# git config --global user.name "SeiroIto"
---
class: inverse
```{r xaringan setup, include=FALSE, warning=FALSE}
xaringanExtra::use_panelset()
library(xaringanthemer)
#style_mono_dark(base_color = "#cbf7ed")
#style_mono_light(base_color = "#23395b")
style_solarized_light(text_color = "black", title_slide_background_color = "darkblue",
  link_color = "blue")
```
```{css echo=FALSE}
.highlight-last-item > ul > li,
.highlight-last-item > ol > li {
  opacity: 0.5;
}
.highlight-last-item > ul > li:last-of-type,
.highlight-last-item > ol > li:last-of-type {
  opacity: 1;
}
.inverse {
  background-color: #272822;
  color: #d6d6d6;
  text-shadow: 0 0 20px #333;
}
```
## For gross production functions, parameters are not indentified: 

## When using profit maximization FOC in material demand inversion and using IVs to estimate other parameters.  

## Material demand variations (say, by input prices which may or maynot be observed), can help identification but weak.

@GNR2020 


---
	## What does it mean if we want to estimate wage markdown using financial statement data?

* You need to use only cost minimization FOCs.  
* But markdown derived from cost minimization FOC generally does not match with markdown derived from profit maximization FOC. 


---
.pull-left[
See how the right box is going down

so down.
]

.pull-right[
```{r}
y <- data.frame(A = LETTERS[1:5],
            B = 1:5,
            C = sqrt(6:10))
```
]
