---
title: "A summary on labour market power literature"
author: "Noreen"
date: "`r format(Sys.time(), '%Y-%m-%d- %R')`"
output:
  tufte::tufte_html:
    citation_package: natbib
  pdf_document:
     latex_engine: xelatex
header-includes:
    - \usepackage{amsmath}
    - \usepackage{amssymb}
    - \usepackage{amsfonts}
    - \usepackage{bookmark} 
bibliography: bibfile.bib
link-citations: yes
#### setwd(path <- "Z:/Workbenches/epguest/seiro_ito/outfiles/"); rmarkdown::render("RmarkdownTemplate.Rmd")
#### rmarkdown::render("RmarkdownTemplate.Rmd", output_format = "all")
---
```{css, echo=FALSE}
.SeiroBenign {
  background-color: #FFEBCD;
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
	}
.SeiroLightGreen {
  background-color: #D0F0C0; /* Tea green */
  padding: 0.5em; /*文字まわり（上下左右）の余白*/
  font-family: Noto Sans;
  /* border: 1px solid yellow; */
  /* font-weight: bold; */
}
```

```{r setup, include=FALSE} 
#### include = F <==> echo=F & results = F
library(tufte)
library(data.table)
library(qs)
#### invalidate cache when the tufte version changes
knitr::opts_chunk$set(
  tidy = FALSE, cache.extra = packageVersion('tufte'), 
  margin_references = TRUE,
  echo = T, 
  #### remove leading hashes in html output
  comment = "", 
  class.source = "SeiroBenign", class.output = "SeiroLightGreen")
options(htmltools.dir.version = FALSE)
```
\def\Perp{\mkern2mu\rotatebox[origin=c]{90}{$\models$}\mkern2mu}
\newcommand{\mpage}[2]{\begin{minipage}[t]{#1}#2\end{minipage}}
\newcommand{\cnvp}{\stackrel{p}{\longrightarrow}}
\newcommand{\cnvd}{\stackrel{d}{\longrightarrow}}
\newcommand{\bfV}{\mathbf{V}}
\newcommand{\bfX}{\mathbf{X}}
\newcommand{\bfx}{\mathbf{x}}
\newcommand{\bfa}{\mathbf{a}}
\newcommand{\bfb}{\mathbf{b}}
\newcommand{\bfp}{\mathbf{p}}
\newcommand{\bfs}{\mathbf{s}}
\newcommand{\bfz}{\mathbf{z}}
\newcommand{\E}{{\Large\varepsilon}}
\newcommand{\NU}{{\Large\nu}}
\newcommand{\0}{{\mathbf{0}}}
\newcommand{\bfalpha}{\boldsymbol{\alpha}}
\newcommand{\bfbeta}{\boldsymbol{\beta}}
\newcommand{\bftheta}{\boldsymbol{\theta}}
\newcommand{\bfeta}{\boldsymbol{\eta}}
\newcommand{\bfmu}{\boldsymbol{\mu}}
\newcommand{\bflambda}{\boldsymbol{\lambda}}
\newcommand{\ind}{\mathrel{\perp\!\!\!\!\perp}}

# This is header level 1
## This is header level 2

```{r data exercise}

```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(x = "rmarkdown", file = 'RmarkdownTemplate.bib')
```

