---
title: "Wage markdown"
subtitle: ""
author: "Seiro Ito"
date: "`r format(Sys.time(), '%Y年%m月%d日 %R')`"
header_includes:
    - \usepackage{tikz}
    - \usepackage{adjustbox}
    - \usepackage{seiro_minimal}
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, default-fonts, xaringan-themer.css]
    nature:
      highlightStyle: github
      countIncrementalSlides: false
      ratio: "16:9"
#    includes:
#      before_body: mathabbreviation.html
    citation_package: natbib
bibliography: c:/seiro/settings/Tex/seiro.bib
self_contained: true
# setwd(path <- "c:/data/MinWageMarketPower/"); rmarkdown::render("docs/WageMarkdown_slides.Rmd")
# renderthis::to_pdf("docs/WageMarkdown_slides.Rmd")
# git init
# git add /docs    ## this adds a folder to track
# git commit -am "initial commit"
# git config --global user.name "SeiroIto"
# git remote add origin https://github.com/SeiroIto/MinWageMarketPower.git  ## specify remote
# git push -f origin main  ## push files to remote
---
class: inverse
```{r xaringan setup, include=FALSE, warning=FALSE}
xaringanExtra::use_panelset()
library(xaringanthemer)
#style_mono_dark(base_color = "#cbf7ed")
#style_mono_light(base_color = "#23395b")
style_solarized_light(text_color = "black", title_slide_background_color = "darkblue",
  link_color = "blue")
# xaringancolor setup
library(xaringancolor)
xaringancolor::setup_colors(
  red ="red",
  green = "green",
  blue = "blue"
)
library(RefManageR)
```
```{r xaringan bib, include=FALSE, cache=T}
# bibliography setup
BibOptions(check.entries = FALSE, style = "markdown", cite.style = "authoryear",
  bib.style = "authoryear", dashed = T, hyperlink = "to.doc")
## (old) Uncomment below for the 1st run of each R session. 2nd run will get cached results.
## (new) Delete cache if you add an entry to seiro.bib.
myBib <- ReadBib("c:/seiro/settings/TeX/seiro.bib")
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{red}{rgb}{1, 0, 0}$$
$$\require{color}\definecolor{green}{rgb}{0, 1, 0}$$
$$\require{color}\definecolor{blue}{rgb}{0, 0, 1}$$
</div>
<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      red: ["{\\color{red}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1],
      blue: ["{\\color{blue}{#1}}", 1]
      E: "{\\Large\\varepsilon}",
      bfx: "{\\mathbf{x}}",
      bfX: "{\\mathbf{X}}",
      bfbeta: "{\\boldsymbol{\\beta}}",
      st: "{\\mbox{s.t.}}"
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>
<style>
.red {color: #FF0000;}
.green {color: #00FF00;}
.blue {color: #0000FF;}
</style>
```{r flair_color, echo=FALSE}
library(flair)
red <- "#FF0000"
green <- "#00FF00"
blue <- "#0000FF"
```
```{css echo=FALSE}
.highlight-last-item > ul > li,
.highlight-last-item > ol > li {
  opacity: 0.5;
}
.highlight-last-item > ul > li:last-of-type,
.highlight-last-item > ol > li:last-of-type {
  opacity: 1;
}
.inverse {
  background-color: #272822;
  color: #d6d6d6;
  text-shadow: 0 0 20px #333;
}
```

`r Citet(myBib, "HashemiKirovTraina2022")` starts with labor wedge $\tau_{i}\in\mathbb R$ in price-taker cost minimization:
\begin{aligned}
\min_{L_{i}, M_{i}}
&\;\;\; (1+\tau_{i})P^{W}_{i}L_{i}+P^{M}_{i}M_{i}\\
\st &\;\;\; Q^{*}_{i}\leqslant F_{i}(L_{i}, M_{i})e^{\omega_{i}+\epsilon_{i}}.
\end{aligned}
\begin{equation*}
\mathcal L=(1+\tau_{i})P^{W}_{i}L_{i}+P^{M}_{i}M_{i}+\lambda_{i}\left[Q^{*}_{i}-F_{i}(L_{i}, M_{i})e^{\omega_{i}+\epsilon_{i}}\right].
\end{equation*}
FOCs:
\begin{equation*}
(1+\tau_{i})P^{W}_{i}
=
\lambda_{i}F_{iL}\E\left[e^{\omega_{i}+\epsilon_{i}}\right],\quad
P^{M}_{i}
=
\lambda_{i}F_{iM}\E\left[e^{\omega_{i}+\epsilon_{i}}\right].
\end{equation*}
Given $\lambda_{i}=MC_{i}$, we get markup $\mu_{i}\equiv P_{i}/MC_{i}$:
\begin{aligned}
\mu_{i}
&=
\frac{P_{i}}{\frac{P^{M}_{i}}{F_{iM}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}}
=
\frac{P_{i}F_{i}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}{P^{M}_{i}M_{i}}
\frac{M_{i}F_{iM}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}{F_{i}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}
=
\frac{R_{i}}{P^{M}_{i}M_{i}}
\frac{\partial F_{i}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}{\partial M_{i}}\frac{M_{i}}{{F_{i}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}}
=
\frac{\theta^{M}_{i}}{\alpha^{M}_{i}},\\
\mu_{i}
&=
\frac{P_{i}}{(1+\tau_{i})\frac{P^{L}_{i}}{F_{iL}\E\left[e^{\omega_{i}+\epsilon_{i}}\right]}}
=
(1+\tau_{i})\frac{\theta^{L}_{i}}{\alpha^{L}_{i}}.
\end{aligned}
---
The given wedge $\tau_{i}$ is passed over to the markup in price-taker cost minimization. 

--

### Why taking the wedge as given? This is OK if there are frictions in market transactions. But not OK if the firms can affect prices. 

--

### Isn't the wedge an outcome of monopsonistic pricing? 

* `r Citet(myBib, "Rubens2023")` derives an estimable markdown formula for a monopsonist. 

---
